---
title: "Turnover covariate data prep and visualization"
output:
  github_document: default
---

```{r setup}
library(data.table)
library(ggplot2)
library(beanplot) # for beanplots
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
require(scales) # for custom axis scales


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.

signedsqrt_trans <- function() trans_new('signedsqrt', 
                                         transform = function(x) sign(x)*sqrt(abs(x)), 
                                         inverse = function(x) sign(x)*x^2)
```

# Load data
```{r load data}
# biotime taxa category and other info
load('data/biotime_blowes/bt_malin.Rdata') # load bt_malin
btinfo <- data.table(bt_malin); rm(bt_malin)
btinfo <- btinfo[!duplicated(rarefyID), .(rarefyID, rarefyID_x, rarefyID_y, Biome, taxa_mod, REALM, STUDY_ID)]

# biotime community dissimilarity data
load('data/biotime_blowes/all_pairs_beta.RData') # load rarefied_beta_medians
bt <- data.table(rarefied_beta_medians); rm(rarefied_beta_medians)
bt[, year1 := as.numeric(year1)] # not sure why it gets read in as character
bt[, year2 := as.numeric(year2)]
bt[, Horn := 1- Hornsim] # convert similarity to dissimilarity
bt[, Hornsim := NULL]
bt <- merge(bt, btinfo[, .(rarefyID, rarefyID_x, rarefyID_y, REALM, STUDY_ID, Biome, taxa_mod)]) # add lat/lon and realm

# Temperature average, changes, and seasonality
temperature <- fread('output/temperature_byrarefyID.csv.gz')

# microclimates
microclim <- fread('output/microclimates.csv.gz', drop = 1)

# NPP
npp <- fread('output/npplandocean.csv.gz')

# Body size
bs <- fread('output/mass_byrarefyid.csv.gz', drop = 1)
bs[, ':='(STUDY_ID = NULL, REALM = NULL, taxa_mod = NULL)] # remove unnecessary columns 

# Mobility
speed <- fread('output/speed_byrarefyID.csv.gz', drop = 1)
speed[, ':='(STUDY_ID = NULL, REALM = NULL, taxa_mod = NULL)] # remove unnecessary columns 

# Lifespan
lsp <- fread('output/lifespan_byrarefyID.csv.gz')

# CTI
cti <- fread('output/cti_byrarefyID.csv.gz')
    
# consumer vs. producer
consfrac <- fread('output/consfrac_byrarefyID.csv.gz')

# richness
rich <- fread('output/richness_by_rarefyID.csv.gz') # number of species

# endotherm vs. ectotherm
endofrac <- fread('output/endofrac_byrarefyID.csv.gz') # endotherm vs. ectotherm classifications

# human impact
human <- fread('output/humanimpact_by_rarefyID.csv.gz')

# %veg
veg <- as.data.table(readRDS('output/vct_by_rarefyID.rds'))
veg[, veg := (`tree cover % (mean)` + 0.5 * `non-tree veg. % (mean)`)/100] # veg index from 0 (all non-veg) to 1 (all tree). Non-tree veg counts as 0.5.
```


## Add covariates to BT data
```{r add covariate data}
# add covariates
bt <- merge(bt, temperature, all.x = TRUE, by = c('rarefyID', 'year1', 'year2')) # temperature ave, ave metabolic, change, and seasonality
bt <- merge(bt, microclim[, .(rarefyID, microclim = Temp_sd20km)], all.x = TRUE, by = 'rarefyID') # microclimates
bt <- merge(bt, npp, all.x = TRUE, by = 'rarefyID') # npp
bt <- merge(bt, bs[, .(rarefyID, mass_mean_weight, mass_sd_weight)], all.x = TRUE) # body size mass (g)
bt <- merge(bt, speed[, .(rarefyID, speed_mean_weight, speed_sd_weight)], all.x = TRUE) # speed (km/hr)
bt <- merge(bt, lsp[, .(rarefyID, lifespan_mean_weight, lifespan_sd_weight)], all.x = TRUE) # lifespan (yr)
bt <- merge(bt, cti[, .(rarefyID, thermal_bias)], all.x = TRUE) # thermal bias (degC)
bt <- merge(bt, consfrac[, .(rarefyID, consfrac)], all.x = TRUE) # fraction consumers
bt <- merge(bt, rich, all.x = TRUE) # species richness
bt <- merge(bt, endofrac[, .(rarefyID, endofrac)], all.x = TRUE) # endotherm vs. ectotherm
bt <- merge(bt, human[, .(rarefyID, human_bowler = atc, human_venter = hfp, human_halpern = himp)], all.x = TRUE) # human impact
bt <- merge(bt, veg[, .(rarefyID, veg = veg)], all.x = TRUE) # vegetation index
bt[REALM == 'Marine', veg := 0] # veg index is 0 at sea
```


## Write out
Only if file doesn't yet exist
```{r write out}
if(!file.exists('output/turnover_w_covariates.csv.gz')){
  write.csv(bt, gzfile('output/turnover_w_covariates.csv.gz'), row.names = FALSE)
}

```



# A bit more prep for visualizing covariate distributions
## Make a data.table summarized by rarefyID
```{r summarize by rarefyID}
btave <- bt[, .(Jtu = mean(Jtu), minyrBT = min(year1), maxyrBT = max(year2), nyrBT = length(Jtu) + 1,
                REALM = unique(REALM), tempchange = mean(tempchange), tempave = mean(tempave),
                tempave_metab = mean(tempave_metab), seas = mean(seas), microclim = mean(microclim),
                mass_mean_weight = mean(mass_mean_weight), speed_mean_weight = mean(speed_mean_weight),
                lifespan_mean_weight = mean(lifespan_mean_weight), consfrac = mean(consfrac),
                endofrac = mean(endofrac), Nspp = mean(Nspp), thermal_bias = mean(thermal_bias),
                npp = mean(npp), veg = mean(veg), human_bowler = mean(human_bowler),
                human_venter = mean(human_venter), human_halpern = mean(human_halpern)),  by = rarefyID]
```


## Add useful variables
```{r useful vars}
# realm that combined Terrestrial and Freshwater, for interacting with human impact
btave[, REALM2 := REALM]
levels(btave$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")
```

## Try a logit-transform of the response variables
```{r logit transform}
logit = function(x) return(log(x/(1-x)))

# have to adjust variables away from 0 and 1
jtumin = bt[Jtu > 0, min(Jtu)]
jtumax = bt[Jtu < 1, max(Jtu)]
bt[Jtu > 0 & Jtu < 1, Jtulogit := logit(Jtu)]
bt[Jtu == 0, Jtulogit := logit(jtumin)]
bt[Jtu == 1, Jtulogit := logit(jtumax)]

jbetamin = bt[Jbeta > 0, min(Jbeta)]
jbetamax = bt[Jbeta < 1, max(Jbeta)]
bt[Jbeta > 0 & Jbeta < 1, Jbetalogit := logit(Jbeta)]
bt[Jbeta == 0, Jbetalogit := logit(jbetamin)]
bt[Jbeta == 1, Jbetalogit := logit(jbetamax)]

hornmin = bt[Horn > 0, min(Horn)]
hornmax = bt[Horn < 1, max(Horn)]
bt[Horn > 0 & Horn < 1, Hornlogit := logit(Horn)]
bt[Horn == 0, Hornlogit := logit(hornmin)]
bt[Horn == 1, Hornlogit := logit(hornmax)]

bt[, summary(Jtulogit)]
bt[, summary(Jbetalogit)]
bt[, summary(Hornlogit)]

```

## Log-transform some explanantory variables, then center and scale. 
```{r center and scale}
btave[, tempave.sc := scale(tempave)]
btave[, tempave_metab.sc := scale(tempave_metab)]
btave[, seas.sc := scale(seas)]
btave[, microclim.sc := scale(log(microclim))]
bt[, tempchange.sc := scale(tempchange, center = FALSE)] # all the raw data
bt[, tempchange_abs.sc := scale(abs(tempchange), center = FALSE)] # do not center, so that 0 is still 0 temperature change
btave[, tempchange.sc := scale(tempchange, center = FALSE)] # averages within studies
btave[, tempchange_abs.sc := scale(abs(tempchange), center = FALSE)]
btave[, mass.sc := scale(log(mass_mean_weight))]
btave[, speed.sc := scale(log(speed_mean_weight+1))]
btave[, lifespan.sc := scale(log(lifespan_mean_weight))]
btave[, consumerfrac.sc := scale(consfrac)]
btave[, endothermfrac.sc := scale(endofrac)]
btave[, nspp.sc := scale(log(Nspp))]
btave[, thermal_bias.sc := scale(thermal_bias)]
btave[, npp.sc := scale(log(npp))]
btave[, veg.sc := scale(log(veg+1))]
btave[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
btave[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
btave[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]
```


# Check variable distributions
## Response variables
```{r histograms response}
bt[, summary(Jbeta)]
bt[, summary(Jtu)]
bt[, summary(Horn)]

# fraction 0 or 1
bt[, sum(Jbeta == 0)/.N]
bt[, sum(Jtu == 0)/.N]
bt[!is.na(Horn), sum(Horn == 0)/.N]

bt[, sum(Jbeta == 1)/.N]
bt[, sum(Jtu == 1)/.N]
bt[!is.na(Horn), sum(Horn == 1)/.N]

# histograms
invisible(bt[, hist(Jbeta, main = 'Jaccard total', breaks = 80)])
invisible(bt[, hist(Jtu, main = 'Jaccard turnover', breaks = 80)])
invisible(bt[, hist(Horn, main = 'Morisita-Horn', breaks = 80)])

invisible(bt[, hist(Jbetalogit, main = 'logit(Jaccard total)', breaks = 80)])
invisible(bt[, hist(Jtulogit, main = 'logit(Jaccard turnover)', breaks = 80)])
invisible(bt[, hist(Hornlogit, main = 'logit(Morisita-Horn)', breaks = 80)])


```

## Unscaled covariates
```{r histograms unscaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btave[, hist(minyrBT, main = 'Start year', cex.main = cexmain)])
invisible(btave[, hist(maxyrBT - minyrBT, main = 'Duration (years)', cex.main = cexmain)])
invisible(btave[, hist(nyrBT, main = 'Number of sampled years', cex.main = cexmain)])
invisible(btave[, hist(mass_mean_weight, main = 'Mass (g)', cex.main = cexmain)])
invisible(btave[, hist(speed_mean_weight, main = 'Speed (km/hr)', cex.main = cexmain)])
invisible(btave[, hist(lifespan_mean_weight, main = 'Lifespan (yr)', cex.main = cexmain)])
invisible(btave[, hist(tempave_metab, main = 'Metabolic temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(consfrac, main = 'Consumers (fraction)', cex.main = cexmain)])
invisible(btave[, hist(endofrac, main = 'Endotherms (fraction)', cex.main = cexmain)])
invisible(btave[, hist(tempave, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(bt[, hist(tempchange, main = 'Temperature trend (°C/yr)', cex.main = cexmain)]) # all the raw data
invisible(btave[, hist(seas, main = 'Seasonality (°C)', cex.main = cexmain)])
invisible(btave[, hist(microclim, main = 'Microclimates (°C)', cex.main = cexmain)])
invisible(btave[, hist(Nspp, main = 'Species richness', cex.main = cexmain)])
invisible(btave[, hist(thermal_bias, main = 'Thermal bias (°C)', cex.main = cexmain)])
invisible(btave[, hist(npp, main = 'Net primary productivity', cex.main = cexmain)])
invisible(btave[, hist(veg, main = 'Vegetation index', cex.main = cexmain)])
invisible(btave[, hist(human_bowler, main = 'Human impact score (Bowler)', cex.main = cexmain)])
invisible(btave[, hist(human_venter, main = 'Human impact score (Venter)', cex.main = cexmain)])
invisible(btave[, hist(human_halpern, main = 'Human impact score (Halpern)', cex.main = cexmain)])

```

## Scaled covariates
```{r histograms scaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btave[, hist(tempave.sc, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(tempave_metab.sc, main = 'Metabolic temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(seas.sc, main = 'Seasonality (°C)', cex.main = cexmain)])
invisible(btave[, hist(microclim.sc, main = 'log Microclimates (°C)', cex.main = cexmain)])
invisible(bt[, hist(tempchange.sc, main = 'Temperature trend (°C/yr)', cex.main = cexmain)])
invisible(bt[, hist(tempchange_abs.sc, main = 'abs(Temperature trend) (°C/yr)', cex.main = cexmain)])
invisible(btave[, hist(mass.sc, main = 'log Mass (g)', cex.main = cexmain)])
invisible(btave[, hist(speed.sc, main = 'log Speed (km/hr)', cex.main = cexmain)])
invisible(btave[, hist(lifespan.sc, main = 'log Lifespan (yr)', cex.main = cexmain)])
invisible(btave[, hist(consumerfrac.sc, main = 'Consumers (fraction)', cex.main = cexmain)])
invisible(btave[, hist(endothermfrac.sc, main = 'Endotherms (fraction)', cex.main = cexmain)])
invisible(btave[, hist(nspp.sc, main = 'log Species richness', cex.main = cexmain)])
invisible(btave[, hist(thermal_bias.sc, main = 'Thermal bias (°C)', cex.main = cexmain)])
invisible(btave[, hist(npp.sc, main = 'log Net primary productivity', cex.main = cexmain)])
invisible(btave[, hist(veg.sc, main = 'log Vegetation index', cex.main = cexmain)])
invisible(btave[, hist(human_bowler.sc, main = 'log Human impact score (Bowler)', cex.main = cexmain)])
invisible(btave[, hist(human_footprint.sc, main = 'log Human impact score (Venter & Halpern)', cex.main = cexmain)])

```



# Check correlations among variables. Pearson's r is on the lower diagonal.
```{r pairs, fig.height=10, fig.width=10}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y, use = 'pairwise.complete.obs')
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt) #, cex = cex.cor * r)
}
pairs(formula = ~ tempave.sc + tempave_metab.sc + seas.sc + microclim.sc + tempchange.sc + tempchange_abs.sc + mass.sc + speed.sc + lifespan.sc + consumerfrac.sc + endothermfrac.sc + nspp.sc + thermal_bias.sc + npp.sc + veg.sc + human_bowler.sc + human_footprint.sc, data = btave, gap = 1/10, cex = 0.2, col = '#00000022', lower.panel = panel.cor)

```

Mass and lifespan look tightly correlated, but r only 0.56...?
Tempave_metab and lifespan don't look tightly correlated, but r= -0.81 
Tempave_metab and speed don't look tightly correlated, but r= -0.83 
Lifespan and speed don't look tightly correlated, but r = 0.73

# Compare covariates across realms
```{r compare across realms, fig.height=12, fig.width=9}
i <- bt[, !duplicated(rarefyID)]; sum(i)
par(mfrow=c(5,3))
beanplot(rarefyID_y ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Latitude (degN)', ll = 0.05)
beanplot(tempave ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature (degC)', ll = 0.05)
beanplot(tempave_metab ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Metabolic Temperature (degC)', ll = 0.05, bw = 'nrd0') # nrd0 bandwidth to calculation gap
beanplot(seas ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Seasonality (degC)', ll = 0.05)
beanplot(microclim ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Microclimates (degC)', ll = 0.05)
beanplot(tempchange ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature change (degC)', ll = 0.05)
beanplot(mass_mean_weight ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Mass (g)', ll = 0.05, log = 'y')
beanplot(speed_mean_weight +1 ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Speed (km/hr)', ll = 0.05, log = 'y')
beanplot(lifespan_mean_weight ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Lifespan (yr)', ll = 0.05, log = 'y')
#beanplot(consfrac ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Consumers (fraction)', ll = 0.05, log = '') # too sparse
#beanplot(endofrac ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Endotherms (fraction)', ll = 0.05, log = '') # too sparse
beanplot(Nspp ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Number of species', ll = 0.05, log = 'y')
beanplot(thermal_bias ~ REALM, data = bt[i & !is.na(thermal_bias),], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Thermal bias (degC)', ll = 0.05)
beanplot(npp ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)
beanplot(veg ~ REALM, data = bt[i & REALM !='Marine',], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)

```

Marine are in generally warmer locations (seawater doesn't freeze)
Marine have much lower seasonality.
Marine and freshwater have some very small masses (plankton), but much of dataset is similar to terrestrial.
Marine has a lot of slow, crawling organisms, but land has plants. Land also has birds (fast).

# Frequency of time differences
```{r plot time differences}
invisible(bt[, hist(year2 - year1, breaks = seq(0.5, 120.5, by = 1))])
```

# Plot dissimilarity vs. time
```{r plot diss vs time, echo=FALSE, fig.height = 3, fig.width = 11}
p1 <- ggplot(bt, aes(year2 - year1, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Jaccard turnover')

p2 <- ggplot(bt, aes(year2 - year1, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Jaccard total')

p3 <- ggplot(bt, aes(year2 - year1, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Morisita-Horn')



grid.arrange(p1, p2, p3, ncol = 3)
```

# Plot dissimilarity vs. temperature change
## All temporal scales
```{r plot diss vs temp change, echo=FALSE, fig.height = 3, fig.width = 11}
p1 <- ggplot(bt, aes(tempchange, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard turnover')

p2 <- ggplot(bt, aes(tempchange, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard total')

p3 <- ggplot(bt, aes(tempchange, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Morisita-Horn')



grid.arrange(p1, p2, p3, ncol = 3)
```

## Only 1 year differences
```{r plot diss vs temp change 1 yr, echo=FALSE, fig.height = 3, fig.width = 11}
bt[, dur := year2 - year1]
p1 <- ggplot(bt[dur == 1, ], aes(tempchange, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard turnover')

p2 <- ggplot(bt[dur == 1, ], aes(tempchange, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard total')

p3 <- ggplot(bt[dur == 1, ], aes(tempchange, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Morisita-Horn')

grid.arrange(p1, p2, p3, ncol = 3)
```

## Only 10 year differences
```{r plot diss vs temp change 10 yr, echo=FALSE, fig.height = 3, fig.width = 11}
bt[, dur := year2 - year1]
p1 <- ggplot(bt[dur == 10, ], aes(tempchange, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard turnover')

p2 <- ggplot(bt[dur == 10, ], aes(tempchange, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard total')

p3 <- ggplot(bt[dur == 10, ], aes(tempchange, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Morisita-Horn')

grid.arrange(p1, p2, p3, ncol = 3)
```

## Only 20 year differences
```{r plot diss vs temp change 20 yr, echo=FALSE, fig.height = 3, fig.width = 11}
bt[, dur := year2 - year1]
p1 <- ggplot(bt[dur == 20, ], aes(tempchange, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard turnover')

p2 <- ggplot(bt[dur == 20, ], aes(tempchange, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Jaccard total')

p3 <- ggplot(bt[dur == 20, ], aes(tempchange, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature change (°C)', y = 'Morisita-Horn')

grid.arrange(p1, p2, p3, ncol = 3)
```


# Temperature trend and time-series length
```{r ts length and temp trend}
ggplot(bt, aes(year2 - year1, tempchange)) +
  geom_point(alpha = 0.2) +
  geom_smooth()
```
# Plot dissimilarity vs. explanatory variables
Lines are ggplot smoother fits
Just Jtu for now, averaged within rarefyID
```{r plot diss v explanatory vars, echo=FALSE, warning = FALSE, message = FALSE, fig.height = 16, fig.width = 9}
p1 <- ggplot(btave, aes(REALM, Jtu)) +
  geom_boxplot(na.rm = TRUE) + 
  labs(x = 'Realm', y = 'Jaccard turnover')

p2 <- ggplot(btave, aes(tempave, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature (°C)', y = 'Jaccard turnover')

p3 <- ggplot(btave, aes(tempave_metab, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Metabolic temperature (°C)', y = 'Jaccard turnover')

p4 <- ggplot(btave, aes(seas, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Seasonality (°C)', y = 'Jaccard turnover')

p5 <- ggplot(btave, aes(microclim, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Microclimate availability (°C)', y = 'Jaccard turnover')

p6 <- ggplot(btave, aes(mass_mean_weight, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Mass (g))', y = 'Jaccard turnover')

p7 <- ggplot(btave, aes(speed_mean_weight+1, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Speed (km / hr))', y = 'Jaccard turnover')

p8 <- ggplot(btave, aes(lifespan_mean_weight, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Lifespan (yr))', y = 'Jaccard turnover')

p9 <- ggplot(btave, aes(consfrac, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction consumers', y = 'Jaccard turnover')

p10 <- ggplot(btave, aes(endofrac, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction endotherms', y = 'Jaccard turnover')

p11 <- ggplot(btave, aes(Nspp, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Number of species', y = 'Jaccard turnover')

p12 <- ggplot(btave, aes(thermal_bias, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Thermal bias (°C)', y = 'Jaccard turnover')

p13 <- ggplot(btave, aes(npp, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Net primary productivity (mg C / m2 / day)', y = 'Jaccard turnover')

p14 <- ggplot(btave, aes(veg, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Vegetation score', y = 'Jaccard turnover')

p15 <- ggplot(btave, aes(human_bowler, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Bowler human impact score', y = 'Jaccard turnover')

p16 <- ggplot(btave, aes(human_venter, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'scaled Human impact score (Venter)', y = 'Jaccard turnover')

p17 <- ggplot(btave, aes(human_halpern, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'scaled Human impact score (Halpern)', y = 'Jaccard turnover')


grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, p16, p17, ncol = 2)
```

Strong trends with temperature change, but trends are pretty symmetric around no trend in temperature, which implies warming or cooling drives similar degree of community turnover.
Some indication of less turnover for larger organisms (mass)
Higher turnover on land with higher seasonality?
More turnover for shorter-lived organisms?
No really clear differences among realms.

