---
title: "Pairwise dissimilarity covariate data prep and visualization"
output:
  github_document: default
---
Trims based on data quality and adds covariate data.

```{r setup}
if(Sys.info()[[4]]=='annotate.sebs.rutgers.edu'){ # for gam smoother
    library('mgcv', lib.loc = '/usr/lib64/R/library') # when running on Annotate. Need to load 1.8-26, not 1.8-33.
} else {
    library(mgcv)
}
library(data.table)
library(ggplot2)
library(beanplot) # for beanplots
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
require(scales) # for custom axis scales
require(here)
source(here('code', 'util.R')) # simple utility functions
logit <- function(x) return(log(x/(1-x)))


```

# Load data
```{r load data}
# biotime taxa category and other info
load(here::here('data', 'biotime_blowes', 'bt_malin.Rdata')) # load bt_malin
btinfo <- data.table(bt_malin); rm(bt_malin)
btinfo[, Nave := mean(N), by = rarefyID] # average number of individuals in the timeseries
btinfo <- btinfo[!duplicated(rarefyID), .(rarefyID, rarefyID_x, rarefyID_y, Biome, taxa_mod, Nave, REALM, STUDY_ID)]

# biotime community dissimilarity data (all pairs)
load(here::here('data', 'biotime_blowes', 'all_pairs_beta.Rdata')) # load rarefied_beta_medians
bt <- data.table(rarefied_beta_medians); rm(rarefied_beta_medians)
bt[, year1 := as.numeric(year1)] # not sure why it gets read in as character
bt[, year2 := as.numeric(year2)]
bt[, Horn := 1- Hornsim] # convert similarity to dissimilarity
bt[, Hornsim := NULL]
bt <- merge(bt, btinfo) # add lat/lon and realm

# Temperature average, changes
temperature <- fread(here('output', 'temperature_byrarefyID.csv.gz')) # from assemble_temp.Rmd

# microclimates
microclim <- fread(here('output', 'microclimates.csv.gz'), drop = 1) # from assemble_microclimates.R

# richness
rich <- fread(here('output', 'richness_by_rarefyID.csv.gz')) # number of species. from extract_richness.R

# human impact
human <- fread(here('output', 'humanimpact_by_rarefyID.csv.gz')) # from extract_human_impacts.R

```


## Add covariates to BT data
```{r add covariate data}
# add covariates
bt <- merge(bt, temperature, all.x = TRUE, by = c('rarefyID', 'year1', 'year2')) # temperature ave, change
bt <- merge(bt, microclim[, .(rarefyID, microclim = Temp_sd20km)], all.x = TRUE, by = 'rarefyID') # microclimates
bt <- merge(bt, rich, all.x = TRUE) # species richness
bt <- merge(bt, human[, .(rarefyID, human_bowler = atc)], all.x = TRUE) # human impact
```



## Trim data
At least 5 species and at least 10 individuals on average.
```{r}
cat('original number of rows'); norig <- nrow(bt); norig
# trim out timeseries with too few species
cat('rows with <5 spp'); bt[Nspp < 5 | is.na(Nspp), .N]
bt <- bt[Nspp >= 5, ]
  
# or few individuals
cat('rows with <10 indivs'); bt[Nave < 10, .N]
bt <- bt[Nave >= 10 | is.na(Nave), ]


cat('rows left'); nrow(bt)
cat('fraction kept'); nrow(bt)/norig
cat('number of studies'); bt[, length(unique(STUDY_ID))]
cat('number of timeseries'); bt[, length(unique(rarefyID))]

```

### Choose the appropriate temperature value for each study and duration
Average temperature for the full duration. Thiel-Sen temperature slope.
```{r}
btall <- bt # legacy from when we divided dataset into different ts durations (3, 5, 10, 20 years)
btall[, tempave := picklongest(year1, year2, tempave), by = rarefyID]
btall[, tempchange := median(tempchange/(year2 - year1)), by = rarefyID] # Thiel-Sen estimator of the slope: median of all pairwise differences

```

## Set up useful variables and transformations
```{r}
# set realm order
btall[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# realm that combines Terrestrial and Freshwater, for interacting with human impact
btall[, REALM2 := REALM]
levels(btall$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")


# duration
btall[, duration := year2 - year1]

# group Marine invertebrates/plants in with All
btall[, taxa_mod2 := taxa_mod]
btall[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']


btall[, tsign := signneg11(tempchange)]

#######################
## Transformations

### Adjust response away from 0-1
btall[, ':='(Jtu.sc = transform01(Jtu), Jne.sc = transform01(Jne), 
           Jbeta.sc = transform01(Jbeta), Horn.sc = transform01(Horn))]


### Log-transform some variables, then center and scale. 
btall[, ':='(tempave.sc = scale(tempave),
           microclim.sc = scale(log(microclim)),
           tempchange.sc = scale(tempchange, center = TRUE), 
           tempchange_abs.sc = scale(abs(tempchange), center = TRUE),
           nspp.sc = scale(log(Nspp)))]
btall[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm


## save the centering and scaling
scalingall <- data.frame(center = t(t(sapply(btall, attr, 'scaled:center'))),
                 scale = t(t(sapply(btall, attr, 'scaled:scale'))))
scalingall$var <- rownames(scalingall)
scalingall <- scalingall[!vapply(scalingall$center, is.null, TRUE) | !vapply(scalingall$scale, is.null, TRUE),]
scalingall$center[vapply(scalingall$center, is.null, TRUE)] <- NA # turn null to NA
scalingall$center <- vapply(scalingall$center, paste, collapse = ", ", character(1L)) # flatten list to character. not sure why it's a list.
scalingall$scale <- vapply(scalingall$scale, paste, collapse = ", ", character(1L))

## save the transformations just entered above
scalingall$log <- FALSE # whether variable was log-transformed before scaling
scalingall$log[scalingall$var %in% c('microclim.sc', 'nspp.sc', 'durationlog.sc', 'human_bowler.sc')] <- TRUE
scalingall$plus <- 0 # whether anything was added to the variable before logtransforming
scalingall$plus[scalingall$var %in% c('human_bowler.sc')] <- 1

```

## Dataset sizes
```{r}
btall[, .N]

btall[, length(unique(STUDY_ID)), by = REALM]

btall[, length(unique(rarefyID)), by = REALM]
```


## Write out
```{r write out}
write.csv(btall[,.(STUDY_ID, rarefyID, rarefyID_x, rarefyID_y, REALM, Biome, taxa_mod, taxa_mod2, Nspp, year1, year2, duration, 
                   Jtu, Jtu.sc, Horn, Horn.sc, 
                   tempave.sc, 
                   microclim.sc, tempchange, tsign, tempchange.sc, tempchange_abs.sc,
                   human_bowler.sc, REALM2)], 
          gzfile(here('output', 'turnover_w_covariates.csv.gz')), row.names = FALSE)
write.csv(scalingall, here('output','turnover_w_covariates_scaling.csv'), row.names = FALSE)
```



# A bit more prep for visualizing covariate distributions
## Make a data.table summarized by rarefyID
```{r summarize by rarefyID}
btaveall <- btall[, .(Jtuslope = lmNAcoef(Jtu, duration), Jbetaslope = lmNAcoef(Jbeta, duration), Hornslope = lmNAcoef(Horn, duration), 
                    minyrBT = min(year1), maxyrBT = max(year2), nyrBT = length(Jtu) + 1,
                    REALM = unique(REALM), REALM2 = unique(REALM2), rarefyID_y = mean(rarefyID_y),
                    tempchange = mean(tempchange), tempave = mean(tempave),
                    microclim = mean(microclim),
                    Nspp = mean(Nspp), 
                    human_bowler = mean(human_bowler),
                    tempave.sc = mean(tempave.sc),
                    tempchange.sc = mean(tempchange.sc), tempchange_abs.sc = mean(tempchange_abs.sc),
                    microclim.sc = mean(microclim.sc),
                    nspp.sc = mean(nspp.sc),
                    human_bowler.sc = mean(human_bowler.sc)),
                by = rarefyID]
```



# Check variable distributions
## Response variables
```{r histograms response}
btall[, summary(Jbeta)]
btall[, summary(Jtu)]
btall[, summary(Horn)]

# fraction 0 or 1
btall[, sum(Jbeta == 0)/.N]
btall[, sum(Jtu == 0)/.N]
btall[!is.na(Horn), sum(Horn == 0)/.N]

btall[, sum(Jbeta == 1)/.N]
btall[, sum(Jtu == 1)/.N]
btall[!is.na(Horn), sum(Horn == 1)/.N]

# histograms
invisible(btall[, hist(Jbeta, main = 'Jaccard total', breaks = 80)])
invisible(btall[, hist(Jtu, main = 'Jaccard turnover', breaks = 80)])
invisible(btall[, hist(Horn, main = 'Morisita-Horn', breaks = 80)])


```
## Unscaled covariate distributions
```{r histograms unscaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btaveall[, hist(minyrBT, main = 'Start year', cex.main = cexmain)])
invisible(btaveall[, hist(maxyrBT - minyrBT, main = 'Duration (years)', cex.main = cexmain)])
invisible(btaveall[, hist(nyrBT, main = 'Number of sampled years', cex.main = cexmain)])
invisible(btaveall[, hist(tempave, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(btaveall[, hist(tempchange, main = 'Temperature trend (°C/yr)', cex.main = cexmain)]) # all the raw data
invisible(btaveall[, hist(microclim, main = 'Microclimates (°C)', cex.main = cexmain)])
invisible(btaveall[, hist(Nspp, main = 'Species richness', cex.main = cexmain)])
invisible(btaveall[, hist(human_bowler, main = 'Human impact score (Bowler)', cex.main = cexmain)])
```

## Scaled covariate distributions
```{r histograms scaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btaveall[, hist(tempave.sc, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(btaveall[, hist(microclim.sc, main = 'log Microclimates (°C)', cex.main = cexmain)])
invisible(btaveall[, hist(tempchange.sc, main = 'Temperature trend (°C/yr)', cex.main = cexmain)])
invisible(btaveall[, hist(tempchange_abs.sc, main = 'abs(Temperature trend) (°C/yr)', cex.main = cexmain)])
invisible(btaveall[, hist(nspp.sc, main = 'log Species richness', cex.main = cexmain)])
invisible(btaveall[, hist(human_bowler.sc, main = 'log Human impact score (Bowler)', cex.main = cexmain)])
```


# Dissimilarity vs. time
## Jaccard turnover
```{r plot diss vs time, echo=FALSE, fig.height = 3, fig.width = 11}
p1 <- ggplot(btall, aes(duration, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Jaccard turnover')

p1
```


## Frequency of time differences
```{r plot time differences}
invisible(btall[, hist(year2 - year1, breaks = seq(0.5, max(year2 - year1)+0.5, by = 1))])
```

## Plot a few long time-series
```{r long ts plots}
p1 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, Jtu)) +
    geom_point() +
    geom_smooth() +
    labs(title = '428_534345')
p2 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, Jtu)) +
    geom_point() +
    geom_smooth() +
    labs(title = '339_1085477')
p3 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, Jtu)) +
    geom_point() +
    geom_smooth() +
    labs(title = '215_450479')
p4 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, Horn)) +
    geom_point() +
    geom_smooth() +
    labs(title = '428_534345')
p5 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, Horn)) +
    geom_point() +
    geom_smooth() +
    labs(title = '339_1085477')
p6 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, Horn)) +
    geom_point() +
    geom_smooth() +
    labs(title = '215_450479')
p1
p2
p3
p4
p5
p6
```

## Logit transformed
```{r long ts plots}
p1 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '428_534345')
p2 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '339_1085477')
p3 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '215_450479')
p4 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, logit(Horn))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '428_534345')
p5 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, logit(Horn))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '339_1085477')
p6 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, logit(Horn))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm', color='red') +
    labs(title = '215_450479')
p1
p2
p3
p4
p5
p6
```



## Log transformed similarity
Do + 0.1 to avoid -Inf
```{r long ts plots}
p1 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, log(1-Jtu+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '428_534345')
p2 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, log(1-Jtu+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '339_1085477')
p3 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, log(1-Jtu+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '215_450479')
p4 <- ggplot(btall[rarefyID == '428_534345'], aes(duration, log(1-Horn+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '428_534345')
p5 <- ggplot(btall[rarefyID == '339_1085477'], aes(duration, log(1-Horn+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '339_1085477')
p6 <- ggplot(btall[rarefyID == '215_450479'], aes(duration, log(1-Horn+0.1))) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = 'lm') +
    labs(title = '215_450479')
p1
p2
p3
p4
p5
p6
```

## Log transformed t and logit(dissimilarity)
```{r long ts plots}
p1 <- ggplot(btall[rarefyID == '428_534345'], aes(log(duration), logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    labs(title = '428_534345')
p2 <- ggplot(btall[rarefyID == '339_1085477'], aes(log(duration), logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    labs(title = '339_1085477')
p3 <- ggplot(btall[rarefyID == '215_450479'], aes(log(duration), logit(Jtu))) +
    geom_point() +
    geom_smooth() +
    labs(title = '215_450479')
p4 <- ggplot(btall[rarefyID == '428_534345'], aes(log(duration), logit(Horn))) +
    geom_point() +
    geom_smooth() +
    labs(title = '428_534345')
p5 <- ggplot(btall[rarefyID == '339_1085477'], aes(log(duration), logit(Horn))) +
    geom_point() +
    geom_smooth() +
    labs(title = '339_1085477')
p6 <- ggplot(btall[rarefyID == '215_450479'], aes(log(duration), logit(Horn))) +
    geom_point() +
    geom_smooth() +
    labs(title = '215_450479')
p1
p2
p3
p4
p5
p6
```
# Correlations among variables 
Pearson's r is in the lower triangle
```{r pairs, fig.height=10, fig.width=10}
pairs(formula = ~ tempave.sc + microclim.sc + tempchange.sc + tempchange_abs.sc + nspp.sc + human_bowler.sc, data = btaveall, gap = 1/10, cex = 0.2, col = '#00000022', lower.panel = panel.cor)

```

