---
title: "Assemble temperature dataset (average, difference, and seasonality) to match with dissimilarity slope. Also consider ecto vs. endotherms."
output: 
    github_document: default

---
```{r setup}
require(data.table)
require(ncdf4)
require(ggplot2)
require(here)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
```

# Read in BioTime and temperature data
Using average daily temperature by month from CRU TS
Using average temperature by month from ERSST. NOAA_ERSST_V5 data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at https://www.esrl.noaa.gov/psd/
```{r read in data}
# biotime community similarity data
btlong <- readRDS(here('temp', 'trendstemp.rds'))

# biotime lat/lon
load(here::here('data', 'biotime_blowes', 'bt_malin.Rdata')) # load bt_malin
btinfo <- data.table(bt_malin); rm(bt_malin)
btinfo <- btinfo[!duplicated(rarefyID), .(rarefyID, rarefyID_x, rarefyID_y, REALM)]
btlong2 <- merge(btlong, btinfo, by = 'rarefyID') # trims out rarefyIDs not in bt_malin
if(nrow(btlong) != nrow(btlong2)){ 
    stop('missing lat/lon for some slopes!')
} else {
    btlong <- btlong2
    rm(btlong2)
}


# biotime species lists
load(here('data', 'biotime_blowes', 'bt_grid_spp_list.Rdata')) # loads bt_grid_spp_list. this has some studies not in bt.
btspp <- data.table(bt_grid_spp_list); rm(bt_grid_spp_list) # rename to btspp

btspp[, length(unique(Species))] # 16315 species

# endo vs. ectotherm classifications by species
endo <- fread(here('output', 'metab_feed_byspecies.csv.gz'))

# CRU TS mean temp by year
n = nc_open(here('dataDL', 'cruts', 'cru_ts4.03.1901.2018.tmp.dat.nc')) # Open the netCDF file. Won't work on .gz for some reason
# print(n) # get information about the file format
cruts = ncvar_get(n, 'tmp') # dim order: 720 lon x 360 lat x 1416 time (observations are month, time measured in days since Jan 1, 1900). 0 lon is at date line.
dim(cruts)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin=as.Date('1900-01-01')), format = '%Y%m')
dimnames(cruts) <- list(lon_cruts = ncvar_get(n, 'lon'), lat_cruts = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

# ERSST temp by month
n = nc_open(here('dataDL', 'ersst', 'sst.mnmean.nc')) # Open the netCDF file
# print(n) # get information about the file format
ersst = ncvar_get(n, 'sst') # dim order: 180 lon x 89 lat x 1994 time (observations are month, time measured in days since Jan 1, 1800). 0 lon is at greenwhich mean.
dim(ersst)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin = as.Date('1800-01-01')), format = '%Y%m')
dimnames(ersst) <- list(lon_ersst = ncvar_get(n, 'lon'), lat_ersst = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

```

Check the temperature data read in correctly
```{r}
# CTU TS
head(dimnames(cruts)[[3]]) # 190101: good
tail(dimnames(cruts)[[3]]) # should be Dec. 2018: good

# ERSST
head(dimnames(ersst)[[3]]) # should be Jan. 1854: good
tail(dimnames(ersst)[[3]]) # should be 2020: good

```

# Average temperature by year
```{r}
# CRU TS
crutsmelt <- as.data.table(cruts, value.name='cruts') # reshape to long format
crutsmelt[, YEAR := as.numeric(substr(time, 1,4))] # extract year
crutsyr <- crutsmelt[, .(cruts = mean(cruts, na.rm=TRUE)), by = .(lat_cruts = as.numeric(lat_cruts), lon_cruts = as.numeric(lon_cruts), YEAR)]

# ERSST
ersstmelt <- as.data.table(ersst, value.name = 'ersst')
ersstmelt[, YEAR := as.numeric(substr(time, 1,4))]
ersstyr <- ersstmelt[, .(ersst = mean(ersst, na.rm=TRUE)), by = .(lat_ersst = as.numeric(lat_ersst), lon_ersst = as.numeric(lon_ersst), YEAR)]
ersstyr[lon_ersst > 180, lon_ersst := lon_ersst - 360] # convert to -180 to 180

```

## Plot CRU TS to make sure it worked
```{r}
ggplot(crutsyr[YEAR == 1901,], aes(lon_cruts, lat_cruts, color = cruts)) + geom_point(size = 0.1)
```
## Plot ERSST to make sure it worked
```{r}
ggplot(ersstyr[YEAR == 1901,], aes(lon_ersst, lat_ersst, color = ersst)) + geom_point(size = 1) # plot to make sure it worked
```

# Calculate seasonality as SD across average months (SD)
```{r}
# CRU TS
crutsmelt[, mo := as.numeric(substr(time, 5,6))] # extract month
crutsmo <- crutsmelt[, .(cruts = mean(cruts, na.rm = TRUE)), by = .(lat_cruts = as.numeric(lat_cruts), lon_cruts = as.numeric(lon_cruts), mo)] # average by month
crutsseas <- crutsmo[, .(seas_cruts = sd(cruts, na.rm = TRUE)), by = .(lat_cruts, lon_cruts)]

# ERSST
ersstmelt[, mo := as.numeric(substr(time, 5,6))] # extract month
ersstmo <- ersstmelt[, .(ersst = mean(ersst, na.rm = TRUE)), by = .(lat_ersst = as.numeric(lat_ersst), lon_ersst = as.numeric(lon_ersst), mo)] # average by month
ersstmo[lon_ersst > 180, lon_ersst := lon_ersst - 360] # convert to -180 to 180
ersstseas <- ersstmo[, .(seas_ersst = sd(ersst, na.rm = TRUE)), by = .(lat_ersst, lon_ersst)]

```

## Plot CRU TS to make sure seasonality worked
```{r}
ggplot(crutsseas, aes(lon_cruts, lat_cruts, color = seas_cruts)) + geom_point(size = 0.1) # plot to make sure it worked
```
## Plot ERSST to make sure seasonality worked
```{r}
ggplot(ersstseas, aes(lon_ersst, lat_ersst, color = seas_ersst)) + geom_point(size = 1) # plot to make sure it worked
```


# Calculate temperature change slope based on biotime year pair/lat/lon
```{r calc slope}
# Set up lat/lon grid in bt
btlong[, lat_cruts := floor(rarefyID_y*2)/2 + 0.25] # find nearest CRU TS lat and lon grid values (0.5)
btlong[, lon_cruts := floor(rarefyID_x*2)/2 + 0.25]

btlong[, lat_ersst := floor((rarefyID_y - 1)*0.5 + 1)/0.5] # find nearest ERSST lat and lon grid values (2).
btlong[, lon_ersst := floor((rarefyID_x + 1)*0.5 + 1)/0.5]

# dataframe with all years
btlongallyrs <- btlong[, .(year = seq(year1, year2)), by = .(rarefyID, year1, year2, lat_cruts, lon_cruts, lat_ersst, lon_ersst, REALM)]
btlongallyrs <- merge(btlongallyrs, crutsyr[, .(year = YEAR, lat_cruts, lon_cruts, cruts)], all.x = TRUE)
btlongallyrs <- merge(btlongallyrs, ersstyr[, .(year = YEAR, lat_ersst, lon_ersst, ersst)], all.x = TRUE, by = c('lon_ersst', 'lat_ersst', 'year'))

# slopes
tslopes_cruts <- btlongallyrs[!is.na(cruts), .(trend_cruts = coef(lm(cruts ~ year))[2]), by = .(rarefyID, year1, year2, lat_cruts, lon_cruts, REALM)]
tslopes_ersst <- btlongallyrs[!is.na(ersst), .(trend_ersst = coef(lm(ersst ~ year))[2]), by = .(rarefyID, year1, year2, lat_ersst, lon_ersst, REALM)]
tslopes <- merge(tslopes_cruts, tslopes_ersst, all = TRUE)

```


# Find fraction ectotherm vs. mammal vs. bird by rarefyID
```{r endotherm fractions}
# summarize by rarefyID
endofrac <- endo[, .(mammfrac = sum(metab == 'mamm')/.N, birdfrac = sum(metab == 'bird')/.N, ectofrac = sum(metab == 'ecto')/.N),
                 by = .(rarefyID)]

# check
endofrac[, all(rowSums(cbind(mammfrac, birdfrac, ectofrac)) == 1)] # TRUE, good
invisible(endofrac[mammfrac != 0, hist(mammfrac)])
invisible(endofrac[birdfrac != 0, hist(birdfrac)])
invisible(endofrac[ectofrac != 0, hist(ectofrac)])

```


## Average temperature and average metabolic temperature (choose body temperature for endotherms)
```{r ave temp and ave metab temp}
# CTU TS
avecruts <- btlongallyrs[, .(ave_cruts = mean(cruts, na.rm = TRUE)), 
          by = .(rarefyID, year1,  year2, lat_cruts, lon_cruts, REALM)] # calculate ave for each pairwise comparison of years
avecruts <- merge(avecruts, endofrac, all.x = TRUE)
avecruts[, ave_metab_cruts := ave_cruts * ectofrac + 38 * mammfrac + 40 * birdfrac] # average temperature, considering endotherm body temps (latter from McCoy & Gillooly 2008 Ecology Letters)

# ERSST
aveersst <- btlongallyrs[, .(ave_ersst = mean(ersst, na.rm = TRUE)), 
          by = .(rarefyID, year1, year2, lat_ersst, lon_ersst, REALM)] # calculate ave
aveersst <- merge(aveersst, endofrac, all.x = TRUE)
aveersst[, ave_metab_ersst := ave_ersst * ectofrac + 38 * mammfrac + 40 * birdfrac]

# examine
avecruts[, plot(ave_metab_cruts ~ ectofrac)]
aveersst[, plot(ave_metab_ersst ~ ectofrac)]

```




## Merge ave, change, and seasonality with bt by rarefyID, year1, year2
```{r merge temp ave trend seas with bt}
# merge seasonality with BT by lat/lon
btsum <- merge(btlong[, .(rarefyID, year1, year2, REALM, lat_cruts, lon_cruts, lat_ersst, lon_ersst)], crutsseas, by = c('lat_cruts', 'lon_cruts'), all.x = TRUE) # CRU TS
btsum <- merge(btsum, ersstseas, by = c('lat_ersst', 'lon_ersst'), all.x = TRUE) # ERSST

# merge in average temperature and metabolically corrected ave temp
btsum <- merge(btsum, avecruts[, .(rarefyID, year1, year2, ave_cruts, ave_metab_cruts)], all.x = TRUE, by = c('rarefyID', 'year1', 'year2'))
btsum <- merge(btsum, aveersst[, .(rarefyID, year1, year2, ave_ersst, ave_metab_ersst)], all.x = TRUE, by = c('rarefyID', 'year1', 'year2'))

# merge in temperature tredd
btsum <- merge(btsum, tslopes[, .(rarefyID, year1, year2, trend_cruts, trend_ersst)], all.x = TRUE, by = c('rarefyID', 'year1', 'year2'))
```

# Examine CRUTS vs. ERSST
Temperature trends are correlated, though non-linear for seasonality (at the poles, low in the ocean and high on land)
```{r basic graphs of temperatures}
# compare temperature trends from CRU TS vs ERSST
ggplot(btsum, aes(trend_cruts, trend_ersst)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()

# average temperature from CRU TS and ERSST
ggplot(btsum, aes(ave_cruts, ave_ersst, color = lat_cruts)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()

# compare seasonality from CRU TS and ERSST
ggplot(btsum, aes(seas_cruts, seas_ersst, color = lat_cruts)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()
```

# Choose CRU TS vs. ERSST
Use CRUTS in general, except ERSST in marine realm
```{r combine cruts and ersst}
# average temperature and metabolically corrected ave temp
btsum[, ':='(tempave = ave_cruts, tempave_metab = ave_metab_cruts)]
btsum[REALM == 'Marine', ':='(tempave = ave_ersst, tempave_metab = ave_metab_ersst)]
btsum[REALM == 'Marine' & is.na(ave_ersst), ':='(tempave = ave_cruts, tempave_metab = ave_metab_cruts)] # fill missing marine with CRU TS
btsum[REALM != 'Marine' & is.na(ave_cruts), ':='(tempave = ave_ersst, tempave_metab = ave_metab_ersst)] # fill missing non-marine with ERSST

# temperature change
btsum[ , temptrend := trend_cruts] # CRU TS for land and freshwater
btsum[REALM == 'Marine', temptrend := trend_ersst] # ERSST for marine
btsum[REALM == 'Marine' & is.na(trend_ersst), ':='(temptrend = trend_cruts)]
btsum[REALM != 'Marine' & is.na(trend_cruts), ':='(temptrend = trend_ersst)]

# seasonality
btsum[, seas := seas_cruts]
btsum[REALM == 'Marine', seas := seas_ersst]
btsum[REALM == 'Marine' & is.na(seas_ersst), ':='(seas = seas_cruts)]
btsum[REALM != 'Marine' & is.na(seas_cruts), ':='(seas = seas_ersst)]

# set column order and remove columns
out <- btsum[, .(rarefyID, year1, year2, tempave, tempave_metab, temptrend, seas)]

# trim to unique rarefyID, year1, year2, combinations
out <- out[!duplicated(out), ]

# check
dim(btlong)
dim(out)
out
```

# Checking missing
```{r check missing}
out[!duplicated(rarefyID) & is.na(tempave), .N]
out[!duplicated(rarefyID) & is.na(tempave_metab), .N]
out[!duplicated(rarefyID) & is.na(temptrend), .N]
out[!duplicated(rarefyID) & is.na(seas), .N]

```

```{r plot missing}
ggplot(btsum, aes(lon_cruts, lat_cruts, color = is.na(tempave))) + geom_point(size=0.2)

ggplot(btsum, aes(lon_cruts, lat_cruts, color = is.na(tempave_metab))) + geom_point(size=0.2)

ggplot(btsum, aes(lon_cruts, lat_cruts, color = is.na(temptrend))) + geom_point(size=0.2)

ggplot(btsum, aes(lon_cruts, lat_cruts, color = is.na(seas))) + geom_point(size=0.2)
```

# Write out 
```{r write}
write.csv(out, file = gzfile('output/temperaturetrend_byrarefyID_by_year1_year2.csv.gz'), row.names = FALSE)

```