---
title: "Assemble temperature dataset (average, trend, and seasonality. Also consider ecto vs. endotherms."
output: html_notebook
---
```{r setup}
require(data.table)
require(ncdf4)
require(ggplot2)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
```

# Read in BioTime and temperature data
Using average daily temperature by month from CRU TS
Using average temperature by month from ERSST. NOAA_ERSST_V5 data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at https://www.esrl.noaa.gov/psd/
```{r}
# biotime community similarity data
load('data/biotime_blowes/bt_malin.Rdata') # load bt_malin
bt <- data.table(bt_malin); rm(bt_malin)

# biotime species lists
load('data/biotime_blowes/bt_grid_spp_list.Rdata') # loads bt_grid_spp_list. this has some studies not in bt.
btspp <- data.table(bt_grid_spp_list); rm(bt_grid_spp_list) # rename to btspp
btspp <- merge(btspp, bt[!duplicated(rarefyID), .(rarefyID, taxa_mod)], by = 'rarefyID') # add taxa_mod to spp list

btspp[, length(unique(Species))] # 26289 species

# endo vs. ectotherm classifications by species
endo <- fread('output/metab_feed.csv.gz')

# CRU TS mean temp by year
n = nc_open('dataDL/cruts/cru_ts4.03.1901.2018.tmp.dat.nc') # Open the netCDF file. Won't work on .gz for some reason
# print(n) # get information about the file format
cruts = ncvar_get(n, 'tmp') # dim order: 720 lon x 360 lat x 1416 time (observations are month, time measured in days since Jan 1, 1900). 0 lon is at date line.
dim(cruts)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin=as.Date('1900-01-01')), format = '%Y%m')
dimnames(cruts) <- list(lon_cruts = ncvar_get(n, 'lon'), lat_cruts = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

# ERSST temp by month
n = nc_open('dataDL/ersst/sst.mnmean.nc') # Open the netCDF file
# print(n) # get information about the file format
ersst = ncvar_get(n, 'sst') # dim order: 180 lon x 89 lat x 1994 time (observations are month, time measured in days since Jan 1, 1800). 0 lon is at greenwhich mean.
dim(ersst)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin = as.Date('1800-01-01')), format = '%Y%m')
dimnames(ersst) <- list(lon_ersst = ncvar_get(n, 'lon'), lat_ersst = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

```

Check the temperature data read in correctly
```{r}
# CTU TS
head(dimnames(cruts)[[3]]) # 190101: good
tail(dimnames(cruts)[[3]]) # should be Dec. 2018: good

# ERSST
head(dimnames(ersst)[[3]]) # should be Jan. 1854: good
tail(dimnames(ersst)[[3]]) # should be 2020: good

```

Average temperature by year
```{r}
# CRU TS
crutsmelt <- as.data.table(cruts, value.name='cruts') # reshape to long format
crutsmelt[, YEAR := as.numeric(substr(time, 1,4))] # extract year
crutsyr <- crutsmelt[, .(cruts = mean(cruts, na.rm=TRUE)), by = .(lat_cruts = as.numeric(lat_cruts), lon_cruts = as.numeric(lon_cruts), YEAR)]

# ERSST
ersstmelt <- as.data.table(ersst, value.name = 'ersst')
ersstmelt[, YEAR := as.numeric(substr(time, 1,4))]
ersstyr <- ersstmelt[, .(ersst = mean(ersst, na.rm=TRUE)), by = .(lat_ersst = as.numeric(lat_ersst), lon_ersst = as.numeric(lon_ersst), YEAR)]

```

Plot CRU TS to make sure it worked
```{r}
ggplot(crutsyr[YEAR == 1901,], aes(lon_cruts, lat_cruts, color = cruts)) + geom_point(size = 0.1)
```
Plot ERSST to make sure it worked
```{r}
ggplot(ersstyr[YEAR == 1901,], aes(lon_ersst, lat_ersst, color = ersst)) + geom_point(size = 1) # plot to make sure it worked
```

Calculate seasonality by year (SD)
```{r}
# CRU TS
crutsmelt[, mo := as.numeric(substr(time, 5,6))] # extract month
crutsmo <- crutsmelt[, .(cruts = mean(cruts, na.rm = TRUE)), by = .(lat_cruts = as.numeric(lat_cruts), lon_cruts = as.numeric(lon_cruts), mo)] # average by month
crutsseas <- crutsmo[, .(seas_cruts = sd(cruts, na.rm = TRUE)), by = .(lat_cruts, lon_cruts)]

# ERSST
ersstmelt[, mo := as.numeric(substr(time, 5,6))] # extract month
ersstmo <- ersstmelt[, .(ersst = mean(ersst, na.rm = TRUE)), by = .(lat_ersst = as.numeric(lat_ersst), lon_ersst = as.numeric(lon_ersst), mo)] # average by month
ersstmo[lon_ersst > 180, lon_ersst := lon_ersst - 360] # convert to -180 to 180
ersstseas <- ersstmo[, .(seas_ersst = sd(ersst, na.rm = TRUE)), by = .(lat_ersst, lon_ersst)]

```

Plot CRU TS to make sure seasonality worked
```{r}
ggplot(crutsseas, aes(lon_cruts, lat_cruts, color = seas_cruts)) + geom_point(size = 0.1) # plot to make sure it worked
```
Plot ERSST to make sure seasonality worked
```{r}
ggplot(ersstseas, aes(lon_ersst, lat_ersst, color = seas_ersst)) + geom_point(size = 1) # plot to make sure it worked
```


# Merge temperature with biotime by year/lat/lon
Set up lat/lon grid in bt
```{r set up lat/lon grid}
bt[, lat_cruts := floor(rarefyID_y*2)/2 + 0.25] # find nearest CRU TS lat and lon grid values (0.5)
bt[, lon_cruts := floor(rarefyID_x*2)/2 + 0.25]

bt[, lat_ersst := floor((rarefyID_y - 1)*0.5 + 1)/0.5] # find nearest ERSST lat and lon grid values (2).
bt[, lon_ersst := floor((rarefyID_x + 1)*0.5 + 1)/0.5]

```

```{r merge temp with bt by year}
# CRU TS
btcruts <- merge(bt, crutsyr, by = c('lat_cruts', 'lon_cruts', 'YEAR'), all.x = TRUE) # CRU TS

# ERSST
ersstyr[lon_ersst > 180, lon_ersst := lon_ersst - 360] # convert to -180 to 180
btersst <- merge(bt, ersstyr, by = c('lat_ersst', 'lon_ersst', 'YEAR'), all.x = TRUE) # ERSST

```

Check for missing CRU TS data in terrestrial, freshwater and marine. na is # missing values. n is # values with  data. Mostly missing in marine, as expected. Study 293 and 216 start before 1900.
```{r}
btcruts[REALM == 'Terrestrial', .(na = sum(is.na(cruts)), n = sum(!is.na(cruts))), by = .(STUDY_ID, rarefyID)][na > 0,]
btcruts[REALM == 'Freshwater', .(na = sum(is.na(cruts)), n = sum(!is.na(cruts))), by = .(STUDY_ID, rarefyID)][na > 0,]
btcruts[REALM == 'Marine', .(na = sum(is.na(cruts)), n = sum(!is.na(cruts))), by = .(STUDY_ID, rarefyID)][na > 0, .N]

ggplot(btcruts[REALM %in% c('Terrestrial', 'Freshwater'), ], aes(rarefyID_x, rarefyID_y, color = is.na(cruts))) +
    geom_point(size = 0.2)

```

Check for missing ERSST data in terrestrial, freshwater and marine. na is # missing values. n is # values with  data. Mostly missing in marine, as expected. Study 293 and 216 start before 1900.
```{r}
btersst[REALM == 'Terrestrial', .(na = sum(is.na(ersst)), n = sum(!is.na(ersst))), by = .(STUDY_ID, rarefyID)][na > 0, .N]
btersst[REALM == 'Freshwater', .(na = sum(is.na(ersst)), n = sum(!is.na(ersst))), by = .(STUDY_ID, rarefyID)][na > 0, .N]
btersst[REALM == 'Marine', .(na = sum(is.na(ersst)), n = sum(!is.na(ersst))), by = .(STUDY_ID, rarefyID, lat_ersst, lon_ersst)][na > 0, ]
btersst[REALM == 'Marine' & is.na(ersst), ][, .(na = sum(is.na(ersst))), by = .(lat_ersst, lon_ersst)]

ggplot(btersst[REALM == 'Marine', ], aes(rarefyID_x, rarefyID_y, color = is.na(ersst))) +
    geom_point(size = 0.2)
```

Fill in temperature for unsampled years in BT
```{r temperature for all years}
# CRU TS
btallyrscruts <- btcruts[!is.na(cruts), .(YEAR = seq(min(YEAR), max(YEAR), by = 1)), by = .(rarefyID, lat_cruts, lon_cruts)] # make a data.table of all years from first to last for each BioTime dataset with cruts data
btallyrscruts <- merge(btallyrscruts, crutsyr, by = c('lat_cruts', 'lon_cruts', 'YEAR'), all.x = TRUE) # merge the data.table with cruts values

# ERSST
btallyrsersst <- btersst[!is.na(ersst), .(YEAR = seq(min(YEAR), max(YEAR), by = 1)), by = .(rarefyID, lat_ersst, lon_ersst)]
btallyrsersst <- merge(btallyrsersst, ersstyr, by = c('lat_ersst', 'lon_ersst', 'YEAR'), all.x = TRUE)

```

# Calculations by rarefyID
Find fraction ectotherm vs. mammal vs. bird by rarefyID
```{r endotherm fractions}
# summarize by rarefyID
endofrac <- endo[, .(mammfrac = sum(metab == 'mamm')/.N, birdfrac = sum(metab == 'bird')/.N, ectofrac = sum(metab == 'ecto')/.N),
                 by = .(rarefyID)]

# check
endofrac[, all(rowSums(cbind(mammfrac, birdfrac, ectofrac)) == 1)] # TRUE, good
endofrac[mammfrac != 0, hist(mammfrac)]
endofrac[birdfrac != 0, hist(birdfrac)]
endofrac[ectofrac != 0, hist(ectofrac)]

```

Average temperature. Choose body temperature for endotherms
```{r ave body temp}
# CTU TS
avecruts <- btallyrscruts[, .(ave_cruts = mean(cruts, na.rm = TRUE), YEAR_min = min(YEAR), YEAR_max = max(YEAR)), 
          by = .(rarefyID)] # calculate ave
avecruts <- merge(avecruts, endofrac, all.x = TRUE)
avecruts[, ave_metab_cruts := ave_cruts * ectofrac + 38 * mammfrac + 40 * birdfrac] # average temperature, considering endotherm body temps (latter from McCoy & Billooly 2008 Ecology Letters)

# ERSST
aveersst <- btallyrsersst[, .(ave_ersst = mean(ersst, na.rm = TRUE), YEAR_min = min(YEAR), YEAR_max = max(YEAR)), 
          by = .(rarefyID)] # calculate ave
aveersst <- merge(aveersst, endofrac, all.x = TRUE)
aveersst[, ave_metab_ersst := ave_ersst * ectofrac + 38 * mammfrac + 40 * birdfrac] # average temperature, considering endotherm body temps (latter from McCoy & Billooly 2008 Ecology Letters)

# examine
avecruts[, plot(ave_metab_cruts ~ ectofrac)]
aveersst[, plot(ave_metab_ersst ~ ectofrac)]

```

Temperature trends for all years from first to last in BioTime.
```{r calc temp trend all years}
# CRU TS
trend_allyrscruts <- btallyrscruts[, .(trend_cruts = coef(lm(cruts ~ YEAR))[2], YEAR_min = min(YEAR), YEAR_max = max(YEAR)), 
          by = .(rarefyID)] # calculate trends

# ERSST
trend_allyrsersst <- btallyrsersst[, .(trend_ersst = coef(lm(ersst ~ YEAR))[2], YEAR_min = min(YEAR), YEAR_max = max(YEAR)), 
          by = .(rarefyID)] # calculate trends

```



Merge ave, trend, and seasonality with bt by rarefyID
```{r merge temp ave, trend, seas with bt}
# merge seasonality with BT by lat/lon
btsum <- merge(bt[!duplicated(rarefyID), .(rarefyID, REALM, rarefyID_x, rarefyID_y, lat_cruts, lon_cruts, lat_ersst, lon_ersst)], crutsseas, by = c('lat_cruts', 'lon_cruts'), all.x = TRUE) # CRU TS
btsum <- merge(btsum, ersstseas, by = c('lat_ersst', 'lon_ersst'), all.x = TRUE) # ERSST

# merge in average temperature and metabolically corrected ave temp
btsum <- merge(btsum, avecruts[, .(rarefyID, ave_cruts, ave_metab_cruts)], all.x = TRUE, by = 'rarefyID')
btsum <- merge(btsum, aveersst[, .(rarefyID, ave_ersst, ave_metab_ersst)], all.x = TRUE, by = 'rarefyID')

# merge in temperature trend
btsum <- merge(btsum, trend_allyrscruts[, .(rarefyID, trend_cruts)], all.x = TRUE, by = 'rarefyID')
btsum <- merge(btsum, trend_allyrsersst[, .(rarefyID, trend_ersst)], all.x = TRUE, by = 'rarefyID')
```

Examine the data
Temperature trends are correlated, though complicated for seasonality (at the poles, low in the ocean and high on land)
```{r basic graphs of temperatures}
# compare temperature trends from CRU TS vs ERSST
ggplot(btsum, aes(trend_cruts, trend_ersst)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()

# average temperature from CRU TS and ERSST
ggplot(btsum, aes(ave_cruts, ave_ersst, color = rarefyID_y)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()

# compare seasonality from CRU TS and ERSST
ggplot(btsum, aes(seas_cruts, seas_ersst, color = rarefyID_y)) + # ERSST more muted at fastest rates of change in CRU TS
    geom_point() +
    geom_smooth()
```

# Choose CRU TS vs. ERSST
```{r combine cruts and ersst}
# average temperature and metabolically corrected ave temp
btsum[, ':='(tempave = ave_cruts, tempave_metab = ave_metab_cruts)]
btsum[REALM == 'Marine', ':='(tempave = ave_ersst, tempave_metab = ave_metab_ersst)]

# temperature trend
btsum[ , temptrend := trend_cruts] # CRU TS for land and freshwater
btsum[REALM == 'Marine', temptrend := trend_ersst] # ERSST for marine

# seasonality
btsum[, seas := seas_cruts]
btsum[REALM == 'Marine', seas := seas_ersst]

# set column order and remove columns
btsum[, ':='(lon_cruts = NULL, lat_cruts = NULL, lon_ersst = NULL, lat_ersst = NULL)]
setcolorder(btsum, c('rarefyID', 'REALM', 'rarefyID_x', 'rarefyID_y', 'tempave', 'ave_cruts', 'ave_ersst', 
                     'tempave_metab', 'ave_metab_cruts', 'ave_metab_ersst', 
                     'temptrend', 'trend_cruts', 'trend_ersst', 'seas', 'seas_cruts', 'seas_ersst'))
# check
btsum
```


# Write out 
```{r write}
write.csv(btsum, file = gzfile('output/temperature_byrarefyID.csv.gz'), row.names = FALSE)

```