---
title: "Turnover covariate data prep and visualization"
output:
  github_document: default
---

If running on Annotate with R 3.5.2, also make sure to manually load mgcv 1.8-26, not the auto-loaded 1.8-33.
```{r setup}
library(data.table)
library(ggplot2)
library(beanplot) # for beanplots
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
require(scales) # for custom axis scales
require(here)


#knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.

signedsqrt_trans <- function() trans_new('signedsqrt', 
                                         transform = function(x) sign(x)*sqrt(abs(x)), 
                                         inverse = function(x) sign(x)*x^2)
```

# Load data
```{r load data}
# biotime taxa category and other info
load(here::here('data', 'biotime_blowes', 'bt_malin.Rdata')) # load bt_malin
btinfo <- data.table(bt_malin); rm(bt_malin)
btinfo <- btinfo[!duplicated(rarefyID), .(rarefyID, rarefyID_x, rarefyID_y, Biome, taxa_mod, REALM, STUDY_ID)]

# biotime community dissimilarity data
load(here::here('data', 'biotime_blowes', 'all_pairs_beta.Rdata')) # load rarefied_beta_medians
bt <- data.table(rarefied_beta_medians); rm(rarefied_beta_medians)
bt[, year1 := as.numeric(year1)] # not sure why it gets read in as character
bt[, year2 := as.numeric(year2)]
bt[, Horn := 1- Hornsim] # convert similarity to dissimilarity
bt[, Hornsim := NULL]
bt <- merge(bt, btinfo[, .(rarefyID, rarefyID_x, rarefyID_y, REALM, STUDY_ID, Biome, taxa_mod)]) # add lat/lon and realm

# Temperature average, changes, and seasonality
temperature <- fread(here('output', 'temperature_byrarefyID.csv.gz'))

# microclimates
microclim <- fread(here('output', 'microclimates.csv.gz'), drop = 1)

# NPP
npp <- fread(here('output', 'npplandocean.csv.gz'))

# Body size
bs <- fread(here('output', 'mass_byrarefyID.csv.gz'), drop = 1)
bs[, ':='(STUDY_ID = NULL, REALM = NULL, taxa_mod = NULL)] # remove unnecessary columns 

# Mobility
speed <- fread(here('output', 'speed_byrarefyID.csv.gz'), drop = 1)
speed[, ':='(STUDY_ID = NULL, REALM = NULL, taxa_mod = NULL)] # remove unnecessary columns 

# Lifespan
lsp <- fread(here('output', 'lifespan_byrarefyID.csv.gz'))

# CTI
cti <- fread(here('output', 'cti_byrarefyID.csv.gz'))
    
# consumer vs. producer
consfrac <- fread(here('output', 'consfrac_byrarefyID.csv.gz'))

# richness
rich <- fread(here('output', 'richness_by_rarefyID.csv.gz')) # number of species

# endotherm vs. ectotherm
endofrac <- fread(here('output', 'endofrac_byrarefyID.csv.gz')) # endotherm vs. ectotherm classifications

# human impact
human <- fread(here('output', 'humanimpact_by_rarefyID.csv.gz'))

# %veg
veg <- as.data.table(readRDS(here('output', 'vct_by_rarefyID.rds')))
veg[, veg := (`tree cover % (mean)` + 0.5 * `non-tree veg. % (mean)`)/100] # veg index from 0 (all non-veg) to 1 (all tree). Non-tree veg counts as 0.5.
```


## Add covariates to BT data
```{r add covariate data}
# add covariates
bt <- merge(bt, temperature, all.x = TRUE, by = c('rarefyID', 'year1', 'year2')) # temperature ave, ave metabolic, change, and seasonality
bt <- merge(bt, microclim[, .(rarefyID, microclim = Temp_sd20km)], all.x = TRUE, by = 'rarefyID') # microclimates
bt <- merge(bt, npp, all.x = TRUE, by = 'rarefyID') # npp
bt <- merge(bt, bs[, .(rarefyID, mass_mean_weight, mass_sd_weight)], all.x = TRUE) # body size mass (g)
bt <- merge(bt, speed[, .(rarefyID, speed_mean_weight, speed_sd_weight)], all.x = TRUE) # speed (km/hr)
bt <- merge(bt, lsp[, .(rarefyID, lifespan_mean_weight, lifespan_sd_weight)], all.x = TRUE) # lifespan (yr)
bt <- merge(bt, cti[, .(rarefyID, thermal_bias)], all.x = TRUE) # thermal bias (degC)
bt <- merge(bt, consfrac[, .(rarefyID, consfrac)], all.x = TRUE) # fraction consumers
bt <- merge(bt, rich, all.x = TRUE) # species richness
bt <- merge(bt, endofrac[, .(rarefyID, endofrac)], all.x = TRUE) # endotherm vs. ectotherm
bt <- merge(bt, human[, .(rarefyID, human_bowler = atc, human_venter = hfp, human_halpern = himp)], all.x = TRUE) # human impact
bt <- merge(bt, veg[, .(rarefyID, veg = veg)], all.x = TRUE) # vegetation index
bt[REALM == 'Marine', veg := 0] # veg index is 0 at sea
```

## Trim data
```{r}
# trim to only data with some temperature change
# important since sign of temperature change is a variable
bt[tempchange == 0, .N] # number to remove
bt <- bt[tempchange != 0, ] # also removes any NA values
```

## Set up useful variables and transformations
```{r}
# set realm order
bt[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# set up sign of temperature change
bt[, tsign := factor(sign(tempchange))]

# realm that combines Terrestrial and Freshwater, for interacting with human impact
bt[, REALM2 := REALM]
levels(bt$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")

# group Marine invertebrates/plants in with All
bt[, taxa_mod2 := taxa_mod]
bt[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']

# calculate duration
bt[, duration := year2 - year1]

#add a comparison id
bt[, compID := paste0(rarefyID, '_', year1, '_', year2)]


#######################
## Transformations

### Adjust response away from 0-1
# transformation for 2 categories. Eq. 1 in Douma & Weedon 2019 MEE
transform01 <- function(x) (x * (length(x) - 1) + 0.5) / (length(x))

bt[, Jtu.sc := transform01(Jtu)]
bt[, Jbeta.sc := transform01(Jbeta)]
bt[, Horn.sc := transform01(Horn)]


### Log-transform some variables, then center and scale. 
bt[, tempave.sc := scale(tempave)]
bt[, tempave_metab.sc := scale(tempave_metab)]
bt[, seas.sc := scale(seas)]
bt[, microclim.sc := scale(log(microclim))]
bt[, tempchange.sc := scale(tempchange, center = TRUE)] # do not center
bt[, tempchange_abs.sc := scale(abs(tempchange), center = TRUE)] # do not center, so that 0 is still 0 temperature change
bt[, mass.sc := scale(log(mass_mean_weight))]
bt[, speed.sc := scale(log(speed_mean_weight+1))]
bt[, lifespan.sc := scale(log(lifespan_mean_weight))]
bt[, consumerfrac.sc := scale(consfrac)]
bt[, endothermfrac.sc := scale(endofrac)]
bt[, nspp.sc := scale(log(Nspp))]
bt[, thermal_bias.sc := scale(thermal_bias)]
bt[, npp.sc := scale(log(npp))]
bt[, veg.sc := scale(log(veg+1))]
bt[, duration.sc := scale(log(duration))]
bt[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
bt[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
bt[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]

## save the centering and scaling
scaling <- data.frame(center = t(t(sapply(bt, attr, 'scaled:center'))),
                 scale = t(t(sapply(bt, attr, 'scaled:scale'))))
scaling$var <- rownames(scaling)
scaling <- scaling[!vapply(scaling$center, is.null, TRUE) | !vapply(scaling$scale, is.null, TRUE),]
scaling$center[vapply(scaling$center, is.null, TRUE)] <- NA # turn null to NA
scaling$center <- vapply(scaling$center, paste, collapse = ", ", character(1L)) # flatten list to character. not sure why it's a list.
scaling$scale <- vapply(scaling$scale, paste, collapse = ", ", character(1L))
```


## Write out
Only if file doesn't yet exist
```{r write out}
if(!file.exists(here('output', 'turnover_w_covariates.csv.gz'))){
  write.csv(bt[,.(STUDY_ID, rarefyID, REALM, Biome, taxa_mod, taxa_mod2, duration, Jtu.sc, Jbeta.sc, Horn.sc, tsign, tempave.sc, tempave_metab.sc, seas.sc, 
                  microclim.sc, tempchange, tempchange.sc, tempchange_abs.sc,
                  mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc,
                  duration.sc, human_bowler.sc, REALM2)], 
            gzfile(here('output', 'turnover_w_covariates.csv.gz')), row.names = FALSE)
  write.csv(scaling, here('output','turnover_w_covariates_scaling.csv'), row.names = FALSE)
}

```



# A bit more prep for visualizing covariate distributions
## Make a data.table summarized by rarefyID
```{r summarize by rarefyID}
btave <- bt[, .(Jtu = mean(Jtu), minyrBT = min(year1), maxyrBT = max(year2), nyrBT = length(Jtu) + 1,
                REALM = unique(REALM), tempchange = mean(tempchange), tempave = mean(tempave),
                tempave_metab = mean(tempave_metab), seas = mean(seas), microclim = mean(microclim),
                mass_mean_weight = mean(mass_mean_weight), speed_mean_weight = mean(speed_mean_weight),
                lifespan_mean_weight = mean(lifespan_mean_weight), consfrac = mean(consfrac),
                endofrac = mean(endofrac), Nspp = mean(Nspp), thermal_bias = mean(thermal_bias),
                npp = mean(npp), veg = mean(veg), human_bowler = mean(human_bowler),
                human_venter = mean(human_venter), human_halpern = mean(human_halpern)),  by = rarefyID]
```


## Add useful variables
```{r useful vars}
# realm that combined Terrestrial and Freshwater, for interacting with human impact
btave[, REALM2 := REALM]
levels(btave$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")
```


## Log-transform some explanantory variables, then center and scale. 
```{r center and scale}
btave[, tempave.sc := scale(tempave)]
btave[, tempave_metab.sc := scale(tempave_metab)]
btave[, seas.sc := scale(seas)]
btave[, microclim.sc := scale(log(microclim))]
bt[, tempchange.sc := scale(tempchange, center = FALSE)] # all the raw data
bt[, tempchange_abs.sc := scale(abs(tempchange), center = FALSE)] # do not center, so that 0 is still 0 temperature change
btave[, tempchange.sc := scale(tempchange, center = FALSE)] # averages within studies
btave[, tempchange_abs.sc := scale(abs(tempchange), center = FALSE)]
btave[, mass.sc := scale(log(mass_mean_weight))]
btave[, speed.sc := scale(log(speed_mean_weight+1))]
btave[, lifespan.sc := scale(log(lifespan_mean_weight))]
btave[, consumerfrac.sc := scale(consfrac)]
btave[, endothermfrac.sc := scale(endofrac)]
btave[, nspp.sc := scale(log(Nspp))]
btave[, thermal_bias.sc := scale(thermal_bias)]
btave[, npp.sc := scale(log(npp))]
btave[, veg.sc := scale(log(veg+1))]
btave[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
btave[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
btave[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]
```


# Check variable distributions
## Response variables
```{r histograms response}
bt[, summary(Jbeta)]
bt[, summary(Jtu)]
bt[, summary(Horn)]

# fraction 0 or 1
bt[, sum(Jbeta == 0)/.N]
bt[, sum(Jtu == 0)/.N]
bt[!is.na(Horn), sum(Horn == 0)/.N]

bt[, sum(Jbeta == 1)/.N]
bt[, sum(Jtu == 1)/.N]
bt[!is.na(Horn), sum(Horn == 1)/.N]

# histograms
invisible(bt[, hist(Jbeta, main = 'Jaccard total', breaks = 80)])
invisible(bt[, hist(Jtu, main = 'Jaccard turnover', breaks = 80)])
invisible(bt[, hist(Horn, main = 'Morisita-Horn', breaks = 80)])

```

## Unscaled covariates
```{r histograms unscaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btave[, hist(minyrBT, main = 'Start year', cex.main = cexmain)])
invisible(btave[, hist(maxyrBT - minyrBT, main = 'Duration (years)', cex.main = cexmain)])
invisible(btave[, hist(nyrBT, main = 'Number of sampled years', cex.main = cexmain)])
invisible(btave[, hist(mass_mean_weight, main = 'Mass (g)', cex.main = cexmain)])
invisible(btave[, hist(speed_mean_weight, main = 'Speed (km/hr)', cex.main = cexmain)])
invisible(btave[, hist(lifespan_mean_weight, main = 'Lifespan (yr)', cex.main = cexmain)])
invisible(btave[, hist(tempave_metab, main = 'Metabolic temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(consfrac, main = 'Consumers (fraction)', cex.main = cexmain)])
invisible(btave[, hist(endofrac, main = 'Endotherms (fraction)', cex.main = cexmain)])
invisible(btave[, hist(tempave, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(bt[, hist(tempchange, main = 'Temperature trend (°C/yr)', cex.main = cexmain)]) # all the raw data
invisible(btave[, hist(seas, main = 'Seasonality (°C)', cex.main = cexmain)])
invisible(btave[, hist(microclim, main = 'Microclimates (°C)', cex.main = cexmain)])
invisible(btave[, hist(Nspp, main = 'Species richness', cex.main = cexmain)])
invisible(btave[, hist(thermal_bias, main = 'Thermal bias (°C)', cex.main = cexmain)])
invisible(btave[, hist(npp, main = 'Net primary productivity', cex.main = cexmain)])
invisible(btave[, hist(veg, main = 'Vegetation index', cex.main = cexmain)])
invisible(btave[, hist(human_bowler, main = 'Human impact score (Bowler)', cex.main = cexmain)])
invisible(btave[, hist(human_venter, main = 'Human impact score (Venter)', cex.main = cexmain)])
invisible(btave[, hist(human_halpern, main = 'Human impact score (Halpern)', cex.main = cexmain)])

```

## Scaled covariates
```{r histograms scaled, fig.height = 9}
# histograms to examine
cexmain = 0.6
par(mfrow = c(5,4))
invisible(btave[, hist(tempave.sc, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(tempave_metab.sc, main = 'Metabolic temperature (°C)', cex.main = cexmain)])
invisible(btave[, hist(seas.sc, main = 'Seasonality (°C)', cex.main = cexmain)])
invisible(btave[, hist(microclim.sc, main = 'log Microclimates (°C)', cex.main = cexmain)])
invisible(bt[, hist(tempchange.sc, main = 'Temperature trend (°C/yr)', cex.main = cexmain)])
invisible(bt[, hist(tempchange_abs.sc, main = 'abs(Temperature trend) (°C/yr)', cex.main = cexmain)])
invisible(btave[, hist(mass.sc, main = 'log Mass (g)', cex.main = cexmain)])
invisible(btave[, hist(speed.sc, main = 'log Speed (km/hr)', cex.main = cexmain)])
invisible(btave[, hist(lifespan.sc, main = 'log Lifespan (yr)', cex.main = cexmain)])
invisible(btave[, hist(consumerfrac.sc, main = 'Consumers (fraction)', cex.main = cexmain)])
invisible(btave[, hist(endothermfrac.sc, main = 'Endotherms (fraction)', cex.main = cexmain)])
invisible(btave[, hist(nspp.sc, main = 'log Species richness', cex.main = cexmain)])
invisible(btave[, hist(thermal_bias.sc, main = 'Thermal bias (°C)', cex.main = cexmain)])
invisible(btave[, hist(npp.sc, main = 'log Net primary productivity', cex.main = cexmain)])
invisible(btave[, hist(veg.sc, main = 'log Vegetation index', cex.main = cexmain)])
invisible(btave[, hist(human_bowler.sc, main = 'log Human impact score (Bowler)', cex.main = cexmain)])
invisible(btave[, hist(human_footprint.sc, main = 'log Human impact score (Venter & Halpern)', cex.main = cexmain)])

```



# Check correlations among variables. Pearson's r is on the lower diagonal.
```{r pairs, fig.height=10, fig.width=10}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y, use = 'pairwise.complete.obs')
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt) #, cex = cex.cor * r)
}
pairs(formula = ~ tempave.sc + tempave_metab.sc + seas.sc + microclim.sc + tempchange.sc + tempchange_abs.sc + mass.sc + speed.sc + lifespan.sc + consumerfrac.sc + endothermfrac.sc + nspp.sc + thermal_bias.sc + npp.sc + veg.sc + human_bowler.sc + human_footprint.sc, data = btave, gap = 1/10, cex = 0.2, col = '#00000022', lower.panel = panel.cor)

```

- Mass and lifespan look tightly correlated, but r only 0.56...?
- Tempave_metab and lifespan don't look tightly correlated, but r= -0.81 
- Tempave_metab and speed don't look tightly correlated, but r= -0.83 
- Lifespan and speed don't look tightly correlated, but r = 0.73

# Compare covariates across realms
```{r compare across realms, fig.height=12, fig.width=9}
i <- bt[, !duplicated(rarefyID)]; sum(i)
par(mfrow=c(5,3))
beanplot(rarefyID_y ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Latitude (degN)', ll = 0.05)
beanplot(tempave ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature (degC)', ll = 0.05)
beanplot(tempave_metab ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Metabolic Temperature (degC)', ll = 0.05, bw = 'nrd0') # nrd0 bandwidth to calculation gap
beanplot(seas ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Seasonality (degC)', ll = 0.05)
beanplot(microclim ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Microclimates (degC)', ll = 0.05)
beanplot(tempchange ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature change (degC)', ll = 0.05)
beanplot(mass_mean_weight ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Mass (g)', ll = 0.05, log = 'y')
beanplot(speed_mean_weight +1 ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Speed (km/hr)', ll = 0.05, log = 'y')
beanplot(lifespan_mean_weight ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Lifespan (yr)', ll = 0.05, log = 'y')
#beanplot(consfrac ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Consumers (fraction)', ll = 0.05, log = '') # too sparse
#beanplot(endofrac ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Endotherms (fraction)', ll = 0.05, log = '') # too sparse
beanplot(Nspp ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Number of species', ll = 0.05, log = 'y')
beanplot(thermal_bias ~ REALM, data = bt[i & !is.na(thermal_bias),], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Thermal bias (degC)', ll = 0.05)
beanplot(npp ~ REALM, data = bt[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)
beanplot(veg ~ REALM, data = bt[i & REALM !='Marine',], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)

```

- Marine are in generally warmer locations (seawater doesn't freeze)
- Marine have much lower seasonality.
- Marine and freshwater have some very small masses (plankton), but much of dataset is similar to terrestrial.
- Marine has a lot of slow, crawling organisms, but land has plants. Land also has birds (fast).

# Frequency of time differences
```{r plot time differences}
invisible(bt[, hist(year2 - year1, breaks = seq(0.5, 120.5, by = 1))])
```

# Plot dissimilarity vs. time
```{r plot diss vs time, echo=FALSE, fig.height = 3, fig.width = 11}
p1 <- ggplot(bt, aes(year2 - year1, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Jaccard turnover')

p2 <- ggplot(bt, aes(year2 - year1, Jbeta, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Jaccard total')

p3 <- ggplot(bt, aes(year2 - year1, Horn, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Years', y = 'Morisita-Horn')



grid.arrange(p1, p2, p3, ncol = 3)
```

# Plot dissimilarity vs. explanatory variables
Lines are ggplot smoother fits
Just Jtu for now, averaged within rarefyID
```{r plot diss v explanatory vars, echo=FALSE, warning = FALSE, message = FALSE, fig.height = 16, fig.width = 9}
p1 <- ggplot(btave, aes(REALM, Jtu)) +
  geom_boxplot(na.rm = TRUE) + 
  labs(x = 'Realm', y = 'Jaccard turnover')

p2 <- ggplot(btave, aes(tempave, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature (°C)', y = 'Jaccard turnover')

p3 <- ggplot(btave, aes(tempave_metab, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Metabolic temperature (°C)', y = 'Jaccard turnover')

p4 <- ggplot(btave, aes(seas, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Seasonality (°C)', y = 'Jaccard turnover')

p5 <- ggplot(btave, aes(microclim, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Microclimate availability (°C)', y = 'Jaccard turnover')

p6 <- ggplot(btave, aes(mass_mean_weight, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Mass (g))', y = 'Jaccard turnover')

p7 <- ggplot(btave, aes(speed_mean_weight+1, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Speed (km / hr))', y = 'Jaccard turnover')

p8 <- ggplot(btave, aes(lifespan_mean_weight, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Lifespan (yr))', y = 'Jaccard turnover')

p9 <- ggplot(btave, aes(consfrac, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction consumers', y = 'Jaccard turnover')

p10 <- ggplot(btave, aes(endofrac, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction endotherms', y = 'Jaccard turnover')

p11 <- ggplot(btave, aes(Nspp, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Number of species', y = 'Jaccard turnover')

p12 <- ggplot(btave, aes(thermal_bias, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Thermal bias (°C)', y = 'Jaccard turnover')

p13 <- ggplot(btave, aes(npp, Jtu)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Net primary productivity (mg C / m2 / day)', y = 'Jaccard turnover')

p14 <- ggplot(btave, aes(veg, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Vegetation score', y = 'Jaccard turnover')

p15 <- ggplot(btave, aes(human_bowler, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Bowler human impact score', y = 'Jaccard turnover')

p16 <- ggplot(btave, aes(human_venter, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'scaled Human impact score (Venter)', y = 'Jaccard turnover')

p17 <- ggplot(btave, aes(human_halpern, Jtu, color = REALM)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'scaled Human impact score (Halpern)', y = 'Jaccard turnover')


grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, p16, p17, ncol = 2)
```

- Strong trends with temperature change, but trends are pretty symmetric around no trend in temperature, which implies warming or cooling drives similar degree of community turnover.
- Some indication of less turnover for larger organisms (mass)
- Higher turnover on land with higher seasonality?
- More turnover for shorter-lived organisms?
- No really clear differences among realms.

## Plot by realm, tempave_metab, npp, human
### Jtu
```{r, fig.height = 7}
bt[tempave_metab <= 10, temp_bin := '<=10 degC']
bt[tempave_metab > 10 & tempave_metab <=25, temp_bin := '(10-25]']
bt[tempave_metab > 25, temp_bin := '>25 degC']
bt[, temp_bin := ordered(temp_bin, c('<=10 degC', '(10-25]', '>25 degC'))]

bt[npp <= 700, npp_bin := '<=700 mgC m-2 day-1']
bt[npp > 700 & npp <= 1100, npp_bin := '(700-1100]']
bt[npp > 1100, npp_bin := '>1100 mgC m-2 day-1']
bt[, npp_bin := ordered(npp_bin, c('<=700 mgC m-2 day-1', '(700-1100]', '>1100 mgC m-2 day-1'))]

bt[human_bowler <= 1, hum_bin := '<=1']
bt[human_bowler > 1 & human_bowler <= 3, hum_bin := '(1-3]']
bt[human_bowler > 3, hum_bin := '>3']
bt[, hum_bin := ordered(hum_bin, c('<=1', '(1-3]', '>3'))]

ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin), ], aes(x=tempave_metab, y=Jtu.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Jaccard turnover', title = 'All data')

 ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 1, ], 
        aes(x=tempave_metab, y=Jtu.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Jaccard turnover', title = '1 yr')

ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 5, ], 
       aes(x=tempave_metab, y=Jtu.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Jaccard turnover', title = '5 yr')

ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 10, ], 
       aes(x=tempave_metab, y=Jtu.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Jaccard turnover', title = '10 yr')
```



### Horn
```{r, fig.height = 7}
ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin), ], aes(x=tempave_metab, y=Horn.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Morisita-Horn dissimilarity', title = 'All data')

 ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 1, ], 
        aes(x=tempave_metab, y=Horn.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Morisita-Horn dissimilarity', title = '1 yr')

ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 5, ], 
       aes(x=tempave_metab, y=Horn.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Morisita-Horn dissimilarity', title = '5 yr')

ggplot(bt[!is.na(npp_bin) & !is.na(hum_bin) & duration == 10, ], 
       aes(x=tempave_metab, y=Horn.sc, color = hum_bin, group = hum_bin)) +
  geom_point(size = 1, alpha = 0.05) +
  #geom_smooth(method = 'gam', formula = y~s(x, k = 5), method.args = list(family = 'quasibinomial')) +
  geom_smooth(method = 'glm', linetype = 'dashed', method.args = list(family = 'quasibinomial')) +
  facet_grid(npp_bin~ REALM, scales = 'free') +
  labs(y = 'Morisita-Horn dissimilarity', title = '10 yr')

```