---
title: "Gains/losses and turnover rate"
output: 
    github_document: default
---
Do the proportion of gains and losses differ by realm, temperature state, or microclimate heterogeneity?

```{r setup, include=FALSE}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(glmmTMB) # for beta regression
library(here) # for navigating directories in the repo
if(Sys.info()['nodename'] == 'annotate.sebs.rutgers.edu'){
  library('mgcv', lib.loc = '/usr/lib64/R/library') # when running on Annotate. Need to load 1.8-26, not 1.8-33, for geom_smooth() in ggplot
} else {
  library('mgcv')
}
source(here('code', 'util.R')) # simple utility functions

# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 

```

# Load data and set up
```{r}
# gains and losses data
load('data/biotime_blowes/bt-rarefy-collate.Rdata') # loads rarefied_alpha_medians and rarefied_beta_medians
rarefied_alpha_medians <- as.data.table(rarefied_alpha_medians)
rarefied_beta_medians <- as.data.table(rarefied_beta_medians)

# biotime community dissimilarity data
bt <- fread('output/turnover_w_covariates.csv.gz') # the timeseries that pass QA/QC. from assemble_turnover_covariates.Rmd
nrow(bt)

# add initial dissimilarities to dissimilarity data
bt[, minduration := min(duration), by = rarefyID]
initdiss <- bt[duration == minduration, .(Jtu.init = mean(Jtu.sc)), by = .(rarefyID)] # initial dissimilarities
bt <- merge(bt, initdiss[, .(rarefyID, Jtu.init)])
nrow(bt)

# add gains and losses to dissimilarity data
bt <- merge(bt, rarefied_alpha_medians[, .(rarefyID, year1 = YEAR, S1 = S)], by = c('rarefyID', 'year1')) # merge in num species in first year
bt <- merge(bt, rarefied_alpha_medians[, .(rarefyID, year2 = YEAR, S2 = S)], by = c('rarefyID', 'year2')) # and second year
bt <- merge(bt, rarefied_beta_medians[, .(rarefyID, year1 = as.numeric(YEAR1), year2 = as.numeric(YEAR2), gains, losses)], by = c('rarefyID', 'year1', 'year2')) # merge in num gains and losses. not clear why YEAR1 and YEAR2 got read as characters, but coercing to numeric seems fine.
nrow(bt)

# missing some gains and losses
bt[, sum(is.na(gains))]
bt[, sum(is.na(losses))]

# make a long version of gains/losses summarized by rarefyID for plotting
btgainloss <- rbind(bt[, .(prop = mean(gains/S1), type = 'gain'), by = .(REALM, tempchange_abs = abs(tempchange), microclim.sc, rarefyID)],
                    bt[, .(prop = mean(losses/S1), type = 'loss'), by = .(REALM, tempchange_abs = abs(tempchange), microclim.sc, rarefyID)],
                    bt[, .(prop = mean(gains/losses), type = 'gainlossratio'), by = .(REALM, tempchange_abs = abs(tempchange), microclim.sc, rarefyID)])

# turnover rate (slope of dissimilarity vs. time)
trends <- fread('output/slope.csv.gz') # linear slopes. from calc_turnover.R

# add turnover rate to initdiss
trends <- merge(initdiss, trends[measure == 'Jtu' & duration_group == 'All'], by = 'rarefyID')
nrow(trends)

# add gains/losses to initdiss
trends <- merge(trends, bt[, .(gainlossprop = mean(gains/losses)), by = .(rarefyID)])
nrow(trends)
```

# Examine gainloss differences across the dataset
## By realm
### Plot
Terrestrial tends to have a smaller proportion of gains and of losses
```{r}
# plots
ggplot(btgainloss[type %in% c('gain', 'loss')], aes(REALM, prop)) +
  geom_boxplot() +
  scale_y_log10() + 
  facet_wrap(~type)

```

```{r}
ggplot(btgainloss[type == 'gainlossratio'], aes(REALM, prop)) +
  geom_boxplot() +
  scale_y_log10() + 
  facet_wrap(~type) +
  labs(y = 'Gain-loss ratio')

```

### Stats
Nonparametric test of gains by realm.
```{r}
kruskal.test(prop ~ REALM, data = btgainloss[type == 'gain' & is.finite(prop), .(prop, REALM = as.factor(REALM))])

```

Nonparametric test of losses by realm.
```{r}
kruskal.test(prop ~ REALM, data = btgainloss[type == 'loss' & is.finite(prop), .(prop, REALM = as.factor(REALM))])

```
Nonparametric test of gains vs. losses by realm.
```{r}
kruskal.test(prop ~ REALM, data = btgainloss[type == 'gainlossratio' & is.finite(prop), .(prop, REALM = as.factor(REALM))])

```

## By tempchange_abs
### Plot
```{r}
ggplot(btgainloss[type == 'gainlossratio' & !is.na(tempchange_abs)], aes(tempchange_abs, prop)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  scale_y_log10() + 
  scale_x_sqrt() + 
  facet_wrap(~type) +
  labs(y = 'Gain-loss ratio')

```
### Stats
Nonparametric test of gains vs. losses by tempchange_abs. Suggests slightly more losses at faster temperature change.
```{r}
cor.test(~ prop + tempchange_abs, method = 'spearman', data = btgainloss[type == 'gainlossratio' & is.finite(prop)])

```

## By microclim
### Plot
```{r}
ggplot(btgainloss[type == 'gainlossratio' & !is.na(microclim.sc)], aes(microclim.sc, prop)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  scale_y_log10() + 
  facet_wrap(~type) +
  labs(y = 'Gain-loss ratio')

```
### Stats
Nonparametric test of gains vs. losses by microclim. Suggests ...
```{r}
cor.test(~ prop + microclim.sc, method = 'spearman', data = btgainloss[type == 'gainlossratio' & is.finite(prop)])

```
## Interaction of gains/losses with initial dissimilarity
### Plot
More losses to the left, more gains to the right. Don't see any clear evidence for the hypothesis that losses have a big impact when initial dissimilarity is low.
```{r}
ggplot(trends, aes(gainlossprop, disstrend, color = year2 - year1)) + 
    geom_point(alpha = 0.5) +
    geom_smooth() +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  facet_wrap(~cut(Jtu.init, breaks = 4), scales = "free_y") +
  scale_color_gradient(trans = 'log10') +
  scale_x_log10()
 
```

# Examine gainlossprop models
Fit by code/turnover_vs_temperature_GLMM_fit_Jtu.gainloss.R.
## Load models
```{r}
# with gainlossprop for Table 1
if(!exists('modGainLossAllJtu')) modGainLossAllJtu <- readRDS(here('temp', 'modGainLossAllJtu.rds')) # Null
#if(!exists('modRealmGainLossAllJtu')) modRealmGainLossAllJtu <- readRDS('temp/modRealmGainLossAllJtu.rds') # Realm. 
#if(!exists('modTaxamod2GainLossAllJtu')) modTaxamod2GainLossAllJtu <- readRDS('temp/modTaxamod2GainLossAllJtu.rds') # Taxon. 
#if(!exists('modsdTtsignGainLossAllJtu')) modsdTtsignGainLossAllJtu <- readRDS(here('temp', 'modsdTtsignGainLossAllJtu.rds')) # tsign, tempchange_abs.
if(!exists('modsdTRealmtsignGainLossAllJtu')) modsdTRealmtsignGainLossAllJtu <- readRDS(here('temp', 'modsdTRealmtsignGainLossAllJtu.rds')) # tsign:tempchange_abs by realm. 
#if(!exists('modabsLatsdTabsLatRealmtsignGainLossAllJtu')) modabsLatsdTabsLatRealmtsignGainLossAllJtu <- readRDS(here('temp', 'modabsLatsdTabsLatRealmtsignGainLossAllJtu.rds')) # tsign:tempchange_abs:absLat. Not yet included since not fit yet
if(!exists('modrawTsdTTRealmtsignGainLossAllJtu')) modrawTsdTTRealmtsignGainLossAllJtu <- readRDS(here('temp','modrawTsdTTRealmtsignGainLossAllJtu.rds')) # tsign:tempchange_abs:tempave:realm. 

# with GainLoss for Table S3
#if(!exists('modGainLossAllHorn')) modGainLossAllHorn <- readRDS(here('temp', 'modGainLossAllHorn.rds')) # Null with only duration
#if(!exists('modRealmGainLossAllHorn')) modRealmGainLossAllHorn <- readRDS('temp/modRealmGainLossAllHorn.rds') # Realm
#if(!exists('modTaxamod2GainLossAllHorn')) modTaxamod2GainLossAllHorn <- readRDS('temp/modTaxamod2GainLossAllHorn.rds') # Taxon
#if(!exists('modsdTtsignGainLossAllHorn')) modsdTtsignGainLossAllHorn <- readRDS(here('temp', 'modsdTtsignGainLossAllHorn.rds')) # tsign, tempchange_abs
#if(!exists('modsdTRealmtsignGainLossAllHorn')) modsdTRealmtsignGainLossAllHorn <- readRDS(here('temp', 'modsdTRealmtsignGainLossAllHorn.rds')) # tsign:tempchange by realm
#if(!exists('modrawTsdTTRealmtsignGainLossAllHorn')) modrawTsdTTRealmtsignGainLossAllHorn <- readRDS(here('temp','modrawTsdTTRealmtsignGainLossAllHorn.rds')) # adds tsign to tempave:tempchange:realm

# with GainLoss for Table S4
#if(!exists('modrawTsdTTRealmtsignmicroclimGainLossAllJtu')) modrawTsdTTRealmtsignmicroclimGainLossAllJtu <- readRDS('temp/modrawTsdTTRealmtsignmicroclimGainLossAllJtu.rds') # has microclimates
#if(!exists('modrawTsdTTRealmtsignhumanGainLossAllJtu')) modrawTsdTTRealmtsignhumanGainLossAllJtu <- readRDS('temp/modrawTsdTTRealmtsignhumanGainLossAllJtu.rds') # has human impact. 


# w/out
if(!exists('modsdTRealmtsignAllJtu')) modsdTRealmtsignAllJtu <- readRDS(here('temp', 'modsdTRealmtsignAllJtu.rds')) # Fit by code/turnover_vs_temperature_GLMM_fit_Jtu.R
if(!exists('modrawTsdTTRealmtsignAllJtu')) modrawTsdTTRealmtsignAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignAllJtu.rds')) # Fit by code/turnover_vs_temperature_GLMM_fit_Jtu.R


```

## Summaries
```{r}
# calculating these are slow, so only do if they don't yet exist
if(!exists('sum.null')) sum.null <- summary(modGainLossAllJtu)
if(!exists('sum.Tchange')) sum.Tchange <- summary(modsdTRealmtsignGainLossAllJtu)
if(!exists('sum.TchangeTave')) sum.TchangeTave <- summary(modrawTsdTTRealmtsignGainLossAllJtu)

# display
sum.null
sum.Tchange
sum.TchangeTave
```


## AIC
### Like Table 1 (main Jtu models)
```{r}

aics <- AIC(modGainLossAllJtu, 
            #modRealmGainLossAllJtu, 
            #modTaxamod2GainLossAllJtu, # simple models w/out tempchange
            #modsdTtsignGainLossAllJtu, 
            modsdTRealmtsignGainLossAllJtu, # tsign:tempchange_abs w/out or w/ realm
            #modabsLatsdTabsLatRealmtsignGainLossAllJtu,
            modrawTsdTTRealmtsignGainLossAllJtu) # add tempave:tempchange_abs
aics$dAIC <- aics$AIC - min(aics$AIC)
aics$dAICnull <- aics$AIC - aics$AIC[rownames(aics)=='modGainLossAllJtu']
aics

```


# Examine Jtu.init:gainlossprop models
Fit by code/turnover_vs_temperature_GLMM_fit_Jtu.init.gainloss.R
## Load models
```{r}
# with Jtu.init:gainlossprop for Table 1
if(!exists('modInitGainLossAllJtu')) modInitGainLossAllJtu <- readRDS(here('temp', 'modInitGainLossAllJtu.rds')) # Null
#if(!exists('modRealmInitGainLossAllJtu')) modRealmInitGainLossAllJtu <- readRDS('temp/modRealmInitGainLossAllJtu.rds') # Realm. 
#if(!exists('modTaxamod2InitGainLossAllJtu')) modTaxamod2InitGainLossAllJtu <- readRDS('temp/modTaxamod2InitGainLossAllJtu.rds') # Taxon. 
#if(!exists('modsdTtsignInitGainLossAllJtu')) modsdTtsignInitGainLossAllJtu <- readRDS(here('temp', 'modsdTtsignInitGainLossAllJtu.rds')) # tsign, tempchange_abs.
if(!exists('modsdTRealmtsignInitGainLossAllJtu')) modsdTRealmtsignInitGainLossAllJtu <- readRDS(here('temp', 'modsdTRealmtsignInitGainLossAllJtu.rds')) # tsign:tempchange_abs by realm. 
#if(!exists('modabsLatsdTabsLatRealmtsignInitGainLossAllJtu')) modabsLatsdTabsLatRealmtsignInitGainLossAllJtu <- readRDS(here('temp', 'modabsLatsdTabsLatRealmtsignInitGainLossAllJtu.rds')) # tsign:tempchange_abs:absLat. Not yet included since not fit yet
if(!exists('modrawTsdTTRealmtsignInitGainLossAllJtu')) modrawTsdTTRealmtsignInitGainLossAllJtu <- readRDS(here('temp','modrawTsdTTRealmtsignInitGainLossAllJtu.rds')) # tsign:tempchange_abs:tempave:realm. 


```

## AIC
### Like Table 1 (main Jtu models)
```{r}

aics <- AIC(modInitGainLossAllJtu, 
            #modRealmInitGainLossAllJtu, 
            #modTaxamod2InitGainLossAllJtu, # simple models w/out tempchange
            #modsdTtsignInitGainLossAllJtu, 
            modsdTRealmtsignInitGainLossAllJtu, # tsign:tempchange_abs w/out or w/ realm
            #modabsLatsdTabsLatRealmtsignInitGainLossAllJtu,
            modrawTsdTTRealmtsignInitGainLossAllJtu) # add tempave:tempchange_abs
aics$dAIC <- aics$AIC - min(aics$AIC)
aics$dAICnull <- aics$AIC - aics$AIC[rownames(aics)=='modInitGainLossAllJtu']
aics

```

## Plot predictions
### Gainloss vs. init effect
Expectation: effect of Jtu.init depends on gainlossprop

- for Jtu.init near 1, gains increase turnover rates and losses decrease them
- for Jtu.init near 0, gains decrease turnover rates and losses increase them

But not what we see. The interaction is the opposite (for Jtu.init near 1, gains decrease turnover rates and losses increase them).

```{r}
# predicted slopes from the tsign model (no tempave)
slopes_modInitGainLossAllJtu <- readRDS(here('temp', 'slopes_modInitGainLossAllJtu.rds')) # made by pred_modrawInitGainLossAllJtu.r

ggplot(slopes_modInitGainLossAllJtu[tempave == -20 & tempchange == 2 & REALM == 'Marine'], # the same across these, so can arbitrarily pick one
       aes(gainlossprop, slope, ymin=slope-1.96*slope.se,  ymax=slope+1.96*slope.se)) +
  geom_ribbon(alpha = 0.5) +
  geom_line() +
  facet_grid(~ Jtu.init)

```

### Like Fig. 2, only tempchange model
```{r, fig.height = 10, fig.width=10}

# predicted slopes from the tsign model (no tempave)
slopes_modsdTRealmtsignInitGainLossAllJtu <- readRDS(here('temp', 'slopes_modsdTRealmtsignInitGainLossAllJtu.rds')) # made by pred_modrawInitGainLossAllJtu.r
slopes_modsdTRealmtsignInitGainLossAllJtu <- slopes_modsdTRealmtsignInitGainLossAllJtu[tempave == -20,] # all the same, so choose one

# plot of change vs. dT
p2 <- ggplot() +
    geom_hline(yintercept = 0, linetype = 'dotted') +
    geom_line(data = slopes_modsdTRealmtsignInitGainLossAllJtu, 
              mapping=aes(tempchange, slope, color = as.factor(gainlossprop), group = as.factor(gainlossprop)), 
              size=0.5) +
    geom_ribbon(data = slopes_modsdTRealmtsignInitGainLossAllJtu, alpha = 0.2, color = NA, 
                aes(tempchange, slope,
                    ymin=slope - slope.se, 
                    ymax=slope + slope.se,
                    color = as.factor(gainlossprop), 
                    group = as.factor(gainlossprop))) +
    scale_color_brewer(palette = 'Dark2') +
    scale_fill_brewer(palette = 'Dark2') +
    facet_grid(Jtu.init ~ REALM, scales = 'free', labeller = label_both)  +
    labs(tag = 'B)', x = 'Temperature change [°C/year]', y = expression(atop('Turnover rate','['~Delta~'Turnover/year]')), 
         size = 'Duration [years]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8)) +
    scale_y_continuous(trans = signedsqrttrans, 
                       breaks = c(seq(-1,-0.2, by = 0.2), -0.1, 0, 0.1, seq(0.2, 1, by=0.2))) +
    scale_size(trans='log', range = c(0.8,3), breaks = c(2, 5, 20, 50)) +
    guides(size = guide_legend(override.aes = list(alpha=1))) # set alpha to 1 for points in the legend
    
p2 <- addSmallLegend(p2, pointSize = 0.8, spaceLegend = 0.1, textSize = 6)
p2
```



```{r, fig.height = 30, fig.width=8}

# predicted slopes from the tempave interaction model
slopes_modrawTsdTTRealmtsignInitGainLossAllJtu <- readRDS(here('temp', 'slopes_modrawTsdTTRealmtsignInitGainLossAllJtu.rds')) # made by pred_modrawXAllJtu.sh/.r
vals <- slopes_modrawTsdTTRealmtsignInitGainLossAllJtu[c(which.min(abs(tempave - 0)), which.min(abs(tempave-25))), unique(tempave)] # show 0 and 25degC (+/-2SD from the mean)
slopes_modrawTsdTTRealmtsignInitGainLossAllJtu <- slopes_modrawTsdTTRealmtsignInitGainLossAllJtu[tempave %in% vals,]
slopes_modrawTsdTTRealmtsignInitGainLossAllJtu[, tempave := factor(as.character(round(tempave)), levels = c('2', '24'))] # re-order factor for nicer plotting

# b) plot of change vs. dT
p2 <- ggplot() +
    geom_hline(yintercept = 0, linetype = 'dotted') +
    geom_line(data = slopes_modrawTsdTTRealmtsignInitGainLossAllJtu, mapping=aes(tempchange, slope, color = tempave, group = tempave), linetype = 'dashed') +
    geom_ribbon(data = slopes_modrawTsdTTRealmtsignInitGainLossAllJtu, alpha = 0.2, color = NA,
                aes(tempchange, slope, fill = tempave,
                    ymin=slope - slope.se,
                    ymax=slope + slope.se)) +
    scale_color_brewer(palette = 'Dark2') +
    scale_fill_brewer(palette = 'Dark2') +
    facet_grid(Jtu.init + gainlossprop ~ REALM, scales = 'free', labeller = label_both)  +
    labs(tag = 'B)', x = 'Temperature change [°C/year]', y = expression(atop('Turnover rate','['~Delta~'Turnover/year]')), 
         size = 'Duration [years]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8)) +
    scale_y_continuous(trans = signedsqrttrans, 
                       breaks = c(seq(-1,-0.2, by = 0.2), -0.1, 0, 0.1, seq(0.2, 1, by=0.2))) +
    scale_size(trans='log', range = c(0.8,3), breaks = c(2, 5, 20, 50)) +
    guides(size = guide_legend(override.aes = list(alpha=1))) # set alpha to 1 for points in the legend
    
p2 <- addSmallLegend(p2, pointSize = 0.8, spaceLegend = 0.1, textSize = 6)
p2
```



