---
title: "Sample global temperature dataset trends"
output:
    github_document: default
---
```{r setup}
require(data.table)
require(ncdf4)
require(ggplot2)
require(gridExtra)
library(scales) # for defining signed sqrt axis transformation
library(here)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.

signedsqrt = function(x) sign(x)*sqrt(abs(x))
signedsq = function(x) sign(x) * x^2
signedsqrttrans <- trans_new(name = 'signedsqrt', transform = signedsqrt, inverse = signedsq) # for plotting

```

# Read in BioTime and temperature data
Using average daily temperature by month from CRU TS
Using average temperature by month from ERSST. NOAA_ERSST_V5 data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at https://www.esrl.noaa.gov/psd/
```{r}
# biotime covariate data, including temperature trends
trends <- fread('output/turnover_w_covariates.csv.gz') # From assemble_turnover_covariates.Rmd
trends[, REALM2 := as.character(REALM)] # group terrestrial and freshwater
trends[REALM %in% c('Terrestrial', 'Freshwater'), REALM2 := 'TerrFresh']

# CRU TS mean temp by year
n = nc_open('dataDL/cruts/cru_ts4.03.1901.2018.tmp.dat.nc') # Open the netCDF file. Won't work on .gz for some reason
# print(n) # get information about the file format
cruts = ncvar_get(n, 'tmp') # dim order: 720 lon x 360 lat x 1416 time (observations are month, time measured in days since Jan 1, 1900). 0 lon is at date line.
dim(cruts)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin=as.Date('1900-01-01')), format = '%Y%m')
dimnames(cruts) <- list(lon_cruts = ncvar_get(n, 'lon'), lat_cruts = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

# ERSST temp by month
n = nc_open('dataDL/ersst/sst.mnmean.nc') # Open the netCDF file
# print(n) # get information about the file format
ersst = ncvar_get(n, 'sst') # dim order: 180 lon x 89 lat x 1994 time (observations are month, time measured in days since Jan 1, 1800). 0 lon is at greenwhich mean.
dim(ersst)
dates <- format(as.Date(as.numeric(ncvar_get(n, 'time')), origin = as.Date('1800-01-01')), format = '%Y%m')
dimnames(ersst) <- list(lon_ersst = ncvar_get(n, 'lon'), lat_ersst = ncvar_get(n, 'lat'), time = dates)
nc_close(n)

```

Check the temperature data read in correctly
```{r}
# CTU TS
head(dimnames(cruts)[[3]]) # 190101: good
tail(dimnames(cruts)[[3]]) # should be Dec. 2018: good

# ERSST
head(dimnames(ersst)[[3]]) # should be Jan. 1854: good
tail(dimnames(ersst)[[3]]) # should be 2020: good

```

# Average temperature by year
```{r}
# CRU TS
crutsmelt <- as.data.table(cruts, value.name='cruts') # reshape to long format
crutsmelt[, YEAR := as.numeric(substr(time, 1,4))] # extract year
crutsyr <- crutsmelt[, .(cruts = mean(cruts, na.rm=TRUE)), by = .(lat_cruts = as.numeric(lat_cruts), lon_cruts = as.numeric(lon_cruts), YEAR)]

# ERSST
ersstmelt <- as.data.table(ersst, value.name = 'ersst')
ersstmelt[, YEAR := as.numeric(substr(time, 1,4))]
ersstyr <- ersstmelt[, .(ersst = mean(ersst, na.rm=TRUE)), by = .(lat_ersst = as.numeric(lat_ersst), lon_ersst = as.numeric(lon_ersst), YEAR)]

```

## Plot CRU TS to make sure it worked
```{r}
ggplot(crutsyr[YEAR == 1901,], aes(lon_cruts, lat_cruts, color = cruts)) + geom_point(size = 0.1)
```
## Plot ERSST to make sure it worked
```{r}
ggplot(ersstyr[YEAR == 1901,], aes(lon_ersst, lat_ersst, color = ersst)) + geom_point(size = 1) # plot to make sure it worked
```



# Pick 50000 locations globally and assign sets of years
```{r random locations}
nloc = 50000 # number of locations to pick
set.seed(1) # to ensure repeatability

# function to calculate lat/lon grid area. size is in degrees (length of one side)
llgridarea <- function(lat, lon, size){
    R <- 6371 # earth's radius
    lat1 <- (lat - size/2)*pi/180 # in rads
    lat2 <- (lat + size/2)*pi/180
    lon1 <- (lon - size/2)*pi/180
    lon2 <- (lon + size/2)*pi/180
    
    # https://www.pmel.noaa.gov/maillists/tmap/ferret_users/fu_2004/msg00023.html
    A = 2 * pi * R^2 * abs(sin(lat1) - sin(lat2)) * abs(lon1 - lon2)/2/pi
    return(A)
}

# Pick locations weighted by area and assign a rarefyID/year1/year2 randomly
crutslls <- crutsyr[!duplicated(cbind(lat_cruts, lon_cruts)), .(lat_cruts, lon_cruts)] # get lat/lons.slow
crutslls[, area := llgridarea(lat_cruts, lon_cruts, 0.5)] # calculate grid area
crutslls <- crutslls[sample(1:.N, nloc, replace = TRUE, prob = area), ] # pick locations weighted by area
crutslls <- cbind(crutslls, trends[REALM %in% c('Terrestrial', 'Freshwater'), 
                                  ][sample(1:.N, nloc, replace = TRUE), .(rarefyID)])  # randomly assign rarefyIDs, which we will next use for deciding in which years to sample
crutslls[, id := 1:.N] # add an id to keep each location sample separate
crutsallyrs <- merge(crutslls, trends[, .(rarefyID, year1, year2)], by = 'rarefyID', allow.cartesian = TRUE) # merge in all the years for each rarefyID

ersstllsall <- ersstyr[!duplicated(cbind(lat_ersst, lon_ersst)), .(lat_ersst, lon_ersst)]
ersstllsall[, area := llgridarea(lat_ersst, lon_ersst, 2)]
ersstlls <- ersstllsall[sample(1:.N, nloc, replace = TRUE, prob = area), ]
ersstlls <- cbind(ersstlls, trends[REALM == 'Marine', 
                                  ][sample(1:.N, nloc, replace = TRUE), .(rarefyID)])  # randomly assign rarefyIDs
ersstlls[, id := 1:.N]
ersstallyrs <- merge(ersstlls, trends[, .(rarefyID, year1, year2)], by = 'rarefyID', allow.cartesian = TRUE) # merge in all the years for each rarefyID

# Fill out with temperature
crutsallyrs <- merge(crutsallyrs, crutsyr[, .(lat_cruts, lon_cruts, year1 = YEAR, cruts1 = cruts)], by = c('lat_cruts', 'lon_cruts', 'year1'), all.x = TRUE) # merge the data.table with cruts values for first year
crutsallyrs <- merge(crutsallyrs, crutsyr[, .(lat_cruts, lon_cruts, year2 = YEAR, cruts2 = cruts)], by = c('lat_cruts', 'lon_cruts', 'year2'), all.x = TRUE) # merge the data.table with cruts values for 2nd year

ersstallyrs <- merge(ersstallyrs, ersstyr[, .(lat_ersst, lon_ersst, year1 = YEAR, ersst1 = ersst)], by = c('lat_ersst', 'lon_ersst', 'year1'), all.x = TRUE)
ersstallyrs <- merge(ersstallyrs, ersstyr[, .(lat_ersst, lon_ersst, year2 = YEAR, ersst2 = ersst)], by = c('lat_ersst', 'lon_ersst', 'year2'), all.x = TRUE)
```

## Temperature change
Calculate for each pair of years
```{r calc temp diff}
# CRU TS
crutsallyrs[, dcruts := cruts2 - cruts1]

# ERSST
ersstallyrs[, dersst := ersst2 - ersst1]

```

## Check zeros
2% of locations are always frozen. Those are the only locations that have no temperature change through time.
```{r check zeros}
ersstallyrs[(ersst1 < -1.79 & ersst2 < -1.79), .N]/nrow(ersstallyrs) # fraction of values that are always frozen
ersstallyrs[dersst == 0, summary(ersst1)] # temperatures of locations that have no change in temperature

# remove comparisons that are at freezing temp all the time?
#ersstallyrs <- ersstallyrs[!(ersst1 < -1.79 & ersst2 < -1.79),]

```

## Merge and calc Thiel-Sen slope of temperature vs. time
Also remove the NA values. Leaves us with about 50k locations each on land and at sea.
```{r merge}
tempchange <- rbind(crutsallyrs[, .(REALM = 'Terrestrial', tempchange = median(dcruts/(year2 - year1))), 
                                by = id],
                    ersstallyrs[, .(REALM = 'Marine', tempchange = median(dersst/(year2 - year1))), 
                                by = id])

tempchange[is.na(tempchange), .N] # some NA values
tempchange <- tempchange[!is.na(tempchange), ] # remove them
tempchange[, .N, by = REALM]
```

# Summaries of the rate of temperature change
In degrees C per year
```{r summary}
# degC/year
tempchange[, .(mean = mean(tempchange), se = sd(tempchange)/sqrt(.N)), by = REALM]
trends[!is.na(tempchange), .(mean = mean(tempchange), se = sd(tempchange)/sqrt(.N)), by = REALM2]
```

# Plot
```{r plot, fig.width=8, fig.height=3}
p1 <- ggplot(tempchange, aes(x = tempchange, fill = REALM)) +
    geom_density(alpha = 0.25) +
    scale_y_sqrt() +
    scale_x_continuous(limits = c(-3, 3), trans = signedsqrttrans) +
    ggtitle('Random locations')

p2 <- ggplot(trends[!is.na(tempchange), ], aes(x = tempchange, fill = REALM2)) +
    geom_density(alpha = 0.25) +
    scale_y_sqrt() +
    scale_x_continuous(limits = c(-3, 3), trans = signedsqrttrans) +
    ggtitle('BioTime')

grid.arrange(p1, p2, nrow = 1)
```

# Write out 
```{r write}
write.csv(tempchange, file = gzfile(here('output', 'temperature_trends_sampled.csv.gz')), row.names = FALSE)

```


