---
title: 'Community dissimilarity through time using beta models: Evaluate and plot models'
output: 
    github_document:
        toc: true
        toc_depth: 3

---
# Prep
First run turnover_vs_temperature_GLMM_fit.R to fit the models.
```{r setup, echo=FALSE}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(glmmTMB) # for ME models
library(lme4) # for plotting random effects dotplot
library(bbmle) # for AICtab
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
library(RColorBrewer)
library(scales) # for defining custom scales in ggplot
library(DHARMa) # for model diagnostics
library(here) # for relative paths
library(performance) # for r2 calculations
library(lavaan) # for SEM models of slopes

options(width=500) # turn off most text wrapping

signedsqrt = function(x) sign(x)*sqrt(abs(x))
signedsq = function(x) sign(x) * x^2
signedsqrttrans <- trans_new(name = 'signedsqrt', transform = signedsqrt, inverse = signedsq)
scaleme <- function(x, nm){
  if(!(nm %in% scalingall[,var])) stop('nm not found in scalingall')
  if(scalingall[var==nm, log]){
    x.sc <- (log(x + scalingall[var==nm, plus]) - scalingall[var == nm, center]) / scalingall[var == nm, scale]  
  } else {
    x.sc <- (x  + scalingall[var==nm, plus] - scalingall[var == nm, center]) / scalingall[var == nm, scale]
  }
  return(x.sc)
}

unscaleme <- function(x.sc, nm){
  if(!(nm %in% scalingall[,var])) stop('nm not found in scalingall')
  if(scalingall[var==nm, log]){
    x <- exp(x.sc * scalingall[var == nm, scale] + scalingall[var == nm, center]) - scalingall[var==nm, plus]
  } else {
    x <- x.sc * scalingall[var == nm, scale] + scalingall[var == nm, center] - scalingall[var==nm, plus]
  }
  return(x)
}

# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 

# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trendsall <- fread(here('output', 'turnover_w_covariates.csv.gz'))

# And the scaling factors
scalingall <- fread(here('output', 'turnover_w_covariates_scaling.csv'))


```


# Summary stats
## Examine how many data points are available
### Just turnover and non-zero tempchange
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trendsall), '\n')
cat('# studies: ', trendsall[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trendsall[, length(unique(rarefyID))], '\n')
trendsall[!duplicated(rarefyID), table(REALM)]
trendsall[!duplicated(rarefyID), table(taxa_mod)]
trendsall[!duplicated(rarefyID), table(taxa_mod, REALM)]
```

### With all covariates
Use Bowler for human impact
```{r sample size Jtu covariates}
# the cases we can compare
apply(trendsall[, .(Jtu.sc, REALM, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trendsall[, complete.cases(Jtu.sc, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
cat('Overall # dissimilarities: ', sum(i), '\n')
cat('# studies: ', trendsall[i, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trendsall[i, length(unique(rarefyID))], '\n')
trendsall[i & !duplicated(rarefyID), table(REALM)]
trendsall[i & !duplicated(rarefyID), table(taxa_mod)]
trendsall[i & !duplicated(rarefyID), table(taxa_mod, REALM)]
```

# Models
## Choose the variance structure
Try combinations of

- beta errors
- variance scaled to Nspp or to Nspp + REALM
- random intercept for taxa_mod2, STUDY_ID, rarefyID
- random slopes for duration

Did not try random intercept for compID (for overdispersion), since had trouble converging.
Use a full set of fixed effects and choose the RE with lowest AIC.
Many of the 3- and 5-year models with random slopes had singular convergence issues.

### 10-years
This chooses beta errors, random slopes (duration.sc) & intercepts for STUDY_ID and rarefyID, and variance scaled to nspp + realm.
We haven't dealt with potential testing on the boundary issues here.
```{r choose variance structure, echo=FALSE}
mods10 <- vector('list', length = 17)
i = 1
if(file.exists(here('temp', 'modRFgauss10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFgauss10.rds')); names(mods10)[i] <- 'modRFgauss10'; i = i+1}
if(file.exists(here('temp', 'modRFbeta10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFbeta10.rds')); names(mods10)[i] <- 'modRFbeta10'; i = i+1}
if(file.exists(here('temp', 'modRFrID10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFrID10.rds')); names(mods10)[i] <- 'modRFrID10'; i = i+1}
if(file.exists(here('temp', 'modRF2lev10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2lev10.rds')); names(mods10)[i] <- 'modRF2lev10'; i = i+1}
if(file.exists(here('temp', 'modRFnestedRE10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFnestedRE10.rds')); names(mods10)[i] <- 'modRFnestedRE10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE10.rds')); names(mods10)[i] <- 'modRFslopeRE10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2lev10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2lev10.rds')); names(mods10)[i] <- 'modRFslopeRE2lev10'; i = i+1}
if(file.exists(here('temp', 'modRFdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdisp10.rds')); names(mods10)[i] <- 'modRFdisp10'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2levdisp10.rds')); names(mods10)[i] <- 'modRF2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisp10.rds')); names(mods10)[i] <- 'modRFslopeRE2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeREdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeREdisp10.rds')); names(mods10)[i] <- 'modRFslopeREdisp10'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2levdisprealm10.rds')); names(mods10)[i] <- 'modRF2levdisprealm10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisprealm10.rds')); names(mods10)[i] <- 'modRFslopeRE2levdisprealm10'; i = i+1}

# with log duration
if(file.exists(here('temp', 'modRFdurlog2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisp10.rds')); names(mods10)[i] <- 'modRFdurlog2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisp10.rds')); names(mods10)[i] <- 'modRFdurlogslopeRE2levdisp10'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeREdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeREdisp10.rds')); names(mods10)[i] <- 'modRFdurlogslopeREdisp10'; i = i+1} # convergence error
if(file.exists(here('temp', 'modRFdurlog2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisprealm10.rds')); names(mods10)[i] <- 'modRFdurlog2levdisprealm10'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisprealm10.rds')); names(mods10)[i] <- 'modRFdurlogslopeRE2levdisprealm10'; i = i+1}

modes <- sapply(X = mods10, FUN = mode)
aics <- sapply(X = mods10[modes == 'list'], FUN = AIC)
aics - min(aics)
```


### Summary of the chosen model

```{r, echo=FALSE}
summary(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]])
```


### Plot random effects
No clear discontinuities. Looks ok.
```{r, echo=FALSE}
lme4:::dotplot.ranef.mer(ranef(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]])$cond)
```

### DHARMa model evaluation
```{r, echo=FALSE, eval=FALSE}
## run the top part by hand if needed. takes ~3 hrs
if(FALSE) {
  i10 <- trends10[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, 
                                   endothermfrac.sc, microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]

  res_RFslopeRE2levdisprealm10 <- simulateResiduals(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]], n = 250)
  
  saveRDS(res_RFslopeRE2levdisprealm10, file = 'temp/res_RFslopeRE2levdisprealm10.rds')
} else {
  if(file.exists('temp/res_RFslopeRE2levdisprealm10.rds')) res_RFslopeRE2levdisprealm10 <- readRDS('temp/res_res_RFslopeRE2levdisprealm10.rds')
  i10 <- trends10[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, 
                                   endothermfrac.sc, microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
}
```

#### Plot residuals for model fit to all data
There are many outliers and signs of underdispersion, but this is probably ok.
```{r, echo=FALSE, eval=FALSE}
if(exists('res_RFslopeRE2levdisprealm10')){
  plot(res_RFslopeRE2levdisprealm10)
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$tempchange_abs.sc[i10], xlab = 'tempchange_abs.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$REALM[i10], xlab = 'REALM', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$tempave_metab.sc[i10], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$duration.sc[i10], xlab = 'duration.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$mass.sc[i10], xlab = 'mass.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$endothermfrac.sc[i10], xlab = 'endothermfrac.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$microclim.sc[i10], xlab = 'microclim.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$npp.sc[i10], xlab = 'npp.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$human_bowler.sc[i10], xlab = 'human_bowler.sc', main = '')
}
```


#### Overdispersion
There is underdispersion.
```{r, echo=FALSE, eval=FALSE}
testDispersion(res_RFslopeRE2levdisprealm10)

```

#### Near-zero-inflation
(Zero values have been transformed to slightly >0, so can't test zero-inflation directly)

Near-zero values are not inflated.
```{r, eval=FALSE}
countNearZero <- function(x) sum(x < 0.0001)
testGeneric(res_RFslopeRE2levdisprealm10, summary = countNearZero, alternative = 'greater')

```

#### Near-one-inflation
Near-one values are not inflated.
```{r, eval=FALSE}
countNearOne <- function(x) sum(x > 0.9999)
testGeneric(res_RFslopeRE2levdisprealm10, summary = countNearOne, alternative = 'greater')

```

## Realm x duration
Compare very basic models w/ and w/out realm as an explanatory factor for differences in turnover (proportion species/yr)
```{r}
modAllJtu <- readRDS('temp/modAllJtu.rds')
modRealmAllJtu <- readRDS('temp/modRealmAllJtu.rds')
```

A model with REALM is worse than a model without (just looking at average slopes, no other covariates)
```{r}
aics <- AIC(modAllJtu, modRealmAllJtu)
aics$dAIC <- aics$AIC - min(aics$AIC)
aics
```

## Temperature models
### AIC for Jtu
```{r, rows.print=20}
modAllJtu <- readRDS(here('temp', 'modAllJtu.rds')) # just duration
modRealmAllJtu <- readRDS(here('temp', 'modRealmAllJtu.rds')) # realm
modTaxamod2AllJtu <- readRDS(here('temp', 'modTaxamod2AllJtu.rds')) # taxa
modsdTRealmtsignAllJtu <- readRDS(here('temp', 'modsdTRealmtsignAllJtu.rds')) # tsign, tempchange_abs, realm
modrawTsdTTRealmtsignAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignAllJtu.rds')) # tempave * tempchange_abs, realm, tsign

aics3 <- AIC(modAllJtu, modRealmAllJtu,modTaxamod2AllJtu, # no tempchange
             modsdTRealmtsignAllJtu, # tempchange with realm
             modrawTsdTTRealmtsignAllJtu)# add tempave*tempchange and realm
aics3$dAIC <- aics3$AIC - min(aics3$AIC)
aics3
```

### AIC for Horn
```{r, rows.print=20}
modAllHorn <- readRDS(here('temp', 'modAllHorn.rds')) # just duration
modRealmAllHorn <- readRDS(here('temp', 'modRealmAllHorn.rds')) # realm
modTaxamod2AllHorn <- readRDS(here('temp', 'modTaxamod2AllHorn.rds')) # realm
modsdTRealmtsignAllHorn <- readRDS(here('temp', 'modsdTRealmtsignAllHorn.rds')) # tsign, tempchange_abs, realm
modrawTsdTTRealmtsignAllHorn <- readRDS(here('temp', 'modrawTsdTTRealmtsignAllHorn.rds')) # tempave * tempchange_abs, realm, tsign

aics3 <- AIC(modAllHorn, modRealmAllHorn, modTaxamod2AllHorn,
             modsdTRealmtsignAllHorn, # tempchange with realm
             modrawTsdTTRealmtsignAllHorn
             )# add tempave*tempchange and realm
aics3$dAIC <- aics3$AIC - min(aics3$AIC)
aics3
```

### Plots for Jtu
#### Make/load predictions
Uses output from scripts.
```{r, echo = FALSE}
# using tsign:tempchange_abs:realm
if(file.exists(here('temp', 'slopes_modsdTRealmtsignAllJtu.rds'))){
  slopessdTRealmtsignAllJtu <- readRDS(here('temp', 'slopes_modsdTRealmtsignAllJtu.rds'))
  print('Loaded sdTRealmtsignAllJtu predictions from file')
} else {
  print('No modsdTRealmtsignAllJtu effects. Need to run code/pred_modrawXAllJtu.sh modsdTRealmtsignAllJtu')
}

# using tsign:tempchange_abs * tempave * realm
if(file.exists(here('temp', 'slopes_rawTsdTTRealmtsign.rds'))){
  slopesTsdTTRealmtsignJtu <- readRDS(here('temp', 'slopes_rawTsdTTRealmtsign.rds'))
  print('Loaded TsdTTRealmtsignJtu predictions from file')
} else {
  print('No TsdTTRealmtsignJtu effects. Need to run pred_GLMMmodrawTsdTTRealmtsignAllJtu.R')
}

```
#### tsign:tempchange_abs:realm plots
##### Confidence bounds
CIs are from a SEM fit to the trends.
```{r, echo = FALSE}
vals <- slopessdTRealmtsignAllJtu[c(which.min(abs(tempave - 0)), which.min(abs(tempave-25))), unique(tempave)] # show 0 and 25degC
ggplot(slopessdTRealmtsignAllJtu[tempave %in% vals,], 
       aes(tempchange, slope, color = factor(round(tempave)), group = tempave)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope-1.96*slope.se, ymax=slope+1.96*slope.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM))  +
    labs(x = 'dT [degC/yr]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8))  

```




#### tsign:tempchange_abs*tempave:realm effects
```{r}
fixef(modTsdTTRealmtsignAllJtu)

```

##### Interaction
modrawTsdTTRealmtsignAllJtu model
```{r, echo = FALSE}
# plot
p1 <- ggplot(slopesTsdTTRealmtsignJtu[REALM=='Marine'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Marine') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Jtu slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)
p2 <- ggplot(slopesTsdTTRealmtsignJtu[REALM=='Terrestrial'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Terrestrial') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Jtu slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)
p3 <- ggplot(slopesTsdTTRealmtsignJtu[REALM=='Freshwater'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Freshwater') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Jtu slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)

p1
p2
p3
```

##### Confidence bounds
CIs are from a SEM fit to the trends.
```{r, echo=FALSE}
vals <- slopesTsdTTRealmtsignJtu[c(which.min(abs(tempave - 0)), which.min(abs(tempave-25))), unique(tempave)] # show 0 and 25degC
ggplot(slopesTsdTTRealmtsignJtu[tempave %in% vals,], 
       aes(tempchange, slope_realmtsign, color = factor(round(tempave)), group = tempave)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_realmtsign-1.96*slope_realmtsign.se, ymax=slope_realmtsign+1.96*slope_realmtsign.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM))  +
    labs(x = 'dT [degC/yr]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8))  

```


### Plots for Horn
#### Make/load predictions
Uses output from scripts.
```{r, echo = FALSE}
# using tsign:tempchange_abs:realm
if(file.exists(here('temp', 'slopes_modsdTRealmtsignAllHorn.rds'))){
  slopessdTRealmtsignAllHorn <- readRDS(here('temp', 'slopes_modsdTRealmtsignAllHorn.rds'))
  print('Loaded sdTRealmtsignAllHorn predictions from file')
} else {
  print('No modsdTRealmtsignAllHorn effects. Need to run code/pred_modrawXAllHorn.sh modsdTRealmtsignAllHorn')
}

# using tsign:tempchange_abs * tempave * realm
if(file.exists(here('temp', 'slopes_rawTsdTTRealmtsignHorn.rds'))){
  slopesTsdTTRealmtsignHorn <- readRDS(here('temp', 'slopes_rawTsdTTRealmtsignHorn.rds'))
  print('Loaded TsdTTRealmtsignHorn predictions from file')
} else {
  print('No TsdTTRealmtsignHorn effects. Need to run pred_GLMMmodrawTsdTTRealmtsignAllHorn.R')
}

```
#### tsign:tempchange_abs:realm plots
##### Confidence bounds
CIs are from a SEM fit to the trends.
```{r, echo = FALSE}
vals <- slopessdTRealmtsignAllHorn[c(which.min(abs(tempave - 0)), which.min(abs(tempave-25))), unique(tempave)] # show 0 and 25degC
ggplot(slopessdTRealmtsignAllHorn[tempave %in% vals,], 
       aes(tempchange, slope, color = factor(round(tempave)), group = tempave)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope-1.96*slope.se, ymax=slope+1.96*slope.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM))  +
    labs(x = 'dT [degC/yr]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8))  

```




#### tsign:tempchange_abs*tempave:realm effects
```{r}
fixef(modTsdTTRealmtsignAllHorn)

```

##### Interaction
modrawTsdTTRealmtsignAllHorn model
```{r, echo = FALSE}
# plot
p1 <- ggplot(slopesTsdTTRealmtsignHorn[REALM=='Marine'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Marine') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Horn slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)
p2 <- ggplot(slopesTsdTTRealmtsignHorn[REALM=='Terrestrial'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Terrestrial') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Horn slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)
p3 <- ggplot(slopesTsdTTRealmtsignHorn[REALM=='Freshwater'], aes(tempchange, tempave, z = slope_realmtsign)) +
  geom_raster(aes(fill = slope_realmtsign)) +
  labs(x = 'Temperature change (degC per year)', y = 'Temperature (degC)', title = 'Freshwater') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0, name = 'Horn slope') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)

p1
p2
p3
```

##### Confidence bounds
CIs are from a SEM fit to the trends.
```{r, echo=FALSE}
vals <- slopesTsdTTRealmtsignHorn[c(which.min(abs(tempave - 0)), which.min(abs(tempave-25))), unique(tempave)] # show 0 and 25degC
ggplot(slopesTsdTTRealmtsignHorn[tempave %in% vals,], 
       aes(tempchange, slope_realmtsign, color = factor(round(tempave)), group = tempave)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_realmtsign-1.96*slope_realmtsign.se, ymax=slope_realmtsign+1.96*slope_realmtsign.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM))  +
    labs(x = 'dT [degC/yr]') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), axis.line = element_line(colour = "black"),
          legend.key=element_blank(),
          axis.text=element_text(size=8),
          axis.title=element_text(size=8),
          plot.title=element_text(size=8))  

```


#### Are endotherms common at high tempave on land?
tempG20 means temperature > 20degC. Endotherms are common at all environmental temperatures on land, not just high temperatures.
```{r, echo = FALSE}
trendsall[!is.na(tempave.sc) & !duplicated(rarefyID), .(mean = round(mean(unscaleme(endothermfrac.sc, 'endothermfrac.sc')),2), sd = signif(sd(unscaleme(endothermfrac.sc, 'endothermfrac.sc')),2)), by = .(REALM, tempG20 = unscaleme(tempave.sc, 'tempave.sc') > 20 )][order(REALM, tempG20),]
```

#### Which species are common in each quadrant?
FALSE is less than 20degC, TRUE is greater than.
```{r}
trendsall[!is.na(tempave.sc) & !duplicated(rarefyID), table(taxa_mod2, unscaleme(tempave.sc, 'tempave.sc') > 20, REALM)]
trendsall[!is.na(tempave.sc) & !duplicated(rarefyID), round(prop.table(table(taxa_mod2, unscaleme(tempave.sc, 'tempave.sc') > 20, REALM), margin = c(2,3)),2)]
```


## Latitude or temperature?
Compare temperature vs. latitude in an interaction with abs(tempchange). Not very useful until we fit abs(latitude) model so that effects are mirrored at each pole.
```{r}
modTsdTTRealmAllJtu <- readRDS(here('temp', 'modTsdTTRealmAllJtu.rds')) # uses temperature
modrawTsdTTRealmAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmAllJtu.rds')) # uses environmental temperature
modLatsdTTRealmAllJtu <- readRDS(here('temp', 'modLatsdTTRealmAllJtu.rds')) # uses latitude
```

### AIC
Clearly chooses environmental temperature.
```{r}
aics <- AIC(modTsdTTRealmAllJtu, modLatsdTTRealmAllJtu, modrawTsdTTRealmAllJtu)
aics$dAIC <- aics$AIC - min(aics$AIC)
aics
```


## Covariate models
### Load models
```{r, echo = FALSE}
# simple to more complex models
modRealmAllJtu <- readRDS(here('temp', 'modRealmAllJtu.rds')) # only realm
modrawTsdTTRealmtsignAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignAllJtu.rds')) # tempave, tempchange_abs, realm, tsign

# covariate models with tsign
modrawTsdTTRealmtsignhumanAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignhumanAllJtu.rds')) # human
modrawTsdTTRealmtsignmicroclimAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignmicroclimAllJtu.rds')) # microclim
modrawTsdTTRealmtsignnppAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignnppAllJtu.rds')) # npp
modrawTsdTTRealmtsignseasAllJtu <- readRDS(here('temp', 'modrawTsdTTRealmtsignseasAllJtu.rds')) # seas
```

### AIC
```{r, rows.print=15}
aics <- AIC(modRealmAllJtu, modrawTsdTTRealmtsignAllJtu,
            #modrawTsdTTRealmtsignnppAllJtu, 
            modrawTsdTTRealmtsignseasAllJtu,
            modrawTsdTTRealmtsignmicroclimAllJtu, modrawTsdTTRealmtsignhumanAllJtu)
aics$dAIC <- aics$AIC - min(aics$AIC)
aics

```


### Interactions as lines
#### Load the interaction effects
```{r, echo = FALSE}

if(file.exists(here('temp', 'slopes_rawTsdTTRealmtsignCovariate.rds'))){
  slopes_rawTsdTTRealmtsignCovariate <- readRDS(here('temp', 'slopes_rawTsdTTRealmtsignCovariate.rds'))
  print('Loaded tsign interaction predictions from file')

} else {
  print('Need to run pred_GLMMmodrawTsdTTRealmCovariate.R with argument tsign')
  
}

if(file.exists(here('temp', 'slopes_rawinteractions2.rds'))){
  slopesraw2 <- readRDS(here('temp', 'slopes_rawinteractions2.rds'))
  print('Loaded complex interaction predictions from file')

} else {
  print('Need to run pred_GLMMmodrawTsdTTRealmCovariate.R with argument full')
  
}

```

#### Plot the interactions for tsign models
Covariate only interacts with tempchange_abs.

Error bars are 95% CIs
```{r echo=FALSE, fig.height=15, fig.width=6}
# plot
p1 <- ggplot(slopes_rawTsdTTRealmtsignCovariate, aes(tempchange, slope_microclim, 
                                                     color = microclim, group = microclim)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_microclim-1.96*slope_microclim.se, 
                        ymax=slope_microclim+1.96*slope_microclim.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM), rows = vars(tempave), labeller = label_both)
p2 <- ggplot(slopes_rawTsdTTRealmtsignCovariate, aes(tempchange, slope_npp, color = npp, group = npp)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_npp-1.96*slope_npp.se, ymax=slope_npp+1.96*slope_npp.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM), rows = vars(tempave), labeller = label_both)
p3 <- ggplot(slopes_rawTsdTTRealmtsignCovariate, aes(tempchange, slope_seas, color = seas, group = seas)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_seas-1.96*slope_seas.se, ymax=slope_seas+1.96*slope_seas.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM), rows = vars(tempave))
p4 <- ggplot(slopes_rawTsdTTRealmtsignCovariate, aes(tempchange, slope_human, 
                                                     color = human_bowler, group = human_bowler)) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_human-1.96*slope_human.se, ymax=slope_human+1.96*slope_human.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM), rows = vars(tempave), labeller = label_both)

p1
p2
p3
p4
```


## Thermal bias test
Does the thermal bias of a community explain the propensity to change when temperature changes? 

### With metabolic temperature
Model with thermal bias is selected
```{r, echo = FALSE}
modTsdTTRealmAllJtu_thermal_biasdata <- readRDS(here('temp','modTsdTTRealmAllJtu_thermal_biasdata_test.rds')) # no thermal bias, but using the thermal bias dataset
modTsdTTRealmtsignAllJtu_thermal_biasdata <- readRDS(here('temp', 'modTsdTTRealmtsignAllJtu_thermal_biasdata_test.rds')) # adding only tsign
modTsdTTRealmthermal_biasAllJtu_thermal_biasdata <- readRDS(here('temp','modTsdTTRealmthermal_biasAllJtu_thermal_biasdata_test.rds')) # adds thermal bias:tempave_metab:tempchange_abs

aics <- AIC(modTsdTTRealmAllJtu_thermal_biasdata, modTsdTTRealmtsignAllJtu_thermal_biasdata, modTsdTTRealmthermal_biasAllJtu_thermal_biasdata)
aics$dAIC <- aics$AIC - min(aics$AIC)
aics
```

### With environmental temperature (rawT)
Model with only tsign is selected
```{r, echo = FALSE}
modrawTsdTTRealmAllJtu_thermal_biasdata <- readRDS(here('temp','modrawTsdTTRealmAllJtu_thermal_biasdata.rds')) # no thermal bias, but using the thermal bias dataset
modrawTsdTTRealmtsignAllJtu_thermal_biasdata <- readRDS(here('temp', 'modrawTsdTTRealmtsignAllJtu_thermal_biasdata.rds')) # adding only tsign
modrawTsdTTRealmthermal_biassdTAllJtu_thermal_biasdata <- readRDS(here('temp','modrawTsdTTRealmthermal_biassdTAllJtu_thermal_biasdata.rds')) # adds thermal bias:tempchange_abs:tsign
modrawTsdTTRealmthermal_biasAllJtu_thermal_biasdata <- readRDS(here('temp','modrawTsdTTRealmthermal_biasAllJtu_thermal_biasdata.rds')) # adds thermal bias:tempave

aics <- AIC(modrawTsdTTRealmAllJtu_thermal_biasdata, modrawTsdTTRealmtsignAllJtu_thermal_biasdata, modrawTsdTTRealmthermal_biassdTAllJtu_thermal_biasdata,  modrawTsdTTRealmthermal_biasAllJtu_thermal_biasdata)
aics$dAIC <- aics$AIC - min(aics$AIC)
aics
```

#### Thermal bias coefs
Model that doesn't have a thermal_bias:tempave interaction.
```{r, echo = FALSE}
fixef(modrawTsdTTRealmthermal_biassdTAllJtu_thermal_biasdata)
```

#### Make or load thermal bias effect
```{r, echo = FALSE}
if(file.exists(here('temp', 'slopes_rawTsdTTRealmthermal_bias.rds'))){
  slopesTB <- readRDS(here('temp', 'slopes_rawTsdTTRealmthermal_bias.rds'))
  print('Loaded thermal bias predictions from file')
} else {
  print('Need to run pred_GLMMmodrawTsdTTRealmthermal_biasAllJtu.R')
}
```



#### Plot thermal bias T:dT interaction
Thermal bias is the Community Temperature Index minus the environmental temperature. Positive means that the species are adapted to warmer conditions than they are experiencing (and presumably are more tolerant of warming but less tolerant of cooling).  
Plot the simpler model (only has thermal_bias:tempchange_abs:tsign and not also thermal_bias:tempave).  
```{r, echo = FALSE, fig.height=8}
# plot
p1 <- ggplot(slopesTB, aes(tempchange, tempave, z = slope_thermal_biassdT)) +
  geom_raster(aes(fill = slope_thermal_biassdT)) +
  labs(x = 'Temperature change (degC per year)', y = 'Environmental temperature (degC)', title = 'Negative (left) or positive (right) thermal bias') +
  scale_fill_gradient2(high= "#B2182B", mid = "white", low= "#2166AC", midpoint = 0) +
  facet_grid(row = vars(REALM), col = vars(thermal_bias)) +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "top",
        legend.margin = margin(c(0, 0.1, -20, 0)),
        legend.justification = c(0.92, 0.9),
        legend.text=element_text(size= 12),
        legend.title=element_text(size= 18),
        legend.title.align = 1)

p1
```

#### Plot interaction as lines
```{r echo=FALSE}
# plot
p1 <- ggplot(slopesTB[tempave %in% c(-20, 30)], aes(tempchange, slope_thermal_biassdT, color = as.factor(thermal_bias), group = as.factor(thermal_bias))) +
    geom_line() +
    geom_pointrange(aes(ymin=slope_thermal_biassdT-1.96*slope_thermal_biassdT.se,
                        ymax=slope_thermal_biassdT+1.96*slope_thermal_biassdT.se), alpha = 0.25) + 
    facet_grid(cols = vars(REALM), rows = vars(tempave), labeller = label_both)

p1
```



## Taxon test
Models with taxon effects interacted with temperature change all returned singular fits (NAs for SEs). Using simplest model here.
```{r, echo = FALSE}
modTaxamod2AllJtu <- readRDS(here('temp', 'modTaxamod2AllJtu.rds')) # has taxa_mod2

# previous models
modAllJtu <- readRDS(here('temp', 'modAllJtu.rds')) # only duration
modRealmAllJtu <- readRDS(here('temp', 'modRealmAllJtu.rds')) # only realm

```

#### AIC
```{r}
aics <- AIC(modAllJtu, modRealmAllJtu, modTaxamod2AllJtu)
aics$dAIC <- aics$AIC - min(aics$AIC, na.rm=TRUE)
aics

```


#### Summary
```{r}
summary(modTaxamod2AllJtu)
```



