---
title: 'Drivers of variation in the community response to temperature change across realms'
subtitle: '(using mixed effects models)'
output: 
    github_document: default
    #html_document: default
    html_notebook: default
---

Collaborators: Shane Blowes, Jon Chase, Helmut Hillebrand, Michael Burrows, Amanda Bates, Uli Brose, Benoit Gauzens, Laura Antao
Assistance: Katherine Lew, Josef Hauser

# Introduction
- Climate change is driving a widespread reorganization of ecological communities around the world (Parmsesan & Yohe 2003, Poloczanska et al. 2013),
- but the impacts of climate change vary substantially from one location to another and among taxa (Molinos et al. 2016 NCC, Antao et al. 2020 NEE).
- Community reorganization is substantially more common than an aggregate loss or gain of species (Dornelas et al. 2014 Science, Blowes et al. 2019 Science, Hillebrand et al. 2017 J Appl Ecol)
- There are many hypotheses for why some communities are more sensitive to warming than others, including differences in
  - metabolic rates (Dillon et al. 2010 Nature), 
  - thermal physiology (Deutsch et al. 2008 PNAS, Pinsky et al. 2019 Nature), 
  - microclimate availability (Burrows et al. 2019 NCC, Suggitt et al. 2018 NCC),
  - species mobility (Poloczanska et al. 2013 NCC, Burrows et al. 2011 Science, Sunday et al. 2012 NCC)
  - or generation time (Beaugrand et al. 2009 DSR II, Poloczanska et al. 2013 NCC),
  - consumers vs. producers (Petchey et al. 1999 Nature)
  - community composition (Stuart-Smith et al. 2015 Nature, Beaugrand et a. 2015 NCC, Trisos et al. 2020 Nature), 
  - ecosystem productivity (Thomas et al. 2017 GCB, Brett 1971 Am Zoo),
  - exposure to human impacts (White & Kerr 2006 Ecography)
  - and among realms (Antao et al. 2020 NEE).
- Scaling up from organismal effects to whole ecological communities is complex, and yet these scales are critical for ecosystem functioning and human well-being. 
- There is a need for a comprehensive test to understand where warming is driving and is likely to drive the most dramatic community turnover

# Methods
- BioTime dataset, gridded to 96 km2 hexagons, summarized as temporal turnover (Blowes)
  - Temporal slope of Jaccard turnover compared to the first year
  - Same for Jaccard total
  - and Morisita-Horn turnover
- Tested explanatory variables for differences in rate of turnover:
  - Temperature trend over the time-frame of each time-series (CRU TS 4.03 on land and in freshwater, ERSST v5 in the ocean)
  - Seasonality as a metric of thermal sensitivity (Deutsch et al. 2008 PNAS). Standard deviation of monthly temperatures.
  - Average temperature as a metric of metabolic rates (Dillon et al. 2010 Nature, Antao et al. 2020 Nat E&E)
  - Mobility calculated from body mass and taxonomic group classifications of mobility mode (fly, run, swim, crawl, sessile). Fly/run/swim followed the allometric relationship in Hirt et al. 2017 Nat E&E. Crawl set at 0.1 km/hr, sessile set to 0 km/hr. Then calculated averaged within each assemblage.
  - Net primary productivity (NPP) from the merged land/ocean product produced by the [Ocean Productivity](http://www.science.oregonstate.edu/ocean.productivity/) group at Oregon State using methods from Zhao et al. 2005 and Behrenfeld & Falkowski 1997. 
 - TO DO:
   - Generation time calculated from body mass and endotherm vs. ectotherm classifications, following McCoy & Gillooly 2008 ELE. Averaged across species within each assemblage. 
   - Human impact calculated from Carsten Meyer's ecosystem cube data
   - Thermal bias calculated from Species Temperature Indices (Mike Burrows)
   - Microclimates calculated from WorldClim and BioOracle (Laura Antao)
   - Consumer vs. producer classification
 - Maybe to do:
   - Species pool richness
- Differences in temporal turnover (response variable) modeled with a linear mixed effects model (nlme package, lme() function). See below for details.

```{r setup}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
#library(lme4)
library(nlme) # for ME models
library(beanplot) # for beanplots
library(maps) # for map
library(ggeffects) # marginal effect plots
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together

# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
```

```{r load data}
# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends <- fread('output/turnover_w_covariates.csv.gz')

# set realm order
trends[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = TRUE)]
```

Log-transform some variables, then center and scale. Do they look ok?
``` {r center and scale}
trends[, temptrend_comb_allyr.sc := scale(temptrend_comb_allyr)]
trends[, temptrend_comb_allyr_abs.sc := scale(log(abs(temptrend_comb_allyr)))]
trends[, seas_comb.sc := scale(seas_comb)]
trends[, tempave_comb.sc := scale(tempave_comb)]
trends[, npp.sc := scale(npp)]
trends[, mass_geomean.sc := scale(log(mass_geomean))]
trends[, speed_geomean.sc := scale(log(speed_geomean+1))]

# histograms to examine
par(mfrow = c(3,3))
invisible(trends[, hist(temptrend_comb_allyr.sc, main = 'Temperature trend (°C/yr)')])
invisible(trends[, hist(temptrend_comb_allyr_abs.sc, main = 'log abs(Temperature trend) (°C/yr)')])
invisible(trends[, hist(seas_comb.sc, main = 'Seasonality (°C)')])
invisible(trends[, hist(tempave_comb.sc, main = 'Temperature (°C)')])
invisible(trends[, hist(npp.sc, main = 'Net primary productivity')])
invisible(trends[, hist(mass_geomean.sc, main = 'log Mass (g)')])
invisible(trends[, hist(speed_geomean.sc, main = 'log Speed (km/hr)')])

```


# Results
## Where do we have data?
```{r map}
world <- map_data('world')
ggplot(world, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = 'lightgray', color = 'white') +
    geom_point(data = trends, aes(rarefyID_x, rarefyID_y, group = REALM, color = REALM), size = 0.5, alpha = 0.4)  +
    scale_color_brewer(palette="Set1")
```
Mostly northern hemisphere, but spread all over. No so much in Africa or much of Asia.


## Plot turnover vs. temperature trends
Lines are ggplot smoother fits by realm.
```{r plot turnover v temp trend, echo=FALSE}
# Jaccard turnover trend vs. temperature trend (across all years)
p1 <- ggplot(trends, aes(temptrend_comb_allyr, Jtutrend, color = REALM)) +
    geom_point(size = 0.2, alpha = 0.5) + 
    geom_smooth() +
    scale_color_brewer(palette="Set1") + 
    labs(x = 'Temperature trend (°C/year)', y = 'Jaccard turnover temporal trend')

# Jaccard turnover trend vs. abs temperature trend (across all years)
# ggplot(trends, aes(abs(temptrend_comb_allyr), Jtutrend, group = REALM, color = REALM)) +
#     geom_point(size = 0.2, alpha = 0.5) + 
#     geom_smooth() +
#     geom_smooth(method = 'lm') +
#     scale_color_brewer(palette="Set1")

# Jaccard total trend vs. temperature trend (across all years)
p3 <- ggplot(trends, aes(temptrend_comb_allyr, Jbetatrend, color = REALM)) +
    geom_point(size = 0.2, alpha = 0.5) + 
    geom_smooth() +
    scale_color_brewer(palette="Set1") + 
    labs(x = 'Temperature trend (°C/year)', y = 'Jaccard total temporal trend')

# Jaccard turnover trend vs. abs temperature trend (across all years)
# ggplot(trends, aes(abs(temptrend_comb_allyr), Jbetatrend, group = REALM, color = REALM)) +
#     geom_point(size = 0.2, alpha = 0.5) + 
#     geom_smooth() +
#     geom_smooth(method = 'lm') +
#     scale_color_brewer(palette="Set1")

# Horn-Morisita turnover trend vs. temperature trend (across all years)
p5 <- ggplot(trends, aes(temptrend_comb_allyr, Horntrend, color = REALM)) +
    geom_point(size = 0.2, alpha = 0.5) + 
    geom_smooth() +
    scale_color_brewer(palette="Set1") + 
    labs(x = 'Temperature trend (°C/year)', y = 'Morisita-Horn temporal turnover')

# Horn-Morisita turnover trend vs. abs temperature trend (across all years)
# ggplot(trends, aes(abs(temptrend_comb_allyr), Horntrend, group = REALM, color = REALM)) +
#     geom_point(size = 0.2, alpha = 0.5) + 
#     geom_smooth() +
#     geom_smooth(method = 'lm') +
#     scale_color_brewer(palette="Set1")

p1
p3
p5
```
Trends are pretty symmetric around no trend in temperature, which implies warming or cooling drives similar degree of community turnover.


## Compare covariates across realms
```{r compare across realms}
i <- trends[, !duplicated(STUDY_ID)]; sum(i)
par(mfrow=c(2,3))
beanplot(rarefyID_y ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Latitude (degN)', ll = 0.05)
beanplot(tempave_comb ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature (degC)', ll = 0.05)
beanplot(seas_comb ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Seasonality (degC)', ll = 0.05)
beanplot(npp ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)
beanplot(mass_geomean ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Mass (g)', ll = 0.05)
beanplot(speed_geomean ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Speed (km/hr)', ll = 0.05, log = '')
```
Marine are in generally warmer locations (seawater doesn't freeze)
Marine have much lower seasonality.
Marine and freshwater have some very small masses (plankton), but much of dataset is similar to terrestrial.
Marine has a lot of slow, crawling organisms, but land has plants. Land also has birds (fast).


## Model choice: Jaccard turnover temporal trend vs. temperature trend
First examine how many data points are available
```{r model setup for Jaccard turnover}
# the cases we can compare
cat('Number of data points available:\n')
apply(trends[, .(Jtutrend, temptrend_comb_allyr, tempave_comb, REALM, seas_comb, npp, mass_geomean, speed_geomean)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends[, complete.cases(Jtutrend, temptrend_comb_allyr, REALM, seas_comb, npp, tempave_comb, mass_geomean, speed_geomean)]
cat('Overall:\n')
sum(i)
```

### Choose the variance structure
Try combinations of

- variance scaled to a power of the number of years in the community time-series
- variance scaled to a power of the abs temperature trend
- random intercept for STUDY_ID
- random slope (abs temperature trend) for STUDY_ID
- random intercept for rarefyID (for overdispersion)

And choose the one with lowest AIC
```{r choose variance structure for Jacard turnover}
# fit models for variance structure
fixed <- formula(Jtutrend ~ temptrend_comb_allyr_abs.sc*REALM + 
                     temptrend_comb_allyr_abs.sc*tempave_comb + 
                     temptrend_comb_allyr_abs.sc*seas_comb + 
                     temptrend_comb_allyr_abs.sc*npp + 
                     temptrend_comb_allyr_abs.sc*mass_geomean + 
                     temptrend_comb_allyr_abs.sc*speed_geomean)
i <- trends[, complete.cases(Jtutrend, temptrend_comb_allyr, REALM, seas_comb, npp, tempave_comb, mass_geomean, speed_geomean)]
mods <- vector('list', 0)
mods[[1]] <- gls(fixed, data = trends[i,])
mods[[2]] <- gls(fixed, data = trends[i,], weights = varPower(-0.5, ~nyrBT))
mods[[3]] <- gls(fixed, data = trends[i,], weights = varPower(0.5, ~ abs(temptrend_comb_allyr)))

mods[[4]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, control = lmeControl(opt = "optim"))
mods[[5]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))

mods[[6]] <- lme(fixed, data = trends[i,], random = ~temptrend_comb_allyr_abs.sc | STUDY_ID)
mods[[7]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_comb_allyr_abs.sc, rarefyID = ~1)) # includes overdispersion. new formula so that random slope is only for study level (not enough data to extend to rarefyID).

mods[[8]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[9]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~nyrBT))
mods[[10]] <- lme(fixed, data = trends[i,], random = ~temptrend_comb_allyr_abs.sc|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[11]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_comb_allyr_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))

mods[[12]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend_comb_allyr)))
mods[[13]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~abs(temptrend_comb_allyr)))
mods[[14]] <- lme(fixed, data = trends[i,], random = ~temptrend_comb_allyr_abs.sc|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend_comb_allyr)))
mods[[15]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_comb_allyr_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~abs(temptrend_comb_allyr)))

aics <- sapply(mods, AIC)
minaics <- aics - min(aics)
minaics

```
Chooses the random slopes & intercepts, overdispersion, and variance scaled to number of years.
We haven't dealt with potential testing on the boundary issues here yet.

### Compare fixed effects among models
Try interactions of abs temperature trend with each covariate:

- realm
- average temperature
- seasonality
- speed
- mass
- NPP

```{r choose fixed effects for Jacard turnover}
i <- trends[, complete.cases(Jtutrend, temptrend_comb_allyr, REALM, seas_comb, npp, tempave_comb, mass_geomean, speed_geomean)]

randef <- list(STUDY_ID = ~ temptrend_comb_allyr_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

mods <- vector('list', 0)
mods[[1]] <- lme(Jtutrend ~ 1, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[2]] <- lme(Jtutrend ~ temptrend_comb_allyr.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[3]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[4]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * REALM, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[5]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * tempave_comb.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[6]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * seas_comb.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[7]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * speed_geomean.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[8]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * mass_geomean.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')
mods[[9]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc * npp.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')

mods[[10]] <- lme(Jtutrend ~ temptrend_comb_allyr_abs.sc + speed_geomean.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')

mods[[11]] <- lme(Jtutrend ~ speed_geomean.sc, random = randef, 
                 weights = varef, data = trends[i,], method = 'ML')

# examine models
cat('AIC:\n')
aics <- sapply(mods, AIC) - min(sapply(mods, AIC)); aics
cat(paste0('\nChose model ', which.min(aics), ': ', paste0(mods[[which.min(aics)]]$call$fixed[c(2,1,3)], collapse = ' '), '\n'))
cat('\nModel terms:\n')
summary(mods[[which.min(aics)]])


```
Chooses a model with temperature trend and speed, but maybe not the interaction


## Examine the chosen model
```{r model examination}
bmod <- update(mods[[7]], method = 'REML') # re-fit with REML
summary(bmod)
hist(residuals(bmod))
qqnorm(bmod, ~ resid(., type = 'p'), abline = c(0,1))
qqnorm(bmod, ~ ranef(., level = 1))
qqnorm(bmod, ~ ranef(., level = 2))

plot(fitted(bmod), resid(bmod, type = 'normalized'), col = '#00000033')
plot(trends[i,nyrBT], residuals(bmod, type = 'normalized'), col = '#00000011', log = 'x')
plot(trends[i,temptrend_comb_allyr_abs.sc], residuals(bmod, type = 'normalized'), col = '#00000011')
plot(trends[i,speed_geomean.sc], residuals(bmod, type = 'normalized'), col = '#00000033')

```


Residuals may still be overdispersed despite the observation-level RE? Not sure what else to try.
Still some heterogeneity (cone) for nyrBT and temperature trend, but may be created by so few data points in certain locations on the axis


## Plot the chosen model
```{r examine model}
ggplot(data = trends[i,], aes(x = speed_geomean.sc, y = Jtutrend, color = REALM)) +
    geom_point(shape = 16, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    scale_color_brewer(palette="Set1")

# fix temptrend at low or high
newdat1 <- trends[i, .(STUDY_ID, rarefyID, REALM, temptrend_comb_allyr_abs.sc = quantile(temptrend_comb_allyr_abs.sc, 0.25), speed_geomean, speed_geomean.sc, grp = 1)]
newdat2 <- trends[i, .(STUDY_ID, rarefyID, REALM, temptrend_comb_allyr_abs.sc = quantile(temptrend_comb_allyr_abs.sc, 0.75), speed_geomean, speed_geomean.sc, grp = 2)]
newdat <- rbind(newdat1, newdat2)
newdat$preds <- predict(bmod, newdata = newdat, level = 0)

ggplot(newdat, aes(speed_geomean.sc, preds, color = grp, group = grp)) +
    geom_line()
ggplot(newdat, aes(speed_geomean, preds, color = grp, group = grp)) +
    geom_line()

# fix speed at low or high
newdat1 <- trends[i, .(STUDY_ID, rarefyID, REALM, temptrend_comb_allyr, temptrend_comb_allyr_abs.sc, speed_geomean.sc = quantile(speed_geomean.sc, 0.25), grp = 1)]
newdat2 <- trends[i, .(STUDY_ID, rarefyID, REALM, temptrend_comb_allyr, temptrend_comb_allyr_abs.sc, speed_geomean.sc = quantile(speed_geomean.sc, 0.75), grp = 2)]
newdat <- rbind(newdat1, newdat2)
newdat$preds <- predict(bmod, newdata = newdat, level = 0)

ggplot(newdat, aes(temptrend_comb_allyr_abs.sc, preds, color = grp, group = grp)) +
    geom_line()

ggplot(newdat, aes(temptrend_comb_allyr, preds, color = grp, group = grp)) +
    geom_line()

```


# To do
- plot residuals on top of temperature relationship
- plot turnover vs. speed, look for outliers
- other temporal turnover metrics (Jaccard total, M-H)
