---
title: 'Community dissimilarity through time using beta models: Evaluate and plot models'
output: 
    github_document: default

---
# Prep
First run turnover_vs_temperature_GLMM_fit.R to fit the models.
```{r setup, echo=FALSE}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(glmmTMB) # for ME models
library(lme4) # for plotting random effects dotplot
library(bbmle) # for AICtab
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
library(RColorBrewer)
library(scales) # for defining custom scales in ggplot
library(DHARMa) # for model diagnostics
library(here) # for relative paths
library(performance) # for r2 calculations

options(width=500) # turn off most text wrapping

signedsqrt = function(x) sign(x)*sqrt(abs(x))
signedsq = function(x) sign(x) * x^2
signedsqrttrans <- trans_new(name = 'signedsqrt', transform = signedsqrt, inverse = signedsq)


# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 

# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends3 <- fread(here('output', 'turnover_w_covariates3.csv.gz'))
trends5 <- fread(here('output', 'turnover_w_covariates5.csv.gz'))
trends10 <- fread(here('output', 'turnover_w_covariates10.csv.gz'))
trends20 <- fread(here('output', 'turnover_w_covariates20.csv.gz'))

```


# Summary stats
## Examine how many data points are available
### Just turnover and non-zero tempchange
#### 3-years
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trends3), '\n')
cat('# studies: ', trends3[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends3[, length(unique(rarefyID))], '\n')
trends3[!duplicated(rarefyID), table(REALM)]
trends3[!duplicated(rarefyID), table(taxa_mod)]
trends3[!duplicated(rarefyID), table(taxa_mod, REALM)]
```
#### 5-years
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trends5), '\n')
cat('# studies: ', trends5[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends5[, length(unique(rarefyID))], '\n')
trends5[!duplicated(rarefyID), table(REALM)]
trends5[!duplicated(rarefyID), table(taxa_mod)]
trends5[!duplicated(rarefyID), table(taxa_mod, REALM)]
```

#### 10-years
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trends10), '\n')
cat('# studies: ', trends10[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends10[, length(unique(rarefyID))], '\n')
trends10[!duplicated(rarefyID), table(REALM)]
trends10[!duplicated(rarefyID), table(taxa_mod)]
trends10[!duplicated(rarefyID), table(taxa_mod, REALM)]
```

#### 20-years
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trends20), '\n')
cat('# studies: ', trends20[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends20[, length(unique(rarefyID))], '\n')
trends20[!duplicated(rarefyID), table(REALM)]
trends20[!duplicated(rarefyID), table(taxa_mod)]
trends20[!duplicated(rarefyID), table(taxa_mod, REALM)]
```

### With all covariates
Use Bowler for human impact
#### 3-year
```{r sample size Jtu covariates}
# the cases we can compare
apply(trends3[, .(Jtu.sc, REALM, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends3[, complete.cases(Jtu.sc, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
cat('Overall # dissimilarities: ', sum(i), '\n')
cat('# studies: ', trends3[i, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends3[i, length(unique(rarefyID))], '\n')
trends3[i & !duplicated(rarefyID), table(REALM)]
trends3[i & !duplicated(rarefyID), table(taxa_mod)]
trends3[i & !duplicated(rarefyID), table(taxa_mod, REALM)]
```

# Models
## Choose the variance structure
Try combinations of

- beta errors
- variance scaled to Nspp or to Nspp + REALM
- random intercept for taxa_mod2, STUDY_ID, rarefyID
- random slopes for duration

Did not try random intercept for compID (for overdispersion), since had trouble converging.
Use a full set of fixed effects and choose the RE with lowest AIC.
Many of the 3- and 5-year models with random slopes had singular convergence issues.

### 3-years
Best is beta errors, durationlog.sc, intercepts for STUDY_ID and rarefyID, variance scaled to nspp + realm.
Models with slopes didn't converge.
We haven't dealt with potential testing on the boundary issues here. Note that a number of models with random slopes did not converge.
```{r choose variance structure, echo=FALSE}
# comment out the ones that didn't converge (see log files from fitting)
mods3 <- vector('list', length = 17)
i = 1
if(file.exists(here('temp', 'modRFgauss3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFgauss3.rds')); names(mods3)[i] <- 'modRFgauss3'; i = i+1}
if(file.exists(here('temp', 'modRFbeta3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFbeta3.rds')); names(mods3)[i] <- 'modRFbeta3'; i = i+1}
if(file.exists(here('temp', 'modRFrID3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFrID3.rds')); names(mods3)[i] <- 'modRFrID3'; i = i+1}
if(file.exists(here('temp', 'modRF2lev3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRF2lev3.rds')); names(mods3)[i] <- 'modRF2lev3'; i = i+1}
if(file.exists(here('temp', 'modRFnestedRE3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFnestedRE3.rds')); names(mods3)[i] <- 'modRFnestedRE3'; i = i+1}
#if(file.exists(here('temp', 'modRFslopeRE3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFslopeRE3.rds')); names(mods3)[i] <- 'modRFslopeRE3'; i = i+1}
# if(file.exists(here('temp', 'modRFslopeRE2lev3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFslopeRE2lev3.rds')); names(mods3)[i] <- 'modRFslopeRE2lev3'; i = i+1}
if(file.exists(here('temp', 'modRFdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdisp3.rds')); names(mods3)[i] <- 'modRFdisp3'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRF2levdisp3.rds')); names(mods3)[i] <- 'modRF2levdisp3'; i = i+1}
# if(file.exists(here('temp', 'modRFslopeRE2levdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisp3.rds')); names(mods3)[i] <- 'modRFslopeRE2levdisp3'; i = i+1}
#if(file.exists(here('temp', 'modRFslopeREdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFslopeREdisp3.rds')); names(mods3)[i] <- 'modRFslopeREdisp3'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisprealm3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRF2levdisprealm3.rds')); names(mods3)[i] <- 'modRF2levdisprealm3'; i = i+1}
# if(file.exists(here('temp', 'modRFslopeRE2levdisprealm3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisprealm3.rds')); names(mods3)[i] <- 'modRFslopeRE2levdisprealm3'; i = i+1}

# with log duration
if(file.exists(here('temp', 'modRFdurlog2levdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisp3.rds')); names(mods3)[i] <- 'modRFdurlog2levdisp3'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisp3.rds')); names(mods3)[i] <- 'modRFdurlogslopeRE2levdisp3'; i = i+1} # convergence error
# if(file.exists(here('temp', 'modRFdurlogslopeREdisp3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdurlogslopeREdisp3.rds')); names(mods3)[i] <- 'modRFdurlogslopeREdisp3'; i = i+1} # convergence error
if(file.exists(here('temp', 'modRFdurlog2levdisprealm3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisprealm3.rds')); names(mods3)[i] <- 'modRFdurlog2levdisprealm3'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisprealm3.rds'))){ mods3[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisprealm3.rds')); names(mods3)[i] <- 'modRFdurlogslopeRE2levdisprealm3'; i = i+1} # convergence error


modes <- sapply(X = mods3, FUN = mode)
aics <- sapply(X = mods3[modes == 'list'], FUN = AIC)
aics - min(aics)
```

### 5-years
Best is beta errors, duration.sc, slope & intercepts for STUDY_ID and rarefyID, variance scaled to nspp + realm.
We haven't dealt with potential testing on the boundary issues here. Note that a number of models with random slopes did not converge.
```{r choose variance structure, echo=FALSE}
mods5 <- vector('list', length = 12)
i = 1
if(file.exists(here('temp', 'modRFgauss5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFgauss5.rds')); names(mods5)[i] <- 'modRFgauss5'; i = i+1}
if(file.exists(here('temp', 'modRFbeta5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFbeta5.rds')); names(mods5)[i] <- 'modRFbeta5'; i = i+1}
if(file.exists(here('temp', 'modRFrID5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFrID5.rds')); names(mods5)[i] <- 'modRFrID5'; i = i+1}
if(file.exists(here('temp', 'modRF2lev5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRF2lev5.rds')); names(mods5)[i] <- 'modRF2lev5'; i = i+1}
if(file.exists(here('temp', 'modRFnestedRE5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFnestedRE5.rds')); names(mods5)[i] <- 'modRFnestedRE5'; i = i+1}
#if(file.exists(here('temp', 'modRFslopeRE5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFslopeRE5.rds')); names(mods5)[i] <- 'modRFslopeRE5'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2lev5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFslopeRE2lev5.rds')); names(mods5)[i] <- 'modRFslopeRE2lev5'; i = i+1}
if(file.exists(here('temp', 'modRFdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdisp5.rds')); names(mods5)[i] <- 'modRFdisp5'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRF2levdisp5.rds')); names(mods5)[i] <- 'modRF2levdisp5'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisp5.rds')); names(mods5)[i] <- 'modRFslopeRE2levdisp5'; i = i+1}
#if(file.exists(here('temp', 'modRFslopeREdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFslopeREdisp5.rds')); names(mods5)[i] <- 'modRFslopeREdisp5'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisprealm5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRF2levdisprealm5.rds')); names(mods5)[i] <- 'modRF2levdisprealm5'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisprealm5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisprealm5.rds')); names(mods5)[i] <- 'modRFslopeRE2levdisprealm5'; i = i+1}

# with log duration
if(file.exists(here('temp', 'modRFdurlog2levdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisp5.rds')); names(mods5)[i] <- 'modRFdurlog2levdisp5'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisp5.rds')); names(mods5)[i] <- 'modRFdurlogslopeRE2levdisp5'; i = i+1} # convergence error
# if(file.exists(here('temp', 'modRFdurlogslopeREdisp5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdurlogslopeREdisp5.rds')); names(mods5)[i] <- 'modRFdurlogslopeREdisp5'; i = i+1} # convergence error
if(file.exists(here('temp', 'modRFdurlog2levdisprealm5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisprealm5.rds')); names(mods5)[i] <- 'modRFdurlog2levdisprealm5'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisprealm5.rds'))){ mods5[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisprealm5.rds')); names(mods5)[i] <- 'modRFdurlogslopeRE2levdisprealm5'; i = i+1} # convergence error

modes <- sapply(X = mods5, FUN = mode)
aics <- sapply(X = mods5[modes == 'list'], FUN = AIC)
aics - min(aics)
```

### 10-years
This chooses beta errors, random slopes (duration.sc) & intercepts for STUDY_ID and rarefyID, and variance scaled to nspp + realm.
We haven't dealt with potential testing on the boundary issues here.
```{r choose variance structure, echo=FALSE}
mods10 <- vector('list', length = 17)
i = 1
if(file.exists(here('temp', 'modRFgauss10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFgauss10.rds')); names(mods10)[i] <- 'modRFgauss10'; i = i+1}
if(file.exists(here('temp', 'modRFbeta10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFbeta10.rds')); names(mods10)[i] <- 'modRFbeta10'; i = i+1}
if(file.exists(here('temp', 'modRFrID10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFrID10.rds')); names(mods10)[i] <- 'modRFrID10'; i = i+1}
if(file.exists(here('temp', 'modRF2lev10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2lev10.rds')); names(mods10)[i] <- 'modRF2lev10'; i = i+1}
if(file.exists(here('temp', 'modRFnestedRE10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFnestedRE10.rds')); names(mods10)[i] <- 'modRFnestedRE10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE10.rds')); names(mods10)[i] <- 'modRFslopeRE10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2lev10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2lev10.rds')); names(mods10)[i] <- 'modRFslopeRE2lev10'; i = i+1}
if(file.exists(here('temp', 'modRFdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdisp10.rds')); names(mods10)[i] <- 'modRFdisp10'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2levdisp10.rds')); names(mods10)[i] <- 'modRF2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisp10.rds')); names(mods10)[i] <- 'modRFslopeRE2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeREdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeREdisp10.rds')); names(mods10)[i] <- 'modRFslopeREdisp10'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRF2levdisprealm10.rds')); names(mods10)[i] <- 'modRF2levdisprealm10'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisprealm10.rds')); names(mods10)[i] <- 'modRFslopeRE2levdisprealm10'; i = i+1}

# with log duration
if(file.exists(here('temp', 'modRFdurlog2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisp10.rds')); names(mods10)[i] <- 'modRFdurlog2levdisp10'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisp10.rds')); names(mods10)[i] <- 'modRFdurlogslopeRE2levdisp10'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeREdisp10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeREdisp10.rds')); names(mods10)[i] <- 'modRFdurlogslopeREdisp10'; i = i+1} # convergence error
if(file.exists(here('temp', 'modRFdurlog2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisprealm10.rds')); names(mods10)[i] <- 'modRFdurlog2levdisprealm10'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisprealm10.rds'))){ mods10[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisprealm10.rds')); names(mods10)[i] <- 'modRFdurlogslopeRE2levdisprealm10'; i = i+1}

modes <- sapply(X = mods10, FUN = mode)
aics <- sapply(X = mods10[modes == 'list'], FUN = AIC)
aics - min(aics)
```

### 20-years
This chooses beta errors, random slopes (duration.sc) & intercepts for STUDY_ID and rarefyID, and variance scaled to nspp + realm.
We haven't dealt with potential testing on the boundary issues here.
```{r choose variance structure, echo=FALSE}
mods20 <- vector('list', length = 17)
i = 1
if(file.exists(here('temp', 'modRFgauss20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFgauss20.rds')); names(mods20)[i] <- 'modRFgauss20'; i = i+1}
if(file.exists(here('temp', 'modRFbeta20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFbeta20.rds')); names(mods20)[i] <- 'modRFbeta20'; i = i+1}
if(file.exists(here('temp', 'modRFrID20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFrID20.rds')); names(mods20)[i] <- 'modRFrID20'; i = i+1}
if(file.exists(here('temp', 'modRF2lev20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRF2lev20.rds')); names(mods20)[i] <- 'modRF2lev20'; i = i+1}
if(file.exists(here('temp', 'modRFnestedRE20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFnestedRE20.rds')); names(mods20)[i] <- 'modRFnestedRE20'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFslopeRE20.rds')); names(mods20)[i] <- 'modRFslopeRE20'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2lev20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFslopeRE2lev20.rds')); names(mods20)[i] <- 'modRFslopeRE2lev20'; i = i+1}
if(file.exists(here('temp', 'modRFdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdisp20.rds')); names(mods20)[i] <- 'modRFdisp20'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRF2levdisp20.rds')); names(mods20)[i] <- 'modRF2levdisp20'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisp20.rds')); names(mods20)[i] <- 'modRFslopeRE2levdisp20'; i = i+1}
# if(file.exists(here('temp', 'modRFslopeREdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFslopeREdisp20.rds')); names(mods20)[i] <- 'modRFslopeREdisp20'; i = i+1}
if(file.exists(here('temp', 'modRF2levdisprealm20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRF2levdisprealm20.rds')); names(mods20)[i] <- 'modRF2levdisprealm20'; i = i+1}
if(file.exists(here('temp', 'modRFslopeRE2levdisprealm20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFslopeRE2levdisprealm20.rds')); names(mods20)[i] <- 'modRFslopeRE2levdisprealm20'; i = i+1}

# with log duration
if(file.exists(here('temp', 'modRFdurlog2levdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisp20.rds')); names(mods20)[i] <- 'modRFdurlog2levdisp20'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisp20.rds')); names(mods20)[i] <- 'modRFdurlogslopeRE2levdisp20'; i = i+1}
# if(file.exists(here('temp', 'modRFdurlogslopeREdisp20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdurlogslopeREdisp20.rds')); names(mods20)[i] <- 'modRFdurlogslopeREdisp20'; i = i+1} # convergence error
if(file.exists(here('temp', 'modRFdurlog2levdisprealm20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdurlog2levdisprealm20.rds')); names(mods20)[i] <- 'modRFdurlog2levdisprealm20'; i = i+1}
if(file.exists(here('temp', 'modRFdurlogslopeRE2levdisprealm20.rds'))){ mods20[[i]] <- readRDS(here('temp', 'modRFdurlogslopeRE2levdisprealm20.rds')); names(mods20)[i] <- 'modRFdurlogslopeRE2levdisprealm20'; i = i+1}

modes <- sapply(X = mods20, FUN = mode)
aics <- sapply(X = mods20[modes == 'list'], FUN = AIC)
aics - min(aics)
```

### Summary of the chosen model
### 3 years
```{r, echo=FALSE}
summary(mods3[[which(names(mods3)=='modRF2levdisprealm3')]])
```

### 5 years
```{r, echo=FALSE}
summary(mods5[[which(names(mods5)=='modRFslopeRE2levdisprealm5')]])
```

### 10 years
```{r, echo=FALSE}
summary(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]])
```

### 20 years
```{r, echo=FALSE}
summary(mods20[[which(names(mods20)=='modRFslopeRE2levdisprealm20')]])
```

### Plot random effects
No clear discontinuities. Looks ok.
```{r, echo=FALSE}
lme4:::dotplot.ranef.mer(ranef(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]])$cond)
```

### DHARMa model evaluation
```{r, echo=FALSE, eval=FALSE}
## run the top part by hand if needed. takes ~3 hrs
if(FALSE) {
  i10 <- trends10[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, 
                                   endothermfrac.sc, microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]

  res_RFslopeRE2levdisprealm10 <- simulateResiduals(mods10[[which(names(mods10)=='modRFslopeRE2levdisprealm10')]], n = 250)
  
  saveRDS(res_RFslopeRE2levdisprealm10, file = 'temp/res_RFslopeRE2levdisprealm10.rds')
} else {
  if(file.exists('temp/res_RFslopeRE2levdisprealm10.rds')) res_RFslopeRE2levdisprealm10 <- readRDS('temp/res_res_RFslopeRE2levdisprealm10.rds')
  i10 <- trends10[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, 
                                   endothermfrac.sc, microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
}
```

#### Plot residuals for model fit to all data
There are many outliers and signs of underdispersion, but this is probably ok.
```{r, echo=FALSE, eval=FALSE}
if(exists('res_RFslopeRE2levdisprealm10')){
  plot(res_RFslopeRE2levdisprealm10)
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$tempchange_abs.sc[i10], xlab = 'tempchange_abs.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$REALM[i10], xlab = 'REALM', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$tempave_metab.sc[i10], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$duration.sc[i10], xlab = 'duration.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$mass.sc[i10], xlab = 'mass.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$endothermfrac.sc[i10], xlab = 'endothermfrac.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$microclim.sc[i10], xlab = 'microclim.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, trends10$npp.sc[i10], xlab = 'npp.sc', main = '')
  plotResiduals(res_RFslopeRE2levdisprealm10, form=trends10$human_bowler.sc[i10], xlab = 'human_bowler.sc', main = '')
}
```


#### Overdispersion
There is underdispersion.
```{r, echo=FALSE, eval=FALSE}
testDispersion(res_RFslopeRE2levdisprealm10)

```

#### Near-zero-inflation
(Zero values have been transformed to slightly >0, so can't test zero-inflation directly)

Near-zero values are not inflated.
```{r, eval=FALSE}
countNearZero <- function(x) sum(x < 0.0001)
testGeneric(res_RFslopeRE2levdisprealm10, summary = countNearZero, alternative = 'greater')

```

#### Near-one-inflation
Near-one values are not inflated.
```{r, eval=FALSE}
countNearOne <- function(x) sum(x > 0.9999)
testGeneric(res_RFslopeRE2levdisprealm10, summary = countNearOne, alternative = 'greater')

```


## Temperature, temperature slope & duration models
### Load the models
```{r temperature, temp slope and duration, eval=FALSE, echo=FALSE}
# all years
# modTdT3 <- readRDS('temp/modTdT3.rds')
modTdT5 <- readRDS('temp/modTdT5.rds')
modTdT10 <- readRDS('temp/modTdT10.rds')
modTdT20 <- readRDS('temp/modTdT20.rds')
```

### Summary
```{r summary of TD mods, eval=FALSE, echo=FALSE}
summary(modTdT5)
summary(modTdT10)
summary(modTdT20)
```



### Plot the coefficients
Not used
```{r modTD coefs, eval=FALSE, echo=FALSE}

coefs1 <- summary(modTDJtu)$coefficients$cond
coefs2 <- summary(modTDJbeta)$coefficients$cond
coefs3 <- summary(modTDHorn)$coefficients$cond

varstoplot <- unique(c(rownames(coefs1), rownames(coefs2), rownames(coefs3)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs1) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
xlims <- range(c(coefs1[rows1_1,1] - 1.96*coefs1[rows1_1,2], coefs1[rows1_1,1] + 1.96*coefs1[rows1_1,2], 
                  coefs2[rows1_2,1] - 1.96*coefs2[rows1_2,2], coefs2[rows1_2,1] + 1.96*coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - 1.96*coefs3[rows1_3,2], coefs3[rows1_3,1] + 1.96*coefs3[rows1_3,2]))

cols <- brewer.pal(3, 'Dark2') # for Jtu, Jbeta and Horn models
pchs <- c(16, 16, 16)
offs <- c(0.1, 0, -0.1) # offset vertically for each model

par(las = 1, mai = c(0.5, 4, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(0.9,length(varstoplot)+0.1), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs1)){
    x = coefs1[rownames(coefs1) == varstoplot[i], 1] # the coef in col 1
    se = coefs1[rownames(coefs1) == varstoplot[i], 2] # the SE in col 2
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
}
legend('bottomright', col = cols, pch = 16, lwd = 1, legend = c('Jtu', 'Jbeta', 'Horn'), cex = 0.5)
```
abs(tempchange) is in degC

duration is in years

### Plot the response
Not used
```{r plot TD response, eval=FALSE, echo=FALSE}
if(!file.exists('temp/modTDJtu_preds.rds')) {
  newdat <- expand.grid(duration = c(1, 50), tempchange = seq(0.01, 5, length.out = 5), 
                      taxa_mod2 = NA, STUDY_ID = NA, rarefyID = NA,
                      Nspp = 60) # no REs, ave number of species
  scaling <- fread('output/turnover_w_covariates_scaling.csv')
  cent <- scaling$center[scaling$var == 'tempchange_abs.sc'] # centering
  scl <- scaling$scale[scaling$var == 'tempchange_abs.sc'] # scaling factor for tempchange
  newdat$tempchange_abs.sc <- (abs(newdat$tempchange)-cent)/scl
  cent <- scaling$center[scaling$var == 'nspp.sc'] # centering
  scl <- scaling$scale[scaling$var == 'nspp.sc'] # scaling factor for tempchange
  newdat$nspp.sc <- (log(newdat$Nspp) - cent)/scl
  Jtu <- predict(modTDJtu, newdata = newdat, se.fit = TRUE, re.form = NA, type = 'response')
  newdat$durationfac <- as.factor(newdat$duration)
  newdat$Jtu <- Jtu$fit
  newdat$Jtu.se <- Jtu$se.fit
  
  saveRDS(newdat, file = 'temp/modTDJtu_preds.rds')
} else {
  newdat <- readRDS('temp/modTDJtu_preds.rds')
}

ggplot(newdat, aes(tempchange, Jtu, color = durationfac, group = durationfac)) +
  geom_line() +
  geom_ribbon(aes(ymin=Jtu - 1.96*Jtu.se, ymax=Jtu + 1.96*Jtu.se), alpha = 0.1, linetype = 'blank')
```

## Temperature*temperature slope & duration models
### Load the models
```{r temperature by temp slope and duration, eval=FALSE, echo=FALSE}
# modTdT3 <- readRDS('temp/modTdT3.rds')
modTdTT5 <- readRDS('temp/modTdTT5.rds')
modTdTT10 <- readRDS('temp/modTdTT10.rds')
modTdTT20 <- readRDS('temp/modTdTT20.rds')

modTdTT5Horn <- readRDS('temp/modTdTT5Horn.rds')
modTdTT10Horn <- readRDS('temp/modTdTT10Horn.rds')
modTdTT20Horn <- readRDS('temp/modTdTT20Horn.rds')
```

### Plot the response
#### Set up and write to file
Only run this once.
```{r, eval=FALSE, echo=FALSE}
scaling5 <- fread('output/turnover_w_covariates_scaling5.csv') # the scalings
scaling10 <- fread('output/turnover_w_covariates_scaling10.csv') # the scalings
scaling20 <- fread('output/turnover_w_covariates_scaling20.csv') # the scalings

# find a study in all 3 datasets to use as the random effect
# ranef(modTdT20Horn))

# set up the variables to plot
vars <- data.frame(vars = c('tempchange_abs', 'tempave_metab'),
                   min =      c(0,    -10), 
                   max =      c(0.7,    40),
                   len =      c(100,  100),
                   discrete = c(F,    F),
                   stringsAsFactors = FALSE)
basetab5 <- data.table(expand.grid(start = 1:4, end = 2:5, nspp.sc = 0, REALM = 'Terrestrial', STUDY_ID = 195L, rarefyID = '195_289043', tempchange_abs.sc = 0, tempave_metab.sc = 0)) # the base case
basetab5[, duration := end - start]
if(scaling5[var == 'duration.sc', log]) basetab5[, duration.sc := (log(duration) - scaling5[var == 'duration.sc', center])/scaling5[var == 'duration.sc', scale]]
if(!scaling5[var == 'duration.sc', log]) basetab5[, duration.sc := (duration - scaling5[var == 'duration.sc', center])/scaling5[var == 'duration.sc', scale]]
basetab5 <- basetab5[duration > 0, ]

basetab10 <- data.table(expand.grid(start = 1:9, end = 2:10, nspp.sc = 0, REALM = 'Terrestrial', STUDY_ID = 195L, rarefyID = '195_289043', tempchange_abs.sc = 0, tempave_metab.sc = 0)) # the base case
basetab10[, duration := end - start]
if(scaling10[var == 'duration.sc', log]) basetab10[, duration.sc := (log(duration) - scaling10[var == 'duration.sc', center])/scaling10[var == 'duration.sc', scale]]
if(!scaling10[var == 'duration.sc', log]) basetab10[, duration.sc := (duration - scaling10[var == 'duration.sc', center])/scaling10[var == 'duration.sc', scale]]
basetab10 <- basetab10[duration > 0, ]

basetab20 <- data.table(expand.grid(start = 1:19, end = 2:20, nspp.sc = 0, REALM = 'Terrestrial', STUDY_ID = 195L, rarefyID = '195_289043', tempchange_abs.sc = 0, tempave_metab.sc = 0)) # the base case
basetab20[, duration := end - start]
if(scaling20[var == 'duration.sc', log]) basetab20[, duration.sc := (log(duration) - scaling20[var == 'duration.sc', center])/scaling20[var == 'duration.sc', scale]]
if(!scaling20[var == 'duration.sc', log]) basetab20[, duration.sc := (duration - scaling20[var == 'duration.sc', center])/scaling20[var == 'duration.sc', scale]]
basetab20 <- basetab20[duration > 0, ]

# make the data frames for each interaction to plot                
for(j in 1:nrow(vars)){
    # set up the main effects
    if(scaling5[var == paste0(vars$vars[j], '.sc'), log]){
        thisdat <- data.frame(new = 10^seq(log10(vars$min[j]), log10(vars$max[j]), length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    } 
    if(!scaling5[var == paste0(vars$vars[j], '.sc'), log]){
        thisdat <- data.frame(new = seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    }
    names(thisdat) <- c(vars$vars[j], 'var')
    
    # scale the variable
    cent <- scaling5[var == paste0(vars$vars[j], '.sc'), center]
    scl <- scaling5[var == paste0(vars$vars[j], '.sc'), scale]
    lg <- scaling5[var == paste0(vars$vars[j], '.sc'), log]
    plus <- scaling5[var == paste0(vars$vars[j], '.sc'), plus]
    if(is.null(cent)) cent <- 0
    if(!is.null(cent) & !is.null(scl)){
        if(lg) thisdat[[paste0(vars$var[j], '.sc')]] <- (log(thisdat[[vars$vars[j]]] + plus) - cent)/scl
        if(!lg) thisdat[[paste0(vars$var[j], '.sc')]] <- (thisdat[[vars$var[j]]] - cent)/scl
    }
    
    # merge with the rest of the columns
    expand.grid.df <- function(...) Reduce(function(...) merge(..., by=NULL), list(...)) # https://stackoverflow.com/questions/11693599/alternative-to-expand-grid-for-data-frames
    colnamestouse <- setdiff(colnames(basetab5), paste0(vars$vars[j], '.sc'))
    thisdat5 <- expand.grid.df(thisdat, basetab5[, ..colnamestouse])
    thisdat10 <- expand.grid.df(thisdat, basetab10[, ..colnamestouse])
    thisdat20 <- expand.grid.df(thisdat, basetab20[, ..colnamestouse])
    
    # merge with the previous iterations
    if(j == 1){
      newdat5 <- thisdat5
      newdat10 <- thisdat10
      newdat20 <- thisdat20
    }
    if(j > 1){
        colstoadd <- setdiff(colnames(thisdat5), colnames(newdat5))
        for(toadd in colstoadd){
            newdat5[[toadd]] <- NA
            newdat10[[toadd]] <- NA
            newdat20[[toadd]] <- NA
        }
        
        colstoadd2 <- setdiff(colnames(newdat5), colnames(thisdat5))
        for(toadd in colstoadd2){
            thisdat5[[toadd]] <- NA
            thisdat10[[toadd]] <- NA
            thisdat20[[toadd]] <- NA
        }
        
        newdat5 <- rbind(newdat5, thisdat5)
        newdat10 <- rbind(newdat10, thisdat10)
        newdat20 <- rbind(newdat20, thisdat20)
    } 
}

# predict (takes a few minutes on Annotate)
preds5 <- predict(modTdTT5Horn, newdat = newdat5, type = 'response', se.fit = TRUE, re.form = NA)
preds10 <- predict(modTdTT10Horn, newdat = newdat10, type = 'response', se.fit = TRUE, re.form = NA)
preds20 <- predict(modTdTT20Horn, newdat = newdat20, type = 'response', se.fit = TRUE, re.form = NA)

newdat5$Horn.sc <- preds5$fit
newdat5$Horn.sc.se <- preds5$se.fit
newdat10$Horn.sc <- preds10$fit
newdat10$Horn.sc.se <- preds10$se.fit
newdat20$Horn.sc <- preds20$fit
newdat20$Horn.sc.se <- preds20$se.fit

# write out
write.csv(newdat5, 'temp/modTdTT5_preds.csv', row.names = FALSE)
write.csv(newdat10, 'temp/modTdTT10_preds.csv', row.names = FALSE)
write.csv(newdat20, 'temp/modTdTT20_preds.csv', row.names = FALSE)

```

#### Plot
```{r}
# load predictions
newdat5 <- fread('temp/modTdTT5_preds.csv')
newdat10 <- fread('temp/modTdTT10_preds.csv')
newdat20 <- fread('temp/modTdTT20_preds.csv')

# calc duration slope per temp:tempchange level
slope <- function(duration, Horn.sc){
  mod <- lm(Horn.sc ~ duration)
  return(coef(mod)[2])
}
slopese <- function(duration, Horn.sc){
  mod <- lm(Horn.sc ~ duration)
  return(sqrt(diag(vcov(mod))))
}
slope5 <- newdat5[, .(Hornslope = slope(duration, Horn.sc), 
                      Hornslopese = slopese(duration, Horn.sc)), by = .(tempchange_abs, tempave_metab)]

p1 <- ggplot(slope5, aes(x=tempchange_abs, y=Hornslope, ymin=Hornslope-Hornslopese, ymax=Hornslope+Hornslopese)) +
  geom_line() + 
  geom_ribbon(alpha=0.5)
  
p2 <- ggplot(slope5, aes(tempave_metab, Hornslope)) +
  geom_line()

grid.arrange(p1, p2, ncol = 2)
```

## Full models w/ only main effects
These are a main focus right now, though the dissimilarities have not been corrected for alpha & gamma diversity as in Chase et al Ecosphere.

### Load the models
```{r, echo=FALSE}
# all years
if(file.exists('temp/modMainTdTSeMiNPNspMaHuJtu.rds')) modMainJtu <- readRDS('temp/modMainTdTSeMiNPNspMaHuJtu.rds') # mass, endoecto, microclimate, NPP, humans
if(file.exists('temp/modMainTdTSeMiNPNspMaHu1yrJtu.rds')) modMain1yrJtu <- readRDS('temp/modMainTdTSeMiNPNspMaHu1yrJtu.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu5yrJtu.rds')) modMain5yrJtu <- readRDS('temp/modMainTdTSeMiNPNspMaHu5yrJtu.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu10yrJtu.rds')) modMain10yrJtu <- readRDS('temp/modMainTdTSeMiNPNspMaHu10yrJtu.rds')

if(file.exists('temp/modMainTdTSeMiNPNspMaHuJbeta.rds')) modMainJbeta <- readRDS('temp/modMainTdTSeMiNPNspMaHuJbeta.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu1yrJbeta.rds')) modMain1yrJbeta <- readRDS('temp/modMainTdTSeMiNPNspMaHu1yrJbeta.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu10yrJbeta.rds')) modMain10yrJbeta <- readRDS('temp/modMainTdTSeMiNPNspMaHu10yrJbeta.rds')

if(file.exists('temp/modMainTdTSeMiNPNspMaHuHorn.rds')) modMainHorn <- readRDS('temp/modMainTdTSeMiNPNspMaHuHorn.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu1yrHorn.rds')) modMain1yrHorn <- readRDS('temp/modMainTdTSeMiNPNspMaHu1yrHorn.rds')
if(file.exists('temp/modMainTdTSeMiNPNspMaHu10yrHorn.rds')) modMain10yrHorn <- readRDS('temp/modMainTdTSeMiNPNspMaHu10yrHorn.rds')

```

### Summary
#### Jtu
##### All
```{r, echo=FALSE}
if(exists('modMainJtu')) summary(modMainJtu)
```

##### 1 yr
```{r, echo=FALSE}
if(exists('modMain1yrJtu')) summary(modMain1yrJtu)
```

##### 10 yr
```{r, echo=FALSE}
if(exists('modMain10yrJtu'))  summary(modMain10yrJtu)
```

#### Horn
##### All
```{r, echo=FALSE}
if(exists('modMainHorn')) summary(modMainHorn)
```

##### 1 yr
```{r, echo=FALSE}
if(exists('modMain1yrHorn')) summary(modMain1yrHorn)
```

##### 10 yr
```{r, echo=FALSE}
if(exists('modMain10yrHorn'))summary(modMain10yrHorn)
```


### Plot the response
#### Set up and write to file
Only run this once.
```{r, eval=FALSE, echo=FALSE}
scaling <- fread('output/turnover_w_covariates_scaling.csv') # the scalings

# set up the variables to plot
# if variable is logged before scaling (see 'center and scale' above), then need to mark it here and express the limits on a log10 scale (even though log transforming is log)
vars <- data.frame(vars = c('tempchange_abs', 'tempchange_abs', 'tempchange_abs', 'tempave_metab', 'seas', 'microclim', 'mass', 'npp', 
                            'duration', 'nspp', 'human_bowler', 'human_bowler'),
                   min =      c(0,    0,   0,   0,   0,   0,   -1,  1.9, 0.5, 0,   0,   0), 
                   max =      c(4,    4,   4,   30,  16,  0.8, 4,   3.7, 2,   2.5, 1,   1),
                   log =      c(F,    F,   F,   F,   F,   T,   T,   T,   T,   T,   T,   T),
                   len =      c(100,  100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100),
                   discrete = c(F,    F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F),
                   plus =     c(0,    0,   0,   0,   0,   0,   0,   0,   0,   0,   1,   1), # what to add before log-scaling
                   REALM = c('Freshwater', 'Marine', 'Terrestrial', rep('Terrestrial', 7), 'Terrestrial', 'Marine'),
                   REALM2 = c('TerrFresh', 'Marine', rep('TerrFresh', 9), 'Marine'),
                   stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', 
                      tempchange = -0.0001,
                      tempchange_abs.sc = 0.0001,
                      tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      seas.sc = mean(seas.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      nspp.sc = mean(nspp.sc, na.rm=TRUE),
                      nspp.sc = 60, 
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     tempchange = -0.0001,
                     tempchange_abs.sc = 0.0001,
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                     seas.sc = mean(seas.sc, na.rm=TRUE), 
                     microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     nspp.sc = mean(nspp.sc, na.rm=TRUE),
                     nspp.sc = 60, 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    tempchange = -0.0001,
                    tempchange_abs.sc = 0.0001,
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    seas.sc = mean(seas.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    nspp.sc = mean(nspp.sc, na.rm=TRUE),
                    nspp.sc = 60, 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)

basetab[, ':='(duration.sc = 0, nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_504467')]

# make the data frames for each interaction to plot                
for(j in 1:nrow(vars)){
    # set up the main effects
    if(vars$log[j]){
        thisdat <- data.frame(new = 10^seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    } 
    if(!vars$log[j]){
        thisdat <- data.frame(new = seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    }
    names(thisdat) <- c(vars$vars[j], 'var')
    
    # scale the variable
    cent <- scaling$center[scaling$var == paste0(vars$vars[j], '.sc')]
    scl <- scaling$scale[scaling$var == paste0(vars$vars[j], '.sc')]
    if(is.null(cent)) cent <- 0
    if(!is.null(cent) & !is.null(scl)){
        if(vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (log(thisdat[[vars$vars[j]]] + vars$plus[j]) - cent)/scl
        if(!vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (thisdat[[vars$var[j]]] - cent)/scl
    }
    
    # merge with the rest of the columns
    # use realm-specific averages for human impacts
    colnamestouse <- setdiff(colnames(basetab), paste0(vars$vars[j], '.sc'))
    if(vars$vars[j] != 'human_bowler'){
        thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
    }
    if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Terrestrial'){
        thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
    }
    if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Marine'){
        thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
    }
    
    # add realm
    thisdat$REALM <- vars$REALM[j]
    thisdat$REALM2 <- vars$REALM2[j]
    
    # merge with the previous iterations
    if(j == 1) newdat <- thisdat
    if(j > 1){
        colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
        for(toadd in colstoadd){
            newdat[[toadd]] <- NA
        }
        
        colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
        for(toadd in colstoadd2){
            thisdat[[toadd]] <- NA
        }
        
        newdat <- rbind(newdat, thisdat)
    } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$tempchange_abs.sc[1:6] <- rep(0.0001, 6)
newdat$tempchange[1:6] <- rep(c(0.0001, -0.0001), 3)
newdat$var[1:6] <- 'test'

# write out
write.csv(newdat, 'temp/newdata_main_only.csv', row.names = FALSE)

```

Now go run on the command line
```
nohup code/turnover_vs_temperature_GLMM_pred.R temp/modMainTdTSeMiNPNspMaHu1yrJtu.rds temp/newdata_main_only.csv temp/modMainTdTSeMiNPNspMaHu1yrJtu_preds.rds >logs/turnover_vs_temperature_GLMMmodMainTdTSeMiNPNspMaHu1yrJtu_pred.Rout &

nohup code/turnover_vs_temperature_GLMM_pred.R temp/modMainTdTSeMiNPNspMaHu1yrHorn.rds temp/newdata_main_only.csv temp/modMainTdTSeMiNPNspMaHu1yrHorn_preds.rds >logs/turnover_vs_temperature_GLMMmodMainTdTSeMiNPNspMaHu1yrHorn_pred.Rout &
```

#### Load and plot Jtu 1yr
```{r, fig.width = 7, fig.height=7, echo=FALSE}
newdat <- readRDS(here('temp', 'modMainTdTSeMiNPNspMaHu1yrJtu_preds.rds'))
newdat[, REALM := factor(REALM, levels = c('Freshwater', 'Terrestrial', 'Marine'))] # set factor order
newdat[, REALM2 := factor(REALM2, levels = c('TerrFresh', 'Marine'))] # set factor order

p1 <- ggplot(newdat[var == 'tempchange_abs',], aes(tempchange_abs, pred, group = REALM, color = REALM, fill=REALM)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))
p2 <- ggplot(newdat[var == 'tempave_metab',], aes(tempave_metab, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p3 <- ggplot(newdat[var == 'seas',], aes(seas, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p4 <- ggplot(newdat[var == 'microclim',], aes(microclim, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p5 <- ggplot(newdat[var == 'npp',], aes(npp, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p6 <- ggplot(newdat[var == 'nspp',], aes(nspp, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p7 <- ggplot(newdat[var == 'mass',], aes(mass, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  scale_x_log10() +
  labs(y = 'Jaccard turnover')
p8 <- ggplot(newdat[var == 'human_bowler',], aes(human_bowler, pred, group = REALM2, color = REALM2, fill = REALM2)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover', x = 'Human impact') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)

```

#### Load and plot Horn 1yr
```{r, fig.width = 7, fig.height=7, echo=FALSE}
newdat <- readRDS(here('temp', 'modMainTdTSeMiNPNspMaHu1yrHorn_preds.rds'))
newdat[, REALM := factor(REALM, levels = c('Freshwater', 'Terrestrial', 'Marine'))] # set factor order
newdat[, REALM2 := factor(REALM2, levels = c('TerrFresh', 'Marine'))] # set factor order

p1 <- ggplot(newdat[var == 'tempchange_abs',], aes(tempchange_abs, pred, group = REALM, color = REALM, fill=REALM)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))
p2 <- ggplot(newdat[var == 'tempave_metab',], aes(tempave_metab, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p3 <- ggplot(newdat[var == 'seas',], aes(seas, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p4 <- ggplot(newdat[var == 'microclim',], aes(microclim, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p5 <- ggplot(newdat[var == 'npp',], aes(npp, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p6 <- ggplot(newdat[var == 'nspp',], aes(nspp, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p7 <- ggplot(newdat[var == 'mass',], aes(mass, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  scale_x_log10() +
  labs(y = 'Jaccard turnover')
p8 <- ggplot(newdat[var == 'human_bowler',], aes(human_bowler, pred, group = REALM2, color = REALM2, fill = REALM2)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover', x = 'Human impact') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 2)

```

## Full models w/ only main effects, averaged by time-series (rarefyID)


### Load the models
```{r, echo=FALSE}
# all years
if(file.exists('temp/modMainAveMaEnMiNPHu1yrJtu.rds')) modMainAveMaEnMiNPHu1yrJtu <- readRDS('temp/modMainAveMaEnMiNPHu1yrJtu.rds')
if(file.exists('temp/modMainAveMaEnMiNPHu5yrJtu.rds')) modMainAveMaEnMiNPHu5yrJtu <- readRDS('temp/modMainAveMaEnMiNPHu5yrJtu.rds')
if(file.exists('temp/modMainAveMaEnMiNPHu10yrJtu.rds')) modMainAveMaEnMiNPHu10yrJtu <- readRDS('temp/modMainAveMaEnMiNPHu10yrJtu.rds')

if(file.exists('temp/modMainAveMaEnMiNPHu1yrJbeta.rds')) modMainAveMaEnMiNPHu1yrJbeta <- readRDS('temp/modMainAveMaEnMiNPHu1yrJbeta.rds')
if(file.exists('temp/modMainAveMaEnMiNPHu5yrJbeta.rds')) modMainAveMaEnMiNPHu5yrJbeta <- readRDS('temp/modMainAveMaEnMiNPHu5yrJbeta.rds')

if(file.exists('temp/modMainAveMaEnMiNPHu1yrHorn.rds')) modMainAveMaEnMiNPHu1yrHorn <- readRDS('temp/modMainAveMaEnMiNPHu1yrHorn.rds')
if(file.exists('temp/modMainAveMaEnMiNPHu10yrHorn.rds')) modMainAveMaEnMiNPHu10yrHorn <- readRDS('temp/modMainAveMaEnMiNPHu10yrHorn.rds')

```

### Summary
#### Jtu
##### 1 yr
```{r, echo=FALSE}
if(exists('modMainAveMaEnMiNPHu1yrJtu')) summary(modMainAveMaEnMiNPHu1yrJtu)
```

##### 10 yr
```{r, echo=FALSE}
if(exists('modMainAveMaEnMiNPHu10yrJtu'))  summary(modMainAveMaEnMiNPHu10yrJtu)
```

#### Horn
##### 1 yr
```{r, echo=FALSE}
if(exists('modMainAveMaEnMiNPHu1yrHorn')) summary(modMainAveMaEnMiNPHu1yrHorn)
```

##### 10 yr
```{r, echo=FALSE}
if(exists('modMainAveMaEnMiNPHu10yrHorn'))summary(modMainAveMaEnMiNPHu10yrHorn)
```


## Full models w/ temperature interactions
There are a number of different models here:

- Three response variables (Jaccard turnover, Jaccard total, and Morisita-Horn)
- Four subsets of the data: all, only 1-year time-steps, only 5-year time-steps, and only 10-year time-steps

Some take-aways for now:

- Larger temperature-change is not consistently associated with greater community composition change through time. The strongest positive associations between temperature change and community change are usually for freshwater ecosystems and those with few available microclimates and low human impact. For the all data subset, the temperature-change effect is stronger for longer durations.
- In contrast, greater community change through time is pretty consistently associated with 
  - higher metabolic temperatures (tempave_metab.sc)
  - higher NPP (npp.sc)
  - low human impacts on land/in freshwater (human_bowler.sc:REALM2TerrFresh)

### Load the models
```{r full, echo=FALSE}
# all years
if(file.exists('temp/modFullendo.rds')) modFullendo <- readRDS('temp/modFullendo.rds')
if(file.exists('temp/modFullmass.rds')) modFullmass <- readRDS('temp/modFullmass.rds')

if(file.exists('temp/modFullMaEnMiNPHuJtu.rds')) modFullMaEnMiNPHuJtu <- readRDS('temp/modFullMaEnMiNPHuJtu.rds') # mass, endoecto, microclimate, NPP, humans
if(file.exists('temp/modFullMaEnMiNPHu1yrJtu.rds')) modFullMaEnMiNPHu1yrJtu <- readRDS('temp/modFullMaEnMiNPHu1yrJtu.rds')
if(file.exists('temp/modFullMaEnMiNPHu5yrJtu.rds')) modFullMaEnMiNPHu5yrJtu <- readRDS('temp/modFullMaEnMiNPHu5yrJtu.rds')
if(file.exists('temp/modFullMaEnMiNPHu10yrJtu.rds')) modFullMaEnMiNPHu10yrJtu <- readRDS('temp/modFullMaEnMiNPHu10yrJtu.rds')

if(file.exists('temp/modFullMaEnMiNPHuJbeta.rds')) modFullMaEnMiNPHuJbeta <- readRDS('temp/modFullMaEnMiNPHuJbeta.rds')
if(file.exists('temp/modFullMaEnMiNPHu1yrJbeta.rds')) modFullMaEnMiNPHu1yrJbeta <- readRDS('temp/modFullMaEnMiNPHu1yrJbeta.rds')
if(file.exists('temp/modFullMaEnMiNPHu5yrJbeta.rds')) modFullMaEnMiNPHu5yrJbeta <- readRDS('temp/modFullMaEnMiNPHu5yrJbeta.rds')

if(file.exists('temp/modFullMaEnMiNPHuHorn.rds')) modFullMaEnMiNPHuHorn <- readRDS('temp/modFullMaEnMiNPHuHorn.rds')
if(file.exists('temp/modFullMaEnMiNPHu1yrHorn.rds')) modFullMaEnMiNPHu1yrHorn <- readRDS('temp/modFullMaEnMiNPHu1yrHorn.rds')
if(file.exists('temp/modFullMaEnMiNPHu5yrHorn.rds')) modFullMaEnMiNPHu5yrHorn <- readRDS('temp/modFullMaEnMiNPHu5yrHorn.rds')

```

### Summary
#### endo and mass models
```{r summary endo and mass, echo=FALSE}
if(exists('modFullendo')) summary(modFullendo)
if(exists('modFullmass')) summary(modFullmass)
```

#### Mass, endo, microclimate, NPP, humans
##### Jtu
###### All
```{r summary MaEnMiNPHu, echo=FALSE}
if(exists('modFullMaEnMiNPHuJtu')) summary(modFullMaEnMiNPHuJtu)
```

###### 1 yr
```{r summary MaEnMiNPHu1, echo=FALSE}
if(exists('modFullMaEnMiNPHu1yrJtu')) summary(modFullMaEnMiNPHu1yrJtu)
```

###### 5 yr
```{r summary MaEnMiNPHu5, echo=FALSE}
if(exists('modFullMaEnMiNPHu5yrJtu'))summary(modFullMaEnMiNPHu5yrJtu)
```

###### 10 yr
```{r summary MaEnMiNPHu10, echo=FALSE}
if(exists('modFullMaEnMiNPHu10yrJtu'))summary(modFullMaEnMiNPHu10yrJtu)
```

##### Jbeta
###### All
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHuJbeta'))summary(modFullMaEnMiNPHuJbeta)
```



###### 1 yr
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHu1yrJbeta')) summary(modFullMaEnMiNPHu1yrJbeta)
```

###### 5 yr
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHu5yrJbeta')) summary(modFullMaEnMiNPHu5yrJbeta)
```

##### Horn
###### All
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHuHorn')) summary(modFullMaEnMiNPHuHorn)
```



###### 1 yr
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHu1yrHorn')) summary(modFullMaEnMiNPHu1yrHorn)
```

###### 5 yr
```{r, echo=FALSE}
if(exists('modFullMaEnMiNPHu5yrHorn')) summary(modFullMaEnMiNPHu5yrHorn)
```


### DHARMa model evaluation
Need to re-run this with newly trimmed dataset (2021-04-05)

#### Mass, endo, microclimate, NPP, humans
##### Simulate residuals
```{r dharma maenminphu, echo=FALSE, eval=FALSE}
## run the top part (if(FALSE)) by hand if needed. takes ~3 hrs for all data
if(FALSE) {
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc)] # needed because model fit with trends[i,]

  res_MaEnMiNPHuJtu <- simulateResiduals(modFullMaEnMiNPHuJtu, n = 250)
  
  saveRDS(res_MaEnMiNPHuJtu, file = 'temp/res_MaEnMiNPHuJtu.rds')
} else {
  if(file.exists(here('temp', 'res_MaEnMiNPHuJtu.rds'))) res_MaEnMiNPHuJtu <- readRDS(here('temp', 'res_MaEnMiNPHuJtu.rds'))
}

if(FALSE) {
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 1] # needed because model fit with trends[i,]

  res_MaEnMiNPHu1yrJtu <- simulateResiduals(modFullMaEnMiNPHu1yrJtu, n = 250)
  
  saveRDS(res_MaEnMiNPHu1yrJtu, file = 'temp/res_MaEnMiNPHu1yrJtu.rds')
} else {
  if(file.exists(here('temp', 'res_MaEnMiNPHu1yrJtu.rds'))) res_MaEnMiNPHu1yrJtu <- readRDS(here('temp', 'res_MaEnMiNPHu1yrJtu.rds'))
}


if(FALSE) {
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 10] # needed because model fit with trends[i,]

  res_MaEnMiNPHu10yrJtu <- simulateResiduals(modFullMaEnMiNPHu10yrJtu, n = 250)
  
  saveRDS(res_MaEnMiNPHu10yrJtu, file = 'temp/res_MaEnMiNPHu10yrJtu.rds')
} else {
  if(file.exists(here('temp', 'res_MaEnMiNPHu10yrJtu.rds'))) res_MaEnMiNPHu10yrJtu <- readRDS(here('temp', 'res_MaEnMiNPHu10yrJtu.rds'))
}
```

##### Plot residuals
All data
```{r plot darma maenminphu, echo=FALSE, eval=FALSE}
if(exists('res_MaEnMiNPHuJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
  plot(res_MaEnMiNPHuJtu)
  plotResiduals(res_MaEnMiNPHuJtu, form=trends$tempchange_abs.sc[i], xlab = 'tempchange_abs.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, form=trends$REALM[i], xlab = 'REALM', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$tempave_metab.sc[i], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$duration.sc[i], xlab = 'duration.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$mass.sc[i], xlab = 'mass.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$endothermfrac.sc[i], xlab = 'endothermfrac.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$microclim.sc[i], xlab = 'microclim.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, trends$npp.sc[i], xlab = 'npp.sc', main = '')
  plotResiduals(res_MaEnMiNPHuJtu, form=trends$human_bowler.sc[i], xlab = 'human_bowler.sc', main = '')
}
```


1yr data
```{r plot darma maenminphu1yr, echo=FALSE, eval=FALSE}
if(exists('res_MaEnMiNPHu1yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 1]
  plot(res_MaEnMiNPHu1yrJtu)
  plotResiduals(res_MaEnMiNPHu1yrJtu, form=trends$tempchange_abs.sc[i], xlab = 'tempchange_abs.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, form=trends$REALM[i], xlab = 'REALM', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, trends$tempave_metab.sc[i], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, trends$mass.sc[i], xlab = 'mass.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, trends$endothermfrac.sc[i], xlab = 'endothermfrac.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, trends$microclim.sc[i], xlab = 'microclim.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, trends$npp.sc[i], xlab = 'npp.sc', main = '')
  plotResiduals(res_MaEnMiNPHu1yrJtu, form=trends$human_bowler.sc[i], xlab = 'human_bowler.sc', main = '')
}
```

10yr data
```{r plot darma maenminphu10yr, echo=FALSE, eval=FALSE}
if(exists('res_MaEnMiNPHu10yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 10]
  plot(res_MaEnMiNPHu10yrJtu)
  plotResiduals(res_MaEnMiNPHu10yrJtu, form=trends$tempchange_abs.sc[i], xlab = 'tempchange_abs.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, form=trends$REALM[i], xlab = 'REALM', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, trends$tempave_metab.sc[i], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, trends$mass.sc[i], xlab = 'mass.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, trends$endothermfrac.sc[i], xlab = 'endothermfrac.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, trends$microclim.sc[i], xlab = 'microclim.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, trends$npp.sc[i], xlab = 'npp.sc', main = '')
  plotResiduals(res_MaEnMiNPHu10yrJtu, form=trends$human_bowler.sc[i], xlab = 'human_bowler.sc', main = '')
}
```

##### Overdispersion
```{r, echo=FALSE, eval=FALSE}
if(exists('res_MaEnMiNPHuJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
  testDispersion(res_MaEnMiNPHuJtu)
}

if(exists('res_MaEnMiNPHu1yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 1]
  testDispersion(res_MaEnMiNPHu1yrJtu)
}

if(exists('res_MaEnMiNPHu10yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 10]
  testDispersion(res_MaEnMiNPHu10yrJtu)
}
```

##### Near-zero-inflation
Zero values have been transformed to slightly >0, so can't test zero-inflation directly
```{r, echo=FALSE, eval=FALSE}
countNearZero <- function(x) sum(x < 0.0001)

if(exists('res_MaEnMiNPHuJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
  testGeneric(res_MaEnMiNPHuJtu, summary = countNearZero, alternative = 'greater')
}

if(exists('res_MaEnMiNPHu1yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 1]
  testGeneric(res_MaEnMiNPHu1yrJtu, summary = countNearZero, alternative = 'greater')
}

if(exists('res_MaEnMiNPHu10yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 10]
  testGeneric(res_MaEnMiNPHu10yrJtu, summary = countNearZero, alternative = 'greater')
}
```

##### Near-one-inflation
```{r, echo=FALSE, eval=FALSE}
countNearOne <- function(x) sum(x > 0.9999)

if(exists('res_MaEnMiNPHuJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc)]
  testGeneric(res_MaEnMiNPHuJtu, summary = countNearOne, alternative = 'greater')
}

if(exists('res_MaEnMiNPHu1yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 1]
  testGeneric(res_MaEnMiNPHu1yrJtu, summary = countNearOne, alternative = 'greater')
}

if(exists('res_MaEnMiNPHu10yrJtu')){
  i <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc, mass.sc, endothermfrac.sc,
                               microclim.sc, npp.sc, human_bowler.sc, nspp.sc) &
                duration == 10]
  testGeneric(res_MaEnMiNPHu10yrJtu, summary = countNearOne, alternative = 'greater')
}
```


### Plot the response
#### Main effects
##### Set up and write to file
Only run this once
```{r, eval=FALSE, echo=FALSE}
scaling <- fread('output/turnover_w_covariates_scaling.csv') # the scalings

# set up the variables to plot
# if variable is logged before scaling (see 'center and scale' above), then need to mark it here and express the limits on a log10 scale (even though log transforming is log)
vars <- data.frame(vars = c('tempchange_abs', 'tempchange_abs', 'tempchange_abs', 'tempave_metab', 'microclim', 'mass', 'npp', 'duration', 'endothermfrac', 'human_bowler', 'human_bowler'),
                   min =      c(0,    0,   0,   0,   0,   -1,  1.9, 0.5, 0,   0,   0), 
                   max =      c(4,    4,   4,   30,  0.8, 4,   3.7, 2,   1,   1,   1),
                   log =      c(F,    F,   F,   F,   T,   T,   T,   T,   F,   T,   T),
                   len =      c(100,  100, 100, 100, 100, 100, 100, 100, 100, 100, 100),
                   discrete = c(F,    F,   F,   F,   F,   F,   F,   F,   F,   F,   F),
                   plus =     c(0,    0,   0,   0,   0,   0,   0,   0,   0,   1,   1), # what to add before log-scaling
                   REALM = c('Freshwater', 'Marine', 'Terrestrial', rep('Terrestrial', 6), 'Terrestrial', 'Marine'),
                   REALM2 = c('TerrFresh', 'Marine', rep('TerrFresh', 8), 'Marine'),
                   stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', 
                      tempchange = -0.0001,
                      tempchange_abs.sc = 0.0001,
                      tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE),
                      nspp.sc = 60, 
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     tempchange = -0.0001,
                     tempchange_abs.sc = 0.0001,
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                     microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE),
                     nspp.sc = 60, 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    tempchange = -0.0001,
                    tempchange_abs.sc = 0.0001,
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE),
                    nspp.sc = 60, 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)
basetab[, ':='(duration.sc = 0, nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_514668')]

# make the data frames for each interaction to plot                
for(j in 1:nrow(vars)){
    # set up the main effects
    if(vars$log[j]){
        thisdat <- data.frame(new = 10^seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    } 
    if(!vars$log[j]){
        thisdat <- data.frame(new = seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                              var = vars$vars[j], stringsAsFactors = FALSE)
    }
    names(thisdat) <- c(vars$vars[j], 'var')
    
    # scale the variable
    cent <- scaling$center[scaling$var == paste0(vars$vars[j], '.sc')]
    scl <- scaling$scale[scaling$var == paste0(vars$vars[j], '.sc')]
    if(is.null(cent)) cent <- 0
    if(!is.null(cent) & !is.null(scl)){
        if(vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (log(thisdat[[vars$vars[j]]] + vars$plus[j]) - cent)/scl
        if(!vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (thisdat[[vars$var[j]]] - cent)/scl
    }
    
    # merge with the rest of the columns
    # use realm-specific averages for human impacts
    colnamestouse <- setdiff(colnames(basetab), paste0(vars$vars[j], '.sc'))
    if(vars$vars[j] != 'human_bowler'){
        thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
    }
    if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Terrestrial'){
        thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
    }
    if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Marine'){
        thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
    }
    
    # add realm
    thisdat$REALM <- vars$REALM[j]
    thisdat$REALM2 <- vars$REALM2[j]
    
    # merge with the previous iterations
    if(j == 1) newdat <- thisdat
    if(j > 1){
        colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
        for(toadd in colstoadd){
            newdat[[toadd]] <- NA
        }
        
        colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
        for(toadd in colstoadd2){
            thisdat[[toadd]] <- NA
        }
        
        newdat <- rbind(newdat, thisdat)
    } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$tempchange_abs.sc[1:6] <- rep(0.0001, 6)
newdat$tempchange[1:6] <- rep(c(0.0001, -0.0001), 3)
newdat$var[1:6] <- 'test'

# write out
write.csv(newdat, 'temp/newdata_main.csv', row.names = FALSE)

```

Now go run on the command line
```
nohup code/turnover_vs_temperature_GLMM_pred.R temp/modFullMaEnMiNPHuJtu.rds temp/newdata_main.csv temp/modFullMaEnMiNPHuJtu_preds.rds >logs/turnover_vs_temperature_GLMMmodFullMaEnMiNPHuJtu_pred.Rout &
```

##### Load and plot
```{r, fig.width = 7, fig.height=7, echo=FALSE}
newdat <- readRDS(here('temp', 'modFullMaEnMiNPHuJtu_preds.rds'))
newdat[, REALM := factor(REALM, levels = c('Freshwater', 'Terrestrial', 'Marine'))] # set factor order
newdat[, REALM2 := factor(REALM2, levels = c('TerrFresh', 'Marine'))] # set factor order

p1 <- ggplot(newdat[var == 'tempchange_abs',], aes(tempchange_abs, pred, group = REALM, color = REALM, fill=REALM)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))
p2 <- ggplot(newdat[var == 'tempave_metab',], aes(tempave_metab, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p3 <- ggplot(newdat[var == 'microclim',], aes(microclim, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p4 <- ggplot(newdat[var == 'mass',], aes(mass, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  scale_x_log10() +
  labs(y = 'Jaccard turnover')
p5 <- ggplot(newdat[var == 'npp',], aes(npp, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p6 <- ggplot(newdat[var == 'endothermfrac',], aes(endothermfrac, pred)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1) +
  geom_line() +
  labs(y = 'Jaccard turnover')
p7 <- ggplot(newdat[var == 'human_bowler',], aes(human_bowler, pred, group = REALM2, color = REALM2, fill = REALM2)) +
  geom_ribbon(aes(ymin=pred-1.96*pred.se, ymax=pred+1.96*pred.se), alpha = 0.1, linetype = 'blank') +
  geom_line() +
  labs(y = 'Jaccard turnover', x = 'Human impact') +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))

grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2)

```




#### Interactions
##### Set up and write to file
Only run this once
```{r, eval = FALSE, echo=FALSE}
scaling <- fread(here('output', 'turnover_w_covariates_scaling.csv')) # the scalings

# set up the interactions to plot
# if variable is logged before scaling (see assmble_turnover_covariates.Rmd), then need to mark it here and express the limits on a log10 scale (even though log transforming is log)
ints <- data.frame(vars = c('tempave_metab', 'microclim', 'mass',
                            'npp', 'duration', 'endothermfrac',
                            'human_bowler', 'human_bowler'),
                   min =      c(0,   -2,  0,   1.9, 0.5, 0,   0,   0), 
                   max =      c(30,  0.8, 8,   3.7, 2,   1,   1,   1),
                   log =      c(F,   T,   T,   T,   T,   F,   T,   T),
                   len =      c(100, 100, 100, 100, 100, 100, 100, 100),
                   discrete = c(F,   F,   F,   F,   F,   F,   F,   F),
                   plus =     c(0,   0,   0,   0,   0,   0,   1,   1), # what to add before log-scaling
                   REALM = c(rep('Freshwater', 6), 'Terrestrial', 'Marine'),
                   REALM2 = c(rep('TerrFresh', 7), 'Marine'),
                   stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      nspp.sc = 60,
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE),
                     microclim.sc = mean(microclim.sc, na.rm=TRUE),
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     nspp.sc = 60, 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    nspp.sc = 60, 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    endothermfrac.sc = mean(endothermfrac.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)
basetab[, ':='(duration.sc = 0, nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_514668')]

# make the data frames for each interaction to plot                
for(j in 1:nrow(ints)){
  # set up a grid of temperature trends and the interacting variable
  if(ints$log[j]) intvars <- list(tempchange = seq(0, 4, length.out = 100), 
                                  new = 10^seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                  var = ints$vars[j])
  if(!ints$log[j]) intvars <- list(tempchange = seq(0, 4, length.out = 100), 
                                   new = seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                   var = ints$vars[j])
  names(intvars) <- c('tempchange', ints$vars[j], 'var')
  thisdat <- expand.grid(intvars)
  
  # scale the interacting variable
  cent <- scaling$center[scaling$var == paste0(ints$vars[j], '.sc')]
  scl <- scaling$scale[scaling$var == paste0(ints$vars[j], '.sc')]
  if(ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (log(thisdat[[ints$vars[j]]] + ints$plus[j]) - cent)/scl
  if(!ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (thisdat[[ints$var[j]]] - cent)/scl
  
  # merge with the rest of the columns
  # use realm-specific averages for human impacts
  colnamestouse <- setdiff(colnames(basetab), paste0(ints$var[j], '.sc'))
  if(ints$vars[j] != 'human_bowler'){
    thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Terrestrial'){
    thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Marine'){
    thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
  }
  
  # add realm
  thisdat$REALM <- ints$REALM[j]
  thisdat$REALM2 <- ints$REALM2[j]
  
  # add plotting information
  thisdat$log <- ints$log[j]
  thisdat$discrete <- ints$discrete[j]
  
  # merge with the previous iterations
  if(j == 1) newdat <- thisdat
  if(j > 1){
    colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
    for(toadd in colstoadd){
      newdat[[toadd]] <- NA
    }
    
    colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
    for(toadd in colstoadd2){
      thisdat[[toadd]] <- NA
    }
    
    newdat <- rbind(newdat, thisdat)
  } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work, not sure about predict.glmmTMB)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$tempchange[1:6] <- c(-0.2, 0.2, -0.2, 0.2, -0.2, 0.2)
newdat$var[1:6] <- 'test'

# trim to at least some temperature change (so that tsign is -1 or 1)
newdat <- newdat[newdat$tempchange != 0,]

# scale the temperature vars
cent <- scaling$center[scaling$var == 'tempchange.sc']
scl <- scaling$scale[scaling$var == 'tempchange.sc']
newdat$tempchange.sc <- (newdat$tempchange-cent)/scl
newdat$tempchange_abs <- abs(newdat$tempchange)

cent <- scaling$center[scaling$var == 'tempchange_abs.sc']
scl <- scaling$scale[scaling$var == 'tempchange_abs.sc']
newdat$tempchange_abs.sc <- (newdat$tempchange_abs-cent)/scl

# write out
write.csv(newdat, 'temp/newdata_interactions.csv', row.names = FALSE)

```

Now go run on the command line
```
nohup code/turnover_vs_temperature_GLMM_pred.R temp/modFullMaEnMiNPHuJtu.rds temp/newdata_interactions.csv temp/modFullMaEnMiNPHuJtu_preds_interactions.rds >logs/turnover_vs_temperature_GLMMmodFullMaEnMiNPHuJtu_pred_interactions.Rout &
```

##### Load and plot
```{r, fig.width = 9, fig.height = 8, echo=FALSE}
newdatint <- readRDS(here('temp', 'modFullMaEnMiNPHuJtu_preds_interactions.rds'))
newdatint[, REALM := factor(REALM, levels = c('Freshwater', 'Terrestrial', 'Marine'))] # set factor order
newdatint[, REALM2 := factor(REALM2, levels = c('TerrFresh', 'Marine'))] # set factor order

ylims = c(0.2, 0.65)

# prep the plots
ints <- unique(newdatint[ ,.(var, REALM2, log, discrete)])
ints <- ints[ints$var != 'test', ] # remove the extra rows
intplots <- vector('list', nrow(ints))
for(j in 1:length(intplots)){
  subs <- newdatint$var == ints$var[j]
  xvar <- 'tempchange_abs'
  title <- ints$var[j]
  if(ints$var[j] %in% c('human_bowler')){
    subs <- newdatint$var == ints$var[j] & newdatint$tempchange > 0 & newdatint$REALM2 == ints$REALM2[j]
    title <- paste0('human:', ints$REALM2[j])
  } 
  
  thisplot <- ggplot(newdatint[subs, ], 
                     aes_string(x = xvar, y = 'pred', 
                                group = ints$var[j], 
                                color = ints$var[j])) +
    geom_line() +
    #coord_cartesian(ylim = ylims) +
    theme(legend.title = element_text(size = 8), 
          legend.text = element_text(size = 6),
          legend.key.width = unit(0.1, 'inch')) +
    labs(title = title)
  if(ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'log')
  }
  if(!ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'identity')
  }
  if(ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_brewer(palette = "Dark2")
  }
}

grid.arrange(grobs = intplots, ncol = 3)

```


## Full models w/ duration interactions
### Load the models
```{r durint, echo=FALSE}
if(file.exists('temp/modDurIntTdTSeMiNPNspMaHu5yrJtu.rds')){
  modDurInt5yrJtu <- readRDS('temp/modDurIntTdTSeMiNPNspMaHu5yrJtu.rds')
  # needed because model fit with trends[i,]
  trends[, maxyr := max(c(year1, year2)), by = rarefyID]
  rowkeep <- trends[, complete.cases(Jtu.sc, tempchange_abs.sc, REALM, tempave_metab.sc, duration.sc,
                                     seas.sc, microclim.sc, npp.sc, mass.sc, human_bowler.sc, nspp.sc) &
                      year1 > maxyr - 5 &
                      year2 > maxyr - 5] # trim to complete cases in the last 5 years
  rkeep <- trends[rowkeep, .(nyrs = length(unique(c(year1, year2)))), by = rarefyID][nyrs == 5, rarefyID] # find timeseries with exactly 5 years
  i <- trends[, rowkeep & rarefyID %in% rkeep]
  
  trends[i, tempchangeTS_abs := abs(median(tempchange/duration)), by = rarefyID] # Thiel-Sen estimator of the abs(slope): median of all pairwise differences
}
```

### Summary
```{r summary durint, echo=FALSE}
if(exists('modDurInt5yrJut')) summary(modDurInt5yrJtu)
```

### DHARMa model evaluation
#### Simulate residuals
```{r, echo=FALSE}
## run the if(FALSE) part by hand if needed (e.g., the first time)
if(FALSE) {
  res_DurInt5yrJtu <- simulateResiduals(modDurInt5yrJtu, n = 250)
  
  saveRDS(res_DurInt5yrJtu, file = 'temp/res_DurInt5yrJtu.rds')
} else {
  if(file.exists(here('temp', 'res_DurInt5yrJtu.rds'))) res_DurInt5yrJtu <- readRDS(here('temp', 'res_DurInt5yrJtu.rds'))
}
```

#### Plot residuals
```{r, echo=FALSE}
if(exists('res_DurInt5yrJtu')){
  plot(res_DurInt5yrJtu)
  plotResiduals(res_DurInt5yrJtu, trends$duration.sc[i], xlab = 'duration.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, form=trends$REALM[i], xlab = 'REALM', main = '')
  plotResiduals(res_DurInt5yrJtu, form=trends$tempchangeTS_abs[i], xlab = 'tempchangeTS_abs', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$tempave_metab.sc[i], xlab = 'tempave_metab.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$seas.sc[i], xlab = 'seas.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$microclim.sc[i], xlab = 'microclim.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$npp.sc[i], xlab = 'npp.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$nspp.sc[i], xlab = 'nspp.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, trends$mass.sc[i], xlab = 'mass.sc', main = '')
  plotResiduals(res_DurInt5yrJtu, form=trends$human_bowler.sc[i], xlab = 'human_bowler.sc', main = '')
}
```


#### Overdispersion
```{r, echo=FALSE}
if(exists('res_DurInt5yrJtu')){
  testDispersion(res_DurInt5yrJtu)
}

```

#### Near-zero-inflation
Zero values have been transformed to slightly >0, so can't test zero-inflation directly
```{r, echo=FALSE,}
countNearZero <- function(x) sum(x < 0.0001)

if(exists('res_DurInt5yrJtu')){
  testGeneric(res_DurInt5yrJtu, summary = countNearZero, alternative = 'greater')
}

```

#### Near-one-inflation
```{r, echo=FALSE, eval=FALSE}
countNearOne <- function(x) sum(x > 0.9999)

if(exists('res_DurInt5yrJtu')){
  testGeneric(res_DurInt5yrJtu, summary = countNearOne, alternative = 'greater')
}
```

### Interactions
#### Set up and write to file
Only run this once
```{r, eval = FALSE, echo=FALSE}
scaling <- fread(here('output', 'turnover_w_covariates_scaling.csv')) # the scalings

# set up the interactions to plot. min and max on a raw scale.
ints <- data.frame(vars = c('tempchangeTS_abs', 'tempave_metab', 'seas', 'microclim', 
                            'npp', 'nspp', 'mass',
                            'human_bowler', 'human_bowler'),
                   min =      c(0,   -10, 0,   0.1,   0.5,  1,    1,   0,   0), 
                   max =      c(1,   40,  15,  6,   5000, 1000, 1e8, 10,  10),
                   len =      c(100, 100, 100, 100, 100,  100,  100, 100, 100),
                   discrete = c(F,   F,   F,   F,   F,    F,    F,   F,   F),
                   REALM = c(rep('Terrestrial', 8), 'Marine'),
                   REALM2 = c(rep('TerrFresh', 8), 'Marine'),
                   stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', 
                      tempchangeTS_abs = mean(tempchangeTS_abs, na.rm=TRUE), 
                      tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      seas.sc = mean(seas.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      nspp.sc = mean(nspp.sc, na.rm=TRUE),
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     tempchangeTS_abs = mean(tempchangeTS_abs, na.rm=TRUE), 
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                     seas.sc = mean(seas.sc, na.rm=TRUE), 
                     microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     nspp.sc = baseall$nspp.sc, # keep this the same across all
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    tempchangeTS_abs = mean(tempchangeTS_abs, na.rm=TRUE), 
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    seas.sc = mean(seas.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    nspp.sc = baseall$nspp.sc,
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)
basetab[, ':='(STUDY_ID = trends[i, STUDY_ID][1], rarefyID = trends[i, rarefyID][1])] # pick the first RE level in the dataframe

# make the data frames for each interaction to plot                
for(j in 1:nrow(ints)){
  # set up a grid of temperature trends and the interacting variable
  if(ints$vars[j] == 'tempchangeTS_abs'){ # this one isn't in scaling, since made anew for each model based on duration
    intvars <- list(duration = seq(1, 5, length.out = 100), 
                    new = seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                    var = ints$vars[j])
  } else {
    log <- scaling$log[scaling$var == paste0(ints$vars[j], '.sc')]
    plus <- scaling$plus[scaling$var == paste0(ints$vars[j], '.sc')]
    if(log) intvars <- list(duration = seq(1, 5, length.out = 100), 
                                    new = exp(seq(log(ints$min[j]+plus), log(ints$max[j]+plus), length.out = ints$len[j])),
                                    var = ints$vars[j])
    if(!log) intvars <- list(duration = seq(1, 5, length.out = 100), 
                                     new = seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                     var = ints$vars[j])
  }
  names(intvars) <- c('duration', ints$vars[j], 'var')
  thisdat <- expand.grid(intvars)
  
  # scale the interacting variable
  if(ints$vars[j] != 'tempchangeTS_abs'){ # this one isn't in scaling, since made anew for each model based on duration0
    cent <- scaling$center[scaling$var == paste0(ints$vars[j], '.sc')]
    scl <- scaling$scale[scaling$var == paste0(ints$vars[j], '.sc')]
    log <- scaling$log[scaling$var == paste0(ints$vars[j], '.sc')]
    plus <- scaling$plus[scaling$var == paste0(ints$vars[j], '.sc')]
    if(log) thisdat[[paste0(ints$var[j], '.sc')]] <- (log(thisdat[[ints$vars[j]]] + plus) - cent)/scl
    if(!log) thisdat[[paste0(ints$var[j], '.sc')]] <- (thisdat[[ints$var[j]]] - cent)/scl
  }
  
  # merge with the rest of the columns
  # use realm-specific averages for human impacts
  colnamestouse <- setdiff(colnames(basetab), c(ints$var[j], paste0(ints$var[j], '.sc')))
  if(ints$vars[j] != 'human_bowler'){
    thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Terrestrial'){
    thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Marine'){
    thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
  }
  
  # add realm
  thisdat$REALM <- ints$REALM[j]
  thisdat$REALM2 <- ints$REALM2[j]
  
  # add plotting information
  if(ints$vars[j] != 'tempchangeTS_abs'){
    thisdat$log <- scaling$log[scaling$var == paste0(ints$vars[j], '.sc')]
  } else {
    thisdat$log <- FALSE
  }
  thisdat$discrete <- ints$discrete[j]
  
  # merge with the previous iterations
  if(j == 1) newdat <- thisdat
  if(j > 1){
    colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
    for(toadd in colstoadd){
      newdat[[toadd]] <- NA
    }
    
    colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
    for(toadd in colstoadd2){
      thisdat[[toadd]] <- NA
    }
    
    newdat <- rbind(newdat, thisdat)
  } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)
newdat$var <- as.character(newdat$var)

# add extra rows so that all factor levels are represented (for predict.lme to work, not sure about predict.glmmTMB)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$tempchange[1:6] <- c(-0.2, 0.2, -0.2, 0.2, -0.2, 0.2)
newdat$var[1:6] <- 'test'

# scale the duration
cent <- scaling$center[scaling$var == 'duration.sc']
scl <- scaling$scale[scaling$var == 'duration.sc']
newdat$duration.sc <- (log(newdat$duration)-cent)/scl

# write out
write.csv(newdat, 'temp/newdata_interactions_DurInt.csv', row.names = FALSE)

```

Now go run on the command line
```
nohup code/turnover_vs_temperature_GLMM_pred.R temp/modDurIntTdTSeMiNPNspMaHu5yrJtu.rds temp/newdata_interactions_DurInt.csv temp/modDurIntTdTSeMiNPNspMaHu5yrJtu_preds_interactions.rds >logs/turnover_vs_temperature_GLMMmodDurIntTdTSeMiNPNspMaHu5yrJtu_pred_interactions.Rout &
```

#### Load and plot
```{r, fig.width = 9, fig.height = 8, echo=FALSE}
newdatint <- readRDS(here('temp', 'modDurIntTdTSeMiNPNspMaHu5yrJtu_preds_interactions.rds'))
newdatint[, REALM := factor(REALM, levels = c('Freshwater', 'Terrestrial', 'Marine'))] # set factor order
newdatint[, REALM2 := factor(REALM2, levels = c('TerrFresh', 'Marine'))] # set factor order

ylims = c(0.2, 0.65)

# prep the plots
ints <- unique(newdatint[ ,.(var, REALM2, log, discrete)])
ints <- ints[ints$var != 'test', ] # remove the extra rows
intplots <- vector('list', nrow(ints))
for(j in 1:length(intplots)){
  subs <- newdatint$var == ints$var[j]
  xvar <- 'duration'
  title <- ints$var[j]
  if(ints$var[j] %in% c('human_bowler')){
    subs <- newdatint$var == ints$var[j] & newdatint$tempchange > 0 & newdatint$REALM2 == ints$REALM2[j]
    title <- paste0('human:', ints$REALM2[j])
  } 
  
  thisplot <- ggplot(newdatint[subs, ], 
                     aes_string(x = xvar, y = 'pred', 
                                group = ints$var[j], 
                                color = ints$var[j])) +
    geom_line() +
    #coord_cartesian(ylim = ylims) +
    theme(legend.title = element_text(size = 8), 
          legend.text = element_text(size = 6),
          legend.key.width = unit(0.1, 'inch')) +
    labs(title = title)
  if(ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'log')
  }
  if(!ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'identity')
  }
  if(ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_brewer(palette = "Dark2")
  }
}

grid.arrange(grobs = intplots, ncol = 3)

```
