---
title: 'Drivers of variation in the community response to temperature change across realms'
subtitle: '(using mixed effects models)'
output: 
    github_document: default
    #html_document: default
    html_notebook: default
---

Collaborators: Shane Blowes, Jon Chase, Helmut Hillebrand, Michael Burrows, Amanda Bates, Uli Brose, Benoit Gauzens, Laura Antao
Assistance: Katherine Lew, Josef Hauser

# Introduction
- Climate change is driving a widespread reorganization of ecological communities around the world (Parmsesan & Yohe 2003, Poloczanska et al. 2013),
- but the impacts of climate change vary substantially from one location to another and among taxa (Molinos et al. 2016 NCC, Antao et al. 2020 NEE).
- Community reorganization is substantially more common than an aggregate loss or gain of species (Dornelas et al. 2014 Science, Blowes et al. 2019 Science, Hillebrand et al. 2017 J Appl Ecol)
- There are many hypotheses for why some communities are more sensitive to warming than others, including differences in
  - metabolic rates (Dillon et al. 2010 Nature), 
  - thermal physiology (Deutsch et al. 2008 PNAS, Pinsky et al. 2019 Nature), 
  - microclimate availability (Burrows et al. 2019 NCC, Suggitt et al. 2018 NCC),
  - species mobility (Poloczanska et al. 2013 NCC, Burrows et al. 2011 Science, Sunday et al. 2012 NCC)
  - or generation time (Beaugrand et al. 2009 DSR II, Poloczanska et al. 2013 NCC),
  - consumers vs. producers (Petchey et al. 1999 Nature)
  - community composition (Stuart-Smith et al. 2015 Nature, Beaugrand et a. 2015 NCC, Trisos et al. 2020 Nature), 
  - ecosystem productivity (Thomas et al. 2017 GCB, Brett 1971 Am Zoo),
  - exposure to human impacts (White & Kerr 2006 Ecography)
  - and among realms (Antao et al. 2020 NEE).
- Scaling up from organismal effects to whole ecological communities is complex, and yet these scales are critical for ecosystem functioning and human well-being. 
- There is a need for a comprehensive test to understand where warming is driving and is likely to drive the most dramatic community turnover

# Methods
- BioTime dataset, gridded to 96 km2 hexagons, summarized as temporal turnover (Blowes)
  - Temporal slope of Jaccard turnover compared to the first year
  - Same for Jaccard total
  - and Morisita-Horn turnover
- Tested explanatory variables for differences in rate of turnover:
  - Temperature trend over the time-frame of each time-series (CRU TS 4.03 on land and in freshwater, ERSST v5 in the ocean)
  - Seasonality as a metric of thermal sensitivity (Deutsch et al. 2008 PNAS). Standard deviation of monthly temperatures.
  - Average temperature as a metric of metabolic rates (Dillon et al. 2010 Nature, Antao et al. 2020 Nat E&E)
  - Mobility calculated from body mass and taxonomic group classifications of mobility mode (fly, run, swim, crawl, sessile). Fly/run/swim followed the allometric relationship in Hirt et al. 2017 Nat E&E. Crawl set at 0.1 km/hr, sessile set to 0 km/hr. Then calculated averaged within each assemblage.
  - Net primary productivity (NPP) from the merged land/ocean product produced by the [Ocean Productivity](http://www.science.oregonstate.edu/ocean.productivity/) group at Oregon State using methods from Zhao et al. 2005 and Behrenfeld & Falkowski 1997. 
 - TO DO:
   - Generation time calculated from body mass and endotherm vs. ectotherm classifications, following McCoy & Gillooly 2008 ELE. Averaged across species within each assemblage. 
   - Human impact calculated from Carsten Meyer's ecosystem cube data
   - Thermal bias calculated from Species Temperature Indices (Mike Burrows)
   - Microclimates calculated from WorldClim and BioOracle (Laura Antao)
   - Consumer vs. producer classification
 - Maybe to do:
   - Species pool richness
- Differences in temporal turnover (response variable) modeled with a linear mixed effects model (nlme package, lme() function). See below for details.

```{r setup}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
#library(lme4)
library(nlme) # for ME models
library(beanplot) # for beanplots
library(maps) # for map
library(ggeffects) # marginal effect plots
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
library(gridExtra)

options(width=500) # turn off most text wrapping

# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
```

```{r load data}
# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends <- fread('output/turnover_w_covariates.csv.gz')

# set realm order
trends[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# group Marine invertebrates/plants in with All
trends[, taxa_mod2 := taxa_mod]
trends[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']
```

Log-transform some variables, then center and scale. 
``` {r center and scale}
trends[, tempave.sc := scale(tempave)]
trends[, tempave_metab.sc := scale(tempave_metab)]
trends[, seas.sc := scale(seas)]
trends[, microclim.sc := scale(log(microclim))]
trends[, temptrend.sc := scale(temptrend)]
trends[, temptrend_abs.sc := scale(log(abs(temptrend)))]
trends[, npp.sc := scale(log(npp))]
trends[, mass.sc := scale(log(mass_mean_weight))]
trends[, speed.sc := scale(log(speed_mean_weight+1))]
trends[, lifespan.sc := scale(log(lifespan_mean_weight))]
trends[, thermal_bias.sc := scale(thermal_bias)]
trends[, consumerfrac.sc := scale(consfrac)]
trends[, endothermfrac.sc := scale(endofrac)]
trends[, nspp.sc := scale(log(Nspp))]
trends[, human.sc := scale(human)]
```

### Do the variables look ok?
```{r histograms}
# histograms to examine
cexmain = 0.6
par(mfrow = c(3,5))
invisible(trends[, hist(tempave.sc, main = 'Environmental temperature (°C)', cex.main = cexmain)])
invisible(trends[, hist(tempave_metab.sc, main = 'Metabolic temperature (°C)', cex.main = cexmain)])
invisible(trends[, hist(seas.sc, main = 'Seasonality (°C)', cex.main = cexmain)])
invisible(trends[, hist(microclim.sc, main = 'log Microclimates (°C)', cex.main = cexmain)])
invisible(trends[, hist(temptrend.sc, main = 'Temperature trend (°C/yr)', cex.main = cexmain)])
invisible(trends[, hist(temptrend_abs.sc, main = 'log abs(Temperature trend) (°C/yr)', cex.main = cexmain)])
invisible(trends[, hist(mass.sc, main = 'log Mass (g)', cex.main = cexmain)])
invisible(trends[, hist(speed.sc, main = 'log Speed (km/hr)', cex.main = cexmain)])
invisible(trends[, hist(lifespan.sc, main = 'log Lifespan (yr)', cex.main = cexmain)])
invisible(trends[, hist(consumerfrac.sc, main = 'Consumers (fraction)', cex.main = cexmain)])
invisible(trends[, hist(endothermfrac.sc, main = 'Endotherms (fraction)', cex.main = cexmain)])
invisible(trends[, hist(nspp.sc, main = 'log Species richness', cex.main = cexmain)])
invisible(trends[, hist(thermal_bias.sc, main = 'Thermal bias (°C)', cex.main = cexmain)])
invisible(trends[, hist(npp.sc, main = 'log Net primary productivity', cex.main = cexmain)])
invisible(trends[, hist(human.sc, main = 'Human impact score', cex.main = cexmain)])

```

### Check correlations among variables. Pearson's r is on the lower diagonal.
```{r pairs, fig.height=10, fig.width=10}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y, use = 'pairwise.complete.obs')
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt) #, cex = cex.cor * r)
}
pairs(formula = ~ REALM + tempave.sc + tempave_metab.sc + seas.sc + microclim.sc + temptrend.sc + temptrend_abs.sc +  mass.sc + speed.sc + lifespan.sc + consumerfrac.sc + endothermfrac.sc + nspp.sc + thermal_bias.sc + npp.sc + human.sc, data = trends, gap = 1/10, cex = 0.2, col = '#00000022', lower.panel = panel.cor)

```

Mass and lifespan look tightly correlated, but r only 0.56...?
Tempave_metab and lifespan don't look tightly correlated, but r= -0.81 
Tempave_metab and speed don't look tightly correlated, but r= -0.83 
Lifespan and speed don't look tightly correlated, but r = 0.73


### Examine how many data points are available
Just turnover
```{r sample size all}
cat('Overall # time-series: ', nrow(trends), '\n')
cat('# studies: ', trends[, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[, sum(nyrBT)], '\n')
trends[, table(REALM)]
trends[, table(taxa_mod)]
trends[, table(taxa_mod, REALM)]
```

With all covariates
```{r sample size for Jaccard turnover}
# the cases we can compare
apply(trends[, .(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, temptrend.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, human.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends[, complete.cases(Jtutrend, temptrend.sc, tempave_metab.sc, REALM, seas.sc, microclim.sc, npp.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, thermal_bias.sc)]
cat('Overall # time-series: ', sum(i), '\n')
cat('# studies: ', trends[i, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[i, sum(nyrBT)], '\n')
trends[i, table(REALM)]
trends[i, table(taxa_mod)]
trends[i, table(taxa_mod, REALM)]
```

### Choose the variance structure for mixed effects modles
Try combinations of

- variance scaled to a power of the number of years in the community time-series
- variance scaled to a power of the abs temperature trend
- random intercept for taxa_mod
- random intercept for STUDY_ID
- random slope (abs temperature trend) for taxa_mod
- random slope (abs temperature trend) for STUDY_ID
- random intercept for rarefyID (for overdispersion)

And choose the one with lowest AIC (not run: takes a long time)
```{r choose variance structure for Jacard turnover, eval = FALSE}
# fit models for variance structure
fixed <- formula(Jtutrend ~ REALM + tempave_metab.sc + seas.sc + microclim.sc + npp.sc + temptrend_abs.sc +
                     mass.sc + speed.sc + lifespan.sc + consumerfrac.sc + thermal_bias.sc)
i <- trends[, complete.cases(Jtutrend, REALM, tempave_metab.sc, seas.sc, microclim.sc, npp.sc, temptrend_abs.sc,
                             mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, thermal_bias.sc)]
mods <- vector('list', 0)
mods[[1]] <- gls(fixed, data = trends[i,])
mods[[2]] <- gls(fixed, data = trends[i,], weights = varPower(-0.5, ~nyrBT))
mods[[3]] <- gls(fixed, data = trends[i,], weights = varPower(0.5, ~ abs(temptrend)))

mods[[4]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, control = lmeControl(opt = "optim"))
mods[[5]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, control = lmeControl(opt = "optim"))
mods[[6]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID, control = lmeControl(opt = "optim"))
mods[[7]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))
mods[[8]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))

mods[[9]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | taxa_mod)
mods[[10]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | STUDY_ID)
mods[[11]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | taxa_mod2/STUDY_ID, control = lmeControl(opt = "optim"))
mods[[12]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)) # includes overdispersion. new formula so that random slope is only for study level (not enough data to extend to rarefyID).
mods[[13]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)) # 30+ min to fit

mods[[14]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[15]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, weights = varPower(-0.5, ~nyrBT))
mods[[16]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[17]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~nyrBT))
mods[[18]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, weights = varPower(-0.5, ~nyrBT))
mods[[19]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[20]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))
mods[[21]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ 1), weights = varPower(-0.5, ~nyrBT))
mods[[22]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ 1, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))
mods[[23]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc), weights = varPower(-0.5, ~nyrBT)) # singular precision warning with lmeControl(opt = 'optim') and convergence error without
mods[[24]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT)) # singular precision warning with lmeControl(opt = 'optim') and convergence error without

mods[[25]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, weights = varPower(-0.5, ~abs(temptrend)))
mods[[26]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[27]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[28]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[29]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[30]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|taxa_mod2/STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)), control = lmeControl(opt = "optim"))
mods[[31]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~abs(temptrend)))
mods[[32]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~abs(temptrend)), control = lmeControl(opt = "optim")) # singular precision warning

aics <- sapply(mods, AIC)
minaics <- aics - min(aics)
minaics
which.min(aics)
```
Chooses the random slopes (temptrend_abs) & intercepts for STUDY_ID, overdispersion, and variance scaled to number of years.
We haven't dealt with potential testing on the boundary issues here yet.

# Results
## Where do we have data?
```{r map}
world <- map_data('world')
ggplot(world, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = 'lightgray', color = 'white') +
    geom_point(data = trends, aes(rarefyID_x, rarefyID_y, group = REALM, color = REALM), size = 0.5, alpha = 0.4)  +
    scale_color_brewer(palette="Set1", name = 'Realm') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=16),
        axis.title=element_text(size=20)) +
  labs(x = 'Longitude (°)', y = 'Latitude (°)')
```

Mostly northern hemisphere, but spread all over. No so much in Africa or much of Asia.


## Plot turnover vs. explanatory variables
Lines are ggplot smoother fits by realm.
```{r plot turnover v temp trend, echo=FALSE, fig.height = 16, fig.width = 9,}

p1 <- ggplot(trends, aes(REALM, Jtutrend)) +
  geom_boxplot(na.rm = TRUE) + 
  labs(x = 'Realm', y = 'Jaccard turnover temporal trend')

p2 <- ggplot(trends, aes(tempave, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature (°C)', y = 'Jaccard turnover temporal trend')

p3 <- ggplot(trends, aes(tempave_metab, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Metabolic temperature (°C)', y = 'Jaccard turnover temporal trend')

p4 <- ggplot(trends, aes(seas, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Seasonality (°C)', y = 'Jaccard turnover temporal trend')

p5 <- ggplot(trends, aes(microclim, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Microclimate availability (°C)', y = 'Jaccard turnover temporal trend')

p6 <- ggplot(trends, aes(mass_mean_weight, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Mass (g))', y = 'Jaccard turnover temporal trend')

p7 <- ggplot(trends, aes(speed_mean_weight+1, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Speed (km / hr))', y = 'Jaccard turnover temporal trend')

p8 <- ggplot(trends, aes(lifespan_mean_weight, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Lifespan (yr))', y = 'Jaccard turnover temporal trend')

p9 <- ggplot(trends, aes(consfrac, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction consumers', y = 'Jaccard turnover temporal trend')

p10 <- ggplot(trends, aes(endofrac, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 3), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Fraction endotherms', y = 'Jaccard turnover temporal trend')

p11 <- ggplot(trends, aes(Nspp, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Number of species', y = 'Jaccard turnover temporal trend')

p12 <- ggplot(trends, aes(thermal_bias, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Thermal bias (°C)', y = 'Jaccard turnover temporal trend')

p13 <- ggplot(trends, aes(npp, Jtutrend, size = nyrBT)) +
  geom_point(aes(color = REALM), size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE, color = 'black') +
  scale_color_brewer(palette="Set1") + 
  scale_x_log10() +
  labs(x = 'Net primary productivity (mg C / m2 / day)', y = 'Jaccard turnover temporal trend')

p14 <- ggplot(trends, aes(human, Jtutrend, color = REALM, size = nyrBT)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", k = 5), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Bowler human impact score', y = 'Jaccard turnover temporal trend')

p15 <- ggplot(trends, aes(temptrend, Jtutrend, color = REALM, size = nyrBT)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature trend (°C/yr)', y = 'Jaccard turnover temporal trend')

# Jaccard total trend vs. temperature trend (across all years)
p16 <- ggplot(trends, aes(temptrend, Jbetatrend, color = REALM, size = nyrBT)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature trend (°C/year)', y = 'Jaccard total temporal trend')


# Horn-Morisita turnover trend vs. temperature trend (across all years)
p17 <- ggplot(trends, aes(temptrend, Horntrend, color = REALM, size = nyrBT)) +
  geom_point(size = 0.2, alpha = 0.5, na.rm = TRUE) + 
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"), na.rm = TRUE) +
  scale_color_brewer(palette="Set1") + 
  labs(x = 'Temperature trend (°C/year)', y = 'Morisita-Horn temporal turnover')


grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, p16, p17, ncol = 2)
```

Strong trends with temperature change, but trends are pretty symmetric around no trend in temperature, which implies warming or cooling drives similar degree of community turnover.
Some indication of less turnover for larger organisms (mass)
Higher turnover on land with higher seasonality?
More turnover for shorter-lived organisms?
No really clear differences among realms.

## Nicer plots of turnover vs. temperature?
Violin plots
```{r turnover vs. temperature vioplot, fig.width=2, fig.height=1.75}
trends[temptrend <= -0.5, temptrendtext := 'Cooling']
trends[abs(temptrend) <= 0.1, temptrendtext := 'Stable']
trends[temptrend >= 0.5, temptrendtext := 'Warming']

trends[abs(rarefyID_y) < 35, latzone := 'Subtropics']
trends[abs(rarefyID_y) >= 35 & abs(rarefyID_x) < 66.56339, latzone := 'Temperate'] 
trends[abs(rarefyID_y) >= 66.56339, latzone := 'Polar']
trends[, latzone := factor(latzone, levels = c('Subtropics', 'Temperate', 'Polar'))]

p1 <- ggplot(trends, aes(temptrend, Jtutrend, color = REALM, fill = REALM, size = nyrBT)) +
  geom_point(na.rm = TRUE, shape = 16, alpha = 0.1) + 
  geom_smooth(data=subset(trends, abs(temptrend) < 0.75), method = 'gam', formula = y ~ s(x, bs = "cs"), 
              na.rm = TRUE) +
  scale_color_brewer(palette="Set1", name = 'Realm') + 
  scale_fill_brewer(palette="Set1", name = 'Realm') + 
  labs(x = 'Temperature trend (°C/yr)', y = 'Jaccard turnover') +
  scale_size_continuous(range = c(1, 8), breaks = c(2, 5, 20)) +
  guides(size = guide_legend(title = 'Years',
                             override.aes = list(linetype=0, fill = NA, alpha = 1))) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        legend.key.size = unit(0.5,"line"), 
        axis.text=element_text(size=8),
        axis.title=element_text(size=10))

p2 <- ggplot(trends[!is.na(temptrendtext), ], aes(temptrendtext, Jtutrend)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') +
  labs(x = '', y = 'Jaccard turnover') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10))

p3 <- ggplot(trends[abs(temptrend) >= 0.5 & !is.na(latzone), ], aes(latzone, Jtutrend)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') + 
  labs(x = '', y = '') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=7),
        axis.title=element_text(size=10))

grid.arrange(p1, p2, p3, ncol = 2, layout_matrix = rbind(c(1,1), c(2,3)),
             heights=c(unit(0.66, "npc"), unit(0.34, "npc")))
```

Average rates of turnover
```{r rates of turnover}
trends[abs(temptrend) >= 0.5, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # turnover per year for locations changing temperature
trends[abs(temptrend) < 0.1, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # not changing temperature
trends[temptrend >= 0.5, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # warming
trends[temptrend <= -0.5, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # cooling

trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) < 35, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # tropics and sub-tropics
trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) >= 35 & abs(rarefyID_y) < 66.56339, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # temperate
trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) >= 66.56339, .(mean(Jtutrend), sd(Jtutrend)/sqrt(.N))] # arctic
```

Hexagonal bins
```{r turnover vs. temperature hexbin, fig.width=2, fig.height=1.5}
require(ggplot2)
ggplot(trends[REALM == 'Terrestrial',], aes(temptrend, Jtutrend)) +
  geom_hex() + 
  scale_fill_gradient(trans = "log")
ggplot(trends[REALM == 'Marine',], aes(temptrend, Jtutrend)) +
  geom_hex() + 
  scale_fill_gradient(trans = "log")
ggplot(trends[REALM == 'Freshwater',], aes(temptrend, Jtutrend)) +
  geom_hex() + 
  scale_fill_gradient(trans = "log")
```


## Compare covariates across realms
```{r compare across realms, fig.height=12, fig.width=9}
i <- trends[, !duplicated(rarefyID)]; sum(i)
par(mfrow=c(5,3))
beanplot(rarefyID_y ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Latitude (degN)', ll = 0.05)
beanplot(tempave ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature (degC)', ll = 0.05)
beanplot(tempave_metab ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Metabolic Temperature (degC)', ll = 0.05, bw = 'nrd0') # nrd0 bandwidth to calculation gap
beanplot(seas ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Seasonality (degC)', ll = 0.05)
beanplot(microclim ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Microclimates (degC)', ll = 0.05)
beanplot(temptrend ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Temperature trend (degC/yr)', ll = 0.05)
beanplot(mass_mean_weight ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Mass (g)', ll = 0.05, log = 'y')
beanplot(speed_mean_weight +1 ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Speed (km/hr)', ll = 0.05, log = 'y')
beanplot(lifespan_mean_weight ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Lifespan (yr)', ll = 0.05, log = 'y')
#beanplot(consfrac ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Consumers (fraction)', ll = 0.05, log = '') # too sparse
#beanplot(endofrac ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Endotherms (fraction)', ll = 0.05, log = '') # too sparse
beanplot(Nspp ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Number of species', ll = 0.05, log = 'y')
beanplot(thermal_bias ~ REALM, data = trends[i & !is.na(thermal_bias),], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Thermal bias (degC)', ll = 0.05)
beanplot(npp ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'NPP', ll = 0.05)
beanplot(human ~ REALM, data = trends[i,], what = c(1,1,1,1), col = c("#CAB2D6", "#33A02C", "#B2DF8A"), border = "#CAB2D6", ylab = 'Human impact score', ll = 0.05)

```

Marine are in generally warmer locations (seawater doesn't freeze)
Marine have much lower seasonality.
Marine and freshwater have some very small masses (plankton), but much of dataset is similar to terrestrial.
Marine has a lot of slow, crawling organisms, but land has plants. Land also has birds (fast).




## Jaccard turnover temporal trend
Try static covariates plus interactions of abs temperature trend with each covariate:

- realm
- environmental temperature
- average metabolic temperature
- seasonality
- microclimates
- NPP
- speed
- mass
- lifespan
- consumer vs. producer
- thermal bias

Except for thermal bias: interact with temperature trend (not abs)

### Temperature-only model (Jtutrend, Jbetatrend, Horntrend)
```{r LME Jacard turnover temperature only}
i <- trends[, complete.cases(Jtutrend, REALM, temptrend)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modonlyTtrend.rds')){
  modonlyTtrend <- readRDS('temp/modonlyTtrend.rds')
} else {
  modonlyTtrend <- lme(Jtutrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrend, file = 'temp/modonlyTtrend.rds')
}

i2 <- trends[, complete.cases(Jbetatrend, REALM, temptrend)]
if(file.exists('temp/modonlyTtrendJbeta.rds')){
  modonlyTtrendJbeta <- readRDS('temp/modonlyTtrendJbeta.rds')
} else {
  modonlyTtrendJbeta <- lme(Jbetatrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i2,], method = 'REML')
  saveRDS(modonlyTtrendJbeta, file = 'temp/modonlyTtrendJbeta.rds')
}

i3 <- trends[, complete.cases(Horntrend, REALM, temptrend)]
if(file.exists('temp/modonlyTtrendHorn.rds')){
  modonlyTtrendHorn <- readRDS('temp/modonlyTtrendHorn.rds')
} else {
  modonlyTtrendHorn <- lme(Horntrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i3,], method = 'REML')
  saveRDS(modonlyTtrendHorn, file = 'temp/modonlyTtrendHorn.rds')
}

summary(modonlyTtrend)
summary(modonlyTtrendJbeta)
summary(modonlyTtrendHorn)


```

#### Try simplifying the temp-only models
```{r simplify modonlyTtrend}
if(file.exists('temp/modonlyTtrendsimpreml.rds')){
  modonlyTtrendsimpreml <- readRDS('temp/modonlyTtrendsimpreml.rds')
} else {
  require(MASS) # for stepAIC
  modonlyTtrendml <- update(modonlyTtrend, method = 'ML')
  modonlyTtrendsimp <- stepAIC(modonlyTtrendml, direction = 'backward')
  modonlyTtrendsimpreml <- update(modonlyTtrendsimp, method = 'REML')
  saveRDS(modonlyTtrendsimpreml, file = 'temp/modonlyTtrendsimpreml.rds')
}


if(file.exists('temp/modonlyTtrendJbetasimpreml.rds')){
  modonlyTtrendJbetasimpreml <- readRDS('temp/modonlyTtrendJbetasimpreml.rds')
} else {
  require(MASS) # for stepAIC
  modonlyTtrendJbetaml <- update(modonlyTtrendJbeta, method = 'ML')
  modonlyTtrendJbetasimp <- stepAIC(modonlyTtrendJbetaml, direction = 'backward')
  modonlyTtrendJbetasimpreml <- update(modonlyTtrendJbetasimp, method = 'REML')
  saveRDS(modonlyTtrendJbetasimpreml, file = 'temp/modonlyTtrendJbetasimpreml.rds')
}

if(file.exists('temp/modonlyTtrendHornsimpreml.rds')){
  modonlyTtrendHornsimpreml <- readRDS('temp/modonlyTtrendHornsimpreml.rds')
} else {
  require(MASS) # for stepAIC
  modonlyTtrendHornml <- update(modonlyTtrendHorn, method = 'ML')
  modonlyTtrendHornsimp <- stepAIC(modonlyTtrendHornml, direction = 'backward')
  modonlyTtrendHornsimpreml <- update(modonlyTtrendHornsimp, method = 'REML')
  saveRDS(modonlyTtrendHornsimpreml, file = 'temp/modonlyTtrendHornsimpreml.rds')
}

summary(modonlyTtrendsimpreml)
summary(modonlyTtrendJbetasimpreml)
summary(modonlyTtrendHornsimpreml)

```

#### Plot the temp-only coefficients
```{r plot modonlyTtrendsimp}
coefs <- summary(modonlyTtrend)$tTable
coefs2 <- summary(modonlyTtrendJbeta)$tTable
coefs3 <- summary(modonlyTtrendHorn)$tTable

par(las = 1, mai = c(0.8, 2, 0.1, 0.1))
rows1 <- which(grepl('temptrend', rownames(coefs)))
plot(0,0, col = 'white', xlim=c(-0.02, 0.85), ylim = c(1,length(rows1)+0.3), 
     yaxt='n', xlab = 'Turnover per |°C/yr|', ylab ='')
axis(2, at = length(rows1):1, labels = c('Freshwater', 'Marine', 'Terrestrial'), cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:length(rows1)){
  x = coefs[rows1[i], 1]
  x2 = coefs2[rows1[i], 1]
  x3 = coefs3[rows1[i], 1]
  if(i>1){ # if an interaction, add the intercept
    x = x + coefs['abs(temptrend)', 1]
    x2 = x2 + coefs2['abs(temptrend)', 1]
    x3 = x3 + coefs3['abs(temptrend)', 1]
  }
  se = coefs[rows1[i], 2]
  se2 = coefs2[rows1[i], 2]
  se3 = coefs3[rows1[i], 2]
  points(x, length(rows1) + 1 - i, pch = 16)
  lines(x = c(x-se, x+se), y = c(length(rows1) + 1 - i, length(rows1) + 1 - i))
  
  points(x2, length(rows1) + 1.1 - i, pch = 16, col = 'light grey')
  lines(x = c(x2-se2, x2+se2), y = c(length(rows1) + 1.1 - i, length(rows1) + 1.1 - i), col = 'light grey')
  
  points(x3, length(rows1) + 1.2 - i, pch = 16, col = 'dark grey')
  lines(x = c(x3-se3, x3+se3), y = c(length(rows1) + 1.2 - i, length(rows1) + 1.2 - i), col = 'dark grey')
}
legend('bottomright', col = c('black', 'dark grey', 'light grey'), lwd = 1, pch = 16, 
       legend = c('Jaccard turnover', 'Jaccard total', 'Horn-Morisita'))

```


### Full model
```{r LME Jacard turnover temperature full}
i <- trends[, complete.cases(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, human.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modTfull1.rds')){
  modTfull1 <- readRDS('temp/modTfull1.rds')
} else {
  modTfull1 <- lme(Jtutrend ~ temptrend_abs.sc*REALM + 
                     temptrend_abs.sc*tempave.sc +
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*lifespan.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*endothermfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend.sc*thermal_bias.sc +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*human.sc*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modTfull1, file = 'temp/modTfull1.rds')
}

summary(modTfull1)


```


#### Try simplifying the model
```{r simplify modTfull1}
i <- trends[, complete.cases(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, human.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modTfull1simpreml.rds')){
  modTfull1simpreml <- readRDS('temp/modTfull1simpreml.rds')
} else {
  require(MASS) # for stepAIC
  modTfull1ml <- update(modTfull1, method = 'ML')
  modTfull1simp <- stepAIC(modTfull1ml, direction = 'backward')
  modTfull1simpreml <- update(modTfull1simp, method = 'REML')
  saveRDS(modTfull1simpreml, file = 'temp/modTfull1simpreml.rds')
}
summary(modTfull1simpreml)

```


#### Plot the coefficients from the simplified model
```{r plot fullTmod1simpreml}

coefs <- summary(modTfull1simpreml)$tTable
par(las = 1, mai = c(0.5, 3, 0.1, 0.1))
rows1 <- which(!grepl('Intercept', rownames(coefs)))
plot(0,0, col = 'white', xlim=c(-0.02, 0.08), ylim = c(1,length(rows1)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(rows1):1, labels = rownames(coefs)[rows1], cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:length(rows1)){
  x = coefs[rows1[i], 1]
  se = coefs[rows1[i], 2]
  points(x, length(rows1) + 1 - i, pch = 16)
  lines(x = c(x-se, x+se), y = c(length(rows1) + 1 - i, length(rows1) + 1 - i))
}
```

#### Plot residuals against each predictor
```{r resids modTfull1simp, fig.height = 10, fig.width=10}
resids <- resid(modTfull1simpreml)
preds <- getData(modTfull1simpreml)
col = '#00000033'
cex = 0.5
par(mfrow = c(4,4))
boxplot(resids ~ preds$REALM, cex = cex, col = col)
plot(preds$temptrend_abs.sc, resids, cex = cex, col = col)
plot(preds$temptrend.sc, resids, cex = cex, col = col)
plot(preds$tempave.sc, resids, cex = cex, col = col)
plot(preds$tempave_metab.sc, resids, cex = cex, col = col)
plot(preds$seas.sc, resids, cex = cex, col = col)
plot(preds$microclim.sc, resids, cex = cex, col = col)
plot(preds$mass.sc, resids, cex = cex, col = col)
plot(preds$speed.sc, resids, cex = cex, col = col)
plot(preds$lifespan.sc, resids, cex = cex, col = col)
plot(preds$consumerfrac.sc, resids, cex = cex, col = col)
plot(preds$endothermfrac.sc, resids, cex = cex, col = col)
plot(preds$nspp.sc, resids, cex = cex, col = col)
plot(preds$thermal_bias.sc, resids, cex = cex, col = col)
plot(preds$npp.sc, resids, cex = cex, col = col)
plot(preds$human.sc, resids, cex = cex, col = col)
```

## Sensitivity analysis: total turnover and Morisita-Horn models
```{r LME Jacard total and MH models, fig.width=10, fig.height=8}
i2 <- trends[, complete.cases(Jbetatrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, human.sc)]
i3 <- trends[, complete.cases(Horntrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, human.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

# full models
if(file.exists('temp/modTfullJbeta.rds')){
  modTfullJbeta <- readRDS('temp/modTfullJbeta.rds')
} else {
  modTfullJbeta <- lme(Jbetatrend ~ temptrend_abs.sc*REALM + 
                     temptrend_abs.sc*tempave.sc +
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*lifespan.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*endothermfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend.sc*thermal_bias.sc +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*human.sc*REALM,
                   random = randef, weights = varef, data = trends[i2,], method = 'REML')
  saveRDS(modTfullJbeta, file = 'temp/modTfullJbeta.rds')
}

if(file.exists('temp/modTfullHorn.rds')){
  modTfullHorn <- readRDS('temp/modTfullHorn.rds')
} else {
  modTfullHorn <- lme(Horntrend ~ temptrend_abs.sc*REALM + 
                     temptrend_abs.sc*tempave.sc +
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*lifespan.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*endothermfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend.sc*thermal_bias.sc +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*human.sc*REALM,
                   random = randef, weights = varef, data = trends[i3,], method = 'REML')
  saveRDS(modTfullHorn, file = 'temp/modTfullHorn.rds')
}

summary(modTfullJbeta)
summary(modTfullHorn)

# simplify
if(file.exists('temp/modTfullJbetasimp.rds')){
  modTfullJbetasimp <- readRDS('temp/modTfullJbetasimp.rds')
} else {
  require(MASS) # for stepAIC
  modTfullJbetaml <- update(modTfullJbeta, method = 'ML')
  modTfullJbetasimp <- stepAIC(modTfullJbetaml, direction = 'backward')
  saveRDS(modTfullJbetasimp, file = 'temp/modTfullJbetasimp.rds')
}

if(file.exists('temp/modTfullHornsimp.rds')){
  modTfullHornsimp <- readRDS('temp/modTfullHornsimp.rds')
} else {
  require(MASS) # for stepAIC
  modTfullHornml <- update(modTfullHorn, method = 'ML')
  modTfullHornsimp <- stepAIC(modTfullHornml, direction = 'backward')
  saveRDS(modTfullHornsimp, file = 'temp/modTfullHornsimp.rds')
}

summary(modTfullJbetasimp)
summary(modTfullHornsimp)

if(file.exists('temp/modTfullJbetasimpreml.rds')){
  modTfullJbetasimpreml <- readRDS('temp/modTfullJbetasimpreml.rds')
} else {
  modTfullJbetasimpreml <- update(modTfullJbetasimp, method = 'REML')
  saveRDS(modTfullJbetasimpreml, file = 'temp/modTfullJbetasimpreml.rds')
}
if(file.exists('temp/modTfullHornsimpreml.rds')){
  modTfullHornsimpreml <- readRDS('temp/modTfullHornsimpreml.rds')
} else {
  modTfullHornsimpreml <- update(modTfullHornsimp, method = 'REML')
  saveRDS(modTfullHornsimpreml, file = 'temp/modTfullHornsimpreml.rds')
}

# plot coefs
coefs2 <- summary(modTfullJbetasimpreml)$tTable
coefs3 <- summary(modTfullHornsimpreml)$tTable
varstoplot <- unique(c(rownames(coefs2), rownames(coefs3)))

rows1 <- which(!grepl('Intercept|REALM', varstoplot) | grepl(':', varstoplot)) # vars to plot in first graph
rows1_2 <- which(rownames(coefs2) %in% varstoplot[rows1]) # rows in coefs2
rows1_3 <- which(rownames(coefs3) %in% varstoplot[rows1]) # rows in coefs3
xlims1 <- range(c(coefs2[rows1_2,1] - coefs2[rows1_2,2], 
                  coefs2[rows1_2,1] + coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - coefs3[rows1_3,2], 
                  coefs3[rows1_3,1] + coefs3[rows1_3,2]))

rows2 <- which(grepl('REALM', varstoplot) & !grepl(':', varstoplot)) # vars to plot in 2nd graph
rows2_2 <- which(rownames(coefs2) %in% varstoplot[rows2]) # rows in coefs2
rows2_3 <- which(rownames(coefs3) %in% varstoplot[rows2]) # rows in coefs3
xlims2 <- range(c(coefs2[rows2_2,1] - coefs2[rows2_2,2], 
                  coefs2[rows2_2,1] + coefs2[rows2_2,2], 
                  coefs3[rows2_3,1] - coefs3[rows2_3,2], 
                  coefs3[rows2_3,1] + coefs3[rows2_3,2]))

cols <- c('black', 'grey') # for Jbeta and Horn models, respectively
offs1 <- 0.1 # offset vertically for the two models
offs2 <- 0.01 # offset vertically for the two models (plot 2)

par(mfrow=c(1,2), las = 1, mai = c(0.5, 3, 0.1, 0.1))

plot(0,0, col = 'white', xlim=xlims1, ylim = c(1,length(rows1)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(rows1):1, labels = varstoplot[rows1], cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:length(rows1)){
  if(varstoplot[rows1[i]] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[rows1[i]], 1]
    se = coefs2[rownames(coefs2) == varstoplot[rows1[i]], 2]
    points(x, length(rows1) + 1 - i + offs1, pch = 16, col = cols[1])
    lines(x = c(x-se, x+se), y = c(length(rows1) + 1 - i + offs1, length(rows1) + 1 - i + offs1), col = cols[1])
  }
  if(varstoplot[rows1[i]] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[rows1[i]], 1]
    se = coefs3[rownames(coefs3) == varstoplot[rows1[i]], 2]
    points(x, length(rows1) + 1 - i - offs1, pch = 16, col = cols[2])
    lines(x = c(x-se, x+se), y = c(length(rows1) + 1 - i - offs1, length(rows1) + 1 - i - offs1), col = cols[2])
  }
}

plot(0,0, col = 'white', xlim=xlims2, ylim = c(0.9, length(rows2) + 0.1), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(rows2):1, labels = varstoplot[rows2], cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:length(rows2)){
  if(varstoplot[rows2[i]] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[rows2[i]], 1]
    se = coefs2[rownames(coefs2) == varstoplot[rows2[i]], 2]
    points(x, length(rows2) + 1 - i + offs2, pch = 16, col = cols[1])
    lines(x = c(x-se, x+se), y = c(length(rows2) + 1 - i + offs2, length(rows2) + 1 - i + offs2), col = cols[1])
  }
  if(varstoplot[rows2[i]] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[rows2[i]], 1]
    se = coefs3[rownames(coefs3) == varstoplot[rows2[i]], 2]
    points(x, length(rows2) + 1 - i - offs2, pch = 16, col = cols[2])
    lines(x = c(x-se, x+se), y = c(length(rows2) + 1 - i - offs2, length(rows2) + 1 - i - offs2), col = cols[2])
  }
}

```

Black is for Jaccard total turnover (pres/abs), grey is for Morisita-Horn turnover (considers abundance)

# To do

