---
title: 'Drivers of variation in the community response to temperature change across realms'
subtitle: '(using mixed effects models)'
output: 
    github_document: default

---
# Prep
```{r setup}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(nlme) # for ME models
library(maps) # for map
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
library(RColorBrewer)
library(MASS) # for stepAIC
library(piecewiseSEM) # for rsquared() for nlme models
library(scales) # for defining custom scales in ggplot

options(width=500) # turn off most text wrapping

signedsqrt = function(x) sign(x)*sqrt(abs(x))
signedsq = function(x) sign(x) * x^2
signedsqrttrans <- trans_new(name = 'signedsqrt', transform = signedsqrt, inverse = signedsq)


# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
```

```{r load data}
# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends <- fread('output/turnover_w_covariates.csv.gz')

# set realm order
trends[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# set up sign of temperature change
trends[, tsign := factor(sign(temptrend))]

# realm that combined Terrestrial and Freshwater, for interacting with human impact
trends[, REALM2 := REALM]
levels(trends$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")

# group Marine invertebrates/plants in with All
trends[, taxa_mod2 := taxa_mod]
trends[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']

# calculate duration
trends[, duration := maxyrBT - minyrBT + 1]

# trim to data with >= 3 yrs
trends <- trends[nyrBT >= 3, ]
```


## Log-transform some variables, then center and scale. 
``` {r center and scale}
trends[, tempave.sc := scale(tempave)]
trends[, tempave_metab.sc := scale(tempave_metab)]
trends[, seas.sc := scale(seas)]
trends[, microclim.sc := scale(log(microclim))]
trends[, temptrend.sc := scale(temptrend, center = FALSE)] # do not center
trends[, temptrend_abs.sc := scale(abs(temptrend), center = FALSE)] # do not center, so that 0 is still 0 temperature change
trends[, mass.sc := scale(log(mass_mean_weight))]
trends[, speed.sc := scale(log(speed_mean_weight+1))]
trends[, lifespan.sc := scale(log(lifespan_mean_weight))]
trends[, consumerfrac.sc := scale(consfrac)]
trends[, endothermfrac.sc := scale(endofrac)]
trends[, nspp.sc := scale(log(Nspp))]
trends[, thermal_bias.sc := scale(thermal_bias)]
trends[, npp.sc := scale(log(npp))]
trends[, veg.sc := scale(log(veg+1))]
trends[, duration.sc := scale(log(duration))]
trends[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
trends[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
trends[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]
```

# Summary plots and stats
## Examine how many data points are available
### Just turnover
```{r sample size all}
cat('Overall # time-series: ', nrow(trends), '\n')
cat('# studies: ', trends[, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[, sum(nyrBT)], '\n')
trends[, table(REALM)]
trends[, table(taxa_mod)]
trends[, table(taxa_mod, REALM)]
```

### With all covariates (Bowler for human)
```{r sample size for Jaccard turnover}
# the cases we can compare
apply(trends[, .(Jtutrendrem0, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, temptrend.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends[, complete.cases(Jtutrendrem0, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, temptrend.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
cat('Overall # time-series: ', sum(i), '\n')
cat('# studies: ', trends[i, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[i, sum(nyrBT)], '\n')
trends[i, table(REALM)]
trends[i, table(taxa_mod)]
trends[i, table(taxa_mod, REALM)]
```

## Average rates of turnover
```{r rates of turnover rem0}
trends[abs(temptrend) >= 0.5, temptrendtext1 := 'Changing >=0.5']
trends[abs(temptrend) <= 0.05, temptrendtext1 := 'Stable <=0.05']
trends[temptrend <= -0.5, temptrendtext2 := 'Cooling <= -0.5']
trends[temptrend >= 0.5, temptrendtext2 := 'Warming >= 0.5']
trends[abs(temptrend) >= 0.5 & abs(temptrend) < 1, temptrendtext3 := 'Slow Changing >=0.5 & <1']
trends[abs(temptrend) >= 1, temptrendtext3 := 'Fast Changing >=1']

# reshape to long format
measurenms <- c('Jtutrendrem0', 'Jtutrendrem0_se', 'Jbetatrendrem0', 'Jbetatrendrem0_se', 'Horntrendrem0', 'Horntrendrem0_se', 'Jtutrendz', 'Jbetatrendz', 'Horntrendz', 'Jtulast', 'Jbetalast', 'Hornlast', 'Jtuexp', 'Jbetaexp', 'Hornexp', 'Jtumm', 'Jbetamm', 'Hornmm')
idnms <- setdiff(names(trends), measurenms)
trends2 <- melt(trends, id.vars = idnms, measure.vars = measurenms)

# changing vs. stable
trendsum1 <- trends2[!is.na(temptrendtext1), 
                    .(ave = mean(value, na.rm=TRUE), 
                      se = sd(value, na.rm=TRUE)/sqrt(sum(!is.na(value))),
                      n = sum(!is.na(value))),
                    by = .(text = temptrendtext1, type = variable)] # turnover per year for locations changing temperature


# warming vs. cooling
trendsum2 <- trends2[!is.na(temptrendtext2), 
                    .(ave = mean(value, na.rm=TRUE), 
                      se = sd(value, na.rm=TRUE)/sqrt(sum(!is.na(value))),
                      n = sum(!is.na(value))),
                    by = .(text = temptrendtext2, type = variable)] # inc. direction

# slow vs. fast changing
trendsum3 <- trends2[!is.na(temptrendtext3), 
                    .(ave = mean(value, na.rm=TRUE), 
                      se = sd(value, na.rm=TRUE)/sqrt(sum(!is.na(value))),
                      n = sum(!is.na(value))),
                    by = .(text = temptrendtext3, type = variable)] # inc. rate

# combine
trendsum4 <- rbind(trendsum1, trendsum2, trendsum3)
setorder(trendsum4, type, text)

write.csv(trendsum4, file = 'output/trendsummary.csv', row.names = FALSE)

trendsum4
```

## Plots of turnover rates binned by warming rates
```{r turnover vs. temperature violin plot, message = FALSE, warning = FALSE, fig.height= 12, fig.width = 6}
p1 <- ggplot(trends[!is.na(temptrendtext1), ], aes(temptrendtext1, Horntrendrem0)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') +
  labs(x = '', y = 'Slope (Horn)', tag = 'A', title = 'Rate of temperature change') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10)) +
  geom_abline(intercept = 0, slope = 0)

p2 <- ggplot(trends[!is.na(text), ], aes(temptrendtext2, Horntrendrem0)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') +
  labs(x = '', y = 'Slope (Horn)', tag = 'B', title = 'Rate of temperature change') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10)) +
  geom_abline(intercept = 0, slope = 0)

p3 <- ggplot(trendsum4[type %in% c('Jtutrendrem0', 'Jbetatrendrem0', 'Horntrendrem0')], 
             aes(text, ave, group = type, color = type)) +
  geom_point(position = position_dodge(width = 0.25), size = 0.5) +
  geom_errorbar(aes(ymin=ave-se, ymax=ave+se), width=.1, position = position_dodge(width = 0.25)) +
  labs(x = '', y = 'Slope', title = 'Slope of dissimilarity vs. time') +
  geom_abline(intercept = 0, slope = 0, linetype = 'dashed') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  coord_cartesian(ylim = c(-0.01,0.08))

p4 <- ggplot(trendsum4[type %in% c('Jtutrendz', 'Jbetatrendz', 'Horntrendz')], 
             aes(text, ave, group = type, color = type)) +
  geom_point(position = position_dodge(width = 0.25), size = 0.5) +
  geom_errorbar(aes(ymin=ave-se, ymax=ave+se), width=.1, position = position_dodge(width = 0.25)) +
  labs(x = '', y = 'Standardized slope', title = 'Null model standardized slope') +
  geom_abline(intercept = 0, slope = 0, linetype = 'dashed') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        axis.text.x = element_text(angle = 45, vjust = 0.7))

p5 <- ggplot(trendsum4[type %in% c('Jtulast', 'Jbetalast', 'Hornlast')], 
             aes(text, ave, group = type, color = type)) +
  geom_point(position = position_dodge(width = 0.25), size = 0.5) +
  geom_errorbar(aes(ymin=ave-se, ymax=ave+se), width=.1, position = position_dodge(width = 0.25)) +
  labs(x = '', y = 'Dissimilarity', title = 'First to last year') +
  geom_abline(intercept = 0, slope = 0, linetype = 'dashed') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        axis.text.x = element_text(angle = 45, vjust = 0.7))

p6 <- ggplot(trendsum4[type %in% c('Jtuexp', 'Jbetaexp', 'Hornexp')], 
             aes(text, ave, group = type, color = type)) +
  geom_point(position = position_dodge(width = 0.25), size = 0.5) +
  geom_errorbar(aes(ymin=ave-se, ymax=ave+se), width=.1, position = position_dodge(width = 0.25)) +
  labs(x = '', y = 'Years', title = 'Exponential half-saturation') +
  geom_abline(intercept = 0, slope = 0, linetype = 'dashed') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        axis.text.x = element_text(angle = 45, vjust = 0.7))

p7 <- ggplot(trendsum4[type %in% c('Jtumm', 'Jbetamm', 'Hornmm')], 
             aes(text, ave, group = type, color = type)) +
  geom_point(position = position_dodge(width = 0.25), size = 0.5) +
  geom_errorbar(aes(ymin=ave-se, ymax=ave+se), width=.1, position = position_dodge(width = 0.25)) +
  labs(x = '', y = 'Years', title = 'MM half-saturation') +
  geom_abline(intercept = 0, slope = 0, linetype = 'dashed') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        axis.text.x = element_text(angle = 45, vjust = 0.7))

grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol = 2, 
             layout_matrix = rbind(c(1,2), c(3,3), c(4,4), c(5,5), c(6,6), c(7,7)),
             heights=c(unit(0.1, "npc"), unit(0.18, "npc"), 
                       unit(0.18, "npc"), unit(0.18, "npc"), 
                       unit(0.18, "npc"), unit(0.18, "npc")))
```

## Plots of turnover rates binned by warming rates and duration
```{r turnover vs. temperature trend vs. duration} #, message = FALSE, warning = FALSE}
ggplot(trends[abs(temptrend) <= 0.05, ], aes(x=duration, y=Jtutrendrem0)) +
  geom_smooth(method = 'gam', aes(color = '<=0.05')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.1 & abs(temptrend) > 0.05,], method = 'loess', aes(color = '0.05-0.1')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.5 & abs(temptrend) > 0.1,], method = 'loess', aes(color = '0.1-0.5')) +
  geom_smooth(data = trends[abs(temptrend) >= 0.5,], method = 'lm', aes(color = '>=0.5')) +
  scale_x_log10() + 
  labs(y = 'Jaccard turnover slope', color = '째C/year')

ggplot(trends[abs(temptrend) <= 0.05, ], aes(x=duration, y=Jbetatrendrem0)) +
  geom_smooth(method = 'gam', aes(color = '<=0.05')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.1 & abs(temptrend) > 0.05,], method = 'loess', aes(color = '0.05-0.1')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.5 & abs(temptrend) > 0.1,], method = 'loess', aes(color = '0.1-0.5')) +
  geom_smooth(data = trends[abs(temptrend) >= 0.5,], method = 'lm', aes(color = '>=0.5')) +
  scale_x_log10() + 
  labs(y = 'Jaccard total slope', color = '째C/year')

ggplot(trends[abs(temptrend) <= 0.05, ], aes(x=duration, y=Horntrendrem0)) +
  geom_smooth(method = 'gam', aes(color = '<=0.05')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.1 & abs(temptrend) > 0.05,], method = 'loess', aes(color = '0.05-0.1')) +
  geom_smooth(data = trends[abs(temptrend) <= 0.5 & abs(temptrend) > 0.1,], method = 'loess', aes(color = '0.1-0.5')) +
  geom_smooth(data = trends[abs(temptrend) >= 0.5,], method = 'lm', aes(color = '>=0.5')) +
  scale_x_log10() + 
  labs(y = 'Horn slope', color = '째C/year')

```


# Models

## Choose the variance structure for mixed effects models
Try combinations of

- variance scaled to a power of the number of years in the community time-series
- variance scaled to a power of the abs temperature trend
- random intercept for taxa_mod
- random intercept for STUDY_ID
- random slope (abs temperature trend) for taxa_mod
- random slope (abs temperature trend) for STUDY_ID
- random intercept for rarefyID (for overdispersion)

And choose the one with lowest AIC (not run: takes a long time)
```{r choose variance structure for Jacard turnover, eval = FALSE}
# fit models for variance structure
fixed <- formula(Jtutrendrem0 ~ temptrend_abs.sc*REALM +
                     temptrend_abs.sc*tsign + 
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend_abs.sc*thermal_bias.sc:tsign +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*veg.sc +
                     temptrend_abs.sc*duration.sc +
                     temptrend_abs.sc*human_bowler.sc:REALM2)
i <- trends[, complete.cases(Jtutrendrem0, temptrend_abs.sc, REALM, tsign, tempave_metab.sc, seas.sc, 
                             microclim.sc, mass.sc, speed.sc, consumerfrac.sc, nspp.sc,
                             thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
mods <- vector('list', 0)
mods[[1]] <- gls(fixed, data = trends[i,])
mods[[2]] <- gls(fixed, data = trends[i,], weights = varPower(-0.5, ~nyrBT))
mods[[3]] <- gls(fixed, data = trends[i,], weights = varPower(0.5, ~temptrend_abs.sc))

mods[[4]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, control = lmeControl(opt = "optim"))
mods[[5]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, control = lmeControl(opt = "optim"))
mods[[6]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID, control = lmeControl(opt = "optim"))
mods[[7]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))
mods[[8]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))

mods[[9]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | taxa_mod)
mods[[10]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)) # includes overdispersion. new formula so that random slope is only for study level (not enough data to extend to rarefyID).

mods[[11]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))
mods[[12]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ 1, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))

aics <- sapply(mods, AIC)
minaics <- aics - min(aics)
minaics
which.min(aics)
```
Chooses the random slopes (temptrend_abs) & intercepts for STUDY_ID, overdispersion, and variance scaled to number of years.
We haven't dealt with potential testing on the boundary issues here yet.


## Temperature-only models
### Fit the models
```{r LME temperature only}

randef <- list(STUDY_ID = ~ abs(temptrend), rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

# *trendrem0
if(file.exists('temp/modonlyTtrendrem0.rds')){
  modonlyTtrendrem0 <- readRDS('temp/modonlyTtrendrem0.rds')
} else {
  i <- trends[, complete.cases(Jtutrendrem0, REALM, temptrend)]
  modonlyTtrendrem0 <- lme(Jtutrendrem0 ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrendrem0, file = 'temp/modonlyTtrendrem0.rds')
}

if(file.exists('temp/modonlyTtrendJbetarem0.rds')){
  modonlyTtrendJbetarem0 <- readRDS('temp/modonlyTtrendJbetarem0.rds')
} else {
  i <- trends[, complete.cases(Jbetatrendrem0, REALM, temptrend)]
  modonlyTtrendJbetarem0 <- lme(Jbetatrendrem0 ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJbetarem0, file = 'temp/modonlyTtrendJbetarem0.rds')
}

if(file.exists('temp/modonlyTtrendHornrem0.rds')){
  modonlyTtrendHornrem0 <- readRDS('temp/modonlyTtrendHornrem0.rds')
} else {
  i <- trends[, complete.cases(Horntrendrem0, REALM, temptrend)]
  modonlyTtrendHornrem0 <- lme(Horntrendrem0 ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrendHornrem0, file = 'temp/modonlyTtrendHornrem0.rds')
}

# *trendz
if(file.exists('temp/modonlyTtrendJtuz.rds')){
  modonlyTtrendJtuz <- readRDS('temp/modonlyTtrendJtuz.rds')
} else {
  i <- trends[, complete.cases(Jtutrendz, REALM, temptrend)]
  modonlyTtrendJtuz <- lme(Jtutrendz ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML',
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJtuz, file = 'temp/modonlyTtrendJtuz.rds')
}

if(file.exists('temp/modonlyTtrendJbetaz.rds')){
  modonlyTtrendJbetaz <- readRDS('temp/modonlyTtrendJbetaz.rds')
} else {
  i <- trends[, complete.cases(Jbetatrendz, REALM, temptrend)]
  modonlyTtrendJbetaz <- lme(Jbetatrendz ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJbetaz, file = 'temp/modonlyTtrendJbetaz.rds')
}

if(file.exists('temp/modonlyTtrendHornz.rds')){
  modonlyTtrendHornz <- readRDS('temp/modonlyTtrendHornz.rds')
} else {
  i <- trends[, complete.cases(Horntrendz, REALM, temptrend)]
  modonlyTtrendHornz <- lme(Horntrendz ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrendHornz, file = 'temp/modonlyTtrendHornz.rds')
}

# *exp
if(file.exists('temp/modonlyTtrendJtuexp.rds')){
  modonlyTtrendJtuexp <- readRDS('temp/modonlyTtrendJtuexp.rds')
} else {
  i <- trends[, complete.cases(Jtuexp, REALM, temptrend)]
  modonlyTtrendJtuexp <- lme(log(Jtuexp) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML',
                   control=lmeControl(msMaxIter = 100, maxIter = 100, opt = 'optim'))
  saveRDS(modonlyTtrendJtuexp, file = 'temp/modonlyTtrendJtuexp.rds')
}

if(file.exists('temp/modonlyTtrendJbetaexp.rds')){
  modonlyTtrendJbetaexp <- readRDS('temp/modonlyTtrendJbetaexp.rds')
} else {
  i <- trends[, complete.cases(Jbetaexp, REALM, temptrend)]
  modonlyTtrendJbetaexp <- lme(log(Jbetaexp) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJbetaexp, file = 'temp/modonlyTtrendJbetaexp.rds')
}

if(file.exists('temp/modonlyTtrendHornexp.rds')){
  modonlyTtrendHornexp <- readRDS('temp/modonlyTtrendHornexp.rds')
} else {
  i <- trends[, complete.cases(Hornexp, REALM, temptrend)]
  modonlyTtrendHornexp <- lme(log(Hornexp) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML',
                   control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modonlyTtrendHornexp, file = 'temp/modonlyTtrendHornexp.rds')
}


# *mm
if(file.exists('temp/modonlyTtrendJtumm.rds')){
  modonlyTtrendJtumm <- readRDS('temp/modonlyTtrendJtumm.rds')
} else {
  i <- trends[, complete.cases(Jtumm, REALM, temptrend)]
  modonlyTtrendJtumm <- lme(log(Jtumm+1) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrendJtumm, file = 'temp/modonlyTtrendJtumm.rds')
}

if(file.exists('temp/modonlyTtrendJbetamm.rds')){
  modonlyTtrendJbetamm <- readRDS('temp/modonlyTtrendJbetamm.rds')
} else {
  i <- trends[, complete.cases(Jbetamm, REALM, temptrend)]
  modonlyTtrendJbetamm <- lme(log(Jbetamm+1) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJbetamm, file = 'temp/modonlyTtrendJbetamm.rds')
}

if(file.exists('temp/modonlyTtrendHornmm.rds')){
  modonlyTtrendHornmm <- readRDS('temp/modonlyTtrendHornmm.rds')
} else {
  i <- trends[, complete.cases(Hornmm, REALM, temptrend)]
  modonlyTtrendHornmm <- lme(log(Hornmm+1) ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendHornmm, file = 'temp/modonlyTtrendHornmm.rds')
}
```

### Summary: Slope of dissimilarity
```{r summary of rem0 mods}
summary(modonlyTtrendrem0)
summary(modonlyTtrendJbetarem0)
summary(modonlyTtrendHornrem0)
```

### Summary: Standardized slope of dissimilarity
```{r summary of z mods}
# summary(modonlyTtrendJtuz)
summary(modonlyTtrendJbetaz)
summary(modonlyTtrendHornz)
```

### Summary: Half-saturation exponential fits
```{r summary of exp mods}
#summary(modonlyTtrendJtuexp)
summary(modonlyTtrendJbetaexp)
summary(modonlyTtrendHornexp)
```

### Summary: Half-saturation MM fits
```{r summary of mm mods}
summary(modonlyTtrendJtumm)
summary(modonlyTtrendJbetamm)
summary(modonlyTtrendHornmm)
```


### Plot the temp-only coefficients
```{r modonlyTtrendsimp coefs}
colors <- brewer.pal(3, 'Dark2')

# make table of coefficients
coefs1 <- as.data.frame(summary(modonlyTtrendrem0)$tTable)
coefs2 <- as.data.frame(summary(modonlyTtrendJbetarem0)$tTable)
coefs3 <- as.data.frame(summary(modonlyTtrendHornrem0)$tTable)
coefs1$mod <- 'Jtu'
coefs2$mod <- 'Jbeta'
coefs3$mod <- 'Horn'
rows1 <- which(grepl('temptrend', rownames(coefs1))) # extract temperature effect
cols <- c('Value', 'Std.Error', 'mod')
allcoefs <- rbind(coefs1[rows1, cols], coefs2[rows1, cols], coefs3[rows1, cols])
allcoefs$Value[grepl('REALMMarine', rownames(allcoefs))] <- 
  allcoefs$Value[grepl('REALMMarine', rownames(allcoefs))] + 
  allcoefs$Value[!grepl('REALM', rownames(allcoefs))] # add intercept to marine effects
allcoefs$Value[grepl('REALMTerrestrial', rownames(allcoefs))] <- 
  allcoefs$Value[grepl('REALMTerrestrial', rownames(allcoefs))] + 
  allcoefs$Value[!grepl('REALM', rownames(allcoefs))] # add intercept to terrestrial effects

allcoefs$lCI <- allcoefs$Value - allcoefs$Std.Error # lower confidence interval
allcoefs$uCI <- allcoefs$Value + allcoefs$Std.Error
allcoefs$y <- c(3, 2, 1) + rep(c(0, -0.1, -0.2), c(3, 3, 3)) # y-values
allcoefs$col <- c(rep(colors[1], 3), rep(colors[2], 3), rep(colors[3], 3))
allcoefs$realm <- rep(c('Freshwater', 'Marine', 'Terrestrial'), 3)

par(las = 1, mai = c(0.8, 2, 0.1, 0.1))
plot(0,0, col = 'white', xlim=c(-0.1, 0.85), ylim = c(0.5,3), 
     yaxt='n', xlab = 'Turnover per |째C/yr|', ylab ='')
axis(2, at = 3:1, labels = c('Freshwater', 'Marine', 'Terrestrial'), cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:nrow(allcoefs)){
  with(allcoefs[i, ], points(Value, y, pch = 16, col = col))
  with(allcoefs[i, ], lines(x = c(lCI, uCI), y = c(y, y), col = col))
}
legend('bottomright', col = colors, lwd = 1, pch = 16, 
       legend = c('Jaccard turnover rem0', 'Jaccard total rem0', 'Horn-Morisita rem0'))

```

## Full models
Try static covariates plus interactions of abs temperature trend with each covariate:

- realm
- speed
- mass
- average metabolic temperature
- consumer fraction
- environmental temperature
- seasonality
- microclimates
- thermal bias
- NPP
- vegetation
- duration
- human footprint

Except for thermal bias: interact with temperature trend (not abs)

### Bowler vs Venter/Halpern human impact
Bowler has lower AIC.
```{r LME Jacard turnover temperature full rem0}
# using Bowler for human impact
i1 <- trends[, complete.cases(Jtutrendrem0, REALM, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend_abs.sc, mass.sc, speed.sc, 
                             consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                             veg.sc, duration.sc, human_bowler.sc, human_footprint.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modTfullbowlerrem0.rds')){
  modTfullbowlerrem0 <- readRDS('temp/modTfullbowlerrem0.rds')
} else {

modTfullbowlerrem0 <- lme(Jtutrendrem0 ~ temptrend_abs.sc*REALM +
                     temptrend_abs.sc*tsign + 
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend_abs.sc*thermal_bias.sc:tsign +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*veg.sc +
                     temptrend_abs.sc*duration.sc +
                     temptrend_abs.sc*human_bowler.sc:REALM2,
                   random = randef, weights = varef, data = trends[i1,], method = 'REML')
  saveRDS(modTfullbowlerrem0, file = 'temp/modTfullbowlerrem0.rds')
}

# using Venter/Halpern for human impact
if(file.exists('temp/modTfullfootprintrem0.rds')){
  modTfullfootprintrem0 <- readRDS('temp/modTfullfootprintrem0.rds')
} else {
modTfullfootprintrem0 <- lme(Jtutrendrem0 ~ temptrend_abs.sc*REALM + 
                     temptrend_abs.sc*tsign + 
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend_abs.sc*thermal_bias.sc:tsign +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*veg.sc +
                     temptrend_abs.sc*duration.sc +
                     temptrend_abs.sc*human_footprint.sc:REALM2,
                   random = randef, weights = varef, data = trends[i1,], method = 'REML',
                   control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullfootprintrem0, file = 'temp/modTfullfootprintrem0.rds')

}
AIC(modTfullbowlerrem0, modTfullfootprintrem0)

```

### Fit/load full models
```{r full models, fig.width=10, fig.height=8}
randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

terms <- c('temptrend_abs.sc*REALM', 
           'temptrend_abs.sc*tsign',
           'temptrend_abs.sc*tempave_metab.sc',
           'temptrend_abs.sc*seas.sc',
           'temptrend_abs.sc*microclim.sc',
           'temptrend_abs.sc*mass.sc',
           'temptrend_abs.sc*speed.sc', 
           'temptrend_abs.sc*consumerfrac.sc',
           'temptrend_abs.sc*nspp.sc',
           'temptrend_abs.sc*thermal_bias.sc:tsign',
           'temptrend_abs.sc*npp.sc',
           'temptrend_abs.sc*veg.sc',
           'temptrend_abs.sc*duration.sc',
           'temptrend_abs.sc*human_bowler.sc:REALM2')

completerows <- trends[, complete.cases(REALM, tempave_metab.sc, 
                                seas.sc, microclim.sc, 
                                temptrend_abs.sc, mass.sc, speed.sc, 
                                consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                                veg.sc, duration.sc, human_bowler.sc)]


# rem0
if(file.exists('temp/modTfullJturem0.rds')){
  modTfullJturem0 <- readRDS('temp/modTfullJturem0.rds')
} else {
  i <- trends[, complete.cases(Jtutrendrem0)] & completerows
  modTfullJturem0 <- lme(formula(paste0('Jtutrendrem0 ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML')
  saveRDS(modTfullJturem0, file = 'temp/modTfullJturem0.rds')
}

if(file.exists('temp/modTfullJbetarem0.rds')){
  modTfullJbetarem0 <- readRDS('temp/modTfullJbetarem0.rds')
} else {
  i <- trends[, complete.cases(Jbetatrendrem0)] & completerows
  modTfullJbetarem0 <- lme(formula(paste0('Jbetatrendrem0 ~ ', paste(terms, collapse = ' + '))),
                           random = randef, weights = varef, data = trends[i,], 
                           method = 'REML')
  saveRDS(modTfullJbetarem0, file = 'temp/modTfullJbetarem0.rds')
}

if(file.exists('temp/modTfullHornrem0.rds')){
  modTfullHornrem0 <- readRDS('temp/modTfullHornrem0.rds')
} else {
  i <- trends[, complete.cases(Horntrendrem0)] & completerows
  modTfullHornrem0 <- lme(formula(paste0('Horntrendrem0 ~ ', paste(terms, collapse = ' + '))),
                      random = randef, weights = varef, data = trends[i,], 
                      method = 'REML')
  saveRDS(modTfullHornrem0, file = 'temp/modTfullHornrem0.rds')
}

# trendz
if(file.exists('temp/modTfullJtuz.rds')){
  modTfullJtuz <- readRDS('temp/modTfullJtuz.rds')
} else {
  i <- trends[, complete.cases(Jtutrendz)] & completerows
  modTfullJtuz <- lme(formula(paste0('Jtutrendz ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML',
                   control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullJtuz, file = 'temp/modTfullJtuz.rds')
}

if(file.exists('temp/modTfullJbetaz.rds')){
  modTfullJbetaz <- readRDS('temp/modTfullJbetaz.rds')
} else {
  i <- trends[, complete.cases(Jbetatrendz)] & completerows
  modTfullJbetaz <- lme(formula(paste0('Jbetatrendz ~ ', paste(terms, collapse = ' + '))),
                           random = randef, weights = varef, data = trends[i,], 
                           method = 'REML')
  saveRDS(modTfullJbetaz, file = 'temp/modTfullJbetaz.rds')
}

if(file.exists('temp/modTfullHornz.rds')){
  modTfullHornz <- readRDS('temp/modTfullHornz.rds')
} else {
  i <- trends[, complete.cases(Horntrendz)] & completerows
  modTfullHornz <- lme(formula(paste0('Horntrendz ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML',
                         control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullHornz, file = 'temp/modTfullHornz.rds')
}

# exp
if(file.exists('temp/modTfullJtuexp.rds')){
  modTfullJtuexp <- readRDS('temp/modTfullJtuexp.rds')
} else {
  i <- trends[, complete.cases(Jtuexp)] & completerows
  modTfullJtuexp <- lme(formula(paste0('log(Jtuexp) ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML',
                   control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullJtuexp, file = 'temp/modTfullJtuexp.rds')
}

if(file.exists('temp/modTfullJbetaexp.rds')){
  modTfullJbetaexp <- readRDS('temp/modTfullJbetaexp.rds')
} else {
  i <- trends[, complete.cases(Jbetaexp)] & completerows
  modTfullJbetaexp <- lme(formula(paste0('log(Jbetaexp) ~ ', paste(terms, collapse = ' + '))),
                           random = randef, weights = varef, data = trends[i,], 
                           method = 'REML')
  saveRDS(modTfullJbetaexp, file = 'temp/modTfullJbetaexp.rds')
}

if(file.exists('temp/modTfullHornexp.rds')){
  modTfullHornexp <- readRDS('temp/modTfullHornexp.rds')
} else {
  i <- trends[, complete.cases(Hornexp)] & completerows
  modTfullHornexp <- lme(formula(paste0('log(Hornexp) ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML',
                         control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullHornexp, file = 'temp/modTfullHornexp.rds')
}

# mm
if(file.exists('temp/modTfullJtumm.rds')){
  modTfullJtumm <- readRDS('temp/modTfullJtumm.rds')
} else {
  i <- trends[, complete.cases(Jtumm)] & completerows
  modTfullJtumm <- lme(formula(paste0('log(Jtumm+1) ~ ', paste(terms, collapse = ' + '))),
                       random = randef, weights = varef, data = trends[i,], 
                       method = 'REML',
                       control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullJtumm, file = 'temp/modTfullJtumm.rds')
}

if(file.exists('temp/modTfullJbetamm.rds')){
  modTfullJbetamm <- readRDS('temp/modTfullJbetamm.rds')
} else {
  i <- trends[, complete.cases(Jbetamm)] & completerows
  modTfullJbetamm <- lme(formula(paste0('log(Jbetamm+1) ~ ', paste(terms, collapse = ' + '))),
                         random = randef, weights = varef, data = trends[i,], 
                         method = 'REML',
                         control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  saveRDS(modTfullJbetamm, file = 'temp/modTfullJbetamm.rds')
}

if(file.exists('temp/modTfullHornmm.rds')){
  modTfullHornmm <- readRDS('temp/modTfullHornmm.rds')
} else {
  i <- trends[, complete.cases(Hornmm)] & completerows
  modTfullHornmm <- lme(formula(paste0('log(Hornmm+1) ~ ', paste(terms, collapse = ' + '))),
                        random = randef, weights = varef, data = trends[i,], 
                        method = 'REML',
                        control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullHornmm, file = 'temp/modTfullHornmm.rds')
}
```

#### Summary rem0
```{r summary rem0 full}
summary(modTfullJturem0)
summary(modTfullJbetarem0)
summary(modTfullHornrem0)

rsquared(modTfullJturem0)
rsquared(modTfullJbetarem0)
rsquared(modTfullHornrem0)

```

#### Summary trendz
```{r summary trendz full}
summary(modTfullJtuz)
summary(modTfullJbetaz)
summary(modTfullHornz)

rsquared(modTfullJtuz)
rsquared(modTfullJbetaz)
rsquared(modTfullHornz)

```

#### Summary exp
```{r summary exp full}
summary(modTfullJtuexp)
summary(modTfullJbetaexp)
summary(modTfullHornexp)

rsquared(modTfullJtuexp)
rsquared(modTfullJbetaexp)
rsquared(modTfullHornexp)

```

#### Summary mm
```{r summary mm full}
summary(modTfullJtumm)
summary(modTfullJbetamm)
summary(modTfullHornmm)

rsquared(modTfullJtumm)
rsquared(modTfullJbetamm)
rsquared(modTfullHornmm)

```


### Plots from the full models
#### Plot the coefficients
```{r plot fullTmods, fig.height=12, fig.width=9}

coefs1 <- summary(modTfullJturem0)$tTable
coefs2 <- summary(modTfullJbetarem0)$tTable
coefs3 <- summary(modTfullHornrem0)$tTable

varstoplot <- unique(c(rownames(coefs1), rownames(coefs2), rownames(coefs3)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs1) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
xlims <- range(c(coefs1[rows1_1,1] - coefs1[rows1_1,2], coefs1[rows1_1,1] + coefs1[rows1_1,2], 
                  coefs2[rows1_2,1] - coefs2[rows1_2,2], coefs2[rows1_2,1] + coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - coefs3[rows1_3,2], coefs3[rows1_3,1] + coefs3[rows1_3,2]))


cols <- brewer.pal(3, 'Dark2') # for Jtu, Jbeta and Horn models
pchs <- c(16, 16, 16)
offs <- c(0.1, 0, -0.1) # offset vertically for each model


par(las = 1, mai = c(0.5, 4, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(1,length(varstoplot)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs1)){
    x = coefs1[rownames(coefs1) == varstoplot[i], 1]
    se = coefs1[rownames(coefs1) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
}
legend('topleft', col = cols, pch = 16, lwd = 1, legend = c('Jtu', 'Jbeta', 'Horn'), cex = 0.5)
```


#### Plot the main effects
##### Manually
```{r main effect plots, fig.width=12, fig.height=12, out.width=24, out.height=24, fig.retina=2}

# set up the variables to plot
# if variable is logged before scaling (see 'center and scale' above), then need to mark it here and express the limits on a log10 scale (even though log transforming is log)
vars <- data.frame(vars = c('temptrend_abs', 'tempave_metab', 'seas', 'microclim', 'mass', 'speed', 
                            'consumerfrac', 'nspp', 'thermal_bias', 'npp', 'veg', 'duration', 
                            'human_bowler', 'human_bowler'),
           min =      c(0,  0,   0.1, -2,  0,   0,   0,   0.3, -10, 1.9, 0,   0.5, 0,   0), 
           max =      c(2,  30,  16,  0.8, 8,   2,   1,   2.6, 10,  3.7, 0.3, 2,   1,   1),
           log =      c(F,  F,   F,   T,   T,   T,   F,   T,   F,   T,   T,   T,   T,   T),
           len =      c(100,  100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100),
           discrete = c(F,  F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F),
           plus =     c(0,  0,   0,   0,   0,   1,   0,   0,   0,   0,   1,   0,   1,   1), # what to add before log-scaling
           REALM = c(rep('Terrestrial', 13), 'Marine'),
           REALM2 = c(rep('TerrFresh', 13), 'Marine'),
           stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', 
                      temptrend = -0.0001,
                      temptrend_abs.sc = 0.0001,
                      tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      seas.sc = mean(seas.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      speed.sc = mean(speed.sc, na.rm=TRUE), 
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      nspp.sc = 0, 
                      thermal_bias.sc = mean(thermal_bias.sc, na.rm=TRUE), 
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                      veg.sc = mean(veg.sc, na.rm=TRUE), 
                      consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     temptrend = -0.0001,
                     temptrend_abs.sc = 0.0001,
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                     seas.sc = mean(seas.sc, na.rm=TRUE), 
                     microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                     speed.sc = mean(speed.sc, na.rm=TRUE), 
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     nspp.sc = 0, 
                     thermal_bias.sc = 0, 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                     veg.sc = mean(veg.sc, na.rm=TRUE), 
                     consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    temptrend = -0.0001,
                    temptrend_abs.sc = 0.0001,
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    seas.sc = mean(seas.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    speed.sc = mean(speed.sc, na.rm=TRUE), 
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    nspp.sc = 0, 
                    thermal_bias.sc = 0, 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                    veg.sc = mean(veg.sc, na.rm=TRUE), 
                    consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)
basetab[, ':='(duration.sc = 0, nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_514668')]

# make the data frames for each interaction to plot                
for(j in 1:nrow(vars)){
  # set up the main effects
  if(vars$log[j]){
    thisdat <- data.frame(new = 10^seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                          var = vars$vars[j], stringsAsFactors = FALSE)
  } 
  if(!vars$log[j]){
    thisdat <- data.frame(new = seq(vars$min[j], vars$max[j], length.out = vars$len[j]),
                          var = vars$vars[j], stringsAsFactors = FALSE)
  }
  names(thisdat) <- c(vars$vars[j], 'var')

  # scale the variable
  cent <- attr(trends[[paste0(vars$vars[j], '.sc')]], 'scaled:center')
  scl <- attr(trends[[paste0(vars$vars[j], '.sc')]], 'scaled:scale')
  if(is.null(cent)) cent <- 0
  if(!is.null(cent) & !is.null(scl)){
    if(vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (log(thisdat[[vars$vars[j]]] + vars$plus[j]) - cent)/scl
    if(!vars$log[j]) thisdat[[paste0(vars$var[j], '.sc')]] <- (thisdat[[vars$var[j]]] - cent)/scl
  }

  # merge with the rest of the columns
  # use realm-specific averages for human impacts
  if(vars$vars[j] != 'tsign') colnamestouse <- setdiff(colnames(basetab), paste0(vars$vars[j], '.sc'))
  if(vars$vars[j] == 'tsign') colnamestouse <- setdiff(colnames(basetab), vars$var[j])
  if(vars$vars[j] != 'human_bowler'){
    thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
  }
  if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Terrestrial'){
    thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
  }
  if(vars$vars[j] == 'human_bowler' & vars$REALM[j] == 'Marine'){
    thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
  }
  
  # add realm
  thisdat$REALM <- vars$REALM[j]
  thisdat$REALM2 <- vars$REALM2[j]
  
  # merge with the previous iterations
  if(j == 1) newdat <- thisdat
  if(j > 1){
    colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
    for(toadd in colstoadd){
      newdat[[toadd]] <- NA
    }
    
    colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
    for(toadd in colstoadd2){
      thisdat[[toadd]] <- NA
    }
    
    newdat <- rbind(newdat, thisdat)
  } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$temptrend_abs.sc[1:6] <- rep(0.0001, 6)
newdat$temptrend[1:6] <- rep(c(0.0001, -0.0001), 3)
newdat$var[1:6] <- 'test'

# trim to at least some temperature change (so that tsign is -1 or 1)
newdat <- newdat[newdat$temptrend_abs.sc != 0,]

# set up tsign
newdat$tsign <- factor(sign(newdat$temptrend))


# make predictions
newdat$predsJtu <- predict(object = modTfullJturem0, newdata = newdat, level = 0)
newdat$predsJbeta <- predict(object = modTfullJbetarem0, newdata = newdat, level = 0)
newdat$predsHorn <- predict(object = modTfullHornrem0, newdata = newdat, level = 0)

#compute standard error for predictions
# from https://stackoverflow.com/questions/14358811/extract-prediction-band-from-lme-fit
DesignmatJtu <- model.matrix(eval(eval(modTfullJturem0$call$fixed)[-2]), newdat)
DesignmatJbeta <- model.matrix(eval(eval(modTfullJbetarem0$call$fixed)[-2]), newdat)
DesignmatHorn <- model.matrix(eval(eval(modTfullHornrem0$call$fixed)[-2]), newdat)

predvarJtu <- diag(DesignmatJtu %*% modTfullJturem0$varFix %*% t(DesignmatJtu))
predvarJbeta <- diag(DesignmatJbeta %*% modTfullJbetarem0$varFix %*% t(DesignmatJbeta))
predvarHorn <- diag(DesignmatHorn %*% modTfullHornrem0$varFix %*% t(DesignmatHorn))

newdat$SE_Jtu <- sqrt(predvarJtu) 
newdat$SE_Jbeta <- sqrt(predvarJbeta) 
newdat$SE_Horn <- sqrt(predvarHorn) 

# prep the plots
varplots <- vector('list', nrow(vars))
for(j in 1:length(varplots)){
  subs <- newdat$var == vars$vars[j] # which rows of newdat
  xvar <- vars$vars[j]
  title <- vars$vars[j]
  if(vars$vars[j] %in% c('human_bowler')){
    subs <- newdat$var == vars$vars[j] & newdat$REALM2 == vars$REALM2[j]
    title <- paste0('human:', vars$REALM2[j])
  } 

  se <- 1
  thisplot <- ggplot(newdat[subs, ], 
                     aes_string(x = xvar, y = 'predsJtu')) +
    geom_line() +
    geom_ribbon(aes(ymin = predsJtu - se*SE_Jtu, ymax = predsJtu + se*SE_Jtu), alpha = 0.5, fill = "grey") +
    geom_line(aes(y = predsJbeta), color = 'red') +
    geom_ribbon(aes(ymin = predsJbeta - se*SE_Jbeta, ymax = predsJbeta + se*SE_Jbeta), alpha = 0.5, fill = "red") +
    geom_line(aes(y = predsHorn), color = 'blue') +
    geom_ribbon(aes(ymin = predsHorn - se*SE_Horn, ymax = predsHorn + se*SE_Horn), alpha = 0.5, fill = "blue") +
    #coord_cartesian(ylim = c(0, 0.4)) +
    theme(plot.margin = unit(c(0.5,0,0.5,0), 'cm')) +
    labs(title = title) 
  varplots[[j]] <- thisplot
  if(vars$log[j] & !vars$discrete[j]){
    varplots[[j]] <- thisplot + scale_x_log10()
  }
}

grid.arrange(grobs = varplots, ncol = 3)

# write out the interactions
write.csv(newdat, file = 'output/maineffects.csv')

```
##### Using sjPlot
Doesn't work now
```{r sjPlot for full model, eval = FALSE}
require(sjPlot)
# p1 <- sjPlot::plot_model(modTfullJturem0, type = 'est', terms = c('temptrend_abs.sc'))

```

#### Plot interactions (Jaccard turnover)
```{r interaction plots modTfullJturem0, fig.height = 13, fig.width = 9}

# set up the interactions to plot
# if variable is logged before scaling (see 'center and scale' above), then need to mark it here and express the limits on a log10 scale (even though log transforming is log)
ints <- data.frame(vars = c('tsign', 'tempave_metab', 'seas', 'microclim', 'mass', 'speed', 
                            'consumerfrac', 'nspp', 'thermal_bias', 'npp', 'veg', 'duration', 
                            'human_bowler', 'human_bowler'),
           min =      c(1,  0,   0.1, -2,  0,   0,   0,   0.3, -10, 1.9, 0,   0.5, 0,   0), 
           max =      c(2,  30,  16,  0.8, 8,   2,   1,   2.6, 10,  3.7, 0.3, 2,   1,   1),
           log =      c(F,  F,   F,   T,   T,   T,   F,   T,   F,   T,   T,   T,   T,   T),
           len =      c(2,  100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100),
           discrete = c(T,  F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F,   F),
           plus =     c(0,  0,   0,   0,   0,   1,   0,   0,   0,   0,   1,   0,   1,   1), # what to add before log-scaling
           REALM = c(rep('Terrestrial', 12), 'Terrestrial', 'Marine'),
           REALM2 = c(rep('TerrFresh', 13), 'Marine'),
           stringsAsFactors = FALSE)
baseall <- trends[, .(type = 'all', tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                      seas.sc = mean(seas.sc, na.rm=TRUE), 
                      microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                      speed.sc = mean(speed.sc, na.rm=TRUE), 
                      mass.sc = mean(mass.sc, na.rm=TRUE), 
                      nspp.sc = 0, 
                      thermal_bias.sc = mean(thermal_bias.sc, na.rm=TRUE), 
                      npp.sc = mean(npp.sc, na.rm=TRUE), 
                      human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                      veg.sc = mean(veg.sc, na.rm=TRUE), 
                      consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
baseterr <- trends[REALM == 'Terrestrial', 
                   .(type = 'Terrestrial', 
                     tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                     seas.sc = mean(seas.sc, na.rm=TRUE), 
                     microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                     speed.sc = mean(speed.sc, na.rm=TRUE), 
                     mass.sc = mean(mass.sc, na.rm=TRUE), 
                     nspp.sc = 0, 
                     thermal_bias.sc = 0, 
                     npp.sc = mean(npp.sc, na.rm=TRUE), 
                     human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                     veg.sc = mean(veg.sc, na.rm=TRUE), 
                     consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
basemar <- trends[REALM == 'Marine', 
                  .(type = 'Marine',
                    tempave_metab.sc = mean(tempave_metab.sc, na.rm=TRUE), 
                    seas.sc = mean(seas.sc, na.rm=TRUE), 
                    microclim.sc = mean(microclim.sc, na.rm=TRUE), 
                    speed.sc = mean(speed.sc, na.rm=TRUE), 
                    mass.sc = mean(mass.sc, na.rm=TRUE), 
                    nspp.sc = 0, 
                    thermal_bias.sc = 0, 
                    npp.sc = mean(npp.sc, na.rm=TRUE), 
                    human_bowler.sc = mean(human_bowler.sc, na.rm=TRUE), 
                    veg.sc = mean(veg.sc, na.rm=TRUE), 
                    consumerfrac.sc = mean(consumerfrac.sc, na.rm=TRUE))]
basetab <- rbind(baseall, baseterr, basemar)
basetab[, ':='(duration.sc = 0, nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_514668')]

# make the data frames for each interaction to plot                
for(j in 1:nrow(ints)){
  # set up a grid of temperature trends and the interacting variable
  if(ints$log[j]) intvars <- list(temptrend = seq(-1.5, 1.5, length.out = 100), 
                                  new = 10^seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                   var = ints$vars[j])
  if(!ints$log[j]) intvars <- list(temptrend = seq(-1.5, 1.5, length.out = 100), 
                                   new = seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                   var = ints$vars[j])
  names(intvars) <- c('temptrend', ints$vars[j], 'var')
  thisdat <- expand.grid(intvars)
  
  # scale the interacting variable
  cent <- attr(trends[[paste0(ints$var[j], '.sc')]], 'scaled:center')
  scl <- attr(trends[[paste0(ints$var[j], '.sc')]], 'scaled:scale')
  if(!is.null(cent) & !is.null(scl)){
    if(ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (log(thisdat[[ints$vars[j]]] + ints$plus[j]) - cent)/scl
    if(!ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (thisdat[[ints$var[j]]] - cent)/scl
  }

  # merge with the rest of the columns
  # use realm-specific averages for human impacts
  if(ints$vars[j] != 'tsign') colnamestouse <- setdiff(colnames(basetab), paste0(ints$var[j], '.sc'))
  if(ints$vars[j] == 'tsign') colnamestouse <- setdiff(colnames(basetab), ints$var[j])
  if(ints$vars[j] != 'human_bowler'){
    thisdat <- cbind(thisdat, basetab[type == 'all', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Terrestrial'){
    thisdat <- cbind(thisdat, basetab[type == 'Terrestrial', ..colnamestouse])
  }
  if(ints$vars[j] == 'human_bowler' & ints$REALM[j] == 'Marine'){
    thisdat <- cbind(thisdat, basetab[type == 'Marine', ..colnamestouse])
  }
  
  # add realm
  thisdat$REALM <- ints$REALM[j]
  thisdat$REALM2 <- ints$REALM2[j]
  
  # merge with the previous iterations
  if(j == 1) newdat <- thisdat
  if(j > 1){
    colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
    for(toadd in colstoadd){
      newdat[[toadd]] <- NA
    }
    
    colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
    for(toadd in colstoadd2){
      thisdat[[toadd]] <- NA
    }
    
    newdat <- rbind(newdat, thisdat)
  } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work)
newdat <- rbind(newdat[1:6, ], newdat)
newdat$REALM[1:6] <- c('Marine', 'Marine', 'Freshwater', 'Freshwater', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:6] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh', 'TerrFresh', 'TerrFresh')
newdat$temptrend[1:6] <- c(-1, 1, -1, 1, -1, 1)

# trim to at least some temperature change (so that tsign is -1 or 1)
newdat <- newdat[newdat$temptrend != 0,]

# scale the temperature vars
newdat$temptrend.sc <- newdat$temptrend/attr(trends$temptrend.sc, 'scaled:scale') 
newdat$temptrend_abs <- abs(newdat$temptrend)
newdat$temptrend_abs.sc <- (newdat$temptrend_abs)/attr(trends$temptrend_abs.sc, 'scaled:scale')
newdat$tsign <- factor(sign(newdat$temptrend))

# make predictions
newdat$preds <- predict(object = modTfullJturem0, newdata = newdat, level = 0)

#remove the extra rows
newdat <- newdat[5:nrow(newdat), ]

# prep the plots
intplots <- vector('list', nrow(ints))
for(j in 1:length(intplots)){
  subs <- newdat$var == ints$vars[j] & newdat$temptrend > 0 # select warming side
  xvar <- 'temptrend_abs'
  title <- ints$vars[j]
  if(ints$vars[j] %in% c('tsign')){
    subs <- newdat$var == ints$vars[j]
  } 
  if(ints$vars[j] %in% c('thermal_bias')){
    subs <- newdat$var == ints$vars[j]
    xvar <- 'temptrend'
  } 
  if(ints$vars[j] %in% c('human_bowler')){
    subs <- newdat$var == ints$vars[j] & newdat$temptrend > 0 & newdat$REALM2 == ints$REALM2[j]
    title <- paste0('human:', ints$REALM2[j])
  } 

  thisplot <- ggplot(newdat[subs, ], 
                     aes_string(x = xvar, y = 'preds', 
                                group = ints$vars[j], 
                                color = ints$vars[j])) +
    geom_line() +
    coord_cartesian(ylim = c(-0.2, 0.4)) +
    theme(plot.margin = unit(c(0.5,0,0.5,0), 'cm')) +
    labs(title = title)
  if(ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'log')
  }
  if(!ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'identity')
  }
  if(ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_brewer(palette = "Dark2")
  }
}

#grid.arrange(grobs = intplots, '+', theme(plot.margin = unit(c(0,0,0,0), 'cm'))), ncol=2)
#do.call('grid.arrange', c(intplots, ncol = 2))
grid.arrange(grobs = intplots, ncol = 3)

# write out the interactions
write.csv(newdat, file = 'temp/interactions.csv')

```

#### Plot residuals against each predictor (Jaccard turnover)
```{r resids modTfull1, fig.height = 10, fig.width=10}
resids <- resid(modTfullJturem0)
preds <- getData(modTfullJturem0)
col = '#00000033'
cex = 0.5
par(mfrow = c(5,4))
boxplot(resids ~ preds$REALM, cex = cex, col = col)
plot(preds$temptrend_abs.sc, resids, cex = cex, col = col)
plot(preds$tsign, resids, cex = cex, col = col)
plot(preds$tempave.sc, resids, cex = cex, col = col)
plot(preds$tempave_metab.sc, resids, cex = cex, col = col)
plot(preds$seas.sc, resids, cex = cex, col = col)
plot(preds$microclim.sc, resids, cex = cex, col = col)
plot(preds$mass.sc, resids, cex = cex, col = col)
plot(preds$speed.sc, resids, cex = cex, col = col)
plot(preds$lifespan.sc, resids, cex = cex, col = col)
plot(preds$consumerfrac.sc, resids, cex = cex, col = col)
plot(preds$endothermfrac.sc, resids, cex = cex, col = col)
plot(preds$nspp.sc, resids, cex = cex, col = col)
plot(preds$thermal_bias.sc, resids, cex = cex, col = col)
plot(preds$npp.sc, resids, cex = cex, col = col)
plot(preds$veg.sc, resids, cex = cex, col = col)
plot(preds$human_bowler.sc, resids, cex = cex, col = col)
```

### Remove each term from the full model
```{r term deletion from modTfull}
AICnas <- function(x){
  if(class(x) == 'NULL'){
    return(NA)
  } else {
    return(AIC(x))
  }
}

# set up the response variables to run
torun = data.frame(var = c('Jtutrendrem0', 'Jbetatrendrem0', 'Horntrendrem0', 
                           'Jtutrendz', 'Jbetatrendz', 'Horntrendz',
                           'Jtuexp', 'Jbetaexp', 'Hornexp',
                           'Jtumm', 'Jbetamm', 'Hornmm'), 
                   modvar = c('Jtutrendrem0', 'Jbetatrendrem0', 'Horntrendrem0', 
                           'Jtutrendz', 'Jbetatrendz', 'Horntrendz',
                           'log(Jtuexp)', 'log(Jbetaexp)', 'log(Hornexp)',
                           'log(Jtumm+1)', 'log(Jbetamm+1)', 'log(Hornmm+1)'),
                   run = TRUE, aics = NA)

# check if any already exist. If so, don't re-run that model.
if(file.exists('output/aics_from_full.csv')){
  aicsfromfull <- read.csv('output/aics_from_full.csv')
  
  for(i in 1:nrow(torun)){
    nm <- paste0('dAIC_', torun$var[i])
    if(nm %in% colnames(aicsfromfull)){ # check for column name
      if(!is.na(aicsfromfull[[nm]][1])){ # check that full model is not NA
        torun$run[i] <- FALSE
      }
    }
  }
}

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

terms <- c('temptrend_abs.sc*REALM', 
           'temptrend_abs.sc*tsign',
           'temptrend_abs.sc*tempave_metab.sc',
           'temptrend_abs.sc*seas.sc',
           'temptrend_abs.sc*microclim.sc',
           'temptrend_abs.sc*mass.sc',
           'temptrend_abs.sc*speed.sc', 
           'temptrend_abs.sc*consumerfrac.sc',
           'temptrend_abs.sc*nspp.sc',
           'temptrend_abs.sc*thermal_bias.sc:tsign',
           'temptrend_abs.sc*npp.sc',
           'temptrend_abs.sc*veg.sc',
           'temptrend_abs.sc*duration.sc',
           'temptrend_abs.sc*human_bowler.sc:REALM2')

completerows <- trends[, complete.cases(REALM, tempave_metab.sc, 
                                seas.sc, microclim.sc, 
                                temptrend_abs.sc, mass.sc, speed.sc, 
                                consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                                veg.sc, duration.sc, human_bowler.sc)]

if(any(torun$run)){
  for(k in which(torun$run)){

    i <- !is.na(trends[[as.character(torun$var[k])]]) & completerows
    
    modTdrops <- vector('list', length(terms)+2)
    names(modTdrops) <- c('full', '-temptrend_abs.sc', paste0('-', terms))
    
    # fit full model with ML for model comparison
    modTdrops[[1]] <- lme(formula(paste0(torun$modvar[k], ' ~ ', paste(terms, collapse = ' + '))),
                          random = randef, weights = varef, data = trends[i,], method = 'ML',
                          control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
    
    # w/out temptrend
    modTdrops[[2]] <- lme(formula(paste0(torun$modvar[k], ' ~ ', paste(gsub('temptrend_abs.sc\\*', '', terms), 
                                                                       collapse = ' + '))),
                          random = list(STUDY_ID = ~ 1, rarefyID = ~1), weights = varef, data = trends[i,], method = 'ML',
                          control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
    
    for(j in 1:length(terms)){
      print(j)
      tryCatch({
        modTdrops[[j+2]] <- lme(formula(paste0(torun$modvar[k], ' ~ ', paste(terms[-j], collapse = ' + '))),
                                random = randef, weights = varef, data = trends[i,], method = 'ML')
        
      }, error = function(e){
        print('going to optim (Jtu)')
        tryCatch({
          modTdrops[[j+2]] <- lme(formula(paste0(torun$modvar[k], ' ~ ', paste(terms[-j], collapse = ' + '))),
                                  random = randef, weights = varef, data = trends[i,], method = 'ML',
                                  control = lmeControl(opt = 'optim'))
          
        }, error = function(e){
          print('going to more iters (Jtu)') 
          tryCatch({
            modTdrops[[j+2]] <- lme(formula(paste0(torun$modvar[k], ' ~ ', paste(terms[-j], collapse = ' + '))),
                                    random = randef, weights = varef, data = trends[i,], method = 'ML',
                                    control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
            
          }, error= function(e){
            print('giving up on this one')
            modTdrops[[j+2]] <- NA
          })
        })
      })
    }
    
    torun$aics[k] <- list(sapply(modTdrops, AICnas))
  }
}



# if there was anything new
if(any(torun$run)){
  if(!exists('aicsfromfull')){
    aicsfromfull <- data.frame(mod = names(torun$aics[which(torun$run)[1]][[1]]))
  }
  
  # subtract full from each model AIC. Negative means term removal is supported. Positive means full is the better model.
  for(j in which(torun$run)){
    nm <- paste0('dAIC_', torun$var[j])
    aics <- torun$aics[j][[1]]
    aicsfromfull[[nm]] <- aics - aics[1]
  }

  
  # write out
  write.csv(aicsfromfull, file = 'output/aics_from_full.csv', row.names = FALSE)
}

aicsfromfull
```

#### Plot deltaAICs for all models
```{r plot dAICs, message = FALSE, warning = FALSE}
# transform for a plot
aicsfromfulllong <- reshape(aicsfromfull, direction = 'long',
                            varying = c('dAIC_Jtutrendrem0', 'dAIC_Jbetatrendrem0', 'dAIC_Horntrendrem0',
                                        'dAIC_Jtutrendz', 'dAIC_Jbetatrendz', 'dAIC_Horntrendz',
                                        'dAIC_Jtuexp', 'dAIC_Jbetaexp', 'dAIC_Hornexp',
                                        'dAIC_Jtumm', 'dAIC_Jbetamm', 'dAIC_Hornmm'),
                            v.names = 'dAIC',
                            idvar = 'mod',
                            timevar = 'type',
                            times = c('Jtutrendrem0', 'Jbetatrendrem0', 'Horntrendrem0',
                                      'Jtutrendz', 'Jbetatrendz', 'Horntrendz',
                                      'Jtuexp', 'Jbetaexp', 'Hornexp',
                                      'Jtumm', 'Jbetamm', 'Hornmm'))
aicsfromfulllong$response <- gsub('trendrem0|trendz|exp|mm', '', aicsfromfulllong$type)
aicsfromfulllong$fit <- gsub('Jtu|Jbeta|Horn', '', aicsfromfulllong$type)

# plot
ggplot(aicsfromfulllong, aes(y=dAIC, x = mod, shape = response, color = fit, group = fit)) +
  geom_point(position = position_dodge(width = 0.5)) +
  scale_y_continuous(trans = signedsqrttrans) +
  coord_flip()
```

Light grey is for Jaccard turnover, dark grey is for Jaccard total, black is for Morisita-Horn.
Clear that removing temperature trend makes the model quite a bit worse and has the biggest effect.



## Simplify the full models
This takes a couple days on a laptop to run if temp/ files not available.
```{r simplify the full models}
randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

terms <- ' ~ temptrend_abs.sc*REALM + temptrend_abs.sc*tsign + temptrend_abs.sc*tempave_metab.sc + temptrend_abs.sc*seas.sc + temptrend_abs.sc*microclim.sc + temptrend_abs.sc*mass.sc + temptrend_abs.sc*speed.sc + temptrend_abs.sc*consumerfrac.sc + temptrend_abs.sc*nspp.sc + temptrend_abs.sc*thermal_bias.sc:tsign + temptrend_abs.sc*npp.sc + temptrend_abs.sc*veg.sc + temptrend_abs.sc*duration.sc + temptrend_abs.sc*human_bowler.sc:REALM2'

completerows <- trends[, complete.cases(REALM, tempave_metab.sc, 
                                        seas.sc, microclim.sc, 
                                        temptrend_abs.sc, mass.sc, speed.sc, 
                                        consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                                        veg.sc, duration.sc, human_bowler.sc)]

# rem0
if(file.exists('temp/modTsimpJturem0.rds')){
  modTsimpJturem0 <- readRDS('temp/modTsimpJturem0.rds')
} else {
  i <- !is.na(trends$Jtutrendrem0) & completerows
  
  modTfullJturem0ML <- lme(formula(paste0('Jtutrendrem0', terms)),
                           random = randef, weights = varef, data = trends[i,], method = 'ML',
                           control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  modTsimpJturem0 <- stepAIC(modTfullJturem0ML, direction = 'backward')
  saveRDS(modTsimpJturem0, file = 'temp/modTsimpJturem0.rds')
}

if(file.exists('temp/modTsimpJbetarem0.rds')){
  modTsimpJbetarem0 <- readRDS('temp/modTsimpJbetarem0.rds')
} else {
  i <- !is.na(trends$Jbetatrendrem0) & completerows
  modTfullJbetarem0ML <- lme(formula(paste0('Jbetatrendrem0', terms)),
                       random = randef, weights = varef, data = trends[i,], method = 'ML',
                       control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  modTsimpJbetarem0 <- stepAIC(modTfullJbetarem0ML, direction = 'backward')
  saveRDS(modTsimpJbetarem0, file = 'temp/modTsimpJbetarem0.rds')
}

if(file.exists('temp/modTsimpHornrem0.rds')){
  modTsimpHornrem0 <- readRDS('temp/modTsimpHornrem0.rds')
} else {
  i <- !is.na(trends$Horntrendrem0) & completerows
  modTfullHornrem0ML <- lme(formula(paste0('Horntrendrem0', terms)),
                            random = randef, weights = varef, data = trends[i,], method = 'ML',
                            control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  modTsimpHornrem0 <- stepAIC(modTfullHornrem0ML, direction = 'backward')
  saveRDS(modTsimpHornrem0, file = 'temp/modTsimpHornrem0.rds')
}

# z
if(file.exists('temp/modTsimpJtuz.rds')){
  modTsimpJtuz <- readRDS('temp/modTsimpJtuz.rds')
} else {
  i <- !is.na(trends$Jtutrendz) & completerows
  
  modTfullJtuzML <- lme(formula(paste0('Jtutrendz', terms)),
                           random = randef, weights = varef, data = trends[i,], method = 'ML',
                           control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  modTsimpJtuz <- stepAIC(modTfullJtuzML, direction = 'backward')
  saveRDS(modTsimpJtuz, file = 'temp/modTsimpJtuz.rds')
}

if(file.exists('temp/modTsimpJbetaz.rds')){
  modTsimpJbetaz <- readRDS('temp/modTsimpJbetaz.rds')
} else {
  i <- !is.na(trends$Jbetatrendz) & completerows
  modTfullJbetazML <- lme(formula(paste0('Jbetatrendz', terms)),
                       random = randef, weights = varef, data = trends[i,], method = 'ML',
                       control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  modTsimpJbetaz <- stepAIC(modTfullJbetazML, direction = 'backward')
  saveRDS(modTsimpJbetaz, file = 'temp/modTsimpJbetaz.rds')
}

if(file.exists('temp/modTsimpHornz.rds')){
  modTsimpHornz <- readRDS('temp/modTsimpHornz.rds')
} else {
  i <- !is.na(trends$Horntrendz) & completerows
  modTfullHornzML <- lme(formula(paste0('Horntrendz', terms)),
                            random = randef, weights = varef, data = trends[i,], method = 'ML',
                            control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500, opt = 'optim'))
  modTsimpHornz <- stepAIC(modTfullHornzML, direction = 'backward')
  saveRDS(modTsimpHornz, file = 'temp/modTsimpHornz.rds')
}

summary(modTsimpJturem0)
summary(modTsimpJbetarem0)
summary(modTsimpHornrem0)


```


## Make realm-specific models
```{r LME Horn models by realm, fig.width=10, fig.height=8}
i1 <- trends[, REALM == 'Terrestrial' & complete.cases(Horntrendrem0, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend_abs.sc, mass.sc, speed.sc, 
                             consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                             veg.sc, duration.sc, human_bowler.sc)]
i2 <- trends[, REALM == 'Freshwater' & complete.cases(Horntrendrem0, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend_abs.sc, mass.sc, speed.sc, 
                             nspp.sc, thermal_bias.sc, npp.sc, 
                             veg.sc, duration.sc, human_bowler.sc)] # no consumerfrac
i3 <- trends[, REALM == 'Marine' & complete.cases(Horntrendrem0, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend_abs.sc, mass.sc, speed.sc, 
                             consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, 
                             duration.sc, human_bowler.sc)] # no veg

print(paste('Terrestrial', sum(i1)))
print(paste('Freshwater', sum(i2)))
print(paste('Marine', sum(i3)))

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

# land
if(file.exists('temp/modTfullHornTerr.rds')){
  modTfullHornTerr <- readRDS('temp/modTfullHornTerr.rds')
} else {
  modTfullHornTerr <- lme(Horntrendrem0 ~ 
                         temptrend_abs.sc*tsign +
                         temptrend_abs.sc*tempave_metab.sc + 
                         temptrend_abs.sc*seas.sc + 
                         temptrend_abs.sc*microclim.sc + 
                         temptrend_abs.sc*mass.sc + 
                         temptrend_abs.sc*speed.sc + 
                         temptrend_abs.sc*consumerfrac.sc +
                         temptrend_abs.sc*nspp.sc +
                         temptrend_abs.sc*thermal_bias.sc:tsign +
                         temptrend_abs.sc*npp.sc +
                         temptrend_abs.sc*veg.sc +
                         temptrend_abs.sc*duration.sc +
                         temptrend_abs.sc*human_bowler.sc,
                       random = randef, weights = varef, data = trends[i1,], method = 'REML',
                       control = lmeControl(maxIter = 100, msMaxIter = 100, niterEM = 50, msMaxEval = 500))
  saveRDS(modTfullHornTerr, file = 'temp/modTfullHornTerr.rds')
}

# freshwater
if(file.exists('temp/modTfullHornFresh.rds')){
  modTfullHornFresh <- readRDS('temp/modTfullHornFresh.rds')
} else {
  modTfullHornFresh <- lme(Horntrendrem0 ~ 
                         temptrend_abs.sc*tsign +
                         temptrend_abs.sc*tempave_metab.sc + 
                         temptrend_abs.sc*seas.sc + 
                         temptrend_abs.sc*microclim.sc + 
                         temptrend_abs.sc*mass.sc + 
                         temptrend_abs.sc*speed.sc + 
                         temptrend_abs.sc*nspp.sc +
                         temptrend_abs.sc*thermal_bias.sc:tsign +
                         temptrend_abs.sc*npp.sc +
                         temptrend_abs.sc*veg.sc +
                         temptrend_abs.sc*duration.sc +
                         temptrend_abs.sc*human_bowler.sc,
                       random = randef, weights = varef, data = trends[i2,], method = 'REML')
  saveRDS(modTfullHornFresh, file = 'temp/modTfullHornFresh.rds')
}

# marine
if(file.exists('temp/modTfullHornMar.rds')){
  modTfullHornMar <- readRDS('temp/modTfullHornMar.rds')
} else {
  modTfullHornMar <- lme(Horntrendrem0 ~ 
                         temptrend_abs.sc*tsign +
                         temptrend_abs.sc*tempave_metab.sc + 
                         temptrend_abs.sc*seas.sc + 
                         temptrend_abs.sc*microclim.sc + 
                         temptrend_abs.sc*mass.sc + 
                         temptrend_abs.sc*speed.sc + 
                         temptrend_abs.sc*consumerfrac.sc +
                         temptrend_abs.sc*nspp.sc +
                         temptrend_abs.sc*thermal_bias.sc:tsign +
                         temptrend_abs.sc*npp.sc +
                         temptrend_abs.sc*duration.sc +
                         temptrend_abs.sc*human_bowler.sc,
                       random = randef, weights = varef, data = trends[i3,], method = 'REML')
  saveRDS(modTfullHornMar, file = 'temp/modTfullHornMar.rds')
}

summary(modTfullHornTerr)
summary(modTfullHornFresh)
summary(modTfullHornMar)
```
### Plot the realm-specific coefficients
Also uses the full models across all realms
```{r plot realm mods, fig.height=12, fig.width=9}

coefs1 <- summary(modTfullHornrem0)$tTable
coefs2 <- summary(modTfullHornTerr)$tTable
coefs3 <- summary(modTfullHornFresh)$tTable
coefs4 <- summary(modTfullHornMar)$tTable

varstoplot <- unique(c(rownames(coefs1), rownames(coefs2), rownames(coefs3), rownames(coefs4)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs1) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
rows1_4 <- which(rownames(coefs4) %in% varstoplot)
xlims <- range(c(coefs1[rows1_1,1] - coefs1[rows1_1,2], coefs1[rows1_1,1] + coefs1[rows1_1,2], 
                 coefs2[rows1_2,1] - coefs2[rows1_2,2], coefs2[rows1_2,1] + coefs2[rows1_2,2], 
                 coefs3[rows1_3,1] - coefs3[rows1_3,2], coefs3[rows1_3,1] + coefs3[rows1_3,2],
                 coefs4[rows1_4,1] - coefs4[rows1_4,2], coefs4[rows1_4,1] + coefs4[rows1_4,2]))


cols <- brewer.pal(4, 'Dark2') # for full, terr, fresh, mar
pchs <- c(1, 16, 16, 16)
offs <- c(0.1, 0, -0.1, -0.2) # offset vertically for each model


par(las = 1, mai = c(0.5, 4, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(1,length(varstoplot)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs1)){
    x = coefs1[rownames(coefs1) == varstoplot[i], 1]
    se = coefs1[rownames(coefs1) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
  if(varstoplot[i] %in% rownames(coefs4)){
    x = coefs4[rownames(coefs4) == varstoplot[i], 1]
    se = coefs4[rownames(coefs4) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[4], pch = pchs[4], col = cols[4])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[4], length(varstoplot) + 1 - i + offs[4]), col = cols[4])
  }
}
legend('bottomleft', col = cols, pch = pchs, lwd = 1, legend = c('All', 'Terestrial', 'Freshwater', 'Marine'))
```

[End text in hopes this helps the last figure show up when knitted]