---
title: 'Temperature change and the community response to temperature change across realms'
subtitle: '(using mixed effects models)'
output: 
    github_document: default

---
# Prep
First run turnover_vs_temperature_GLMM.R to fit the models.
<details>
<summary>Click to expand code</summary>

```{r setup}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(glmmTMB) # for ME models
library(bbmle) # for AICtab
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together
library(RColorBrewer)
library(scales) # for defining custom scales in ggplot

options(width=500) # turn off most text wrapping

signedsqrt = function(x) sign(x)*sqrt(abs(x))
signedsq = function(x) sign(x) * x^2
signedsqrttrans <- trans_new(name = 'signedsqrt', transform = signedsqrt, inverse = signedsq)


# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
```


```{r load, trim, set up data}
# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends <- fread('output/turnover_w_covariates.csv.gz')

# trim to only data with some temperature change
# important since sign of temperature change is a variable
trends[tempchange == 0, .N] # number to remove
trends <- trends[tempchange != 0, ] # also removes any NA values

# set realm order
trends[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# set up sign of temperature change
trends[, tsign := factor(sign(tempchange))]

# realm that combines Terrestrial and Freshwater, for interacting with human impact
trends[, REALM2 := REALM]
levels(trends$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")

# group Marine invertebrates/plants in with All
trends[, taxa_mod2 := taxa_mod]
trends[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']

# calculate duration
trends[, duration := year2 - year1]

#add a comparison id
trends[, compID := paste0(rarefyID, '_', year1, '_', year2)]

# transformation for 2 categories. Eq. 1 in Douma & Weedon 2019
transform01 <- function(x) (x * (length(x) - 1) + 0.5) / (length(x))

trends[, Jtu.sc := transform01(Jtu)]
trends[, Jbeta.sc := transform01(Jbeta)]
trends[, Horn.sc := transform01(Horn)]

# Log-transform some variables, then center and scale. 
trends[, tempave.sc := scale(tempave)]
trends[, tempave_metab.sc := scale(tempave_metab)]
trends[, seas.sc := scale(seas)]
trends[, microclim.sc := scale(log(microclim))]
trends[, tempchange.sc := scale(tempchange, center = FALSE)] # do not center
trends[, tempchange_abs.sc := scale(abs(tempchange), center = FALSE)] # do not center, so that 0 is still 0 temperature change
trends[, mass.sc := scale(log(mass_mean_weight))]
trends[, speed.sc := scale(log(speed_mean_weight+1))]
trends[, lifespan.sc := scale(log(lifespan_mean_weight))]
trends[, consumerfrac.sc := scale(consfrac)]
trends[, endothermfrac.sc := scale(endofrac)]
trends[, nspp.sc := scale(log(Nspp))]
trends[, thermal_bias.sc := scale(thermal_bias)]
trends[, npp.sc := scale(log(npp))]
trends[, veg.sc := scale(log(veg+1))]
trends[, duration.sc := scale(log(duration))]
trends[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
trends[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
trends[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]
```

</details>

# Summary stats
## Examine how many data points are available
### Just turnover and non-zero tempchange
```{r sample size all}
cat('Overall # dissimilarities: ', nrow(trends), '\n')
cat('# studies: ', trends[, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends[, length(unique(rarefyID))], '\n')
trends[!duplicated(rarefyID), table(REALM)]
trends[!duplicated(rarefyID), table(taxa_mod)]
trends[!duplicated(rarefyID), table(taxa_mod, REALM)]
```

### With all covariates
Use Bowler for human impact
```{r sample size Jtu covariates}
# the cases we can compare
apply(trends[, .(Jtu, REALM, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends[, complete.cases(Jtu, tempave_metab.sc, seas.sc, microclim.sc, tempchange.sc, mass.sc, speed.sc, consumerfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
cat('Overall # dissimilarities: ', sum(i), '\n')
cat('# studies: ', trends[i, length(unique(STUDY_ID))], '\n')
cat('# timeseries: ', trends[i, length(unique(rarefyID))], '\n')
trends[i & !duplicated(rarefyID), table(REALM)]
trends[i & !duplicated(rarefyID), table(taxa_mod)]
trends[i & !duplicated(rarefyID), table(taxa_mod, REALM)]
```

# Models
## Choose the variance structure
Try combinations of

- beta errors
- variance scaled to Nspp
- random intercept for taxa_mod2, STUDY_ID, rarefyID
- random slopes (abs temperature trend)
- random intercept for compID (for overdispersion)?

And choose the one with lowest AIC
```{r choose variance structure}

if(file.exists('temp/modRFgauss.rds')) modRFgauss <- readRDS('temp/modRFgauss.rds')
if(file.exists('temp/modRFbeta.rds')) modRFbeta <- readRDS('temp/modRFbeta.rds')
if(file.exists('temp/modRFrID.rds')) modRFrID <- readRDS('temp/modRFrID.rds')
if(file.exists('temp/modRFnestedRE.rds')) modRFnestedRE <- readRDS('temp/modRFnestedRE.rds')
if(file.exists('temp/modRFslopeRE.rds')) modRFslopeRE <- readRDS('temp/modRFslopeRE.rds')
if(file.exists('temp/modRFdisp.rds')) modRFdisp <- readRDS('temp/modRFdisp.rds')


AICtab(modRFgauss, modRFbeta, modRFrID, modRFnestedRE, modRFslopeRE, modRFdisp)
summary(modRFdisp)
```
Chooses beta errors, random slopes (tempchange_abs.sc) & intercepts for taxa_mod2, STUDY_ID, rarefyID, and variance scaled to nspp.
We haven't dealt with potential testing on the boundary issues here yet.


## Temperature-only models
### Load the models
```{r LME temperature only}
# all years
modonlyTchangeJtu <- readRDS('temp/modonlyTchangeJtu.rds')
modonlyTchangeJbeta <- readRDS('temp/modonlyTchangeJbeta.rds')
modonlyTchangeHorn <- readRDS('temp/modonlyTchangeHorn.rds')

# 1 year
modonlyTchange1yrJtu <- readRDS('temp/modonlyTchange1yrJtu.rds')
modonlyTchange1yrJbeta <- readRDS('temp/modonlyTchange1yrJbeta.rds')
modonlyTchange1yrHorn <- readRDS('temp/modonlyTchange1yrHorn.rds')

# 10 year
modonlyTchange10yrJtu <- readRDS('temp/modonlyTchange10yrJtu.rds')
modonlyTchange10yrJbeta <- readRDS('temp/modonlyTchange10yrJbeta.rds')
modonlyTchange10yrHorn <- readRDS('temp/modonlyTchange10yrHorn.rds')

```

### Summary (all years)
```{r summary onlyTchange}
summary(modonlyTchangeJtu)
summary(modonlyTchangeJbeta)
summary(modonlyTchangeHorn)
```

### Summary (1 year)
```{r summary onlyTchange1yr}
summary(modonlyTchange1yrJtu)
summary(modonlyTchange1yrJbeta)
summary(modonlyTchange1yrHorn)
```
### Summary (10 year)
```{r summary onlyTchange10yr}
summary(modonlyTchange10yrJtu)
summary(modonlyTchange10yrJbeta)
summary(modonlyTchange10yrHorn)
```
### Plot the temp-only coefficients
```{r modonlyT coefs}
# make table of coefficients
coefs1 <- as.data.frame(summary(modonlyTchangeJtu)$coefficients$cond)
coefs2 <- as.data.frame(summary(modonlyTchangeJbeta)$coefficients$cond)
coefs3 <- as.data.frame(summary(modonlyTchangeHorn)$coefficients$cond)
coefs4 <- as.data.frame(summary(modonlyTchange1yrJtu)$coefficients$cond)
coefs5 <- as.data.frame(summary(modonlyTchange1yrJbeta)$coefficients$cond)
coefs6 <- as.data.frame(summary(modonlyTchange1yrHorn)$coefficients$cond)
coefs7 <- as.data.frame(summary(modonlyTchange10yrJtu)$coefficients$cond)
coefs8 <- as.data.frame(summary(modonlyTchange10yrJbeta)$coefficients$cond)
coefs9 <- as.data.frame(summary(modonlyTchange10yrHorn)$coefficients$cond)
coefs1$response <- coefs4$response <- coefs7$response <- 'Jtu'
coefs2$response <- coefs5$response <- coefs8$response <- 'Jbeta'
coefs3$response <- coefs6$response <- coefs9$response <- 'Horn'
coefs1$fit <- coefs2$fit <- coefs3$fit <- 'all'
coefs4$fit <- coefs5$fit <- coefs6$fit <- '1yr'
coefs7$fit <- coefs8$fit <- coefs9$fit <- '10yr'
rows1 <- which(grepl('tempchange', rownames(coefs1))) # extract temperature effect
cols <- c('Estimate', 'Std. Error', 'response', 'fit')
allcoefs <- rbind(coefs1[rows1, cols], coefs2[rows1, cols], coefs3[rows1, cols],
                  coefs4[rows1, cols], coefs5[rows1, cols], coefs6[rows1, cols],
                  coefs7[rows1, cols], coefs8[rows1, cols], coefs9[rows1, cols])

allcoefs$lCI <- allcoefs$Estimate - 1.96 * allcoefs$`Std. Error` # lower confidence interval
allcoefs$uCI <- allcoefs$Estimate + 1.96 * allcoefs$`Std. Error`
allcoefs$y <- c(3, 2, 1) + rep(c(0, -0.1, -0.2), c(3, 3, 3)) # y-values
allcoefs$group <- paste0(allcoefs$response, allcoefs$fit)

pd <- position_dodge(width = 0.5)
ggplot(allcoefs, aes(x = response, y = Estimate, color = fit, group = group)) +
  geom_point(position = pd) + 
  geom_errorbar(aes(ymin=lCI, ymax=uCI), width=.1, position = pd) +
  labs(y = 'Dissimilarity per |°C change|') #+ 
  #coord_cartesian(ylim = c(-0.01, 0.05))
```
CIs are 1.96*SE

## Temperature&duration models
### Load the models
```{r temperature and duration}
# all years
modTDJtu <- readRDS('temp/modTDJtu.rds')
modTDJbeta <- readRDS('temp/modTDJbeta.rds')
modTDHorn <- readRDS('temp/modTDHorn.rds')
```

### Summary
```{r summary of TD mods}
summary(modTDJtu)
summary(modTDJbeta)
summary(modTDHorn)
```



### Plot the coefficients
```{r modTD coefs}

coefs1 <- summary(modTDJtu)$coefficients$cond
coefs2 <- summary(modTDJbeta)$coefficients$cond
coefs3 <- summary(modTDHorn)$coefficients$cond

varstoplot <- unique(c(rownames(coefs1), rownames(coefs2), rownames(coefs3)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs1) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
xlims <- range(c(coefs1[rows1_1,1] - 1.96*coefs1[rows1_1,2], coefs1[rows1_1,1] + 1.96*coefs1[rows1_1,2], 
                  coefs2[rows1_2,1] - 1.96*coefs2[rows1_2,2], coefs2[rows1_2,1] + 1.96*coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - 1.96*coefs3[rows1_3,2], coefs3[rows1_3,1] + 1.96*coefs3[rows1_3,2]))

cols <- brewer.pal(3, 'Dark2') # for Jtu, Jbeta and Horn models
pchs <- c(16, 16, 16)
offs <- c(0.1, 0, -0.1) # offset vertically for each model

par(las = 1, mai = c(0.5, 4, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(0.9,length(varstoplot)+0.1), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs1)){
    x = coefs1[rownames(coefs1) == varstoplot[i], 1] # the coef in col 1
    se = coefs1[rownames(coefs1) == varstoplot[i], 2] # the SE in col 2
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
}
legend('bottomright', col = cols, pch = 16, lwd = 1, legend = c('Jtu', 'Jbeta', 'Horn'), cex = 0.5)
```
abs(tempchange) is in degC
duration is in years

### Plot the response
```{r plot TD response}
newdat <- expand.grid(duration = c(1, 50), tempchange = seq(0.01, 5, length.out = 5), 
                      taxa_mod2 = NA, STUDY_ID = NA, rarefyID = NA,
                      Nspp = 60) # no REs, ave number of species
scl <- attr(trends$tempchange_abs.sc, 'scaled:scale') # scaling factor for tempchange
newdat$tempchange_abs.sc <- abs(newdat$tempchange)/scl
cent <- attr(trends$nspp.sc, 'scaled:center') 
scl <- attr(trends$nspp.sc, 'scaled:scale') 
newdat$nspp.sc <- (log(newdat$Nspp) - cent)/scl
Jtu <- predict(modTDJtu, newdata = newdat, se.fit = TRUE, re.form = NA, type = 'response')
newdat$durationfac <- as.factor(newdat$duration)
newdat$Jtu <- Jtu$fit
newdat$Jtu.se <- Jtu$se.fit

saveRDS(newdat, file = 'temp/modTDJtu_preds.rds')

ggplot(newdat, aes(tempchange, Jtu, color = durationfac, group = durationfac)) +
  geom_line() +
  geom_ribbon(aes(ymin=Jtu - 1.96*Jtu.se, ymax=Jtu + 1.96*Jtu.se), alpha = 0.1, linetype = 'blank')
```

## Temperature and duration by REALM
### Load the models
```{r LME temperature duration realm}

if(file.exists('temp/modTDrealmJtu.rds')) modTDJrealmtu <- readRDS('temp/modTDrealmJtu.rds')
if(file.exists('temp/modTDrealmJbeta.rds')) modTDrealmJbeta <- readRDS('temp/modTDrealmJbeta.rds')
if(file.exists('temp/modTDrealmHorn.rds')) modTDrealmHorn <- readRDS('temp/modTDrealmHorn.rds')

```

### Summary
```{r summary of TD mods}
summary(modTDrealmJtu)
summary(modTDrealmJbeta)
summary(modTDrealmHorn)
```


### Plot the temp coefficients from TD realm models
```{r modTD coefs}

coefs1 <- summary(modTDrealmJtu)$tTable
coefs2 <- summary(modTDrealmJbeta)$tTable
coefs3 <- summary(modTDrealmHorn)$tTable

varstoplot <- unique(c(rownames(coefs1), rownames(coefs2), rownames(coefs3)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs1) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
xlims <- range(c(coefs1[rows1_1,1] - 1.96*coefs1[rows1_1,2], coefs1[rows1_1,1] + 1.96*coefs1[rows1_1,2], 
                  coefs2[rows1_2,1] - 1.96*coefs2[rows1_2,2], coefs2[rows1_2,1] + 1.96*coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - 1.96*coefs3[rows1_3,2], coefs3[rows1_3,1] + 1.96*coefs3[rows1_3,2]))

cols <- brewer.pal(3, 'Dark2') # for Jtu, Jbeta and Horn models
pchs <- c(16, 16, 16)
offs <- c(0.1, 0, -0.1) # offset vertically for each model

par(las = 1, mai = c(0.5, 4, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(1,length(varstoplot)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs1)){
    x = coefs1[rownames(coefs1) == varstoplot[i], 1]
    se = coefs1[rownames(coefs1) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-1.96*se, x+1.96*se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
}
legend('topleft', col = cols, pch = 16, lwd = 1, legend = c('Jtu', 'Jbeta', 'Horn'), cex = 0.5)
```


