---
title: "simulations to test effects of random sampling and species richness interactions"
output:
    github_document: default
---
```{r setup, echo=TRUE}
# set parameters for simulations
require(vegan)
require(reshape2)
require(data.table)
lown <- 100 # number of species in low richness treatment
highn <- 1000 # number of species in high richness treatment
ngen <- 10 # number of sampling periods
frac <- 0.05 # fraction of species added and removed each sampling period
prob <- 0.5 # probability of seeing a species that is present
reps <- 1000 # number of times to repeat the sampling
set.seed(1001)
```

# simple scenario of species turnover
Add 5% new species and remove 5% of existing species each time step. A high and a low richness treatment

## 100% of species sampled
Results don't depend on species richness.
```{r}
# make the communities
low <- matrix(c(rep(1, lown), rep(0, 2*lown*ngen - lown)), nrow=2*lown, ncol = ngen) # rows are species. in first time step, half are present
high <- matrix(c(rep(1, highn), rep(0, 2*highn*ngen - highn)), nrow = 2*highn, ncol = ngen)
for(i in 2:10){
    low[,i] <- low[,i-1] * c(rep(0, (i-1)*frac*lown), rep(1, 2*lown - (i-1)*frac*lown)) + c(rep(0, lown + (i-2)*frac*lown), rep(1, frac*lown), rep(0, lown - (i-1)*frac*lown)) # remove frac existing species and add frac new species
    high[,i] <- high[,i-1] * c(rep(0, (i-1)*frac*highn), rep(1, 2*highn - (i-1)*frac*highn)) + c(rep(0, highn + (i-2)*frac*highn), rep(1, frac*highn), rep(0, highn - (i-1)*frac*highn))
}

# calculate dissimilarities
lowdist <- vegdist(t(low), method='jaccard', binary=TRUE) # vegan expects spp in columns
highdist <- vegdist(t(high), method='jaccard', binary=TRUE)

# examine distances
print('low species richness')
lowdist
print('high species richness')
highdist
```



## 50% of species sampled
This adds sampling noise
```{r}
for(i in 1:reps){
    if(i %% 50 == 0) cat(paste0('#', i, " "))
    # make the observed communities with only X% of species observed
    lowobs <- low * sample(c(rep(1, prob*2*lown*ngen), rep(0, (1-prob)*2*lown*ngen)), 2*lown*ngen, replace=FALSE) # turn X% of obs to absent, randomly
    highobs <- high * sample(c(rep(1, prob*2*highn*ngen), rep(0, (1-prob)*2*highn*ngen)), 2*highn*ngen, replace=FALSE)
    
    
    # calculate dissimilarities
    lowobsdist <- as.matrix(vegdist(t(lowobs), method='jaccard', binary=TRUE)) # vegan expects spp in columns
    highobsdist <- as.matrix(vegdist(t(highobs), method='jaccard', binary=TRUE))
    
    # examine distances (first example only)
    if(i == 1){
        print('low species richness with obs error')
        print(lowobsdist)
        print('high species richness')
        print(highobsdist)
    }
    
    # convert to long format (only one triangle)
    dimnames(lowobsdist) <- list(1:ngen, 1:ngen)
    lownms <- t(combn(colnames(lowobsdist), 2))
    thislowobslong <- data.frame(lownms, dist=lowobsdist[lownms])
    thislowobslong$rep <- i
    
    dimnames(highobsdist) <- list(1:ngen, 1:ngen)
    highnms <- t(combn(colnames(highobsdist), 2))
    thishighobslong <- data.frame(highnms, dist=highobsdist[highnms])
    thishighobslong$rep <- i
    
    if(i == 1){
        lowobslong <- thislowobslong
        highobslong <- thishighobslong
    } else {
        lowobslong <- rbind(lowobslong, thislowobslong)
        highobslong <- rbind(highobslong, thishighobslong)
    }
}

lowobslong$X1 <- as.numeric(lowobslong$X1)
lowobslong$X2 <- as.numeric(lowobslong$X2)

highobslong$X1 <- as.numeric(highobslong$X1)
highobslong$X2 <- as.numeric(highobslong$X2)

lowobslong <- data.table(lowobslong)
highobslong <- data.table(highobslong)
```


# calc turnover rates
same central tendencies in both high and low species richness scenarios
```{r}
lowslopes <- lowobslong[, as.list(coef(lm(dist ~ I(X2-X1)))), by=rep]
highslopes <- highobslong[, as.list(coef(lm(dist ~ I(X2-X1)))), by=rep]
setnames(lowslopes, c('rep', 'int', 'slope'))
setnames(highslopes, c('rep', 'int', 'slope'))

print('low richness slopes')
lowslopes[, summary(slope)]

print('high richness slopes')
highslopes[, summary(slope)]
```

# plot turnover rates as lines
More variance in turnover rates for the lower richness scenario. This doesn't plot correctly in the github document.
```{r}
par(mfrow=c(1,2))
plot(c(0,10), c(0,1), col='white', xlab = 'temporal distance', ylab='dissimilarity', main = 'low richness')
lowslopes[, abline(a = int, b = slope, col = '#00000033'), by = rep]

plot(c(0,10), c(0,1), col='white', xlab = 'temporal distance', ylab='dissimilarity', main = 'high richness')
highslopes[, abline(a = int, b = slope, col = '#00000033'), by = rep]
```

# plot turnover rates as densities
More variance in turnover rates for the lower richness scenario (black) compared to higher richness (red)
```{r}
lowslopes[, plot(density(slope), ylim =c(0,500), main = 'Turnover rates')]
highslopes[, lines(density(slope), col='red')]
```


