---
title: 'Drivers of variation in the community response to temperature change across realms'
subtitle: '(using mixed effects models)'
output: 
    github_document: default
    #html_document: default
    html_notebook: default
---

Collaborators: Shane Blowes, Jon Chase, Helmut Hillebrand, Michael Burrows, Amanda Bates, Uli Brose, Benoit Gauzens, Laura Antao
Assistance: Katherine Lew, Josef Hauser

# Introduction
- Climate change is driving a widespread reorganization of ecological communities around the world (Parmsesan & Yohe 2003, Poloczanska et al. 2013),
- but the impacts of climate change vary substantially from one location to another and among taxa (Molinos et al. 2016 NCC, Antao et al. 2020 NEE).
- Community reorganization is substantially more common than an aggregate loss or gain of species (Dornelas et al. 2014 Science, Blowes et al. 2019 Science, Hillebrand et al. 2017 J Appl Ecol)
- There are many hypotheses for why some communities are more sensitive to warming than others, including differences in
  - metabolic rates (Dillon et al. 2010 Nature), 
  - thermal physiology (Deutsch et al. 2008 PNAS, Pinsky et al. 2019 Nature), 
  - microclimate availability (Burrows et al. 2019 NCC, Suggitt et al. 2018 NCC),
  - species mobility (Poloczanska et al. 2013 NCC, Burrows et al. 2011 Science, Sunday et al. 2012 NCC)
  - or generation time (Beaugrand et al. 2009 DSR II, Poloczanska et al. 2013 NCC),
  - consumers vs. producers (Petchey et al. 1999 Nature)
  - community composition (Stuart-Smith et al. 2015 Nature, Beaugrand et a. 2015 NCC, Trisos et al. 2020 Nature), 
  - ecosystem productivity (Thomas et al. 2017 GCB, Brett 1971 Am Zoo),
  - exposure to human impacts (White & Kerr 2006 Ecography)
  - and among realms (Antao et al. 2020 NEE).
- Scaling up from organismal effects to whole ecological communities is complex, and yet these scales are critical for ecosystem functioning and human well-being. 
- There is a need for a comprehensive test to understand where warming is driving and is likely to drive the most dramatic community turnover

# Methods
- BioTime dataset, gridded to 96 km2 hexagons, summarized as temporal turnover (Blowes)
  - Temporal slope of Jaccard turnover compared to the first year
  - Same for Jaccard total
  - and Morisita-Horn turnover
- Explanatory variables considered for differences in rate of turnover:
  - Temperature trend over the time-frame of each time-series (CRU TS 4.03 on land and in freshwater, ERSST v5 in the ocean)
  - Seasonality as a metric of thermal sensitivity (Deutsch et al. 2008 PNAS). Standard deviation of monthly temperatures.
  - Microclimates calculated from WorldClim and BioOracle (Laura Antao)
  - Average temperature
  - Body mass, collated from databases and literature searches
  - Metabolic temperature, from average temperature if ectotherms (Dillon et al. 2010 Nature, Antao et al. 2020 Nat E&E)
  - Mobility calculated from body mass and taxonomic group classifications of mobility mode (fly, run, swim, crawl, sessile). Fly/run/swim followed the allometric relationship in Hirt et al. 2017 Nat E&E. Crawl set at 0.1 km/hr, sessile set to 0 km/hr. Then calculated averaged within each assemblage.
  - Generation time calculated from body mass and endotherm vs. ectotherm classifications, following McCoy & Gillooly 2008 ELE. Averaged across species within each assemblage. 
  - Consumer vs. producer classification by species
  - Endotherm vs. ectotherm classification by species
  - Species richness, calculated as the number of species in the assemblage
  - Net primary productivity (NPP) from the merged land/ocean product produced by the [Ocean Productivity](http://www.science.oregonstate.edu/ocean.productivity/) group at Oregon State using methods from Zhao et al. 2005 and Behrenfeld & Falkowski 1997. 
  - Human impact calculated from Bowler et al. 2020 (also try data from Venter et al. 2016 and Halpern et al. 2008)
  - Thermal bias calculated from Species Temperature Indices (Mike Burrows)
  - Vegetation cover index, calculated from %tree cover and %non-tree veg cover (latter counted as 1/2), from vegetation continuous fields (Ruben Remelgado)
- Differences in temporal turnover (response variable) modeled with a linear mixed effects model (nlme package, lme() function). See below for details.

```{r setup}
library(data.table) # for handling large datasets
library(ggplot2) # for some plotting
library(nlme) # for ME models
library(maps) # for map
library(gridExtra) # to combine ggplots together
library(grid) # to combine ggplots together

options(width=500) # turn off most text wrapping

# tell RStudio to use project root directory as the root for this notebook. Needed since we are storing code in a separate directory.
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
```

```{r load data}
# Turnover and covariates assembled by turnover_vs_temperature_prep.Rmd
trends <- fread('output/turnover_w_covariates.csv.gz')

# set realm order
trends[, REALM := factor(REALM, levels = c('Freshwater', 'Marine', 'Terrestrial'), ordered = FALSE)]

# set up sign of temperature change
trends[, tsign := factor(sign(temptrend))]

# realm that combined Terrestrial and Freshwater, for interacting with human impact
trends[, REALM2 := REALM]
levels(trends$REALM2) = list(TerrFresh = "Freshwater", TerrFresh = "Terrestrial", Marine = "Marine")

# group Marine invertebrates/plants in with All
trends[, taxa_mod2 := taxa_mod]
trends[taxa_mod == 'Marine invertebrates/plants', taxa_mod2 := 'All']
```


### Log-transform some variables, then center and scale. 
``` {r center and scale}
trends[, tempave.sc := scale(tempave)]
trends[, tempave_metab.sc := scale(tempave_metab)]
trends[, seas.sc := scale(seas)]
trends[, microclim.sc := scale(log(microclim))]
trends[, temptrend.sc := scale(temptrend, center = FALSE)]
trends[, temptrend_abs.sc := scale(abs(temptrend), center = FALSE)] # do not center, so that 0 is still 0 temperature change
trends[, mass.sc := scale(log(mass_mean_weight))]
trends[, speed.sc := scale(log(speed_mean_weight+1))]
trends[, lifespan.sc := scale(log(lifespan_mean_weight))]
trends[, consumerfrac.sc := scale(consfrac)]
trends[, endothermfrac.sc := scale(endofrac)]
trends[, nspp.sc := scale(log(Nspp))]
trends[, thermal_bias.sc := scale(thermal_bias)]
trends[, npp.sc := scale(log(npp))]
trends[, veg.sc := scale(log(veg+1))]
trends[, human_bowler.sc := scale(log(human_bowler+1)), by = REALM2] # separate scaling by realm
trends[REALM2 == 'TerrFresh', human_footprint.sc := scale(log(human_venter+1))]
trends[REALM2 == 'Marine', human_footprint.sc := scale(log(human_halpern))]
```

### Examine how many data points are available
Just turnover
```{r sample size all}
cat('Overall # time-series: ', nrow(trends), '\n')
cat('# studies: ', trends[, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[, sum(nyrBT)], '\n')
trends[, table(REALM)]
trends[, table(taxa_mod)]
trends[, table(taxa_mod, REALM)]
```

With all covariates (Bowler for human)
```{r sample size for Jaccard turnover}
# the cases we can compare
apply(trends[, .(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, temptrend.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)], MARGIN = 2, FUN = function(x) sum(!is.na(x)))
i <- trends[, complete.cases(Jtutrend, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, temptrend.sc, mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
cat('Overall # time-series: ', sum(i), '\n')
cat('# studies: ', trends[i, length(unique(STUDY_ID))], '\n')
cat('Data points: ', trends[i, sum(nyrBT)], '\n')
trends[i, table(REALM)]
trends[i, table(taxa_mod)]
trends[i, table(taxa_mod, REALM)]
```

### Choose the variance structure for mixed effects modles
Try combinations of

- variance scaled to a power of the number of years in the community time-series
- variance scaled to a power of the abs temperature trend
- random intercept for taxa_mod
- random intercept for STUDY_ID
- random slope (abs temperature trend) for taxa_mod
- random slope (abs temperature trend) for STUDY_ID
- random intercept for rarefyID (for overdispersion)

And choose the one with lowest AIC (not run: takes a long time)
```{r choose variance structure for Jacard turnover, eval = FALSE}
# fit models for variance structure
fixed <- formula(Jtutrend ~ REALM + tempave_metab.sc + seas.sc + microclim.sc + npp.sc + temptrend_abs.sc +
                     mass.sc + speed.sc + lifespan.sc + consumerfrac.sc + thermal_bias.sc)
i <- trends[, complete.cases(Jtutrend, REALM, tempave_metab.sc, seas.sc, microclim.sc, npp.sc, temptrend_abs.sc,
                             mass.sc, speed.sc, lifespan.sc, consumerfrac.sc, thermal_bias.sc)]
mods <- vector('list', 0)
mods[[1]] <- gls(fixed, data = trends[i,])
mods[[2]] <- gls(fixed, data = trends[i,], weights = varPower(-0.5, ~nyrBT))
mods[[3]] <- gls(fixed, data = trends[i,], weights = varPower(0.5, ~ abs(temptrend)))

mods[[4]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, control = lmeControl(opt = "optim"))
mods[[5]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, control = lmeControl(opt = "optim"))
mods[[6]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID, control = lmeControl(opt = "optim"))
mods[[7]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))
mods[[8]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, control = lmeControl(opt = "optim"))

mods[[9]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | taxa_mod)
mods[[10]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | STUDY_ID)
mods[[11]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc | taxa_mod2/STUDY_ID, control = lmeControl(opt = "optim"))
mods[[12]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)) # includes overdispersion. new formula so that random slope is only for study level (not enough data to extend to rarefyID).
mods[[13]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)) # 30+ min to fit

mods[[14]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[15]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, weights = varPower(-0.5, ~nyrBT))
mods[[16]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[17]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~nyrBT))
mods[[18]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, weights = varPower(-0.5, ~nyrBT))
mods[[19]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|STUDY_ID, weights = varPower(-0.5, ~nyrBT))
mods[[20]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))
mods[[21]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ 1), weights = varPower(-0.5, ~nyrBT))
mods[[22]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ 1, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT))
mods[[23]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc), weights = varPower(-0.5, ~nyrBT)) # singular precision warning with lmeControl(opt = 'optim') and convergence error without
mods[[24]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~nyrBT)) # singular precision warning with lmeControl(opt = 'optim') and convergence error without

mods[[25]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2, weights = varPower(-0.5, ~abs(temptrend)))
mods[[26]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[27]] <- lme(fixed, data = trends[i,], random = ~1|STUDY_ID/rarefyID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[28]] <- lme(fixed, data = trends[i,], random = ~1|taxa_mod2/STUDY_ID/rarefyID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[29]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)))
mods[[30]] <- lme(fixed, data = trends[i,], random = ~temptrend_abs.sc|taxa_mod2/STUDY_ID, weights = varPower(-0.5, ~abs(temptrend)), control = lmeControl(opt = "optim"))
mods[[31]] <- lme(fixed, data = trends[i,], random = list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~abs(temptrend)))
mods[[32]] <- lme(fixed, data = trends[i,], random = list(taxa_mod2 = ~ temptrend_abs.sc, STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1), weights = varPower(-0.5, ~abs(temptrend)), control = lmeControl(opt = "optim")) # singular precision warning

aics <- sapply(mods, AIC)
minaics <- aics - min(aics)
minaics
which.min(aics)
```
Chooses the random slopes (temptrend_abs) & intercepts for STUDY_ID, overdispersion, and variance scaled to number of years.
We haven't dealt with potential testing on the boundary issues here yet.

# Results
## Where do we have data?
```{r map}
world <- map_data('world')
ggplot(world, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = 'lightgray', color = 'white') +
    geom_point(data = trends, aes(rarefyID_x, rarefyID_y, group = REALM, color = REALM), size = 0.5, alpha = 0.4)  +
    scale_color_brewer(palette="Set1", name = 'Realm') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=16),
        axis.title=element_text(size=20)) +
  labs(x = 'Longitude (째)', y = 'Latitude (째)')
```

Mostly northern hemisphere, but spread all over. No so much in Africa or much of Asia.




Average rates of turnover
```{r rates of turnover}
trends[abs(temptrend) >= 0.5, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # turnover per year for locations changing temperature
trends[abs(temptrend) < 0.1, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # not changing temperature
trends[temptrend >= 0.5, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # warming
trends[temptrend <= -0.5, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # cooling

trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) < 35, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # tropics and sub-tropics
trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) >= 35 & abs(rarefyID_y) < 66.56339, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # temperate
trends[abs(temptrend) >= 0.5 & abs(rarefyID_y) >= 66.56339, .(ave = mean(Jtutrend), sd = sd(Jtutrend)/sqrt(.N))] # arctic
```




## Temperature-only model (Jtutrend, Jbetatrend, Horntrend)
```{r LME Jacard turnover temperature only}
i <- trends[, complete.cases(Jtutrend, REALM, temptrend)]

randef <- list(STUDY_ID = ~ abs(temptrend), rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modonlyTtrend.rds')){
  modonlyTtrend <- readRDS('temp/modonlyTtrend.rds')
} else {
  modonlyTtrend <- lme(Jtutrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modonlyTtrend, file = 'temp/modonlyTtrend.rds')
}

i2 <- trends[, complete.cases(Jbetatrend, REALM, temptrend)]
if(file.exists('temp/modonlyTtrendJbeta.rds')){
  modonlyTtrendJbeta <- readRDS('temp/modonlyTtrendJbeta.rds')
} else {
  modonlyTtrendJbeta <- lme(Jbetatrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i2,], method = 'REML', 
                   control=lmeControl(msMaxIter = 100, maxIter = 100))
  saveRDS(modonlyTtrendJbeta, file = 'temp/modonlyTtrendJbeta.rds')
}

i3 <- trends[, complete.cases(Horntrend, REALM, temptrend)]
if(file.exists('temp/modonlyTtrendHorn.rds')){
  modonlyTtrendHorn <- readRDS('temp/modonlyTtrendHorn.rds')
} else {
  modonlyTtrendHorn <- lme(Horntrend ~ abs(temptrend)*REALM,
                   random = randef, weights = varef, data = trends[i3,], method = 'REML')
  saveRDS(modonlyTtrendHorn, file = 'temp/modonlyTtrendHorn.rds')
}

summary(modonlyTtrend)
summary(modonlyTtrendJbeta)
summary(modonlyTtrendHorn)


```



### Plot the temp-only coefficients
```{r modonlyTtrendsimp coefs}
# make table of coefficients
coefs <- as.data.frame(summary(modonlyTtrend)$tTable)
coefs2 <- as.data.frame(summary(modonlyTtrendJbeta)$tTable)
coefs3 <- as.data.frame(summary(modonlyTtrendHorn)$tTable)
coefs$mod <- 'Jtu'
coefs2$mod <- 'Jbeta'
coefs3$mod <- 'Horn'
rows1 <- which(grepl('temptrend', rownames(coefs))) # extract temperature effect
cols <- c('Value', 'Std.Error', 'mod')
allcoefs <- rbind(coefs[rows1, cols], coefs2[rows1, cols], coefs3[rows1, cols])
allcoefs$Value[grepl('REALMMarine', rownames(allcoefs))] <- 
  allcoefs$Value[grepl('REALMMarine', rownames(allcoefs))] + 
  allcoefs$Value[!grepl('REALM', rownames(allcoefs))] # add intercept to marine effects
allcoefs$Value[grepl('REALMTerrestrial', rownames(allcoefs))] <- 
  allcoefs$Value[grepl('REALMTerrestrial', rownames(allcoefs))] + 
  allcoefs$Value[!grepl('REALM', rownames(allcoefs))] # add intercept to terrestrial effects

allcoefs$lCI <- allcoefs$Value - allcoefs$Std.Error # lower confidence interval
allcoefs$uCI <- allcoefs$Value + allcoefs$Std.Error
allcoefs$y <- c(3, 2, 1, 2.9, 1.9, 0.9, 2.8, 1.8, 0.8) # y-values
allcoefs$col <- c(rep('black', 3), rep('light grey', 3), rep('dark grey', 3))
allcoefs$realm <- rep(c('Freshwater', 'Marine', 'Terrestrial'), 3)

par(las = 1, mai = c(0.8, 2, 0.1, 0.1))
plot(0,0, col = 'white', xlim=c(-0.02, 0.85), ylim = c(0.7,3), 
     yaxt='n', xlab = 'Turnover per |째C/yr|', ylab ='')
axis(2, at = 3:1, labels = c('Freshwater', 'Marine', 'Terrestrial'), cex.axis = 0.7)
abline(v = 0, col = 'grey')
for(i in 1:nrow(allcoefs)){
  with(allcoefs[i, ], points(Value, y, pch = 16, col = col))
  with(allcoefs[i, ], lines(x = c(lCI, uCI), y = c(y, y), col = col))
}
legend('bottomright', col = c('black', 'dark grey', 'light grey'), lwd = 1, pch = 16, 
       legend = c('Jaccard turnover', 'Jaccard total', 'Horn-Morisita'))

```

### Nicer plots of turnover vs. temperature data and model fit
Scatterplot, violin plots, and coefficient plots all together
```{r turnover vs. temperature big plot, fig.width=6, fig.height=4.5}
# on macbook: fig.width=3, fig.height=2.375, fig.retina=3, out.width=3, out.height=2.375
# on external monitor: fig.width=6, fig.height=4.5
trends[temptrend <= -0.5, temptrendtext := 'Cooling']
trends[abs(temptrend) <= 0.1, temptrendtext := 'Stable']
trends[temptrend >= 0.5, temptrendtext := 'Warming']

trends[abs(rarefyID_y) < 35, latzone := 'Subtropics']
trends[abs(rarefyID_y) >= 35 & abs(rarefyID_x) < 66.56339, latzone := 'Temperate'] 
trends[abs(rarefyID_y) >= 66.56339, latzone := 'Polar']
trends[, latzone := factor(latzone, levels = c('Subtropics', 'Temperate', 'Polar'))]

p1 <- ggplot(trends, aes(temptrend, Jtutrend, color = REALM, fill = REALM, size = nyrBT)) +
  geom_point(na.rm = TRUE, shape = 16, alpha = 0.1) + 
  geom_smooth(data=subset(trends, abs(temptrend) < 0.75), method = 'gam', formula = y ~ s(x, bs = "cs"), 
              na.rm = TRUE) +
  scale_color_brewer(palette="Set1", name = 'Realm') + 
  scale_fill_brewer(palette="Set1", name = 'Realm') + 
  labs(x = 'Temperature trend (째C/yr)', y = 'Jaccard turnover', tag = 'A') +
  scale_size_continuous(range = c(1, 8), breaks = c(2, 5, 20)) +
  guides(size = guide_legend(title = 'Years',
                             override.aes = list(linetype=0, fill = NA, alpha = 1))) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        legend.key.size = unit(0.5,"line"), 
        axis.text=element_text(size=8),
        axis.title=element_text(size=10))

p2 <- ggplot(trends[!is.na(temptrendtext), ], aes(temptrendtext, Jtutrend)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') +
  labs(x = '', y = 'Jaccard turnover', tag = 'B') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10))

p3 <- ggplot(trends[abs(temptrend) >= 0.5 & !is.na(latzone), ], aes(latzone, Jtutrend)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), fill = 'grey') + 
  labs(x = '', y = '', tag = 'C') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.key=element_blank(),
        axis.text=element_text(size=7),
        axis.title=element_text(size=10))

p4 <- ggplot(allcoefs, aes(Value, y, group = mod, color = mod)) +
  geom_errorbarh(aes(xmin = lCI, xmax = uCI, height = 0)) + 
  geom_point() + 
  labs(x = expression(atop('Temperature change effect', '(Turnover '~degree*'C'^'-1'*')')), y = '', tag = 'D') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_cartesian(xlim =c(0, 1)) +
  scale_y_continuous(name = '', breaks = c(1, 2, 3), labels = c('Terrestrial', 'Marine', 'Freshwater'))

grid.arrange(p1, p2, p3, p4, ncol = 3, layout_matrix = rbind(c(1,1,1), c(2,3,4)),
             heights=c(unit(0.66, "npc"), unit(0.34, "npc")))
```

## Full models
Try static covariates plus interactions of abs temperature trend with each covariate:

- realm
- speed
- mass
- lifespan
- average metabolic temperature
- consumer fraction
- endotherm fraction
- environmental temperature
- seasonality
- microclimates
- thermal bias
- NPP
- vegetation
- human footprint

Except for thermal bias: interact with temperature trend (not abs)

### Fit full models
#### Full model for Jaccard total
Try Bowler or Venter/Halpern human impact
```{r LME Jacard turnover temperature full}
# using Bowler for human impact
i <- trends[, complete.cases(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modTfull1.rds')){
  modTfull1 <- readRDS('temp/modTfull1.rds')
} else {
  modTfull1 <- lme(Jtutrend ~ temptrend_abs.sc*REALM +
                     temptrend_abs.sc*tsign + 
                     temptrend_abs.sc*tempave.sc +
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*lifespan.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*endothermfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend_abs.sc*thermal_bias.sc:tsign +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*veg.sc +
                     temptrend_abs.sc*human_bowler.sc:REALM2,
                   random = randef, weights = varef, data = trends[i,], method = 'REML')
  saveRDS(modTfull1, file = 'temp/modTfull1.rds')
}

# using Venter/Halpern for human impact
i1.2 <- trends[, complete.cases(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_footprint.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

if(file.exists('temp/modTfullfootprint.rds')){
  modTfullfootprint <- readRDS('temp/modTfullfootprint.rds')
} else {
  modTfullfootprint <- lme(Jtutrend ~ temptrend_abs.sc*REALM + 
                     temptrend_abs.sc*tsign + 
                     temptrend_abs.sc*tempave.sc +
                     temptrend_abs.sc*tempave_metab.sc + 
                     temptrend_abs.sc*seas.sc + 
                     temptrend_abs.sc*microclim.sc + 
                     temptrend_abs.sc*mass.sc + 
                     temptrend_abs.sc*speed.sc + 
                     temptrend_abs.sc*lifespan.sc + 
                     temptrend_abs.sc*consumerfrac.sc +
                     temptrend_abs.sc*endothermfrac.sc +
                     temptrend_abs.sc*nspp.sc +
                     temptrend_abs.sc*thermal_bias.sc:tsign +
                     temptrend_abs.sc*npp.sc +
                     temptrend_abs.sc*veg.sc +
                     temptrend_abs.sc*human_footprint.sc:REALM2,
                   random = randef, weights = varef, data = trends[i1.2,], method = 'REML')
  saveRDS(modTfullfootprint, file = 'temp/modTfullfootprint.rds')
}

summary(modTfull1)
summary(modTfullfootprint)


```

#### Full models for total and HM
```{r LME Jacard total and MH models, fig.width=10, fig.height=8}
i2 <- trends[, complete.cases(Jbetatrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
i3 <- trends[, complete.cases(Horntrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                             temptrend.sc, temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                             consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]

randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
varef <- varPower(-0.5, ~nyrBT)

# full models
if(file.exists('temp/modTfullJbeta.rds')){
  modTfullJbeta <- readRDS('temp/modTfullJbeta.rds')
} else {
  modTfullJbeta <- lme(Jbetatrend ~ temptrend_abs.sc*REALM + 
                         temptrend_abs.sc*tsign +
                         temptrend_abs.sc*tempave.sc +
                         temptrend_abs.sc*tempave_metab.sc + 
                         temptrend_abs.sc*seas.sc + 
                         temptrend_abs.sc*microclim.sc + 
                         temptrend_abs.sc*mass.sc + 
                         temptrend_abs.sc*speed.sc + 
                         temptrend_abs.sc*lifespan.sc + 
                         temptrend_abs.sc*consumerfrac.sc +
                         temptrend_abs.sc*endothermfrac.sc +
                         temptrend_abs.sc*nspp.sc +
                         temptrend_abs.sc*thermal_bias.sc:tsign +
                         temptrend_abs.sc*npp.sc +
                         temptrend_abs.sc*veg.sc +
                         temptrend_abs.sc*human_bowler.sc:REALM2,
                       random = randef, weights = varef, data = trends[i2,], method = 'REML')
  saveRDS(modTfullJbeta, file = 'temp/modTfullJbeta.rds')
}

if(file.exists('temp/modTfullHorn.rds')){
  modTfullHorn <- readRDS('temp/modTfullHorn.rds')
} else {
  modTfullHorn <- lme(Horntrend ~ temptrend_abs.sc*REALM + 
                        temptrend_abs.sc*tsign +
                        temptrend_abs.sc*tempave.sc +
                        temptrend_abs.sc*tempave_metab.sc + 
                        temptrend_abs.sc*seas.sc + 
                        temptrend_abs.sc*microclim.sc + 
                        temptrend_abs.sc*mass.sc + 
                        temptrend_abs.sc*speed.sc + 
                        temptrend_abs.sc*lifespan.sc + 
                        temptrend_abs.sc*consumerfrac.sc +
                        temptrend_abs.sc*endothermfrac.sc +
                        temptrend_abs.sc*nspp.sc +
                        temptrend_abs.sc*thermal_bias.sc:tsign +
                        temptrend_abs.sc*npp.sc +
                        temptrend_abs.sc*veg.sc +
                        temptrend_abs.sc*human_bowler.sc:REALM2,
                      random = randef, weights = varef, data = trends[i3,], method = 'REML')
  saveRDS(modTfullHorn, file = 'temp/modTfullHorn.rds')
}

summary(modTfullJbeta)
summary(modTfullHorn)
```


### Plots from the full models
#### Plot the coefficients
```{r plot fullTmod1, fig.height=8, fig.width=7}

coefs <- summary(modTfull1)$tTable
coefs2 <- summary(modTfullfootprint)$tTable
coefs3 <- summary(modTfullJbeta)$tTable
coefs4 <- summary(modTfullHorn)$tTable
varstoplot <- unique(c(rownames(coefs), rownames(coefs2), rownames(coefs3), rownames(coefs4)))
varstoplot <- varstoplot[which(!grepl('Intercept', varstoplot) | grepl(':', varstoplot))] # vars to plot

rows1_1 <- which(rownames(coefs) %in% varstoplot) # rows in coefs
rows1_2 <- which(rownames(coefs2) %in% varstoplot)
rows1_3 <- which(rownames(coefs3) %in% varstoplot)
rows1_4 <- which(rownames(coefs4) %in% varstoplot)
xlims <- range(c(coefs[rows1_1,1] - coefs[rows1_1,2], coefs[rows1_1,1] + coefs[rows1_1,2], 
                  coefs2[rows1_2,1] - coefs2[rows1_2,2], coefs2[rows1_2,1] + coefs2[rows1_2,2], 
                  coefs3[rows1_3,1] - coefs3[rows1_3,2], coefs3[rows1_3,1] + coefs3[rows1_3,2],
                  coefs4[rows1_4,1] - coefs4[rows1_4,2], coefs4[rows1_4,1] + coefs4[rows1_4,2]))

cols <- c('light grey', 'light grey', 'dark grey', 'black') # for Jtu Bowler, Jtu footprint, Jbeta and Horn models, respectively
pchs <- c(16, 15, 16, 16)
offs <- c(0.3, 0.1, -0.1, -0.3) # offset vertically for each model



par(las = 1, mai = c(0.5, 3, 0.1, 0.1))

plot(0,0, col = 'white', xlim = xlims, ylim = c(1,length(varstoplot)), yaxt='n', xlab = '', ylab ='')
axis(2, at = length(varstoplot):1, labels = varstoplot, cex.axis = 0.7)
abline(v = 0, col = 'grey', lty = 2)
abline(h = 1:length(varstoplot), col = 'grey', lty = 3)
for(i in 1:length(varstoplot)){
  if(varstoplot[i] %in% rownames(coefs)){
    x = coefs[rownames(coefs) == varstoplot[i], 1]
    se = coefs[rownames(coefs) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[1], pch = pchs[1], col = cols[1])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[1], length(varstoplot) + 1 - i + offs[1]), col = cols[1])
  }
  if(varstoplot[i] %in% rownames(coefs2)){
    x = coefs2[rownames(coefs2) == varstoplot[i], 1]
    se = coefs2[rownames(coefs2) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[2], pch = pchs[2], col = cols[2])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[2], length(varstoplot) + 1 - i + offs[2]), col = cols[2])
  }
  if(varstoplot[i] %in% rownames(coefs3)){
    x = coefs3[rownames(coefs3) == varstoplot[i], 1]
    se = coefs3[rownames(coefs3) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[3], pch = pchs[3], col = cols[3])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[3], length(varstoplot) + 1 - i + offs[3]), col = cols[3])
  }
  if(varstoplot[i] %in% rownames(coefs4)){
    x = coefs4[rownames(coefs4) == varstoplot[i], 1]
    se = coefs4[rownames(coefs4) == varstoplot[i], 2]
    points(x, length(varstoplot) + 1 - i + offs[4], pch = pchs[4], col = cols[4])
    lines(x = c(x-se, x+se), y = c(length(varstoplot) + 1 - i + offs[4], length(varstoplot) + 1 - i + offs[4]), col = cols[4])
  }
}
```


Jaccard turnover in grey (Bowler human footprint circles, Venter/Halpern in squares)
Jaccard total dark grey
Horn-Morisita in black


#### Plot interactions (Jaccard turnover)
```{r interaction plots modTfull1, fig.height = 13, fig.width = 9}
# set up the interactions to plot
ints <- data.frame(vars = c('tsign', 'tempave', 'tempave_metab', 'seas', 'microclim', 'mass', 'speed', 'lifespan', 'consumerfrac', 'endothermfrac', 'nspp', 'thermal_bias', 'npp', 'veg', 'human_bowler', 'human_bowler'),
           min = c(1, -10, 10, 0.1, 0, 0, 0, 0.3, 0, 0, 0.3, -10, 1.9, 0, 0, 0), 
           max = c(2, 30, 40, 16, 6, 8, 2, 2, 1, 1, 2.6, 10, 3.7, 1, 9, 9),
           log = c(F, F, F, F, F, T, T, T, F, F, T, F, T, F, F, F),
           len = c(2, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100, 100),
           discrete = c(T, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F),
           REALM = c(rep('Freshwater', 15), 'Marine'),
           REALM2 = c(rep('TerrFresh', 15), 'Marine'),
           stringsAsFactors = FALSE)
basetab <- data.frame(tempave.sc = 0, tempave_metab.sc = 0, 
                      seas.sc = 0, microclim.sc = 0, mass.sc = 0, 
                      speed.sc = 0, lifespan.sc = 0, endothermfrac.sc = 0, 
                      nspp.sc = 0, thermal_bias.sc = 0, npp.sc = 0, human_bowler.sc = 0, veg.sc = 0,
                      consumerfrac.sc = 0,
                      nyrBT = 20, STUDY_ID = 127L, rarefyID = '127_514668')

# make the data frames for each interaction to plot                
for(j in 1:nrow(ints)){
  # set up a grid of temperature trends and the interacting variable
  if(ints$log[j]) intvars <- list(temptrend = seq(-1.5, 1.5, length.out = 100), 
                                  new = 10^seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                   var = ints$vars[j])
  if(!ints$log[j]) intvars <- list(temptrend = seq(-1.5, 1.5, length.out = 100), 
                                   new = seq(ints$min[j], ints$max[j], length.out = ints$len[j]),
                                   var = ints$vars[j])
  names(intvars) <- c('temptrend', ints$vars[j], 'var')
  thisdat <- expand.grid(intvars)
  
  # scale the interacting variable
  cent <- attr(trends[[paste0(ints$var[j], '.sc')]], 'scaled:center')
  scl <- attr(trends[[paste0(ints$var[j], '.sc')]], 'scaled:scale')
  if(!is.null(cent) & !is.null(scl)){
    if(ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (log(thisdat[[ints$var[j]]]) - cent)/scl
    if(!ints$log[j]) thisdat[[paste0(ints$var[j], '.sc')]] <- (thisdat[[ints$var[j]]] - cent)/scl
  }

  # merge with the rest of the columns
  if(ints$var[j] != 'tsign') colnamestouse <- setdiff(colnames(basetab), paste0(ints$var[j], '.sc'))
  if(ints$var[j] == 'tsign') colnamestouse <- setdiff(colnames(basetab), ints$var[j])
  thisdat <- cbind(thisdat, basetab[, colnamestouse])

  # add realm
  thisdat$REALM <- ints$REALM[j]
  thisdat$REALM2 <- ints$REALM2[j]
  
  # merge with the previous iterations
  if(j == 1) newdat <- thisdat
  if(j > 1){
    colstoadd <- setdiff(colnames(thisdat), colnames(newdat))
    for(toadd in colstoadd){
      newdat[[toadd]] <- NA
    }
    
    colstoadd2 <- setdiff(colnames(newdat), colnames(thisdat))
    for(toadd in colstoadd2){
      thisdat[[toadd]] <- NA
    }
    
    newdat <- rbind(newdat, thisdat)
  } 
}

# character so that new levels can be added
newdat$REALM <- as.character(newdat$REALM)
newdat$REALM2 <- as.character(newdat$REALM2)

# add extra rows so that all factor levels are represented (for predict.lme to work)
newdat <- rbind(newdat[1:4, ], newdat)
newdat$REALM[1:4] <- c('Marine', 'Marine', 'Terrestrial', 'Terrestrial')
newdat$REALM2[1:4] <- c('Marine', 'Marine', 'TerrFresh', 'TerrFresh')
newdat$temptrend[1:4] <- c(-1, 1, -1, 1)

# trim to at least some temperature change (so that tsign is -1 or 1)
newdat <- newdat[newdat$temptrend != 0,]

# scale the temperature vars
newdat$temptrend.sc <- newdat$temptrend/attr(trends$temptrend.sc, 'scaled:scale') 
newdat$temptrend_abs <- abs(newdat$temptrend)
newdat$temptrend_abs.sc <- (newdat$temptrend_abs)/attr(trends$temptrend_abs.sc, 'scaled:scale')
newdat$tsign <- factor(sign(newdat$temptrend))

# make predictions
newdat$preds <- predict(object = modTfull1, newdata = newdat, level = 0)

#remove the extra rows
newdat <- newdat[5:nrow(newdat), ]

# prep the plots
intplots <- vector('list', length(intstoplot))
for(j in 1:length(intplots)){
  subs <- newdat$var == ints$vars[j] & newdat$temptrend > 0 # select warming side
  xvar <- 'temptrend_abs'
  title <- ints$vars[j]
  if(ints$vars[j] %in% c('tsign')){
    subs <- newdat$var == ints$vars[j]
  } 
  if(ints$vars[j] %in% c('thermal_bias')){
    subs <- newdat$var == ints$vars[j]
    xvar <- 'temptrend'
  } 
  if(ints$vars[j] %in% c('human_bowler')){
    subs <- newdat$var == ints$vars[j] & newdat$temptrend > 0 & newdat$REALM2 == ints$REALM2[j]
    title <- paste0('human:', ints$REALM2[j])
  } 

  thisplot <- ggplot(newdat[subs, ], 
                     aes_string(x = xvar, y = 'preds', 
                                group = ints$vars[j], 
                                color = ints$vars[j])) +
    geom_line() +
    coord_cartesian(ylim = c(0, 1)) +
    theme(plot.margin = unit(c(0.5,0,0.5,0), 'cm')) +
    labs(title = title)
  if(ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'log')
  }
  if(!ints$log[j] & !ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_distiller(palette = "YlGnBu", trans = 'identity')
  }
  if(ints$discrete[j]){
    intplots[[j]] <- thisplot + scale_color_brewer(palette = "Dark2")
  }
}

#grid.arrange(grobs = intplots, '+', theme(plot.margin = unit(c(0,0,0,0), 'cm'))), ncol=2)
#do.call('grid.arrange', c(intplots, ncol = 2))
grid.arrange(grobs = intplots, ncol = 3)

```
Positive thermal bias means that the species in the assemblage have warmer thermal niches than the environment (and vice versa). 

#### Plot residuals against each predictor (Jaccard turnover)
```{r resids modTfull1, fig.height = 10, fig.width=10}
resids <- resid(modTfull1)
preds <- getData(modTfull1)
col = '#00000033'
cex = 0.5
par(mfrow = c(5,4))
boxplot(resids ~ preds$REALM, cex = cex, col = col)
plot(preds$temptrend_abs.sc, resids, cex = cex, col = col)
plot(preds$tsign, resids, cex = cex, col = col)
plot(preds$tempave.sc, resids, cex = cex, col = col)
plot(preds$tempave_metab.sc, resids, cex = cex, col = col)
plot(preds$seas.sc, resids, cex = cex, col = col)
plot(preds$microclim.sc, resids, cex = cex, col = col)
plot(preds$mass.sc, resids, cex = cex, col = col)
plot(preds$speed.sc, resids, cex = cex, col = col)
plot(preds$lifespan.sc, resids, cex = cex, col = col)
plot(preds$consumerfrac.sc, resids, cex = cex, col = col)
plot(preds$endothermfrac.sc, resids, cex = cex, col = col)
plot(preds$nspp.sc, resids, cex = cex, col = col)
plot(preds$thermal_bias.sc, resids, cex = cex, col = col)
plot(preds$npp.sc, resids, cex = cex, col = col)
plot(preds$veg.sc, resids, cex = cex, col = col)
plot(preds$human_bowler.sc, resids, cex = cex, col = col)
```

### Remove each term from the full model
#### Jaccard turnover
```{r term deletion from modTfull1}

if(file.exists('output/aics_from_full.csv')){
  aicsfromfull <- read.csv('output/aics_from_full.csv')

} else {
  i <- trends[, complete.cases(Jtutrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                               temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                               consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
  
  randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
  varef <- varPower(-0.5, ~nyrBT)
  
  terms <- c('temptrend_abs.sc*REALM', 
             'temptrend_abs.sc*tsign',
             'temptrend_abs.sc*tempave.sc',
             'temptrend_abs.sc*tempave_metab.sc',
             'temptrend_abs.sc*seas.sc',
             'temptrend_abs.sc*microclim.sc',
             'temptrend_abs.sc*mass.sc',
             'temptrend_abs.sc*speed.sc', 
             'temptrend_abs.sc*lifespan.sc', 
             'temptrend_abs.sc*consumerfrac.sc',
             'temptrend_abs.sc*endothermfrac.sc',
             'temptrend_abs.sc*nspp.sc',
             'temptrend_abs.sc*thermal_bias.sc:tsign',
             'temptrend_abs.sc*npp.sc',
             'temptrend_abs.sc*veg.sc',
             'temptrend_abs.sc*human_bowler.sc:REALM2')
  modTdrops <- vector('list', length(terms)+2)
  names(modTdrops) <- c('full', '-temptrend_abs.sc', paste0('-', terms))
  
  # fit full model with ML for model comparison
  modTdrops[[1]] <- lme(formula(paste0('Jtutrend ~ ', paste(terms, collapse = ' + '))),
                        random = randef, weights = varef, data = trends[i,], method = 'ML')
  
  # w/out temptrend
  modTdrops[[2]] <- lme(formula(paste0('Jtutrend ~ ', paste(gsub('temptrend_abs.sc\\*', '', terms), collapse = ' + '))),
                        random = list(STUDY_ID = ~ 1, rarefyID = ~1), weights = varef, data = trends[i,], method = 'ML')
  
  for(j in 1:length(terms)){
    print(j)
    modTdrops[[j+2]] <- lme(formula(paste0('Jtutrend ~ ', paste(terms[-j], collapse = ' + '))),
                            random = randef, weights = varef, data = trends[i,], method = 'ML')
  }
  
  aics <- sapply(modTdrops, AIC)
  
  aicsfromfull <- data.frame(mod = names(aics), dAIC_Jtu = aics - aics[1])
  
  write.csv(aicsfromfull, file = 'output/aics_from_full.csv', row.names = FALSE)
}

aicsfromfull
```


#### Jbeta and Horn models
```{r term deletion from Jbeta and Horn models}

if(file.exists('output/aics_from_full.csv')){
  aicsfromfull <- read.csv('output/aics_from_full.csv')
  
  if(all(c('dAIC_Jbeta', 'dAIC_Horn') %in% colnames(aicsfromfull))){
    runJbetaHorn <- FALSE
  } else {
    runJbetaHorn <- TRUE
  }
} else {
  runJbetaHorn <- TRUE
}

if(runJbetaHorn){
  i <- trends[, complete.cases(Jbetatrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                               temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                               consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
  i2 <- trends[, complete.cases(Horntrend, REALM, tempave.sc, tempave_metab.sc, seas.sc, microclim.sc, 
                               temptrend_abs.sc, mass.sc, speed.sc, lifespan.sc, 
                               consumerfrac.sc, endothermfrac.sc, nspp.sc, thermal_bias.sc, npp.sc, veg.sc, human_bowler.sc)]
  
  randef <- list(STUDY_ID = ~ temptrend_abs.sc, rarefyID = ~1)
  varef <- varPower(-0.5, ~nyrBT)
  
  terms <- c('temptrend_abs.sc*REALM', 
             'temptrend_abs.sc*tsign',
             'temptrend_abs.sc*tempave.sc',
             'temptrend_abs.sc*tempave_metab.sc',
             'temptrend_abs.sc*seas.sc',
             'temptrend_abs.sc*microclim.sc',
             'temptrend_abs.sc*mass.sc',
             'temptrend_abs.sc*speed.sc', 
             'temptrend_abs.sc*lifespan.sc', 
             'temptrend_abs.sc*consumerfrac.sc',
             'temptrend_abs.sc*endothermfrac.sc',
             'temptrend_abs.sc*nspp.sc',
             'temptrend_abs.sc*thermal_bias.sc:tsign',
             'temptrend_abs.sc*npp.sc',
             'temptrend_abs.sc*veg.sc',
             'temptrend_abs.sc*human_bowler.sc:REALM2')
  
  modTJbetadrops <- vector('list', length(terms)+2)
  modTHorndrops <- vector('list', length(terms)+2)

  names(modTJbetadrops) <- c('full', '-temptrend_abs.sc', paste0('-', terms))
  names(modTHorndrops) <- c('full', '-temptrend_abs.sc', paste0('-', terms))
  
  # fit full model with ML for model comparison
  modTJbetadrops[[1]] <- lme(formula(paste0('Jbetatrend ~ ', paste(terms, collapse = ' + '))),
                        random = randef, weights = varef, data = trends[i,], method = 'ML')
  modTHorndrops[[1]] <- lme(formula(paste0('Horntrend ~ ', paste(terms, collapse = ' + '))),
                        random = randef, weights = varef, data = trends[i2,], method = 'ML')
  
  # w/out temptrend
  modTJbetadrops[[2]] <- lme(formula(paste0('Jbetatrend ~ ', paste(gsub('temptrend_abs.sc\\*', '', terms), collapse = ' + '))),
                        random = list(STUDY_ID = ~ 1, rarefyID = ~1), weights = varef, data = trends[i,], method = 'ML')
  modTHorndrops[[2]] <- lme(formula(paste0('Horntrend ~ ', paste(gsub('temptrend_abs.sc\\*', '', terms), collapse = ' + '))),
                        random = list(STUDY_ID = ~ 1, rarefyID = ~1), weights = varef, data = trends[i2,], method = 'ML')
  
  for(j in 1:length(terms)){
    print(j)
    modTJbetadrops[[j+2]] <- lme(formula(paste0('Jbetatrend ~ ', paste(terms[-j], collapse = ' + '))),
                                 random = randef, weights = varef, data = trends[i,], method = 'ML', 
                                 control = lmeContrl(returnObject = TRUE)) # return fitted object even if a convergence error
    modTHorndrops[[j+2]] <- lme(formula(paste0('Horntrend ~ ', paste(terms[-j], collapse = ' + '))),
                            random = randef, weights = varef, data = trends[i2,], method = 'ML')
  }
  
  aicsJbeta <- sapply(modTJbetadrops, AIC)
  aicsHorn <- sapply(modTHorndrops, AIC)

  if(exists('aicsfromfull')){
    aicsfromfull$dAIC_Jbeta <- aicsJbeta - aicsJbeta[1]
    aicsfromfull$dAIC_Horn <- aicsHorn - aicsHorn[1]
  } else {
    aicsfromfull <- data.frame(mod = names(aics), 
                               dAIC_Jbeta = aicsJbeta - aicsJbeta[1],
                               dAIC_Horn <- aicsHorn - aicsHorn[1])
  }
  
  write.csv(aicsfromfull, file = 'output/aics_from_full.csv', row.names = FALSE)
}

aicsfromfull
```

#### Plot deltaAICs for all 3 models
```{r plot dAICs}
# transform for a plot
aicsfromfulllong <- reshape(aicsfromfull, direction = 'long',
                            varying = c('dAIC_Jtu', 'dAIC_Jbeta', 'dAIC_Horn'),
                            v.names = 'dAIC',
                            idvar = 'mod',
                            timevar = 'type',
                            times = c('Jtu', 'Jbeta', 'Horn'))

trans = function(x) sign(x)*sqrt(abs(x))
aicsfromfulllong$dAIC_tr <- trans(aicsfromfulllong$dAIC)

# plot
xlims <- range(aicsfromfulllong$dAIC_tr)
xticks <- c(-10, 0, 10, 100, 1000, 10000)
par(mai = c(0.5, 3, 0.1, 0.1))
with(aicsfromfulllong[aicsfromfulllong$type == 'Jtu',], plot(dAIC_tr, nrow(aicsfromfull):1, 
                                                           col = 'light grey', xlim = xlims, yaxt = 'n', ylab = '', xaxt = 'n'))
with(aicsfromfulllong[aicsfromfulllong$type == 'Jbeta',], points(dAIC_tr, nrow(aicsfromfull):1 - 0.1, col = 'dark grey'))
with(aicsfromfulllong[aicsfromfulllong$type == 'Horn',], points(dAIC_tr, nrow(aicsfromfull):1 - 0.2, col = 'black'))
axis(2, at = nrow(aicsfromfull):1, labels = aicsfromfull$mod, las = 1, cex.axis = 0.7)
axis(1, at = trans(xticks), labels = xticks, cex.axis = 0.5)
abline(v = 0, lty =2, col = 'grey')
```
Light grey is for Jaccard turnover, dark grey is for Jaccard total, black is for Morisita-Horn


### Plot interaction coefficients from all full models
```{r plot coefs for all models, fig.height=5, fig.width=6, out.height=5, out.width=6, fig.retina=2}
# fig.width = 3, fig.height = 5, out.height=2.5, out.width=3, fig.retina =3 for macbook screen
# double that for external monitor
coefs <- as.data.table(summary(modTfull1)$tTable)
coefs2 <- as.data.table(summary(modTfullJbeta)$tTable)
coefs3 <- as.data.table(summary(modTfullHorn)$tTable)

coefs$mod <- 'Jtu'
coefs2$mod <- 'Jbeta'
coefs3$mod <- 'Horn'

coefs$var <- rownames(summary(modTfull1)$tTable)
coefs2$var <- rownames(summary(modTfullJbeta)$tTable)
coefs3$var <- rownames(summary(modTfullHorn)$tTable)

# extract temperature effects and bind model coefs together
cols <- c('var', 'Value', 'Std.Error', 'mod')

allcoefsfull <- rbind(coefs[grep('temptrend|REALM', var), ..cols], 
                      coefs2[grep('temptrend|REALM', var), ..cols], 
                      coefs3[grep('temptrend|REALM', var), ..cols])
allcoefsfull$var[allcoefsfull$var == 'temptrend_abs.sc'] <- 'temptrend_abs.sc:REALMFreshwater'

# add average temperature effect (across realms) to realm-specific temperature effects
meantempeffect <- allcoefsfull[var == 'temptrend_abs.sc:REALMFreshwater', mean(Value), by = mod]$V1
allcoefsfull[var == 'temptrend_abs.sc:REALMMarine', Value := Value + meantempeffect]
allcoefsfull[var == 'temptrend_abs.sc:REALMTerrestrial', Value := Value + meantempeffect]

# remove non-temperature effects
allcoefsfull <- allcoefsfull[grepl(':', allcoefsfull$var) & grepl('temptrend', allcoefsfull$var), ]

# add info for plotting
allcoefsfull$lCI <- allcoefsfull$Value - allcoefsfull$Std.Error # lower confidence interval
allcoefsfull$uCI <- allcoefsfull$Value + allcoefsfull$Std.Error
nvar <- nrow(allcoefsfull)/3
allcoefsfull$y <- 1:nvar + rep(c(0, 0.1, 0.2), c(nvar, nvar, nvar)) # y-values

# clean up some variable names for nicer plotting
allcoefsfull$varname <- gsub('temptrend_abs.sc:|temptrend.sc:', '', allcoefsfull$var)
allcoefsfull$varname <- gsub('REALM|REALM2', '', allcoefsfull$varname)
allcoefsfull$varname <- gsub('.sc', '', allcoefsfull$varname)
allcoefsfull$varname <- gsub('_bowler', '', allcoefsfull$varname)
allcoefsfull$varname <- gsub('tsign1', 'warming', allcoefsfull$varname)
allcoefsfull$varname <- gsub('tsign-1', 'cooling', allcoefsfull$varname)

# set x-axes
xlims1 <- c(-0.01, 0.11) # for realms
xlims2 <- c(-0.007, 0.017) # for traits
xlims3 <- c(-0.01, 0.02) # for environment
xlims4 <- c(-0.016, 0.01) # for community
xlims5 <- c(-0.005, 0.0025) # for human

ddg <- 0.5 # vertical dodge for each model

# choose which variables are in which graph
set1 <- c('Terrestrial', 'Marine', 'Freshwater')
set2 <- c('mass', 'speed', 'lifespan', 'consumerfrac', 'endothermfrac', 'tempave_metab')
set3 <- c('seas', 'microclim', 'tempave')
set4 <- c('npp', 'nspp', 'tsign-1:thermal_bias', 'tsign1:thermal_bias')
set5 <- c('human_bowler:TerrFresh', 'human_bowler:Marine')

p1 <- ggplot(subset(allcoefsfull, varname %in% set1), 
                    aes(varname, Value, group = mod, color = mod)) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'light grey') +
  geom_errorbar(aes(ymin = lCI, ymax = uCI), width = 0, position = position_dodge(ddg)) + 
  geom_point(position = position_dodge(ddg)) + 
  labs(y = 'Temperature change effect', x = '', tag = 'A') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_flip(ylim = xlims1)

p2 <- ggplot(subset(allcoefsfull, varname %in% set2), 
                    aes(varname, Value, group = mod, color = mod)) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'light grey') +
  geom_errorbar(aes(ymin = lCI, ymax = uCI), width = 0, position = position_dodge(ddg)) + 
  geom_point(position = position_dodge(ddg)) + 
  labs(y = 'Interaction with temperature change effect', x = '', tag = 'B') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_flip(ylim = xlims2)

p3 <- ggplot(subset(allcoefsfull, varname %in% set3), 
                    aes(varname, Value, group = mod, color = mod)) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'light grey') +
  geom_errorbar(aes(ymin = lCI, ymax = uCI), width = 0, position = position_dodge(ddg)) + 
  geom_point(position = position_dodge(ddg)) + 
  labs(y = 'Interaction with temperature change effect', x = '', tag = 'C') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_flip(ylim = xlims3)

p4 <- ggplot(subset(allcoefsfull, varname %in% set4), 
                    aes(varname, Value, group = mod, color = mod)) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'light grey') +
  geom_errorbar(aes(ymin = lCI, ymax = uCI), width = 0, position = position_dodge(ddg)) + 
  geom_point(position = position_dodge(ddg)) + 
  labs(y = 'Interaction with temperature change effect', x = '', tag = 'D') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_flip(ylim = xlims4)

p5 <- ggplot(subset(allcoefsfull, varname %in% set5), 
                    aes(varname, Value, group = mod, color = mod)) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'light grey') +
  geom_errorbar(aes(ymin = lCI, ymax = uCI), width = 0, position = position_dodge(ddg)) + 
  geom_point(position = position_dodge(ddg)) + 
  labs(y = 'Interaction with temperature change effect', x = '', tag = 'E') +
  scale_color_grey() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position='none',
        axis.text=element_text(size=7),
        axis.title=element_text(size=7)) + 
  coord_flip(ylim = xlims5)

grid.arrange(p1, p2, p3, p4, p5, ncol = 2, layout_matrix = rbind(c(1,2), c(3,4), c(5, NA)))


```

